### 2021-03-11 paper/paper.tex
Update.

\documentclass{llncs}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[hidelinks]{hyperref}

\usepackage[links]{agda}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

\usepackage{newunicodechar}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
\newunicodechar{∘}{\ensuremath{\circ}}

\usepackage{cite}

\title{Kit Theory}

\begin{document}
  \maketitle

  \begin{abstract}
    Variable binding and substitution are a long-standing nuisance for mechanized
    metatheory of programming languages. A common approach is to represent
    variables as DeBruijn Indices, where substitution is defined in terms of
    variable renaming. If a language has $n$ different kinds of variables, and
    $m$ different kinds of terms, we potentially need to define both
    substitution and renaming for all $m\cdot n$ combinations and prove the
    theory of those operations both in isolation and with respect to each other.
    McBride introduced kits to define both substitution and renaming by a
    single term traversal for intrinsically typed terms.\cite{mcbride2005kits}

    We propose a technique to derive the implementation and metatheory of
    substitution and renaming for all $m\cdot n$ combinations of variables and terms
    from a single term traversal and a few simple lemmas.
    We use a slight generalization of McBride's kits but for extrinsic typing,
    which decouples type preservation from scope preservation and the
    theory of substitution. Hence, the technique is also suitable for
    non-standard typing relations like dependent types and type state.
    We provide a well-documented implementation of the technique as an Agda
    library and show its use in several case studies including System F.
  \end{abstract}

  % The following line will be replaced by make with an \input{} for each
  % .lagda.tex file:

  % SED MARKER

  % \bibliographystyle{ACM-Reference-Format}
  \bibliographystyle{abbrv}
  \bibliography{paper}

  \clearpage
  \appendix
\end{document}

### 2021-03-12 paper/Paper/Section2-Kits.lagda.tex
Customized agda.sty and started section 2.

\begin{code}[hide]
module Paper.Section2-Kits where

open import Data.List using (List; []; _∷_)

infix 4 _∋_ _⊢_
infixr 5 _⇒_ λx_
infixl 6 _·_
\end{code}

\newpage
\section{Kits}

McBride proposed the kit technique for intrinsically typed formalizations.

\begin{code}
data Type : Set where
  ∅    : Type
  _⇒_  : Type → Type → Type

variable
  t t₁ t₂  : Type
  Γ        : List Type

-- Well-typed DeBruijn variables
data _∋_ : List Type → Type → Set where
  zero  : (t ∷ Γ ∋ t)
  suc   : (Γ ∋ t₁) → (t₂ ∷ Γ ∋ t₁)

-- Well-typed terms
data _⊢_ : List Type → Type → Set where
  `_   : (Γ ∋ t) → (Γ ⊢ t)                    -- Variables
  λx_  : (t₁ ∷ Γ ⊢ t₂) → (Γ ⊢ t₁ ⇒ t₂)        -- Functions
  _·_  : (Γ ⊢ t₁ ⇒ t₂) → (Γ ⊢ t₁) → (Γ ⊢ t₂)  -- Applications

-- The application of the identity function to itself is a well-typed term.
example₁ : [] ⊢ ∅ ⇒ ∅
example₁ = (λx (` zero)) · (λx (` zero))
\end{code}

### 2021-03-12 paper/Paper/Section3-Kits.lagda.tex
Worked on paper.

\begin{code}[hide]
module Paper.Section3-Kits where

open import Paper.Section2-Intrinsic
open import Data.List using (List; []; _∷_)

infix  4  _∋_  _⊢_  _–[_]→_
infixr 5  _⇒_  λx_
infixl 6  _·_
\end{code}

\newpage
\section{Kits}

McBride proposed the kit technique for intrinsically typed formalizations.

\begin{code}
record Kit : Set₁ where
  infix 4 _◆_
  field
    _◆_  : List Type → Type → Set
    vr   : ∀ t → Γ ∋ t → Γ ◆ t
    tm   : ∀ t → Γ ◆ t → Γ ⊢ t
    wk   : ∀ {t'} t → Γ ◆ t → (t' ∷ Γ) ◆ t

_–[_]→_ : List Type → Kit → List Type → Set
Γ₁ –[ K ]→ Γ₂ = ∀ t → Γ₁ ∋ t → Γ₂ ◆ t where open Kit K

lift : (K : Kit) → Γ₁ –[ K ]→ Γ₂ → (t ∷ Γ₁) –[ K ]→ (t ∷ Γ₂)
lift K f _ zero    = vr _ zero     where open Kit K
lift K f _ (suc x) = wk _ (f _ x)  where open Kit K

kitᵣ : Kit
kitᵣ = record {  _◆_  =  _∋_
              ;  vr   = λ t x → x
              ;  tm   = λ t x → ` x
              ;  wk   = λ t x → suc x
              }

traverse : ∀ (K : Kit) → (Γ₁ –[ K ]→ Γ₂) → Γ₁ ⊢ t → Γ₂ ⊢ t
traverse K f (` x)      = Kit.tm K _ (f _ x)
traverse K f (λx e)     = λx (traverse K (lift K f) e)
traverse K f (e₁ · e₂)  = traverse K f e₁ · traverse K f e₂

kitₛ : Kit
kitₛ = record {  _◆_  = _⊢_
              ;  vr   = λ t x → ` x
              ;  tm   = λ t x → x
              ;  wk   = λ t x → traverse kitᵣ (Kit.wk kitᵣ) x
              }

rename : (Γ₁ –[ kitᵣ ]→ Γ₂) → Γ₁ ⊢ t → Γ₂ ⊢ t
rename ρ e = traverse kitᵣ ρ e

subst : (Γ₁ –[ kitₛ ]→ Γ₂) → Γ₁ ⊢ t → Γ₂ ⊢ t
subst σ e = traverse kitₛ σ e
\end{code}

### 2021-03-12 paper/Paper/Section3-Kits.lagda.tex
Worked on paper.

@@
     _◆_  : List Type → Type → Set
-    vr   : ∀ t → Γ ∋ t → Γ ◆ t
-    tm   : ∀ t → Γ ◆ t → Γ ⊢ t
-    wk   : ∀ {t'} t → Γ ◆ t → (t' ∷ Γ) ◆ t
+    vr   : ∀ t → (Γ ∋ t) → (Γ ◆ t)
+    tm   : ∀ t → (Γ ◆ t) → (Γ ⊢ t)
+    wk   : ∀ {t'} t → (Γ ◆ t) → (t' ∷ Γ ◆ t)
 
 _–[_]→_ : List Type → Kit → List Type → Set
-Γ₁ –[ K ]→ Γ₂ = ∀ t → Γ₁ ∋ t → Γ₂ ◆ t where open Kit K
+Γ₁ –[ K ]→ Γ₂ = ∀ t → (Γ₁ ∋ t) → (Γ₂ ◆ t) where open Kit K
 
-lift : (K : Kit) → Γ₁ –[ K ]→ Γ₂ → (t ∷ Γ₁) –[ K ]→ (t ∷ Γ₂)
+lift : (K : Kit) → (Γ₁ –[ K ]→ Γ₂) → (t ∷ Γ₁ –[ K ]→ t ∷ Γ₂)
 lift K f _ zero     = vr _ zero     where open Kit K
@@
 
-traverse : ∀ (K : Kit) → (Γ₁ –[ K ]→ Γ₂) → Γ₁ ⊢ t → Γ₂ ⊢ t
+traverse : ∀ (K : Kit) → (Γ₁ –[ K ]→ Γ₂) → (Γ₁ ⊢ t) → (Γ₂ ⊢ t)
 traverse K f (` x)      = Kit.tm K _ (f _ x)
@@
 
-rename : (Γ₁ –[ kitᵣ ]→ Γ₂) → Γ₁ ⊢ t → Γ₂ ⊢ t
+rename : (Γ₁ –[ kitᵣ ]→ Γ₂) → (Γ₁ ⊢ t) → (Γ₂ ⊢ t)
 rename ρ e = traverse kitᵣ ρ e
 
-subst : (Γ₁ –[ kitₛ ]→ Γ₂) → Γ₁ ⊢ t → Γ₂ ⊢ t
+subst : (Γ₁ –[ kitₛ ]→ Γ₂) → (Γ₁ ⊢ t) → (Γ₂ ⊢ t)
 subst σ e = traverse kitₛ σ e

### 2021-03-16 paper/paper.tex
Customized agda.sty and started section 2.
Worked on paper.
Switched to acmart/sigplan document class for tyde/cpp submission.
Fixed progress.

@@
-\documentclass{llncs}
+\documentclass[sigplan, screen]{acmart}
+% \documentclass[sigplan, screen, anonymous, review, authordraft]{acmart}
+
+\bibliographystyle{ACM-Reference-Format}
+
+\title{Kit Theory}
+
+\author{Hannes Saffrich}
+\orcid{0000-0002-1825-0097}                 %% \orcid is optional
+\affiliation{
+  % \department{Department1}                %% \department is recommended
+  \institution{University of Freiburg}      %% \institution is required
+  \country{Germany}                         %% \country is recommended
+}
+\email{saffrich@informatik.uni-freiburg.de} %% \email is recommended
+
+\author{Peter Thiemann}
+\orcid{0000-0002-9000-1239}                 %% \orcid is optional
+\affiliation{
+  % \department{Department2a}               %% \department is recommended
+  \institution{University of Freiburg}      %% \institution is required
+  \country{Germany}                         %% \country is recommended
+}
+\email{thiemann@informatik.uni-freiburg.de} %% \email is recommended
 
@@
 
-\usepackage[hidelinks]{hyperref}
-
 \usepackage[links]{agda}
 
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
+% \usepackage{amsmath}
+% \usepackage{amsfonts}
+% \usepackage{amssymb}
 
@@
 \newunicodechar{∘}{\ensuremath{\circ}}
+\newunicodechar{∅}{\ensuremath{\varnothing}}
+\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
+\newunicodechar{∋}{\ensuremath{\ni}}
+\newunicodechar{∷}{\ensuremath{\mathop{::}}}
+\newunicodechar{⊢}{\ensuremath{\vdash}}
+\newunicodechar{∀}{\ensuremath{\forall}}
+\newunicodechar{₁}{\ensuremath{{}_1}}
+\newunicodechar{₂}{\ensuremath{{}_2}}
+\newunicodechar{ᵣ}{\ensuremath{{}_{\textsf{r}}}}
+\newunicodechar{ₛ}{\ensuremath{{}_{\textsf{s}}}}
+\newunicodechar{→}{\ensuremath{\to}}
+\newunicodechar{◆}{\ensuremath{\diamond}}
+% \newunicodechar{Γ}{\ensuremath{\texttt{\Gamma}}}
+% \newunicodechar{λ}{\ensuremath{\texttt{\lambda}}}
 
-\usepackage{cite}
+% \usepackage{cite}
+% \usepackage{DejaVuSansMono}
+\usepackage{alphabeta} % Make greek letter macros also work in text-mode.
 
-\title{Kit Theory}
+\usepackage{microtype}
+\DisableLigatures[-]{encoding=T1}
 
 \begin{document}
-  \maketitle
-
   \begin{abstract}
@@
     theory of substitution. Hence, the technique is also suitable for
-    non-standard typing relations like dependent types and type state.
+    non-standard typing relations like dependent types, type state, or resource
+    environments.
     We provide a well-documented implementation of the technique as an Agda
@@
   \end{abstract}
+
+  \maketitle
 
@@
   % \bibliographystyle{ACM-Reference-Format}
-  \bibliographystyle{abbrv}
   \bibliography{paper}

### 2021-03-24 paper/paper.tex
Changed abstract.

@@
     theory of those operations both in isolation and with respect to each other.
-    McBride introduced kits to define both substitution and renaming by a
-    single term traversal for intrinsically typed terms.\cite{mcbride2005kits}
+
+    % McBride introduced kits to define both substitution and renaming by a
+    % single term traversal for intrinsically typed terms.\cite{mcbride2005kits}
 
@@
     substitution and renaming for all $m\cdot n$ combinations of variables and terms
-    from a single term traversal and a few simple lemmas.
-    We use a slight generalization of McBride's kits but for extrinsic typing,
-    which decouples type preservation from scope preservation and the
-    theory of substitution. Hence, the technique is also suitable for
-    non-standard typing relations like dependent types, type state, or resource
-    environments.
+    from a single term traversal and three simple lemmas for extrinsically typed
+    but intrinsically scoped terms.
+    We demonstrate the technique by formalizing type soundness of System F with
+    very parallel substitutions, which replace both type and term variables at
+    once, enabling a single substitution-preserves-typing lemma.
     We provide a well-documented implementation of the technique as an Agda
-    library and show its use in several case studies including System F.
+    library.
+
+    Since the technique is based on extrinsic typing, the metatheory of renaming
+    and substitution is decoupled from type preservation.
+    This has the downside, that type preservation of renaming and substitution
+    are not for free, but also the upside, that the library can be directly
+    instantiated for object languages with non-standard typing relations like
+    dependent types, type state, or resource environments.
   \end{abstract}

### 2021-03-24 paper/paper.tex
Changed abstract.

@@
     $m$ different kinds of terms, we potentially need to define both
-    substitution and renaming for all $m\cdot n$ combinations and prove the
+    renaming and substitution for all $m\cdot n$ combinations and prove the
     theory of those operations both in isolation and with respect to each other.
@@
     We propose a technique to derive the implementation and metatheory of
-    substitution and renaming for all $m\cdot n$ combinations of variables and terms
+    renaming and substitution for all $m\cdot n$ combinations of variables and terms
     from a single term traversal and three simple lemmas for extrinsically typed

### 2022-11-14 presentation/Presentation/Section3-Kits.lagda.tex
Update.

--- 2021-03-12 paper/Paper/Section3-Kits.lagda.tex
@@
 \begin{code}[hide]
-module Paper.Section3-Kits where
+module Presentation.Section3-Kits where
 
@@
 
-infix  4  _∋_  _⊢_  _–[_]→_
-infixr 5  _⇒_  λx_
-infixl 6  _·_
+infix  4  _–[_]→_
 \end{code}
 
-\newpage
 \section{Kits}
 
+\begin{frame}
 McBride proposed the kit technique for intrinsically typed formalizations.
+\end{frame}
 
+\begin{frame}
 \begin{code}
@@
     wk   : ∀ {t'} t → (Γ ◆ t) → (t' ∷ Γ ◆ t)
+\end{code}
+\end{frame}
 
+\begin{frame}
+\begin{code}
 _–[_]→_ : List Type → Kit → List Type → Set
@@
 lift K f _ (suc x)  = wk _ (f _ x)  where open Kit K
+\end{code}
+\end{frame}
 
+\begin{frame}
+\begin{code}
 kitᵣ : Kit
@@
               }
+\end{code}
+\end{frame}
 
+\begin{frame}
+\begin{code}
 traverse : ∀ (K : Kit) → (Γ₁ –[ K ]→ Γ₂) → (Γ₁ ⊢ t) → (Γ₂ ⊢ t)
@@
 traverse K f (e₁ · e₂)  = traverse K f e₁ · traverse K f e₂
+\end{code}
+\end{frame}
 
+\begin{frame}
+\begin{code}
 kitₛ : Kit
@@
               }
+\end{code}
+\end{frame}
 
+\begin{frame}
+\begin{code}
 rename : (Γ₁ –[ kitᵣ ]→ Γ₂) → (Γ₁ ⊢ t) → (Γ₂ ⊢ t)
@@
 \end{code}
+\end{frame}

### 2023-06-03 paper/Paper/Definitions.lagda.tex
Added new paper directory.

\begin{code}
module Paper.Definitions where

open import Paper.Kits
open import Data.List using (List; []; _∷_)
open import Data.Product using (_,_)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; sym; cong; cong₂; subst; module ≡-Reasoning)
open ≡-Reasoning

-- Fixities --------------------------------------------------------------------

infix   3  _⊢_  _↪_  _⊢_∶_
infixr  5  λx_  Λα_  ∀[α∶_]_
infixr  6  _⇒_
infixl  6  _·_  _∙_
infix   7  `_

-- Modes -----------------------------------------------------------------------

data Mode : ModeTy → Set where
  𝕖  : Mode Var    -- Expressions
  𝕥  : Mode Var    -- Types
  𝕜  : Mode NoVar  -- Kinds

variable
  mt                         : ModeTy
  m m₁ m₂ m₃ m' m₁' m₂' m₃'  : Mode mt
  µ µ₁ µ₂ µ₃ µ' µ₁' µ₂' µ₃'  : List (Mode Var)
  x y z                      : µ ∋ m

-- Syntax ----------------------------------------------------------------------

-- Expressions, Types, and Kinds
data _⊢_ : List (Mode Var) → Mode mt → Set where
  `_        : ∀ {m} → µ ∋ m → µ ⊢ m        -- Term and Type Variables
  λx_       : (𝕖 ∷ µ) ⊢ 𝕖 → µ ⊢ 𝕖          -- Term Abstraction
  Λα_       : (𝕥 ∷ µ) ⊢ 𝕖 → µ ⊢ 𝕖          -- Type Abstraction
  ∀[α∶_]_   : µ ⊢ 𝕜 → (𝕥 ∷ µ) ⊢ 𝕥 → µ ⊢ 𝕥  -- Universal Quantification
  _·_       : µ ⊢ 𝕖 → µ ⊢ 𝕖 → µ ⊢ 𝕖        -- Term Application
  _∙_       : µ ⊢ 𝕖 → µ ⊢ 𝕥 → µ ⊢ 𝕖        -- Type Application
  _⇒_       : µ ⊢ 𝕥 → µ ⊢ 𝕥 → µ ⊢ 𝕥        -- Function Type
  ★         : µ ⊢ 𝕜                        -- Type Kind

variable
  e e₁ e₂ e₃ e' e₁' e₂'  : µ ⊢ 𝕖
  t t₁ t₂ t₃ t' t₁' t₂'  : µ ⊢ 𝕥
  k k₁ k₂ k₃ k' k₁' k₂'  : µ ⊢ 𝕜
  E E₁ E₂ E₃ E' E₁' E₂'  : µ ⊢ m

-- Substitution & Lemmas -------------------------------------------------------

terms : Terms
terms = record
  { Mode         = Mode
  ; _⊢_          = _⊢_
  ; `_           = `_
  ; `-injective  = λ { refl → refl }
  }

open Terms terms hiding (Mode; _⊢_; `_)

_⋯_ :
  ∀ {_∋/⊢_ : Scoped} ⦃ K : Kit _∋/⊢_ ⦄ {µ₁ µ₂} 
  → µ₁ ⊢ m → µ₁ –[ K ]→ µ₂ → µ₂ ⊢ m
(` x)           ⋯ ϕ = `/id (ϕ _ x)
(λx t)          ⋯ ϕ = λx (t ⋯ (ϕ ↑ 𝕖))
(Λα t)          ⋯ ϕ = Λα (t ⋯ (ϕ ↑ 𝕥))
(∀[α∶ t₁ ] t₂)  ⋯ ϕ = ∀[α∶ t₁ ⋯ ϕ ] (t₂ ⋯ (ϕ ↑ 𝕥))
(t₁ · t₂)       ⋯ ϕ = (t₁ ⋯ ϕ) · (t₂ ⋯ ϕ)
(t₁ ∙ t₂)       ⋯ ϕ = (t₁ ⋯ ϕ) ∙ (t₂ ⋯ ϕ)
(t₁ ⇒ t₂)       ⋯ ϕ = (t₁ ⋯ ϕ) ⇒ (t₂ ⋯ ϕ)
★               ⋯ ϕ = ★

⋯-id :
  ∀ {_∋/⊢_ : Scoped} ⦃ K : Kit _∋/⊢_ ⦄ {µ} (t : µ ⊢ m)
  → t ⋯ id ⦃ K ⦄ ≡ t
⋯-id ⦃ K ⦄ (` x)     = id/`/id ⦃ K ⦄ x
⋯-id (λx t)          = cong λx_ (t ⋯ (id ↑ 𝕖) ≡⟨ cong (t ⋯_) (~-ext id↑~id) ⟩
                                 t ⋯ id       ≡⟨ ⋯-id t ⟩
                                 t            ∎)
⋯-id (Λα t)          = cong Λα_ (t ⋯ (id ↑ 𝕥) ≡⟨ cong (t ⋯_) (~-ext id↑~id) ⟩
                                 t ⋯ id       ≡⟨ ⋯-id t ⟩
                                 t            ∎)
⋯-id (∀[α∶ t₁ ] t₂)  = cong₂ ∀[α∶_]_ (⋯-id t₁)
                                     (t₂ ⋯ (id ↑ 𝕥) ≡⟨ cong (t₂ ⋯_) (~-ext id↑~id) ⟩
                                      t₂ ⋯ id       ≡⟨ ⋯-id t₂ ⟩
                                      t₂            ∎)
⋯-id (t₁ · t₂)       = cong₂ _·_ (⋯-id t₁) (⋯-id t₂)
⋯-id (t₁ ∙ t₂)       = cong₂ _∙_ (⋯-id t₁) (⋯-id t₂)
⋯-id (t₁ ⇒ t₂)       = cong₂ _⇒_ (⋯-id t₁) (⋯-id t₂)
⋯-id ★               = refl

traversal : Traversal
traversal = record
  { _⋯_   = _⋯_
  ; ⋯-var = λ x ϕ → refl
  ; ⋯-id  = ⋯-id
  }

open Traversal traversal hiding (_⋯_; ⋯-id)

⋯-assoc :
  ∀ {_∋/⊢_ _∋/⊢₁_ _∋/⊢₂_ : Scoped}
    ⦃ K₁ : Kit _∋/⊢₁_ ⦄ ⦃ K₂ : Kit _∋/⊢₂_ ⦄ ⦃ K : Kit _∋/⊢_ ⦄
    ⦃ W₁ : WkKit K₁ ⦄ ⦃ C : ComposeKit K₁ K₂ K ⦄
    {µ₁ µ₂ µ₃}
    (t : µ₁ ⊢ m) (ϕ₁ : µ₁ –[ K₁ ]→ µ₂) (ϕ₂ : µ₂ –[ K₂ ]→ µ₃)
  → (t ⋯ ϕ₁) ⋯ ϕ₂ ≡ t ⋯ (ϕ₁ ·ₘ ϕ₂)
⋯-assoc (` x)          ϕ₁ ϕ₂ = sym (&/⋯-⋯ (ϕ₁ _ x) ϕ₂)
⋯-assoc (λx t)         ϕ₁ ϕ₂ = cong λx_ ((t ⋯ (ϕ₁ ↑ 𝕖)) ⋯ (ϕ₂ ↑ 𝕖)  ≡⟨ ⋯-assoc t (ϕ₁ ↑ 𝕖) (ϕ₂ ↑ 𝕖) ⟩
                                         t ⋯ ((ϕ₁ ↑ 𝕖) ·ₘ (ϕ₂ ↑ 𝕖)) ≡⟨ cong (t ⋯_) (sym (~-ext (dist-↑-· 𝕖 ϕ₁ ϕ₂))) ⟩
                                         t ⋯ ((ϕ₁ ·ₘ ϕ₂) ↑ 𝕖)       ∎)
⋯-assoc (Λα t)         ϕ₁ ϕ₂ = cong Λα_ ((t ⋯ (ϕ₁ ↑ 𝕥)) ⋯ (ϕ₂ ↑ 𝕥)  ≡⟨ ⋯-assoc t (ϕ₁ ↑ 𝕥) (ϕ₂ ↑ 𝕥) ⟩
                                         t ⋯ ((ϕ₁ ↑ 𝕥) ·ₘ (ϕ₂ ↑ 𝕥)) ≡⟨ cong (t ⋯_) (sym (~-ext (dist-↑-· 𝕥 ϕ₁ ϕ₂))) ⟩
                                         t ⋯ ((ϕ₁ ·ₘ ϕ₂) ↑ 𝕥)       ∎)
⋯-assoc (∀[α∶ t₁ ] t₂) ϕ₁ ϕ₂ = cong₂ ∀[α∶_]_ (⋯-assoc t₁ ϕ₁ ϕ₂)
                                        ((t₂ ⋯ (ϕ₁ ↑ 𝕥)) ⋯ (ϕ₂ ↑ 𝕥)  ≡⟨ ⋯-assoc t₂ (ϕ₁ ↑ 𝕥) (ϕ₂ ↑ 𝕥) ⟩
                                         t₂ ⋯ ((ϕ₁ ↑ 𝕥) ·ₘ (ϕ₂ ↑ 𝕥)) ≡⟨ cong (t₂ ⋯_) (sym (~-ext (dist-↑-· 𝕥 ϕ₁ ϕ₂))) ⟩
                                         t₂ ⋯ ((ϕ₁ ·ₘ ϕ₂) ↑ 𝕥)       ∎)
⋯-assoc (t₁ · t₂)      ϕ₁ ϕ₂ = cong₂ _·_ (⋯-assoc t₁ ϕ₁ ϕ₂) (⋯-assoc t₂ ϕ₁ ϕ₂)
⋯-assoc (t₁ ∙ t₂)      ϕ₁ ϕ₂ = cong₂ _∙_ (⋯-assoc t₁ ϕ₁ ϕ₂) (⋯-assoc t₂ ϕ₁ ϕ₂)
⋯-assoc (t₁ ⇒ t₂)      ϕ₁ ϕ₂ = cong₂ _⇒_ (⋯-assoc t₁ ϕ₁ ϕ₂) (⋯-assoc t₂ ϕ₁ ϕ₂)
⋯-assoc ★              ϕ₁ ϕ₂ = refl

compose-traversal : ComposeTraversal
compose-traversal = record { ⋯-assoc = ⋯-assoc }

open ComposeTraversal compose-traversal hiding (⋯-assoc)

-- Type System -----------------------------------------------------------------

types : Types
types = record { ↑ᵗ = λ { 𝕖 → _ , 𝕥 ; 𝕥 → _ , 𝕜 ; 𝕜 → _ , 𝕜 } }

open Types types

variable
  Γ Γ₁ Γ₂ Γ' Γ₁' Γ₂' : Ctx µ
  T T₁ T₂ T' T₁' T₂' : µ ∶⊢ m

data _⊢_∶_ : Ctx µ → µ ⊢ m → µ ∶⊢ m → Set where
  ⊢` : ∀ {x : µ ∋ m} {T : µ ∶⊢ m} →
    Γ ∋ x ∶ T →
    Γ ⊢ ` x ∶ T
  ⊢λ : ∀ {e : (𝕖 ∷ µ) ⊢ 𝕖} →
    (t₁ ∷ₜ Γ) ⊢ e ∶ (wk _ t₂) →
    Γ ⊢ λx e ∶ t₁ ⇒ t₂
  ⊢Λ :
    (k ∷ₜ Γ) ⊢ e ∶ t₂ →
    Γ ⊢ Λα e ∶ ∀[α∶ k ] t₂
  ⊢· :
    Γ ⊢ e₁ ∶ t₁ ⇒ t₂ →
    Γ ⊢ e₂ ∶ t₁ →
    Γ ⊢ e₁ · e₂ ∶ t₂
  ⊢∙ : {Γ : Ctx µ} →
    (k₂ ∷ₜ Γ) ⊢ t₁ ∶ k₁ →
    Γ ⊢ t₂ ∶ k₂ →
    Γ ⊢ e₁ ∶ ∀[α∶ k₂ ] t₁ →
    Γ ⊢ e₁ ∙ t₂ ∶ t₁ ⋯ ⦅ t₂ ⦆
  ⊢τ :
    Γ ⊢ t ∶ ★

typing : Typing
typing = record { _⊢_∶_ = _⊢_∶_ ; ⊢` = ⊢` }

open Typing typing hiding (_⊢_∶_; ⊢`)

_⊢⋯_ :
  ∀ {_∋/⊢_ : Scoped} ⦃ K : Kit _∋/⊢_ ⦄ ⦃ W : WkKit K ⦄
    ⦃ C₁ : ComposeKit K Kᵣ K ⦄ ⦃ C₂ : ComposeKit K K K ⦄
    ⦃ C₃ : ComposeKit K Kₛ Kₛ ⦄
    ⦃ TK : TypingKit K W C₁ C₂ ⦄
    {µ₁ µ₂ mt} {Γ₁ : Ctx µ₁} {Γ₂ : Ctx µ₂} {m : Mode mt}
    {e : µ₁ ⊢ m} {t : µ₁ ∶⊢ m} {ϕ : µ₁ –[ K ]→ µ₂} →
  Γ₁ ⊢ e ∶ t →
  Γ₂ ∋*/⊢*[ TK ] ϕ ∶ Γ₁ →
  Γ₂ ⊢ e ⋯ ϕ ∶ t ⋯ ϕ
⊢` ⊢x ⊢⋯ ⊢ϕ = ⊢`/id (⊢ϕ _ _ ⊢x)
⊢λ {t₂ = t₂} ⊢e ⊢⋯ ⊢ϕ = ⊢λ (subst (_ ⊢ _ ∶_) (sym (⋯-↑-wk t₂ _ _)) (⊢e ⊢⋯ (⊢ϕ ∋↑/⊢↑ _)))
⊢Λ ⊢e ⊢⋯ ⊢ϕ = ⊢Λ (⊢e ⊢⋯ (⊢ϕ ∋↑/⊢↑ _))
⊢· ⊢e₁ ⊢e₂ ⊢⋯ ⊢ϕ = ⊢· (⊢e₁ ⊢⋯ ⊢ϕ) (⊢e₂ ⊢⋯ ⊢ϕ)
⊢∙ {t₁ = t₁} {t₂ = t₂} ⊢t₁ ⊢t₂ ⊢e₁ ⊢⋯ ⊢ϕ = subst (_ ⊢ _ ∶_) (sym (dist-↑-⦅⦆-⋯ t₁ t₂ _))
                                                 (⊢∙ (⊢t₁ ⊢⋯ (⊢ϕ ∋↑/⊢↑ _)) (⊢t₂ ⊢⋯ ⊢ϕ) (⊢e₁ ⊢⋯ ⊢ϕ))
⊢τ ⊢⋯ ⊢ϕ = ⊢τ

typing-traversal : TypingTraversal
typing-traversal = record { _⊢⋯_ = _⊢⋯_ }

open TypingTraversal typing-traversal hiding (_⊢⋯_)

-- Semantics -------------------------------------------------------------------

mutual
  data Neutral : µ ⊢ m → Set where
    `_  : ∀ (x : µ ∋ m) → Neutral (` x)
    _·_ : Neutral e₁ → Value e₂ → Neutral (e₁ · e₂)
    _∙t : Neutral e₁ → Neutral (e₁ ∙ t₂)

  data Value : µ ⊢ m → Set where
    λx_     : ∀ (e : (𝕖 ∷ µ) ⊢ 𝕖) → Value (λx e)
    Λα_     : ∀ (e : (𝕥 ∷ µ) ⊢ 𝕖) → Value (Λα e)
    neutral : Neutral e → Value e

data _↪_ : µ ⊢ m → µ ⊢ m → Set where
  β-λ : ∀ {e₂ : µ ⊢ 𝕖} →
    (λx e₁) · e₂ ↪ e₁ ⋯ ⦅ e₂ ⦆
  β-Λ : ∀ {t₂ : µ ⊢ 𝕥} →
    (Λα e₁) ∙ t₂ ↪ e₁ ⋯ ⦅ t₂ ⦆
  ξ-λ :
    e ↪ e' →
    λx e ↪ λx e'
  ξ-Λ :
    e ↪ e' →
    Λα e ↪ Λα e'
  ξ-·₁ :
    e₁ ↪ e₁' →
    e₁ · e₂ ↪ e₁' · e₂
  ξ-·₂ :
    e₂ ↪ e₂' →
    e₁ · e₂ ↪ e₁ · e₂'
  ξ-∙₁ :
    e₁ ↪ e₁' →
    e₁ ∙ t₂ ↪ e₁' ∙ t₂

-- Subject Reduction -----------------------------------------------------------

subject-reduction :
  Γ ⊢ e ∶ t →
  e ↪ e' →
  Γ ⊢ e' ∶ t
subject-reduction (⊢· {t₂ = t₂} (⊢λ ⊢e₁) ⊢e₂)   β-λ          = subst (_ ⊢ _ ∶_) (wk-cancels-⦅⦆-⋯ t₂ _) (⊢e₁ ⊢⋯ₛ ⊢⦅ ⊢e₂ ⦆ₛ)
subject-reduction (⊢∙ ⊢t₁ ⊢t₂ (⊢Λ ⊢e₁))         β-Λ          = ⊢e₁ ⊢⋯ₛ ⊢⦅ ⊢t₂ ⦆ₛ
subject-reduction (⊢λ ⊢e)                      (ξ-λ e↪e')    = ⊢λ (subject-reduction ⊢e e↪e')
subject-reduction (⊢Λ ⊢e)                      (ξ-Λ e↪e')    = ⊢Λ (subject-reduction ⊢e e↪e')
subject-reduction (⊢· ⊢e₁ ⊢e₂)                 (ξ-·₁ e₁↪e₁') = ⊢· (subject-reduction ⊢e₁ e₁↪e₁') ⊢e₂
subject-reduction (⊢· ⊢e₁ ⊢e₂)                 (ξ-·₂ e₂↪e₂') = ⊢· ⊢e₁ (subject-reduction ⊢e₂ e₂↪e₂')
subject-reduction (⊢∙ ⊢t₁ ⊢t₂ ⊢e₁)             (ξ-∙₁ e₁↪e₁') = ⊢∙ ⊢t₁ ⊢t₂ (subject-reduction ⊢e₁ e₁↪e₁')

\end{code}

### 2023-06-03 paper/Paper/Kits.lagda.tex
Added new paper directory.

\begin{code}
module Paper.Kits where

open import Data.Nat using (ℕ; zero; suc)
open import Data.List using (List; []; _∷_; drop)
open import Data.List.Membership.Propositional public using (_∈_)
open import Data.List.Relation.Unary.Any public using (here; there)
open import Data.List.Relation.Unary.All as All public using (All; []; _∷_)
open import Data.Product using (∃-syntax; _,_; proj₁; proj₂)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; sym; cong; subst; module ≡-Reasoning)
open ≡-Reasoning

infix  4  _∋_

_∋_ : ∀ {ℓ} {A : Set ℓ} → List A → A → Set _
xs ∋ x = x ∈ xs

pattern Z = here refl
pattern S x = there x

data ModeTy : Set where
  Var NoVar : ModeTy

record Terms : Set₁ where
  field
    Mode        : ModeTy → Set
    _⊢_         : ∀ {t} → List (Mode Var) → Mode t → Set
    `_          : ∀ {µ} {m : Mode Var} → µ ∋ m → µ ⊢ m
    `-injective : ∀ {µ m} {x₁ x₂ : µ ∋ m} → ` x₁ ≡ ` x₂ → x₁ ≡ x₂

  Scoped : Set₁
  Scoped = List (Mode Var) → Mode Var → Set

  variable _∋/⊢_ _∋/⊢₁_  _∋/⊢₂_ : Scoped

  record Kit (_∋/⊢_ : List (Mode Var) → Mode Var → Set) : Set where
    field
      id/`            : ∀ {µ m} → µ ∋ m → µ ∋/⊢ m
      `/id            : ∀ {µ m} → µ ∋/⊢ m → µ ⊢ m
      id/`/id         : ∀ {µ m} (x : µ ∋ m) → `/id (id/` x) ≡ ` x

      id/`-injective  : ∀ {µ m} {x₁ x₂ : µ ∋ m} → id/` x₁ ≡ id/` x₂ → x₁ ≡ x₂
      `/id-injective  : ∀ {µ m} {x/t₁ x/t₂ : µ ∋/⊢ m} → `/id x/t₁ ≡ `/id x/t₂ → x/t₁ ≡ x/t₂

      wk              : ∀ {µ m} m' → µ ∋/⊢ m → (m' ∷ µ) ∋/⊢ m
      wk-id/`         : ∀ {µ m} m' (x : µ ∋ m) → wk m' (id/` x) ≡ id/` (there x)

    Map : (µ₁ µ₂ : List (Mode Var)) → Set
    Map µ₁ µ₂ = ∀ m → µ₁ ∋ m → µ₂ ∋/⊢ m

    infixl  8  _&_
    _&_ : ∀ {µ₁ µ₂ m} → µ₁ ∋ m → Map µ₁ µ₂ → µ₂ ∋/⊢ m
    x & f = f _ x

    wkm : ∀ {µ₁ µ₂} m → Map µ₁ µ₂ → Map µ₁ (m ∷ µ₂)
    wkm m f _ x = wk m (f _ x)

    _∷ₘ_ :
      ∀ {µ₁ µ₂ m} → µ₂ ∋/⊢ m → Map µ₁ µ₂ → Map (m ∷ µ₁) µ₂
    (x/t ∷ₘ f) _ Z     = x/t
    (x/t ∷ₘ f) _ (S x) = f _ x

    _↑_ : ∀ {µ₁ µ₂} → Map µ₁ µ₂ → ∀ m → Map (m ∷ µ₁) (m ∷ µ₂)
    f ↑ m = id/` (here refl) ∷ₘ wkm m f

    id : ∀ {µ} → Map µ µ
    id m x = id/` x

    ⦅_⦆ : ∀ {µ m} → µ ∋/⊢ m → Map (m ∷ µ) µ
    ⦅ x/t ⦆ = x/t ∷ₘ id

    weaken : ∀ {µ} m → Map µ (m ∷ µ)
    weaken m = wkm m id

    _~_ :
      ∀ {µ₁ µ₂}
      → (ϕ₁ ϕ₂ : Map µ₁ µ₂)
      → Set
    _~_ {µ₁ = µ₁} ϕ₁ ϕ₂ = ∀ m (x : µ₁ ∋ m) → ϕ₁ m x ≡ ϕ₂ m x

    postulate
      ~-ext : 
        ∀ {µ₁ µ₂} {ϕ₁ ϕ₂ : Map µ₁ µ₂}
        → ϕ₁ ~ ϕ₂
        → ϕ₁ ≡ ϕ₂

    id↑~id : ∀ {µ m} → (id ↑ m) ~ id {m ∷ µ}
    id↑~id m Z     = refl
    id↑~id m (S x) =
      (id ↑ _) m (S x) ≡⟨⟩
      wk _ (id/` x)    ≡⟨ wk-id/` _ x ⟩
      id/` (S x)       ≡⟨⟩
      id m (S x)       ∎

  _∋/⊢[_]_ : ∀ {_∋/⊢_ : Scoped} → List (Mode Var) → Kit _∋/⊢_ → Mode Var → Set
  _∋/⊢[_]_ {_∋/⊢_} µ K m = µ ∋/⊢ m

  _–[_]→_ : ∀ {_∋/⊢_ : Scoped} → List (Mode Var) → Kit _∋/⊢_ → List (Mode Var) → Set
  µ₁ –[ K ]→ µ₂ = Map µ₁ µ₂ where open Kit K

  open Kit ⦃ … ⦄ public

  record Traversal : Set₁ where
    infixl   5  _⋯_

    field
      _⋯_   :
        ∀ {_∋/⊢_ : Scoped} ⦃ K : Kit _∋/⊢_ ⦄ {µ₁ µ₂ t} {m : Mode t} 
        → µ₁ ⊢ m → µ₁ –[ K ]→ µ₂ → µ₂ ⊢ m

      ⋯-var :
        ∀ {_∋/⊢_ : Scoped} ⦃ K : Kit _∋/⊢_ ⦄ {µ₁ µ₂} {m : Mode Var} 
          (x : µ₁ ∋ m) (ϕ : µ₁ –[ K ]→ µ₂)
        → (` x) ⋯ ϕ ≡ `/id (x & ϕ)

      ⋯-id :
        ∀ {_∋/⊢_ : Scoped} ⦃ K : Kit _∋/⊢_ ⦄ {µ t} {m : Mode t} 
          (t : µ ⊢ m)
        → t ⋯ id ⦃ K ⦄ ≡ t

    instance
      Kᵣ : Kit _∋_
      Kᵣ = record
        { id/`           = λ x → x
        ; `/id           = `_
        ; id/`/id        = λ x → refl
        ; id/`-injective = λ eq → eq
        ; `/id-injective = `-injective
        ; wk             = λ m' x → there x
        ; wk-id/`        = λ m' x → refl
        }

      Kₛ : Kit _⊢_
      Kₛ = record
        { id/`           = `_
        ; `/id           = λ t → t
        ; id/`/id        = λ x → refl
        ; id/`-injective = `-injective
        ; `/id-injective = λ eq → eq
        ; wk             = λ m' t → t ⋯ weaken m'
        ; wk-id/`        = λ m' x → (` x) ⋯ weaken m' ≡⟨ ⋯-var x (weaken m') ⟩
                                    ` (x & weaken m') ≡⟨⟩
                                    ` S x             ∎
        }

    open Kit Kᵣ public using () renaming 
      (Map to _→ᵣ_; wkm to wkmᵣ; _∷ₘ_ to _∷ᵣ_; _↑_ to _↑ᵣ_; id to idᵣ; ⦅_⦆ to ⦅_⦆ᵣ; weaken to weakenᵣ)
    open Kit Kₛ public using () renaming 
      (Map to _→ₛ_; wkm to wkmₛ; _∷ₘ_ to _∷ₛ_; _↑_ to _↑ₛ_; id to idₛ; ⦅_⦆ to ⦅_⦆ₛ; weaken to weakenₛ)

    -- Counterpart to wk-id/`
    record WkKit {_∋/⊢_ : Scoped} (K : Kit _∋/⊢_): Set₁ where
      private instance _ = K
      field
        wk-`/id :
          ∀ m {µ m'} (x/t : µ ∋/⊢ m')
          → `/id x/t ⋯ weakenᵣ m ≡ `/id (wk m x/t)

    instance
      Wᵣ : WkKit Kᵣ
      Wᵣ = record { wk-`/id = λ m x → ⋯-var x (weaken m) }

      Wₛ : WkKit Kₛ
      Wₛ = record { wk-`/id = λ m t → refl }

    open WkKit ⦃ … ⦄ public

    record ComposeKit {_∋/⊢_ _∋/⊢₁_ _∋/⊢₂_ : Scoped} (K₁ : Kit _∋/⊢₁_) (K₂ : Kit _∋/⊢₂_) (K₁⊔K₂ : Kit _∋/⊢_) : Set where
      infixl  8  _&/⋯_

      private instance _ = K₁; _ = K₂; _ = K₁⊔K₂

      field
        _&/⋯_ :
          ∀ {µ₁} {µ₂} {m} → µ₁ ∋/⊢[ K₁ ] m → µ₁ –[ K₂ ]→ µ₂ → µ₂ ∋/⊢[ K₁⊔K₂ ] m

        &/⋯-⋯ :
          ∀ {µ₁} {µ₂} {m} (x/t : µ₁ ∋/⊢[ K₁ ] m) (ϕ : µ₁ –[ K₂ ]→ µ₂) 
          → `/id (x/t &/⋯ ϕ) ≡ `/id x/t ⋯ ϕ

        &/⋯-wk-↑ :
          ∀ {µ₁} {µ₂} {m'} {m} (x/t : µ₁ ∋/⊢[ K₁ ] m) (ϕ : µ₁ –[ K₂ ]→ µ₂)
          → wk m' (x/t &/⋯ ϕ) ≡ wk m' x/t &/⋯ (ϕ ↑ m')

      _·ₘ_ : ∀ {µ₁ µ₂ µ₃} → µ₁ –[ K₁ ]→ µ₂ → µ₂ –[ K₂ ]→ µ₃ → µ₁ –[ K₁⊔K₂ ]→ µ₃
      (ϕ₁ ·ₘ ϕ₂) _ x = x & ϕ₁ &/⋯ ϕ₂

      &/⋯-& :
        ∀ {µ₁} {µ₂} {m} (x : µ₁ ∋ m) (ϕ : µ₁ –[ K₂ ]→ µ₂) 
        → `/id (id/` ⦃ K₁ ⦄ x &/⋯ ϕ) ≡ `/id (x & ϕ)
      &/⋯-& {µ₁} {µ₂} {m} x ϕ = 
          `/id (id/` x &/⋯ ϕ)      ≡⟨ &/⋯-⋯ (id/` x) ϕ ⟩
          `/id ⦃ K₁ ⦄ (id/` x) ⋯ ϕ ≡⟨ cong (_⋯ ϕ) (id/`/id ⦃ K₁ ⦄ x) ⟩
          ` x ⋯ ϕ                  ≡⟨ ⋯-var ⦃ K₂ ⦄ x ϕ ⟩
          `/id ⦃ K₂ ⦄  (x & ϕ)     ∎

      dist-↑-· :
        ∀ {µ₁} {µ₂} {µ₃} m
          (ϕ₁ : µ₁ –[ K₁ ]→ µ₂)
          (ϕ₂ : µ₂ –[ K₂ ]→ µ₃)
        → ((ϕ₁ ·ₘ ϕ₂) ↑ m) ~ ((ϕ₁ ↑ m) ·ₘ (ϕ₂ ↑ m))
      dist-↑-· m ϕ₁ ϕ₂ m₁ x@Z = `/id-injective (
        `/id ⦃ K₁⊔K₂ ⦄ (x & ((ϕ₁ ·ₘ ϕ₂) ↑ m))       ≡⟨⟩
        `/id ⦃ K₁⊔K₂ ⦄ (id/` Z)                     ≡⟨ id/`/id ⦃ K₁⊔K₂ ⦄ Z ⟩
        ` Z                                         ≡⟨ sym (id/`/id ⦃ K₂ ⦄ Z) ⟩
        `/id ⦃ K₂ ⦄ (id/` Z)                        ≡⟨⟩
        `/id ⦃ K₂ ⦄ (Z & (ϕ₂ ↑ m))                  ≡⟨ sym (&/⋯-& (id/` Z) (ϕ₂ ↑ m)) ⟩
        `/id ⦃ K₁⊔K₂ ⦄ (id/` Z &/⋯ (ϕ₂ ↑ m))        ≡⟨⟩
        `/id ⦃ K₁⊔K₂ ⦄ (x & (ϕ₁ ↑ m) &/⋯ (ϕ₂ ↑ m))  ≡⟨⟩
        `/id ⦃ K₁⊔K₂ ⦄ (x & ((ϕ₁ ↑ m) ·ₘ (ϕ₂ ↑ m))) ∎
        )
      dist-↑-· m ϕ₁ ϕ₂ m₁ x@(S y) = `/id-injective (
        `/id ⦃ K₁⊔K₂ ⦄ (x & ((ϕ₁ ·ₘ ϕ₂) ↑ m))       ≡⟨⟩
        `/id ⦃ K₁⊔K₂ ⦄ (wk _ (y & (ϕ₁ ·ₘ ϕ₂)))      ≡⟨⟩
        `/id ⦃ K₁⊔K₂ ⦄ (wk _ (y & ϕ₁ &/⋯ ϕ₂))       ≡⟨ cong `/id (&/⋯-wk-↑ (y & ϕ₁) ϕ₂) ⟩
        `/id ⦃ K₁⊔K₂ ⦄ (wk _ (y & ϕ₁) &/⋯ (ϕ₂ ↑ m)) ≡⟨⟩
        `/id ⦃ K₁⊔K₂ ⦄ (x & (ϕ₁ ↑ m) &/⋯ (ϕ₂ ↑ m))  ≡⟨⟩
        `/id ⦃ K₁⊔K₂ ⦄ (x & ((ϕ₁ ↑ m) ·ₘ (ϕ₂ ↑ m))) ∎
        )

    _·[_]_ :
      ∀ {_∋/⊢_ _∋/⊢₁_ _∋/⊢₂_ : Scoped} {K₁ : Kit _∋/⊢₁_} {K₂ : Kit _∋/⊢₂_} {K₁⊔K₂ : Kit _∋/⊢_} {µ₁ µ₂ µ₃}
      → µ₁ –[ K₁ ]→ µ₂ → ComposeKit K₁ K₂ K₁⊔K₂ → µ₂ –[ K₂ ]→ µ₃ → µ₁ –[ K₁⊔K₂ ]→ µ₃
    ϕ₁ ·[ C ] ϕ₂ = ϕ₁ ·ₘ ϕ₂ where open ComposeKit C

    open ComposeKit ⦃ … ⦄ public

    record ComposeTraversal : Set₁ where
      field
        ⋯-assoc :
          ∀ {_∋/⊢_ _∋/⊢₁_ _∋/⊢₂_ : Scoped}
            ⦃ K₁ : Kit _∋/⊢₁_ ⦄ ⦃ K₂ : Kit _∋/⊢₂_ ⦄ ⦃ K : Kit _∋/⊢_ ⦄
            ⦃ W₁ : WkKit K₁ ⦄ ⦃ C : ComposeKit K₁ K₂ K ⦄
            {µ₁ µ₂ µ₃ t} {m : Mode t}
            (t : µ₁ ⊢ m) (ϕ₁ : µ₁ –[ K₁ ]→ µ₂) (ϕ₂ : µ₂ –[ K₂ ]→ µ₃)
          → (t ⋯ ϕ₁) ⋯ ϕ₂ ≡ t ⋯ (ϕ₁ ·ₘ ϕ₂)

      ↑-wk :
        ∀ {_∋/⊢_ : Scoped} ⦃ K : Kit _∋/⊢_ ⦄ ⦃ W : WkKit K ⦄
           ⦃ C : ComposeKit K Kᵣ K ⦄ ⦃ C : ComposeKit Kᵣ K K ⦄ 
          {µ₁ µ₂} (ϕ : µ₁ –[ K ]→ µ₂) m
        → (ϕ ·ₘ weaken m) ~ (weaken m ·ₘ (ϕ ↑ m))
      ↑-wk {µ₁} {µ₂} ϕ m mx x = `/id-injective (
          `/id ((ϕ ·ₘ weakenᵣ m) mx x)       ≡⟨⟩
          `/id (x & ϕ &/⋯ weakenᵣ m)         ≡⟨ &/⋯-⋯ (x & ϕ) (weakenᵣ m) ⟩
          `/id (`/id (x & ϕ) ⋯ weakenᵣ m)    ≡⟨ wk-`/id m (x & ϕ) ⟩
          `/id (S x & (ϕ ↑ m))               ≡⟨ sym (&/⋯-& (S x) (ϕ ↑ m)) ⟩
          `/id (S x &/⋯ (ϕ ↑ m))             ≡⟨⟩
          `/id (x & weakenᵣ m &/⋯ (ϕ ↑ m))   ≡⟨⟩
          `/id ((weakenᵣ m ·ₘ (ϕ ↑ m)) mx x) ∎
        )

      ⋯-↑-wk :
        ∀ {_∋/⊢_ : Scoped} ⦃ K : Kit _∋/⊢_ ⦄ ⦃ W : WkKit K ⦄
          ⦃ C₁ : ComposeKit K Kᵣ K ⦄ ⦃ C₂ : ComposeKit Kᵣ K K ⦄ 
          {µ₁ µ₂ mt} {m : Mode mt} (t : µ₁ ⊢ m) (ϕ : µ₁ –[ K ]→ µ₂) m
        → t ⋯ ϕ ⋯ weakenᵣ m ≡ t ⋯ weakenᵣ m ⋯ (ϕ ↑ m)
      ⋯-↑-wk t ϕ m =
        t ⋯ ϕ ⋯ weakenᵣ m          ≡⟨ ⋯-assoc t ϕ (weakenᵣ m) ⟩
        t ⋯ (ϕ ·ₘ weakenᵣ m)       ≡⟨ cong (t ⋯_) (~-ext (↑-wk ϕ m)) ⟩
        t ⋯ (weakenᵣ m ·ₘ (ϕ ↑ m)) ≡⟨ sym (⋯-assoc t (weakenᵣ m) (ϕ ↑ m)) ⟩
        t ⋯ weakenᵣ m ⋯ (ϕ ↑ m)    ∎

      instance
        Cᵣ : ∀ {_∋/⊢_ : Scoped} ⦃ K₂ : Kit _∋/⊢_ ⦄
             → ComposeKit Kᵣ K₂ K₂
        Cᵣ = record
          { _&/⋯_ = _&_
          ; &/⋯-⋯ = λ x ϕ →
            `/id (x & ϕ) ≡⟨ sym (⋯-var x ϕ) ⟩
            (` x) ⋯ ϕ    ∎
          ; &/⋯-wk-↑ = λ x ϕ → refl
          }

        Cₛ : ∀ {_∋/⊢_ : Scoped} ⦃ K₂ : Kit _∋/⊢_ ⦄ ⦃ W₂ : WkKit K₂ ⦄ ⦃ C : ComposeKit K₂ Kᵣ K₂ ⦄
             → ComposeKit Kₛ K₂ Kₛ
        Cₛ ⦃ C = C ⦄ = record
          { _&/⋯_    = _⋯_
          ; &/⋯-⋯    = λ t ϕ → refl
          ; &/⋯-wk-↑ = λ t ϕ → ⋯-↑-wk t ϕ _
          }

      Cᵣᵣ : ComposeKit Kᵣ Kᵣ Kᵣ
      Cᵣₛ : ComposeKit Kᵣ Kₛ Kₛ
      Cₛᵣ : ComposeKit Kₛ Kᵣ Kₛ
      Cₛₛ : ComposeKit Kₛ Kᵣ Kₛ
      Cᵣᵣ = Cᵣ
      Cᵣₛ = Cᵣ
      Cₛᵣ = Cₛ
      Cₛₛ = Cₛ

      wk-cancels-⦅⦆ :
        ∀ {_∋/⊢_  : Scoped} ⦃ K : Kit _∋/⊢_ ⦄
          {µ m} (x/t : µ ∋/⊢[ K ] m) 
        → (weakenᵣ m ·[ Cᵣ ] ⦅ x/t ⦆) ~ id
      wk-cancels-⦅⦆ ⦃ K ⦄ x/t mx x = `/id-injective (
          `/id ⦃ K ⦄ (x & (weakenᵣ _ ·[ Cᵣ ] ⦅ x/t ⦆)) ≡⟨⟩
          `/id ⦃ K ⦄ (id/` (S x) &/⋯ ⦅ x/t ⦆)          ≡⟨ &/⋯-& ⦃ Cᵣ ⦃ K ⦄ ⦄ (S x) ⦅ x/t ⦆ ⟩
          `/id ⦃ K ⦄ (id/` x)                          ≡⟨⟩
          `/id ⦃ K ⦄ (x & id)                          ∎
        )

      wk-cancels-⦅⦆-⋯ :
        ∀ {_∋/⊢_  : Scoped} ⦃ K : Kit _∋/⊢_ ⦄
          {µ m mt} {m' : Mode mt} (t : µ ⊢ m') (x/t : µ ∋/⊢[ K ] m) 
        → t ⋯ weakenᵣ m ⋯ ⦅ x/t ⦆ ≡ t
      wk-cancels-⦅⦆-⋯ t x/t =
        t ⋯ weakenᵣ _ ⋯ ⦅ x/t ⦆    ≡⟨ ⋯-assoc t (weakenᵣ _) ⦅ x/t ⦆ ⟩
        t ⋯ (weakenᵣ _ ·ₘ ⦅ x/t ⦆) ≡⟨ cong (t ⋯_) (~-ext (wk-cancels-⦅⦆ x/t)) ⟩
        t ⋯ id                     ≡⟨ ⋯-id t ⟩
        t                          ∎

      dist-↑-⦅⦆ :
        ∀ {_∋/⊢_ _∋/⊢₁_ _∋/⊢₂_ : Scoped}
          ⦃ K₁ : Kit _∋/⊢₁_ ⦄ ⦃ K₂ : Kit _∋/⊢₂_ ⦄ ⦃ K : Kit _∋/⊢_ ⦄
          ⦃ W₂ : WkKit K₂ ⦄ ⦃ C₁ : ComposeKit K₁ K₂ K ⦄ ⦃ C₂ : ComposeKit K₂ K K ⦄
          {µ₁ µ₂ m} (x/t : µ₁ ∋/⊢[ K₁ ] m) (ϕ : µ₁ –[ K₂ ]→ µ₂) →
        (⦅ x/t ⦆ ·[ C₁ ] ϕ) ~ ((ϕ ↑ m) ·[ C₂ ] ⦅ (x/t &/⋯ ϕ) ⦆)
      dist-↑-⦅⦆ ⦃ K₁ ⦄ ⦃ K₂ ⦄ ⦃ K ⦄ ⦃ W₂ ⦄ ⦃ C₁ ⦄ ⦃ C₂ ⦄ {µ₁} {µ₂} {m} x/t ϕ mx x@Z = `/id-injective (
          `/id ⦃ K ⦄ (x & (⦅ x/t ⦆ ·[ C₁ ] ϕ))               ≡⟨⟩
          `/id ⦃ K ⦄ (x/t &/⋯ ϕ)                             ≡⟨⟩
          `/id ⦃ K ⦄ (Z & ⦅ (x/t &/⋯ ϕ) ⦆)                   ≡⟨ sym (&/⋯-& ⦃ C₂ ⦄ Z ⦅ (x/t &/⋯ ϕ) ⦆) ⟩
          `/id ⦃ K ⦄ (id/` ⦃ K₂ ⦄ Z &/⋯ ⦅ (x/t &/⋯ ϕ) ⦆)     ≡⟨⟩
          `/id ⦃ K ⦄ (x & ((ϕ ↑ m) ·[ C₂ ] ⦅ (x/t &/⋯ ϕ) ⦆)) ∎
        )
      dist-↑-⦅⦆ ⦃ K₁ ⦄ ⦃ K₂ ⦄ ⦃ K ⦄ ⦃ W₂ ⦄ ⦃ C₁ ⦄ ⦃ C₂ ⦄ {µ₁} {µ₂} {m} x/t ϕ mx x@(S y) = `/id-injective (
          `/id (x & (⦅ x/t ⦆ ·[ C₁ ] ϕ))               ≡⟨⟩
          `/id (id/` ⦃ K₁ ⦄ y &/⋯ ϕ)                   ≡⟨ &/⋯-& ⦃ C₁ ⦄ y ϕ ⟩
          `/id (y & ϕ)                                 ≡⟨ sym (wk-cancels-⦅⦆-⋯ (`/id (y & ϕ)) (x/t &/⋯ ϕ)) ⟩
          `/id (y & ϕ) ⋯ weakenᵣ m ⋯ ⦅ (x/t &/⋯ ϕ) ⦆   ≡⟨ cong (_⋯ ⦅ x/t &/⋯ ϕ ⦆) (wk-`/id m (y & ϕ)) ⟩
          `/id (wk m (y & ϕ)) ⋯ ⦅ (x/t &/⋯ ϕ) ⦆        ≡⟨ sym (&/⋯-⋯ (wk m (y & ϕ)) ⦅ (x/t &/⋯ ϕ) ⦆) ⟩
          `/id (wk m (y & ϕ) &/⋯ ⦅ (x/t &/⋯ ϕ) ⦆)      ≡⟨⟩
          `/id (x & ((ϕ ↑ m) ·[ C₂ ] ⦅ (x/t &/⋯ ϕ) ⦆)) ∎
        )

      dist-↑-⦅⦆-⋯ :
        ∀ {_∋/⊢_ _∋/⊢₁_ _∋/⊢₂_ : Scoped}
          ⦃ K₁ : Kit _∋/⊢₁_ ⦄ ⦃ K₂ : Kit _∋/⊢₂_ ⦄ ⦃ K : Kit _∋/⊢_ ⦄
          ⦃ W₁ : WkKit K₁ ⦄ ⦃ W₂ : WkKit K₂ ⦄ ⦃ C₁ : ComposeKit K₁ K₂ K ⦄ ⦃ C₂ : ComposeKit K₂ K K ⦄
          {µ₁ µ₂ m mt} {m' : Mode mt} (t : (m ∷ µ₁) ⊢ m')
          (x/t : µ₁ ∋/⊢[ K₁ ] m) (ϕ : µ₁ –[ K₂ ]→ µ₂) →
        t ⋯ ⦅ x/t ⦆ ⋯ ϕ ≡ t ⋯ (ϕ ↑ m) ⋯ ⦅ (x/t &/⋯ ϕ) ⦆
      dist-↑-⦅⦆-⋯ t x/t ϕ =
        t ⋯ ⦅ x/t ⦆ ⋯ ϕ                  ≡⟨ ⋯-assoc t ⦅ x/t ⦆ ϕ ⟩
        t ⋯ (⦅ x/t ⦆ ·ₘ ϕ)               ≡⟨ cong (t ⋯_) (~-ext (dist-↑-⦅⦆ x/t ϕ)) ⟩
        t ⋯ ((ϕ ↑ _) ·ₘ ⦅ (x/t &/⋯ ϕ) ⦆) ≡⟨ sym (⋯-assoc t (ϕ ↑ _) ⦅ x/t &/⋯ ϕ ⦆ ) ⟩
        t ⋯ (ϕ ↑ _) ⋯ ⦅ (x/t &/⋯ ϕ) ⦆    ∎

      record Types : Set₁ where
        field
          ↑ᵗ : ∀ {t} → Mode t → ∃[ t ] Mode t

        _∶⊢_ : ∀ {t} → List (Mode Var) → Mode t → Set
        µ ∶⊢ m = µ ⊢ proj₂ (↑ᵗ m)

        depth : ∀ {ℓ} {A : Set ℓ} {x : A} {xs : List A} → xs ∋ x → ℕ
        depth (here px) = zero
        depth (there x) = suc (depth x)

        -- We need to drop one extra using `suc`, because otherwise the types in a
        -- context are allowed to use themselves.
        drop-∈ : ∀ {ℓ} {A : Set ℓ} {x : A} {xs : List A} → xs ∋ x → List A → List A
        drop-∈ e xs = drop (suc (depth e)) xs

        Ctx : List (Mode Var) → Set
        Ctx µ = ∀ m → (x : µ ∋ m) → drop-∈ x µ ∶⊢ m

        -- []ₜ : Ctx []
        -- []ₜ _ ()

        _∷ₜ_ : ∀ {µ m} → µ ∶⊢ m → Ctx µ → Ctx (m ∷ µ)
        (t ∷ₜ Γ) _ (here refl) = t
        (t ∷ₜ Γ) _ (there x)   = Γ _ x

        wk-drop-∈ : ∀ {µ m t} {m' : Mode t} → (x : µ ∋ m) → drop-∈ x µ ⊢ m' → µ ⊢ m'
        wk-drop-∈ (here _)  t = t ⋯ weakenᵣ _
        wk-drop-∈ (there x) t = wk-drop-∈ x t ⋯ weakenᵣ _

        wk-telescope : ∀ {µ m} → Ctx µ → µ ∋ m → µ ∶⊢ m
        wk-telescope Γ x = wk-drop-∈ x (Γ _ x)

        infix   4  _∋_∶_
        _∋_∶_ : ∀ {µ m} → Ctx µ → µ ∋ m → µ ∶⊢ m → Set
        Γ ∋ x ∶ t = wk-telescope Γ x ≡ t

        record Typing : Set₁ where
          infix   4  _⊢_∶_
          field
            _⊢_∶_ : ∀ {µ t} {m : Mode t} → Ctx µ → µ ⊢ m → µ ∶⊢ m → Set

            ⊢` : ∀ {µ m} {Γ : Ctx µ} {x : µ ∋ m} {t} →
                Γ ∋ x ∶ t → Γ ⊢ ` x ∶ t

          record TypingKit {_∋/⊢_ : Scoped} (K : Kit _∋/⊢_) (W : WkKit K) (C₁ : ComposeKit K Kᵣ K) (C₂ : ComposeKit K K K) : Set₁ where
            infix   4  _∋/⊢_∶_  _∋*/⊢*_∶_
            infixl  6  _∋↑/⊢↑_

            private instance _ = K; _ = W; _ = C₁; _ = C₂

            field
              -- Variable/Term Typing
              _∋/⊢_∶_ : ∀ {µ m} → Ctx µ → µ ∋/⊢ m → µ ∶⊢ m → Set

              ∋/⊢∶-lookup : ∀ {µ} {Γ : Ctx µ} {m} (x : µ ∋ m)
                            → Γ ∋/⊢ id/` x ∶ wk-telescope Γ x

              -- Conditional Applications of Variable-Typing-Rule
              id/⊢` : ∀ {µ m} {x : µ ∋ m} {t : µ ∶⊢ m} {Γ : Ctx µ}
                      → Γ ∋ x ∶ t
                      →  Γ ∋/⊢ id/` x ∶ t
              ⊢`/id : ∀ {µ m} {e : µ ∋/⊢ m} {t : µ ∶⊢ m} {Γ : Ctx µ}
                      → Γ ∋/⊢ e ∶ t
                      → Γ ⊢ `/id e ∶ t

              -- Weakening preserveres variable/term typings.
              ∋wk/⊢wk  : ∀ {µ m m'} (Γ : Ctx µ) (t' : µ ∶⊢ m) (e : µ ∋/⊢ m') (t : µ ∶⊢ m')
                        → Γ ∋/⊢ e ∶ t
                        → (t' ∷ₜ Γ) ∋/⊢ wk _ e ∶ (t ⋯ weakenᵣ _)

            _∋*/⊢*_∶_ : ∀ {µ₁ µ₂} → Ctx µ₂ → µ₁ –[ K ]→ µ₂ → Ctx µ₁ → Set
            _∋*/⊢*_∶_ {µ₁} {µ₂} Γ₂ ϕ Γ₁ =
              ∀ {m₁} (x : µ₁ ∋ m₁) (t : µ₁ ∶⊢ m₁) (⊢x : Γ₁ ∋ x ∶ t)
              → Γ₂ ∋/⊢ (x & ϕ) ∶ (t ⋯ ϕ)

            _∋↑/⊢↑_ : ∀ {µ₁} {µ₂} {Γ₁ : Ctx µ₁} {Γ₂ : Ctx µ₂} {ϕ : µ₁ –[ K ]→ µ₂} {m} →
              Γ₂             ∋*/⊢* ϕ       ∶ Γ₁ →
              (t : µ₁ ∶⊢ m) →
              ((t ⋯ ϕ) ∷ₜ Γ₂) ∋*/⊢* (ϕ ↑ m) ∶ (t ∷ₜ Γ₁)
            _∋↑/⊢↑_ {µ₁} {µ₂} {Γ₁} {Γ₂} {ϕ} {m} ⊢ϕ t {mx} x@Z _ refl =
              subst (((t ⋯ ϕ) ∷ₜ Γ₂) ∋/⊢ (K Kit.& Z) (ϕ ↑ m) ∶_)
                    (t ⋯ ϕ ⋯ weakenᵣ m                  ≡⟨ ⋯-↑-wk t ϕ m ⟩
                     t ⋯ weakenᵣ m ⋯ (ϕ ↑ m)            ≡⟨⟩
                     wk-telescope (t ∷ₜ Γ₁) Z ⋯ (ϕ ↑ m) ∎)
                    (id/⊢` {x = here refl} {Γ = (t ⋯ ϕ) ∷ₜ Γ₂} refl)
            _∋↑/⊢↑_ {µ₁} {µ₂} {Γ₁} {Γ₂} {ϕ} {m} ⊢ϕ t {mx} x@(S y) _ refl =
              subst (((t ⋯ ϕ) ∷ₜ Γ₂) ∋/⊢ (K Kit.& S y) (ϕ ↑ m) ∶_)
                    (wk-telescope Γ₁ y ⋯ ϕ ⋯ weakenᵣ m       ≡⟨ ⋯-↑-wk _ ϕ m ⟩
                     wk-telescope Γ₁ y ⋯ weakenᵣ m ⋯ (ϕ ↑ m) ≡⟨⟩
                     wk-telescope (t ∷ₜ Γ₁) (S y) ⋯ (ϕ ↑ m)  ∎)
                    (∋wk/⊢wk _ _ _ _ (⊢ϕ y _ refl))

            ⊢⦅_⦆ : ∀ {m µ} {Γ : Ctx µ} {t : µ ∋/⊢ m} {T : µ ∶⊢ m}
              → Γ ∋/⊢ t ∶ T 
              → Γ ∋*/⊢* ⦅ t ⦆ ∶ (T ∷ₜ Γ)
            ⊢⦅_⦆ {m} {µ} {Γ} {t} {T} ⊢x/t x@Z _ refl =
              subst (Γ ∋/⊢ t ∶_)
                    (T                               ≡⟨ sym (wk-cancels-⦅⦆-⋯ T t) ⟩
                     T ⋯ weakenᵣ _ ⋯ ⦅ t ⦆           ≡⟨⟩
                     wk-telescope (T ∷ₜ Γ) Z ⋯ ⦅ t ⦆ ∎)
                    ⊢x/t
            ⊢⦅_⦆ {m} {µ} {Γ} {t} {T} ⊢x/t x@(S y) _ refl =
              subst (Γ ∋/⊢ id/` y ∶_)
                    (wk-telescope Γ y                     ≡⟨ sym (wk-cancels-⦅⦆-⋯ _ t) ⟩
                     wk-telescope Γ y ⋯ weakenᵣ _ ⋯ ⦅ t ⦆ ≡⟨⟩
                     wk-telescope (T ∷ₜ Γ) (S y) ⋯ ⦅ t ⦆  ∎)
                    (id/⊢` refl)

          open TypingKit ⦃ … ⦄ public

          infixl  5  _∋*/⊢*[_]_∶_
          _∋*/⊢*[_]_∶_ :
            ∀ {_∋/⊢_ : Scoped} {K : Kit _∋/⊢_}
              {W : WkKit K} {C₁ : ComposeKit K Kᵣ K} {C₂ : ComposeKit K K K}
              {µ₁ µ₂}
            → Ctx µ₂ → TypingKit K W C₁ C₂ → µ₁ –[ K ]→ µ₂ → Ctx µ₁ → Set
          Γ₂ ∋*/⊢*[ TK ] f ∶ Γ₁ = Γ₂ ∋*/⊢* f ∶ Γ₁ where instance _ = TK

          record TypingTraversal : Set₁ where
            infixl  5  _⊢⋯_  _⊢⋯ᵣ_  _⊢⋯ₛ_

            field
              _⊢⋯_ :
                ∀ {_∋/⊢_ : Scoped} ⦃ K : Kit _∋/⊢_ ⦄ ⦃ W : WkKit K ⦄
                  ⦃ C₁ : ComposeKit K Kᵣ K ⦄ ⦃ C₂ : ComposeKit K K K ⦄
                  ⦃ C₃ : ComposeKit K Kₛ Kₛ ⦄
                  ⦃ TK : TypingKit K W C₁ C₂ ⦄
                  {µ₁ µ₂ mt} {Γ₁ : Ctx µ₁} {Γ₂ : Ctx µ₂} {m : Mode mt}
                  {e : µ₁ ⊢ m} {t : µ₁ ∶⊢ m} {ϕ : µ₁ –[ K ]→ µ₂} →
                Γ₁ ⊢ e ∶ t →
                Γ₂ ∋*/⊢*[ TK ] ϕ ∶ Γ₁ →
                Γ₂ ⊢ e ⋯ ϕ ∶ t ⋯ ϕ

            instance
              TKᵣ : TypingKit Kᵣ Wᵣ Cᵣ Cᵣ
              TKᵣ = record
                { _∋/⊢_∶_     = _∋_∶_
                ; ∋/⊢∶-lookup = λ x → refl
                ; id/⊢`       = λ ⊢x → ⊢x
                ; ⊢`/id       = ⊢`
                ; ∋wk/⊢wk     = λ { Γ t' x t refl → refl }
                }

              TKₛ : TypingKit Kₛ Wₛ Cₛ Cₛ
              TKₛ = record
                { _∋/⊢_∶_     = _⊢_∶_
                ; ∋/⊢∶-lookup = λ x → ⊢` refl
                ; id/⊢`       = ⊢`
                ; ⊢`/id       = λ ⊢x → ⊢x
                ; ∋wk/⊢wk     = λ Γ t' e t ⊢e → ⊢e ⊢⋯ ∋wk/⊢wk Γ t'
                }

            open TypingKit TKᵣ public using () renaming (∋wk/⊢wk to ⊢wk; _∋*/⊢*_∶_ to _∋*_∶_; ⊢⦅_⦆ to ⊢⦅_⦆ᵣ)
            open TypingKit TKₛ public using () renaming (∋wk/⊢wk to ∋wk; _∋*/⊢*_∶_ to _⊢*_∶_; ⊢⦅_⦆ to ⊢⦅_⦆ₛ)

            -- Renaming preserves typing

            _⊢⋯ᵣ_ : ∀ {µ₁ µ₂ mt} {Γ₁ : Ctx µ₁} {Γ₂ : Ctx µ₂} {m : Mode mt}
                      {e : µ₁ ⊢ m} {t : µ₁ ∶⊢ m} {ρ : µ₁ →ᵣ µ₂} →
                    Γ₁ ⊢ e ∶ t →
                    Γ₂ ∋* ρ ∶ Γ₁ →
                    Γ₂ ⊢ e ⋯ ρ ∶ t ⋯ ρ
            _⊢⋯ᵣ_ = _⊢⋯_

            -- Substitution preserves typing

            _⊢⋯ₛ_ : ∀ {µ₁ µ₂ mt} {Γ₁ : Ctx µ₁} {Γ₂ : Ctx µ₂} {m : Mode mt}
                      {e : µ₁ ⊢ m} {t : µ₁ ∶⊢ m} {σ : µ₁ →ₛ µ₂} →
                    Γ₁ ⊢ e ∶ t →
                    Γ₂ ⊢* σ ∶ Γ₁ →
                    Γ₂ ⊢ e ⋯ σ ∶ t ⋯ σ
            _⊢⋯ₛ_ = _⊢⋯_

\end{code}

### 2023-06-03 paper_old/paper.tex
Added new paper directory.

--- 2021-03-24 paper/paper.tex
@@
 
+\usepackage{bbold} % lower case mathbb letters
+
 \usepackage{newunicodechar}
@@
 \newunicodechar{◆}{\ensuremath{\diamond}}
+\newunicodechar{⊤}{\ensuremath{\top}}
+\newunicodechar{★}{\ensuremath{\star}}
+\newunicodechar{ˣ}{\ensuremath{{}^{\textsf{x}}}}
+\newunicodechar{ᵅ}{\ensuremath{{}^{\textsf{\alpha}}}}
+\newunicodechar{ᵥ}{\ensuremath{{}_{\textsf{v}}}}
+\newunicodechar{ₜ}{\ensuremath{{}_{\textsf{t}}}}
+\newunicodechar{𝕖}{\ensuremath{\mathbb{e}}}
+\newunicodechar{𝕥}{\ensuremath{\mathbb{t}}}
+\newunicodechar{𝕜}{\ensuremath{\mathbb{k}}}
+\newunicodechar{∙}{\ensuremath{\bullet}}
+
 % \newunicodechar{Γ}{\ensuremath{\texttt{\Gamma}}}
@@
     but intrinsically scoped terms.
-    We demonstrate the technique by formalizing type soundness of System F with
+    We demonstrate this technique by formalizing type soundness of System F with
     very parallel substitutions, which replace both type and term variables at
@@
 
-    Since the technique is based on extrinsic typing, the metatheory of renaming
+    As the technique is based on extrinsic typing, the metatheory of renaming
     and substitution is decoupled from type preservation.

### 2023-06-03 paper/paper.tex
Update.

@@
-\documentclass[sigplan, screen]{acmart}
+% \documentclass[sigplan, screen]{acmart}
 % \documentclass[sigplan, screen, anonymous, review, authordraft]{acmart}
+% \documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[sigplan,10pt]{acmart}
 
@@
 \usepackage[T1]{fontenc}
+\usepackage{stmaryrd}
 
@@
 % \usepackage{amssymb}
+
+\usepackage{bbold} % lower case mathbb letters
 
@@
 \newunicodechar{∀}{\ensuremath{\forall}}
+\newunicodechar{ˣ}{\ensuremath{{}^{\textsf{x}}}}
+\newunicodechar{ᵅ}{\ensuremath{{}^{\textsf{\alpha}}}}
+\newunicodechar{ᵗ}{\ensuremath{{}^{\textsf{t}}}}
+\newunicodechar{ᵣ}{\ensuremath{{}_{\textsf{r}}}}
+\newunicodechar{ᵥ}{\ensuremath{{}_{\textsf{v}}}}
 \newunicodechar{₁}{\ensuremath{{}_1}}
 \newunicodechar{₂}{\ensuremath{{}_2}}
-\newunicodechar{ᵣ}{\ensuremath{{}_{\textsf{r}}}}
+\newunicodechar{₃}{\ensuremath{{}_3}}
+\newunicodechar{ₖ}{\ensuremath{{}_{\textsf{k}}}}
+\newunicodechar{ₘ}{\ensuremath{{}_{\textsf{m}}}}
 \newunicodechar{ₛ}{\ensuremath{{}_{\textsf{s}}}}
+\newunicodechar{ₜ}{\ensuremath{{}_{\textsf{t}}}}
 \newunicodechar{→}{\ensuremath{\to}}
 \newunicodechar{◆}{\ensuremath{\diamond}}
+\newunicodechar{⊤}{\ensuremath{\top}}
+\newunicodechar{★}{\ensuremath{\star}}
+\newunicodechar{𝕖}{\ensuremath{\mathbb{e}}}
+\newunicodechar{𝕥}{\ensuremath{\mathbb{t}}}
+\newunicodechar{𝕜}{\ensuremath{\mathbb{k}}}
+\newunicodechar{∙}{\ensuremath{\bullet}}
+\newunicodechar{↪}{\ensuremath{\hookrightarrow}}
+\newunicodechar{∶}{\ensuremath{:}}
+\newunicodechar{⋯}{\ensuremath{\cdot\cdot\cdot}}
+\newunicodechar{⦅}{\ensuremath{\llparenthesis}}
+\newunicodechar{⦆}{\ensuremath{\rrparenthesis}}
+\newunicodechar{⟦}{\ensuremath{\llbracket}}
+\newunicodechar{⟧}{\ensuremath{\rrbracket}}
+\newunicodechar{⦃}{\ensuremath{\{\!\!\{}}
+\newunicodechar{⦄}{\ensuremath{\}\!\!\}}}
+\newunicodechar{α}{\ensuremath{\alpha}}
+\newunicodechar{λ}{\ensuremath{\lambda}}
+\newunicodechar{Λ}{\ensuremath{\Lambda}}
+\newunicodechar{ϕ}{\ensuremath{\phi}}
+\newunicodechar{Γ}{\ensuremath{\Gamma}}
+\newunicodechar{∎}{\ensuremath{\blacksquare}}
+% \newunicodechar{∎}{\ensuremath{\qed}}
+\newunicodechar{↑}{\ensuremath{\uparrow}}
+\newunicodechar{∋}{\ensuremath{\ni}}
+\newunicodechar{∈}{\ensuremath{\in}}
+\newunicodechar{∃}{\ensuremath{\exists}}
+\newunicodechar{ℓ}{\ensuremath{\ell}}
+\newunicodechar{⊔}{\ensuremath{\sqcup}}
+\newunicodechar{ρ}{\ensuremath{\rho}}
+\newunicodechar{σ}{\ensuremath{\sigma}}
+\newunicodechar{τ}{\ensuremath{\tau}}
+\newunicodechar{ξ}{\ensuremath{\xi}}
+\newunicodechar{β}{\ensuremath{\beta}}
+
 % \newunicodechar{Γ}{\ensuremath{\texttt{\Gamma}}}
@@
 % \usepackage{DejaVuSansMono}
-\usepackage{alphabeta} % Make greek letter macros also work in text-mode.
+% \usepackage{alphabeta} % Make greek letter macros also work in text-mode.
 
@@
 
+  \FMode
+  \FSyntax
+  \FTerms
+  \FTraversalOp
+  \FTraversalId
+  \FTraversal
+
   % \bibliographystyle{ACM-Reference-Format}

### 2023-06-03 paper/agdatex.py

@@ -1,0 +1,94 @@
+#!/usr/bin/env python3
+
+import os
+import sys
+from pathlib import Path
+import subprocess
+
+cache_dir = sys.argv[1]
+src_paths = [Path(x) for x in sys.argv[2:]]
+
+tgt_paths = [ p.parents[0] / (p.stem + ".lagda.tex") for p in src_paths ]
+bak_paths = [ p.parents[0] / (p.name + ".bak") for p in src_paths ]
+
+commands = []
+
+for src_path, tgt_path in zip(src_paths, tgt_paths):
+    with open(src_path, 'r') as f:
+        src = f.read();
+
+    tgt = ""
+    prefixes = []
+    mode = "none"
+    stop_command_on_empty_line = False
+    def start_command(name):
+        global mode, tgt, prefixes
+        for p in prefixes:
+            name = p + name
+        commands.append(name)
+        if mode == "hide":
+            tgt += "\\end{code}\n"
+        elif mode == "command":
+            print(f"ERROR: Line {line_num} starts a nested command:\n  {line}")
+        mode = "command"
+        tgt += "\\newcommand*\\" + name + "{\\begin{code}\n"
+    def stop_command():
+        global mode, tgt, prefixes
+        tgt += "\\end{code}}\n"
+        mode = "none"
+    for line_num, line in enumerate(src.splitlines()):
+        l = line.strip()
+        if l.startswith("--!") or l.startswith("-- !"):
+            l = l.split("!", 1)[1].strip()
+            if "{" in l:
+                name = l.split(" ", 1)[0]
+                start_command(name)
+            elif "}" in l:
+                stop_command()
+            elif ">" in l:
+                prefixes.append(l.split(" ", 1)[0])
+            elif "<" in l:
+                prefixes.pop()
+            else:
+                name = l.split(" ", 1)[0]
+                start_command(name)
+                stop_command_on_empty_line = True
+                # print(f"ERROR: Line {line_num} contains invalid agdatex command:\n  {line}")
+        elif line.strip() == "":
+            if mode == "command" and stop_command_on_empty_line:
+                stop_command_on_empty_line = False
+                stop_command()
+            tgt += "\n"
+        else:
+            if mode == "none":
+                tgt += "\\begin{code}[hide]\n"
+                mode = "hide"
+            tgt += line + "\n"
+    if mode == "hide":
[…]

### 2023-06-05 paper/agdatex.py
Update.

@@ -39,6 +39,9 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
     for line_num, line in enumerate(src.splitlines()):
         l = line.strip()
         if l.startswith("--!") or l.startswith("-- !"):
+            if mode == "command" and stop_command_on_empty_line:
+                stop_command_on_empty_line = False
+                stop_command()
             l = l.split("!", 1)[1].strip()
             if "{" in l:
                 name = l.split(" ", 1)[0]

### 2023-06-05 paper/agdatex.py
Update.

@@ -92,6 +92,12 @@ print("Restoring original files...")
 for src_path, bak_path in zip(src_paths, bak_paths):
     bak_path.rename(src_path)
 
+# print("Deleting .lagda.tex files...")
+# for tgt_path in tgt_paths:
+#     tgt_path.unlink()
+
 print("Deleting .lagda.tex files...")
-for tgt_path in tgt_paths:
-    tgt_path.unlink()
+tgt_paths2 = [ cache_dir + "/" + p.stem + ".lagda.tex" for p in src_paths ]
+for tgt_path, tgt_path2 in zip(tgt_paths, tgt_paths2):
+    Path(tgt_path2).parents[0].mkdir(exist_ok=True, parents=True)
+    tgt_path.rename(tgt_path2)

### 2023-06-20 paper/paper.tex
Update.
Paper: added references and renamed Mode to Sort.

--- 2021-03-24 paper/paper.tex
@@
 
-\usepackage{newunicodechar}
-\newunicodechar{≡}{\ensuremath{\equiv}}
-\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
-\newunicodechar{∘}{\ensuremath{\circ}}
-\newunicodechar{∅}{\ensuremath{\varnothing}}
-\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
-\newunicodechar{∋}{\ensuremath{\ni}}
-\newunicodechar{∷}{\ensuremath{\mathop{::}}}
-\newunicodechar{⊢}{\ensuremath{\vdash}}
-\newunicodechar{∀}{\ensuremath{\forall}}
-\newunicodechar{₁}{\ensuremath{{}_1}}
-\newunicodechar{₂}{\ensuremath{{}_2}}
-\newunicodechar{ᵣ}{\ensuremath{{}_{\textsf{r}}}}
-\newunicodechar{ₛ}{\ensuremath{{}_{\textsf{s}}}}
-\newunicodechar{→}{\ensuremath{\to}}
-\newunicodechar{◆}{\ensuremath{\diamond}}
-% \newunicodechar{Γ}{\ensuremath{\texttt{\Gamma}}}
-% \newunicodechar{λ}{\ensuremath{\texttt{\lambda}}}
+\usepackage{bbold} % lower case mathbb letters
+
+\usepackage{agda-unicode}
 
@@
 
+  \section{Introduction}
+  \newpage
+
+  \section{Syntax}
+  \subsection{Library}
+  \KVariables
+  \KSortTy
+  \KTerms
+  \subsection{System F}
+  \FSort
+  \FSyntax
+  \FTerms
+
+  \section{Maps}
+  \subsection{Library}
+  \KKit
+  \KMap
+  \KAp
+  \KExt
+  \KLift
+  \KId
+  \KSingle
+  \KWeaken
+  \KEq
+  \KFunExt
+  \KIdLift
+  % \KIdLiftProof
+  \KKitNotation
+  \KTraversal
+  \KKitInstances
+  \KKitOpen
+  \subsection{System F}
+  \FTraversalOp
+  \FTraversalId
+  \FTraversalIdProofInteresting
+  % \FTraversalIdProofRest
+  \FTraversal
+
+  \section{Map Composition}
+  \subsection{Library}
+  \KWkKit
+  \KWkKitInstances
+  \KComposeKit
+  \KComposition
+  \KComposeKitAp
+  % \KComposeKitApProof
+  \KDistLiftCompose
+  % \KDistLiftComposeProof
+  \KComposeKitNotation
+  \KComposeTraversal
+  \KCommLiftWeaken
+  % \KCommLiftWeakenProof
+  \KCommLiftWeakenTraverse
+  % \KCommLiftWeakenTraverseProof
+  \KComposeKitInstances
+  \KComposeKitInstancesConcrete
+  \KWeakenCancelsSingle
+  % \KWeakenCancelsSingleProof
+  \KWeakenCancelsSingleTraverse
+  % \KWeakenCancelsSingleTraverseProof
+  \KDistLiftSingle
+  % \KDistLiftSingleProof
+  \KDistLiftSingleTraverse
+  % \KDistLiftSingleTraverseProof
+  \subsection{System F}
+  \FAssoc
+  \FAssocProofInteresting
+  % \FAssocProofRest
+  \FComposeTraversal
+
+  \section{Types \& Contexts}
+  \subsection{Library}
+  \KTypeSorts
+  \KTypes
+  \KContextHelper
+  \KContexts
+  \KContextLookup
+  \subsection{System F}
+  \FTypes
+
+  \section{Typing}
+  \subsection{Library}
+  \KVariableTyping
+  \KTyping
+  \KTypingKit
+  \KMapTyping
+  \KLiftTyping
+  % \KLiftTypingProof
+  \KSingleTyping
+  % \KSingleTypingProof
+  \KTypingNotation
+  \KTypingTraversal
+  \KTypingInstances
+  % \KTypingTraversalNotation
+  \subsection{System F}
+  \FTyping
+  \FTypingInst
+  \FPreserve
+  \FTypingTraversal
+
+  \section{Semantics}
+  % \FValues
+  \FReduction
+
+  \section{Subject Reduction}
+  \FSubjectReduction
+  \FSubjectReductionProofInteresting
+  % \FSubjectReductionProofRest
+
+  \section{Related Work}
+  Autosubst \cite{DBLP:conf/cpp/StarkSK19, DBLP:conf/itp/SchaferTS15}.
+  Kits \cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}.
+  Kits linear \cite{DBLP:journals/corr/abs-2005-02247}.
+  Universe \cite{DBLP:journals/pacmpl/AllaisA0MM18}.
+
   % \bibliographystyle{ACM-Reference-Format}

### 2023-06-21 paper/paper.tex
Update.

@@
 \documentclass[sigplan,10pt]{acmart}
+
+\usepackage{agda-unicode}
 
@@
 
-\author{Peter Thiemann}
-\orcid{0000-0002-9000-1239}                 %% \orcid is optional
-\affiliation{
-  % \department{Department2a}               %% \department is recommended
-  \institution{University of Freiburg}      %% \institution is required
-  \country{Germany}                         %% \country is recommended
-}
-\email{thiemann@informatik.uni-freiburg.de} %% \email is recommended
+% \author{Peter Thiemann}
+% \orcid{0000-0002-9000-1239}                 %% \orcid is optional
+% \affiliation{
+%   % \department{Department2a}               %% \department is recommended
+%   \institution{University of Freiburg}      %% \institution is required
+%   \country{Germany}                         %% \country is recommended
+% }
+% \email{thiemann@informatik.uni-freiburg.de} %% \email is recommended
 
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{stmaryrd}
-
-\usepackage[links]{agda}
-
-% \usepackage{amsmath}
-% \usepackage{amsfonts}
-% \usepackage{amssymb}
-
-\usepackage{bbold} % lower case mathbb letters
-
-\usepackage{agda-unicode}
-
-% \usepackage{cite}
-% \usepackage{DejaVuSansMono}
-% \usepackage{alphabeta} % Make greek letter macros also work in text-mode.
-
-\usepackage{microtype}
-\DisableLigatures[-]{encoding=T1}
+\usepackage{tcolorbox}
+% \newenvironment{LibCode}{%
+%   \begin{tcolorbox}[boxrule=0.5pt,top=0pt,left=0pt,right=2.5pt,boxsep=0pt]\vspace{-0.5\baselineskip}}{%
+%   \end{tcolorbox}}
+\newenvironment{LibCode}{%
+  \begin{tcolorbox}[boxrule=0.0pt,frame hidden,top=0pt,left=0pt,right=2.5pt,boxsep=0pt]\vspace{-0.5\baselineskip}}{%
+  \end{tcolorbox}}
 
@@
   \section{Introduction}
+  Formalizing metatheory of programming languages in a proof assistant 
+  requires to deal with variables, binders and substitution.
+
+  In most languages those constructs behave uniformly,
+  i.e. substitution is only concerned with variables and binders and
+  behaves homomorphic on all other syntax constructors.
+
+  Intrinsic typing
+
+  \begin{enumerate}
+  \item DeBruijn
+  \item Intrinsic vs Extrinsic
+  \item Reduction under binder
+  \end{enumerate}
+
+  \subsection{Contributions}
+  \begin{enumerate}
+  \item
+    a multi-sorted approach to extrinsic typing;
+  \item
+    a multi-sorted extension of McBride's kits to unify renamings and substitutions and their metatheory;
+  \item
+    a kit abstraction for composition and it's metatheory, unifying the four combinations of renamings and substitutions;
+  \item
+    a kit abstraction for typing relations, unifying type preservation of renaming and substitution into a single lemma;
+  \item
+    an Agda-implementation featuring derived term-kits, representation
+    independence for substitutions and typing contexts, heterogeneous
+    equality, independence of functional extensionality, case studies
+    on System F, dependent lambda calculus, pattern matching, linear
+    types.
+  \end{enumerate}
+
   \newpage
@@
   \section{Syntax}
+  The following shows a typical intrinsically-scoped syntax of System F:
+  \FUnsortedSyntax
+
+  \texttt{Type}s are indexed by the number of free type variables \texttt{n}.
+  \texttt{Expr}essions are additionally indexed by the number of free expression variables \texttt{m}.
+  Variables are represented as DeBruijn-indices, where \texttt{Fin n} is the type of \texttt{n} elements.
+
+  This style of syntax comes with multiple drawbacks:
+  \begin{itemize}
+  \item Bad scaling for automation
+  \item Multiple substitions require interaction lemmas
+  \end{itemize}
+
+  % Instead we define a single term type which is indexed by the sort of
+  % syntax, i.e.\ whether the term is a kind, type, or expression:
+  To avoid these drawbacks, we define a single term type \texttt{S~⊢~s} that is indexed by its sort \texttt{s},
+  i.e.\ whether the term is a kind, type, or expression, and the sorts of its free variables \texttt{S}:
+
+  \begin{LibCode}
+  \KSortTy
+  \end{LibCode}
+
+  \FSort
+  \FSyntax
+  The variable constructor \texttt{'\_} takes a sorted DeBruijn-index \texttt{S~∋~s}.
+  Since the \texttt{Sort}s in \texttt{S} are restricted to the sort-type \texttt{Var}, it is not possible to construct kind-variables.
+
   \subsection{Library}

### 2023-06-21 paper/paper.tex
Update.

@@
 \usepackage{tcolorbox}
-% \newenvironment{LibCode}{%
-%   \begin{tcolorbox}[boxrule=0.5pt,top=0pt,left=0pt,right=2.5pt,boxsep=0pt]\vspace{-0.5\baselineskip}}{%
-%   \end{tcolorbox}}
-\newenvironment{LibCode}{%
-  \begin{tcolorbox}[boxrule=0.0pt,frame hidden,top=0pt,left=0pt,right=2.5pt,boxsep=0pt]\vspace{-0.5\baselineskip}}{%
-  \end{tcolorbox}}
+
+\newenvironment{LibCode*}{%
+  \begin{tcolorbox}[%
+    colframe=white,%
+    boxrule=0.0pt,%
+    top=2.5pt,%
+    left=2.5pt,%
+    bottom=2.5pt,%
+    right=2.5pt,%
+    boxsep=0pt%
+  ]\vspace{-0.2\baselineskip}%
+}{%
+  \vspace{-1\baselineskip}%
+  \end{tcolorbox}%
+}
+
+\newcommand*\LibCode[1]{\begin{LibCode*}{#1}\end{LibCode*}}
+\newcommand*\AppCode[1]{{#1}}
 
@@
 
-  \begin{LibCode}
-  \KSortTy
-  \end{LibCode}
+  \LibCode\KSortTy
 
-  \FSort
-  \FSyntax
+  \AppCode\FSort
+
+  \LibCode\KVariables
+  \AppCode\FSyntax
   The variable constructor \texttt{'\_} takes a sorted DeBruijn-index \texttt{S~∋~s}.
@@
   \subsection{Library}
-  \KVariables
-  \KSortTy
-  \KTerms
+  \LibCode\KTerms
   \subsection{System F}
@@
   \subsection{Library}
-  \KKit
-  \KMap
-  \KAp
-  \KExt
-  \KLift
-  \KId
-  \KSingle
-  \KWeaken
-  \KEq
-  \KFunExt
-  \KIdLift
-  % \KIdLiftProof
-  \KKitNotation
-  \KTraversal
-  \KKitInstances
-  \KKitOpen
+  \LibCode\KKit
+  \LibCode\KMap
+  \LibCode\KAp
+  \LibCode\KExt
+  \LibCode\KLift
+  \LibCode\KId
+  \LibCode\KSingle
+  \LibCode\KWeaken
+  \LibCode\KEq
+  \LibCode\KFunExt
+  \LibCode\KIdLift
+  % \LibCode\KIdLiftProof
+  \LibCode\KKitNotation
+  \LibCode\KTraversal
+  \LibCode\KKitInstances
+  \LibCode\KKitOpen
   \subsection{System F}
-  \FTraversalOp
-  \FTraversalId
-  \FTraversalIdProofInteresting
-  % \FTraversalIdProofRest
-  \FTraversal
+  \AppCode\FTraversalOp
+  \AppCode\FTraversalId
+  \AppCode\FTraversalIdProofInteresting
+  % \LibCode\FTraversalIdProofRest
+  \AppCode\FTraversal
 
@@
   \subsection{Library}
-  \KWkKit
-  \KWkKitInstances
-  \KComposeKit
-  \KComposition
-  \KComposeKitAp
-  % \KComposeKitApProof
-  \KDistLiftCompose
-  % \KDistLiftComposeProof
-  \KComposeKitNotation
-  \KComposeTraversal
-  \KCommLiftWeaken
-  % \KCommLiftWeakenProof
-  \KCommLiftWeakenTraverse
-  % \KCommLiftWeakenTraverseProof
-  \KComposeKitInstances
-  \KComposeKitInstancesConcrete
-  \KWeakenCancelsSingle
-  % \KWeakenCancelsSingleProof
-  \KWeakenCancelsSingleTraverse
-  % \KWeakenCancelsSingleTraverseProof
-  \KDistLiftSingle
-  % \KDistLiftSingleProof
-  \KDistLiftSingleTraverse
-  % \KDistLiftSingleTraverseProof
+  \LibCode\KWkKit
+  \LibCode\KWkKitInstances
+  \LibCode\KComposeKit
+  \LibCode\KComposition
+  \LibCode\KComposeKitAp
+  % \LibCode\KComposeKitApProof
+  \LibCode\KDistLiftCompose
+  % \LibCode\KDistLiftComposeProof
+  \LibCode\KComposeKitNotation
+  \LibCode\KComposeTraversal
+  \LibCode\KCommLiftWeaken
+  % \LibCode\KCommLiftWeakenProof
+  \LibCode\KCommLiftWeakenTraverse
+  % \LibCode\KCommLiftWeakenTraverseProof
+  \LibCode\KComposeKitInstances
+  \LibCode\KComposeKitInstancesConcrete
+  \LibCode\KWeakenCancelsSingle
+  % \LibCode\KWeakenCancelsSingleProof
+  \LibCode\KWeakenCancelsSingleTraverse
+  % \LibCode\KWeakenCancelsSingleTraverseProof
+  \LibCode\KDistLiftSingle
+  % \LibCode\KDistLiftSingleProof
+  \LibCode\KDistLiftSingleTraverse
+  % \LibCode\KDistLiftSingleTraverseProof
   \subsection{System F}
-  \FAssoc
-  \FAssocProofInteresting
-  % \FAssocProofRest
-  \FComposeTraversal
+  \AppCode\FAssoc
+  \AppCode\FAssocProofInteresting
+  % \AppCode\FAssocProofRest
+  \AppCode\FComposeTraversal
 
@@
   \subsection{Library}
-  \KTypeSorts
-  \KTypes
-  \KContextHelper
-  \KContexts
-  \KContextLookup
+  \LibCode\KTypeSorts
+  \LibCode\KTypes
+  \LibCode\KContextHelper
+  \LibCode\KContexts
+  \LibCode\KContextLookup
   \subsection{System F}
-  \FTypes
+  \AppCode\FTypes
 
@@
   \subsection{Library}
-  \KVariableTyping
-  \KTyping
-  \KTypingKit
-  \KMapTyping
-  \KLiftTyping
-  % \KLiftTypingProof
-  \KSingleTyping
-  % \KSingleTypingProof
-  \KTypingNotation
-  \KTypingTraversal
-  \KTypingInstances
-  % \KTypingTraversalNotation
+  \LibCode\KVariableTyping
+  \LibCode\KTyping
+  \LibCode\KTypingKit
+  \LibCode\KMapTyping
+  \LibCode\KLiftTyping
+  % \LibCode\KLiftTypingProof
+  \LibCode\KSingleTyping
+  % \LibCode\KSingleTypingProof
+  \LibCode\KTypingNotation
+  \LibCode\KTypingTraversal
+  \LibCode\KTypingInstances
+  % \LibCode\KTypingTraversalNotation
   \subsection{System F}
-  \FTyping
-  \FTypingInst
-  \FPreserve
-  \FTypingTraversal
+  \AppCode\FTyping
+  \AppCode\FTypingInst
+  \AppCode\FPreserve
+  \AppCode\FTypingTraversal
 
   \section{Semantics}
-  % \FValues
-  \FReduction
+  % \AppCode\FValues
+  \AppCode\FReduction
 
   \section{Subject Reduction}
-  \FSubjectReduction
-  \FSubjectReductionProofInteresting
+  \AppCode\FSubjectReduction
+  \AppCode\FSubjectReductionProofInteresting
   % \FSubjectReductionProofRest

### 2023-06-21 paper/paper.tex
Update.

@@
 
+\newenvironment{ExampleCode*}{%
+  \begin{tcolorbox}[%
+    colframe=white,%
+    colback=yellow!5,%
+    boxrule=0.0pt,%
+    top=2.5pt,%
+    left=2.5pt,%
+    bottom=2.5pt,%
+    right=2.5pt,%
+    boxsep=0pt%
+  ]\vspace{-0.2\baselineskip}%
+}{%
+  \vspace{-1\baselineskip}%
+  \end{tcolorbox}%
+}
+
 \newcommand*\LibCode[1]{\begin{LibCode*}{#1}\end{LibCode*}}
 \newcommand*\AppCode[1]{{#1}}
+\newcommand*\ExampleCode[1]{\begin{ExampleCode*}{#1}\end{ExampleCode*}}
 
@@
 
+  \AgdaNoSpaceAroundCode{}
+
   \section{Introduction}
@@
   The following shows a typical intrinsically-scoped syntax of System F:
-  \FUnsortedSyntax
+  \ExampleCode\FUnsortedSyntax
 
@@
 
-  \subsection{Library}
   \LibCode\KTerms
-  \subsection{System F}
-  \FSort
-  \FSyntax
-  \FTerms
+  \AppCode\FTerms
 
-  \section{Maps}
+  \section{Substitution \& Renamings}
+  Our sort-indexed syntax allows us to define substitutions and
+  renamings which replace variables of all sorts simultaneously:
+
+  \ExampleCode\FExampleSubRen
+
+  A renaming \texttt{S₁ →ᵣ S₂} maps variables from \texttt{S₁} to
+  variables from \texttt{S₂}.
+  Similarly a substitution \texttt{S₁ →ₛ S₂} maps variables from \texttt{S₁} to
+  terms from \texttt{S₂}.
+  We consider these operations \emph{very parallel}, because they do not just replace all variables of one sort in parallel, but they replace all variables of all sorts in parallel.
+
+  \ExampleCode\FExampleSub
+
+  This representation has the benefit that there is no combinatory
+  explosion of substitutions and renamings, e.g.\ no extra lemmas have to be
+  proved between an expression-in-expression and a type-in-expression
+  substitution, because both are simply substitutions.
+
   \subsection{Library}

### 2023-06-21 paper/paper.tex
Update.

@@
   Our sort-indexed syntax allows us to define substitutions and
-  renamings which replace variables of all sorts simultaneously:
+  renamings which replace all variables of all sorts simultaneously:
 
@@
   variables from \texttt{S₂}.
-  Similarly a substitution \texttt{S₁ →ₛ S₂} maps variables from \texttt{S₁} to
+  A substitution \texttt{S₁ →ₛ S₂} maps variables from \texttt{S₁} to
   terms from \texttt{S₂}.
-  We consider these operations \emph{very parallel}, because they do not just replace all variables of one sort in parallel, but they replace all variables of all sorts in parallel.
 

### 2023-06-21 paper/paper.tex
Update.

@@
   \section{Substitution \& Renamings}
-  Our sort-indexed syntax allows us to define substitutions and
+  Working with a sort-indexed syntax allows us to define substitutions and
   renamings which replace all variables of all sorts simultaneously:

### 2023-06-21 paper/paper.tex
Update.

--- 2023-06-21 paper/paper.tex
@@
   \item
-    a kit abstraction for composition and it's metatheory, unifying the four combinations of renamings and substitutions;
+    an abstraction for composition and it's metatheory, unifying the four combinations of renamings and substitutions;
   \item
-    a kit abstraction for typing relations, unifying type preservation of renaming and substitution into a single lemma;
+    an abstraction for typing relations, unifying type preservation of renaming and substitution into a single lemma;
   \item
@@
 
+  For example a term of type \texttt{[𝕖, 𝕥] ⊢ 𝕖} is an expression with two free variables, where variable 0 is a placeholder for an expression, and variable 1 for a type.
   \LibCode\KSortTy
@@
 
-  \section{Substitution \& Renamings}
-  Our sort-indexed syntax allows us to define substitutions and
-  renamings which replace all variables of all sorts simultaneously:
+  \section{Renamings \& Substitutions}
+  \subsection{Multi-Sorted Renamings \& Substitutions}
+  Working with a sort-indexed syntax allows us to define renamings and substitutions
+  which replace all variables of all sorts simultaneously:
 
@@
   A substitution \texttt{S₁ →ₛ S₂} maps variables from \texttt{S₁} to
-  terms from \texttt{S₂}.
+  terms in \texttt{S₂}.
 
@@
   substitution, because both are simply substitutions.
+
+  \subsection{Unifying Renamings \& Substitutions}
+  A well-known problem with DeBruijn-indices is that full substitution
+  cannot be defined directly by structural recursion on the term.
+  The typical workaround is to define renamings and substitutions
+  separately and use renamings to weaken substitutions when going under binders.
+
+  While this keeps lemmas simple (requiring only structural induction),
+  it duplicates the amount of work as all operations and lemmas have
+  to be defined and proved for both renamings and substitutions.
+
+  To avoid this duplication,
+  McBride\cite{DBLP:journals/jar/BentonHKM12,
+  unpublished:mcbride2005kits} introduced \emph{Kits} - an abstraction
+  which allows to talk about both renamings and substitutions
+  simultaneously.
+
+  The core idea behind kits is that renamings and substitutions only
+  differ in what they return:
+  \ExampleCode\FExampleSubRen
 
@@
   \section{Typing}
+  % TODO move to typing section
+  The only drawback we found, is that types can now also be indexed
+  with free expression variables, even though types can never contain
+  free variables as assured by the variable constructor. However, this
+  only causes minor issues.
+
   \subsection{Library}

### 2023-06-21 paper/paper.tex
Update.

@@
 
-  For example a term of type \texttt{[𝕖, 𝕥] ⊢ 𝕖} is an expression with two free variables, where variable 0 is a placeholder for an expression, and variable 1 for a type.
+  For example a term of type \texttt{[𝕥, 𝕥] ⊢ 𝕖} is an expression (\texttt{𝕖}) with two free type-variables (\texttt{𝕥}).
   \LibCode\KSortTy
@@
   The typical workaround is to define renamings and substitutions
-  separately and use renamings to weaken substitutions when going under binders.
+  separately and use renamings to weaken the terms in substitutions
+  when going under binders.
 
@@
 
+  In both cases \texttt{\_∋\_} and \texttt{\_⊢\_} are instanciated to type
+  \texttt{List (Sort Var) → Sort Var → Set}, so we can create a structure
+  to axiomatize the common theory behind both of them.
+  \LibCode\KKit
+
   \subsection{Library}
-  \LibCode\KKit
   \LibCode\KMap

### 2023-06-21 paper/paper.tex
Update.

@@
   \end{enumerate}
+
+  \subsection{Structure}
+  The rest of this paper will introduce our framework using System F
+  as a running example for a type soundness proof.
+
+  Code of the framework is displayed in gray boxes.
+  Code of examples is written in yellow boxes.
+  Code of the System F formalization is written without boxes. This is
+  the only code a user of our framework has to write.
+
+  In the paper we state all definitions and lemmas, but omit some proofs
+  in the interest of space. However the omitted proofs exclusively
+  depend on lemmas and definitions stated in this paper. The full code
+  from the paper can be found at TODO.
+
+  In the paper we only show the part of the framework necessary for
+  the System F soundness proof. The actual implementation is more
+  general and complete and can be found at TODO.
 

### 2023-07-04 paper/paper.tex
Update.

@@
 
-  In both cases \texttt{\_∋\_} and \texttt{\_⊢\_} are instanciated to type
-  \texttt{List (Sort Var) → Sort Var → Set}, so we can create a structure
-  to axiomatize the common theory behind both of them.
+  In both cases \texttt{\_∋\_} and \texttt{\_⊢\_} are instanciated to the type
+  \LibCode\KScoped
+  which allows us to create a structure to axiomatize the common
+  theory behind both of them:
   \LibCode\KKit

### 2023-07-11 paper/paper.tex
Update.

@@
 
-\title{Kit Theory}
+\title{Abstractions for Multi-Sorted Substitutions}
 
@@
   \section{Syntax}
+  \subsection{Single-Sorted Syntax}
   The following shows a typical intrinsically-scoped syntax of System F:
@@
   \texttt{Expr}essions are additionally indexed by the number of free expression variables \texttt{m}.
-  Variables are represented as DeBruijn-indices, where \texttt{Fin n} is the type of \texttt{n} elements.
+  Variables \texttt{`\_} are represented as DeBruijn-indices, where \texttt{Fin n} is the type of \texttt{n} elements.
 
-  This style of syntax comes with multiple drawbacks:
-  \begin{itemize}
-  \item Bad scaling for automation
-  \item Multiple substitions require interaction lemmas
-  \end{itemize}
+  We identify two drawbacks with this style of syntax:
+  \begin{enumerate}
+  \item the syntactic categories (\texttt{Kind}, \texttt{Type}, and \texttt{Expr})
+    have different types, which makes it difficult to treat them uniformly; and
+  \item the different sorts of variables are modeled separately, which requires to define not just
+    type-in-type and expression-in-expression substitution, but also type-in-expression substitution.
+    Consequently, interaction lemmas between those substitutions are required.
+    In general, this leads to a combinatory explosion of
+    substitutions, e.g.\ adding kind variables would lead to a total of 6
+    different substitutions.
+  \end{enumerate}
 
+  \subsection{Multi-Sorted Syntax}
   % Instead we define a single term type which is indexed by the sort of
@@
 
+  \subsection{A Structure for Multi-Sorted Syntax}
   \LibCode\KTerms
@@
   A substitution \texttt{S₁ →ₛ S₂} maps variables from \texttt{S₁} to
-  terms in \texttt{S₂}.
+  terms containing variables from \texttt{S₂}.
 
@@
   cannot be defined directly by structural recursion on the term.
-  The typical workaround is to define renamings and substitutions
+  The usual workaround is to define renamings and substitutions
   separately and use renamings to weaken the terms in substitutions
@@
   To avoid this duplication,
-  McBride\cite{DBLP:journals/jar/BentonHKM12,
-  unpublished:mcbride2005kits} introduced \emph{Kits} - an abstraction
-  which allows to talk about both renamings and substitutions
-  simultaneously.
+  McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
+  introduced an abstraction called \emph{Kits}, which allows to talk about
+  both renamings and substitutions simultaneously.
 
-  The core idea behind kits is that renamings and substitutions only
-  differ in what they return:
+  The core observation leading to this abstraction is that renamings
+  and substitutions only differ in their return type:
   \ExampleCode\FExampleSubRen
@@
   \subsection{Library}
-  \LibCode\KMap
   \LibCode\KAp

### 2023-07-11 paper/paper.tex
Update.

@@
   \subsection{Multi-Sorted Syntax}
-  % Instead we define a single term type which is indexed by the sort of
-  % syntax, i.e.\ whether the term is a kind, type, or expression:
-  To avoid these drawbacks, we define a single term type \texttt{S~⊢~s} that is indexed by its sort \texttt{s},
-  i.e.\ whether the term is a kind, type, or expression, and the sorts of its free variables \texttt{S}:
+  To avoid these drawbacks, we define the whole syntax as a single sort-indexed type of terms.
 
-  For example a term of type \texttt{[𝕥, 𝕥] ⊢ 𝕖} is an expression (\texttt{𝕖}) with two free type-variables (\texttt{𝕥}).
+  A sort describes to which syntactic category a term belongs and is
+  itself indexed by a sort type which describes whether the syntax
+  permits variables of this sort:
   \LibCode\KSortTy
-
   \AppCode\FSort
 
+  The term type \texttt{S~⊢~s} is then indexed by its sort \texttt{s} and the sorts of its free variables \texttt{S}.
+  For example, \texttt{[𝕥, 𝕥] ⊢ 𝕖} is the type of expressions (\texttt{𝕖}) with two free type-variables (\texttt{𝕥}).
+  \newpage
+  \AppCode\FSyntax
+  The notation \texttt{\_⊢\_} is often used for well-typed terms in
+  intrinsically-typed languages. This is no accident: in effect, we
+  defined an intrinsically-typed language with the twist that the
+  typing relation assures exactly that the syntactic categories are followed.
+  Sorts \texttt{s} correspond to types, and lists of sorts \texttt{S}
+  correspond to type environments.
+
+  As it is typical in intrinsic typing, variables are represented as
+  typed (in our case sorted) DeBruijn indices \texttt{S~∋~s}, i.e.\
+  values of the usual proof-relevant list-membership relation:
   \LibCode\KVariables
-  \AppCode\FSyntax
-  The variable constructor \texttt{'\_} takes a sorted DeBruijn-index \texttt{S~∋~s}.
-  Since the \texttt{Sort}s in \texttt{S} are restricted to the sort-type \texttt{Var}, it is not possible to construct kind-variables.
+
+  Contrary to the usual intrinsic typing, we limit the environment
+  \texttt{S} to sorts with index \texttt{Var}.
+  This ensures that it is impossible to construct kind variables, as
+  it is not possible to extend \texttt{S} with the sort of kinds
+  \texttt{𝕜}.
 
   \subsection{A Structure for Multi-Sorted Syntax}
+  The rest of our formalization builds on top of a multi-sorted
+  syntax, which we capture in the following structure:
   \LibCode\KTerms
+  Aside from capturing our syntactic definitions, we require a proof that the
+  variable introduction \texttt{`\_} is injective. This is trivially true
+  for all constructors. The instantiation for our System F syntax is straightforward:
   \AppCode\FTerms

### 2023-07-14 paper/paper.tex
Update.

@@
   \subsection{A Structure for Multi-Sorted Syntax}
-  The rest of our formalization builds on top of a multi-sorted
-  syntax, which we capture in the following structure:
+  The rest of our framework builds on top of a multi-sorted
+  syntax for which we define the following structure:
   \LibCode\KTerms
-  Aside from capturing our syntactic definitions, we require a proof that the
-  variable introduction \texttt{`\_} is injective. This is trivially true
-  for all constructors. The instantiation for our System F syntax is straightforward:
+  The first three fields record the definitions of sorts, terms, and variable introduction.
+  The last field records that variable introduction
+  \texttt{`\_} is injective, which is trivially true for
+  constructors. The instantiation for our System F syntax is
+  straightforward:
   \AppCode\FTerms
@@
   \subsection{Unifying Renamings \& Substitutions}
+  A well-known problem with DeBruijn-indices is that the operation of
+  applying a substitution \texttt{σ} to a term \texttt{t} cannot be
+  defined directly by structural recursion on \texttt{t}.
+  If \texttt{t} is a term like \texttt{λx e},
+  where the subterm \texttt{e} is under a binder, we need to adjust \texttt{σ}
+  for the shifted DeBruijn-indices in \texttt{e}:
+  \begin{itemize}
+  \item The variable \texttt{zero} refers in \texttt{e} to the variable bound by the binder.
+  \item a variable \texttt{suc x} refers to variable \texttt{x} in \texttt{σ}.
+  \end{itemize}
+
+  \subsection{Unifying Renamings \& Substitutions}
   A well-known problem with DeBruijn-indices is that full substitution
   cannot be defined directly by structural recursion on the term.
-  The usual workaround is to define renamings and substitutions
-  separately and use renamings to weaken the terms in substitutions
-  when going under binders.
+  The usual workaround is to define renamings first and use them
+  to weaken the terms in substitutions when going under binders.
 

### 2023-08-01 paper/paper.tex
Update.

@@
 
+  For example, the substitution, which replaces the term variable 0 with the
+  identity function and the type variable 1 with the bottom type, is defined
+  as followed:
   \ExampleCode\FExampleSub
@@
   If \texttt{t} is a term like \texttt{λx e},
-  where the subterm \texttt{e} is under a binder, we need to adjust \texttt{σ}
-  for the shifted DeBruijn-indices in \texttt{e}:
-  \begin{itemize}
-  \item The variable \texttt{zero} refers in \texttt{e} to the variable bound by the binder.
-  \item a variable \texttt{suc x} refers to variable \texttt{x} in \texttt{σ}.
-  \end{itemize}
+  where the subterm \texttt{e} is under a binder, then the variables
+  in \texttt{e} are shifted, i.e.  the variable \texttt{zero} refers
+  to the variable from the binder, whereas a variable \texttt{suc x}
+  refers to the variable \texttt{x} from outside of the binder.
+  As the terms contained in \texttt{σ} may themselves contain free variables,
+  they need to be shifted, too, before they can be substituted
+  for variables in \texttt{e}. 
+  Shifting the variables in a term is itself a substitution, which maps
+  each variable \texttt{x} to the term \texttt{` (suc x)}.
+  However, applying this shifting substitution to the terms in
+  \texttt{σ} is not structurally recursive, as the terms in \texttt{σ}
+  are not subterms of \texttt{λx e}.
 

### 2023-08-08 paper/paper.tex
Update.

@@
   A substitution \texttt{S₁ →ₛ S₂} maps variables from \texttt{S₁} to
-  terms containing variables from \texttt{S₂}.
+  terms with free variables from \texttt{S₂}.
 
@@
 
-  \subsection{Unifying Renamings \& Substitutions}
-  A well-known problem with DeBruijn-indices is that full substitution
-  cannot be defined directly by structural recursion on the term.
   The usual workaround is to define renamings first and use them
   to weaken the terms in substitutions when going under binders.
-
   While this keeps lemmas simple (requiring only structural induction),

### 2023-08-22 paper/paper.tex
Update.

@@
   \subsection{Multi-Sorted Syntax}
-  To avoid these drawbacks, we define the whole syntax as a single sort-indexed type of terms.
+  To avoid these drawbacks, we define the syntax as a single
+  sort-indexed type of terms.
 
   A sort describes to which syntactic category a term belongs and is
-  itself indexed by a sort type which describes whether the syntax
+  itself indexed by a sort type, which describes whether the syntax
   permits variables of this sort:
@@
   \AppCode\FSyntax
-  The notation \texttt{\_⊢\_} is often used for well-typed terms in
+  The notation \texttt{\_⊢\_} is often used for terms in
   intrinsically-typed languages. This is no accident: in effect, we
@@
   \subsection{A Structure for Multi-Sorted Syntax}
-  The rest of our framework builds on top of a multi-sorted
-  syntax for which we define the following structure:
+  The multi-sorted syntax makes it easy to capture general syntax definitions
+  in a structure:
   \LibCode\KTerms
@@
   Working with a sort-indexed syntax allows us to define renamings and substitutions
-  which replace all variables of all sorts simultaneously:
+  that replace all variables of all sorts simultaneously:
 
@@
 
-  \subsection{Unifying Renamings \& Substitutions}
-  A well-known problem with DeBruijn-indices is that the operation of
-  applying a substitution \texttt{σ} to a term \texttt{t} cannot be
-  defined directly by structural recursion on \texttt{t}.
+  \subsection{The Need for Renamings}
+  A problem with mechanizing languages with DeBruijn-indices is that
+  the operation of applying a substitution \texttt{σ} to a term
+  \texttt{t} cannot be defined directly by structural recursion on
+  \texttt{t}.
   If \texttt{t} is a term like \texttt{λx e},
   where the subterm \texttt{e} is under a binder, then the variables
-  in \texttt{e} are shifted, i.e.  the variable \texttt{zero} refers
+  in \texttt{e} are shifted, i.e.\ the variable \texttt{zero} refers
   to the variable from the binder, whereas a variable \texttt{suc x}
@@
 
-  The usual workaround is to define renamings first and use them
-  to weaken the terms in substitutions when going under binders.
+  The usual workaround is to first define renamings,
+  i.e.\ substitutions that replace variables with variables.
+  In contrast to general substitutions, renamings can be easily
+  shifted: they only contain variables \texttt{x}, which can be
+  shifted by simply taking their successor \texttt{suc x}.
+  As the shifting substitution is a renaming, we can then define substitution,
+  where we apply the shifting \emph{renaming} to the terms in \texttt{σ} when
+  the substitution needs to go under a binder.
+
+  % The usual workaround is to define renamings first and use them
+  % to weaken the terms in substitutions when going under binders.
   While this keeps lemmas simple (requiring only structural induction),
@@
 
-  To avoid this duplication,
+  \subsection{Unifying Renamings \& Substitutions}
+  To avoid the duplication between renamings and substitutions,
   McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
-  introduced an abstraction called \emph{Kits}, which allows to talk about
-  both renamings and substitutions simultaneously.
+  introduced an abstraction called \emph{Kits}.
+
+  % While Kits were originally formulated for intrinsically-typed languages,
+  % they nicely extend to our extrinsically-typed multi-sorted syntax, whose
+  % sort-indexing is a form of intrinsic typing.
+
+  A kit is a structure that abstracts over the difference between variables and terms.
+  There are always exactly two instances for Kits, one for variables and one for terms,
 
   The core observation leading to this abstraction is that renamings
-  and substitutions only differ in their return type:
-  \ExampleCode\FExampleSubRen
+  \texttt{S₁ →ᵣ S₂}
+  and substitutions \texttt{S₁ →ₛ S₂} only differ in their return type:
+  while a substitution returns a term \texttt{S₂ ⊢ s}, a renaming returns
+  a variable \texttt{S₂ ∋ s}.
+
+  A Kit is a structure that abstracts precisely over the commonalities
+  between variables and terms.
+
+  % \ExampleCode\FExampleSubRen
 
@@
 
+  \section{Reflection \& Generics}
+
+  \section{Future Work}
+
   \section{Related Work}
@@
   Universe \cite{DBLP:journals/pacmpl/AllaisA0MM18}.
+
+  \section{Conclusion}
 

### 2023-08-22 paper/paper.tex
Update.

@@
   McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
-  introduced an abstraction called \emph{Kits}.
+  introduced an abstraction called \emph{kits}.\footnote{
+    While kits were originally formulated for intrinsically-typed languages,
+    they nicely extend to our extrinsically-typed multi-sorted syntax, whose
+    sort-indexing is a form of intrinsic typing.
+  }
 
-  % While Kits were originally formulated for intrinsically-typed languages,
-  % they nicely extend to our extrinsically-typed multi-sorted syntax, whose
-  % sort-indexing is a form of intrinsic typing.
-
-  A kit is a structure that abstracts over the difference between variables and terms.
-  There are always exactly two instances for Kits, one for variables and one for terms,
-
-  The core observation leading to this abstraction is that renamings
-  \texttt{S₁ →ᵣ S₂}
-  and substitutions \texttt{S₁ →ₛ S₂} only differ in their return type:
-  while a substitution returns a term \texttt{S₂ ⊢ s}, a renaming returns
-  a variable \texttt{S₂ ∋ s}.
-
-  A Kit is a structure that abstracts precisely over the commonalities
-  between variables and terms.
-
-  % \ExampleCode\FExampleSubRen
-
-  In both cases \texttt{\_∋\_} and \texttt{\_⊢\_} are instanciated to the type
-  \LibCode\KScoped
-  which allows us to create a structure to axiomatize the common
-  theory behind both of them:
+  % A \texttt{Kit} is a structure that captures the commonalities
+  % between variables and terms:
+  A \texttt{Kit} is a structure that allows to abstract over whether
+  something is a term or a variable. The intention is to instanciate
+  this structure exactly twice (once for variables and once for terms),
+  and then write definitions, which are parameterized over a kit and
+  consequently can be used for both variables and terms.
   \LibCode\KKit
 
+  As we intend to have exactly two kit instances, we choose names of
+  the form \texttt{x/y}, where \texttt{x} is the name we choose for the
+  variable instance, and \texttt{y} is the name we choose for the term
+  instance.
+  For example the parameter type \texttt{\_∋/⊢\_} will be instanciated
+  to \texttt{\_∋\_} for the variable kit, and to \texttt{\_⊢\_} for the
+  term kit.
+
+  A kit provides the following operations:
+  \begin{itemize}
+  \item 
+    \texttt{id/`} converts a variable \texttt{S ∋ s} into a
+    \texttt{S ∋/⊢ s}. For the variable kit, \texttt{S ∋/⊢ s} is instanciated to \texttt{S ∋ s}, so this
+    operation is the identity. For the term kit, \texttt{S ∋/⊢ s} is
+    instanciated to \texttt{S ⊢ s}, so this operation is the
+    variable constructor \texttt{`\_}.
+  \item 
+    \texttt{`/id} converts a \texttt{S ∋/⊢ s} into a term \texttt{S ⊢ s}
+    and is dual to the \texttt{id/`} operation.
+  \item 
+    \texttt{wk} shifts the DeBruin-Indices in a \texttt{S ∋/⊢ s}.
+    The new, unused variable \texttt{zero} can assume any sort \texttt{s'}.
+    For variables \texttt{wk} is the successor \texttt{suc}.
+    For terms \texttt{wk} means applying a shifting renaming to the term.
+  \end{itemize}
+
+  The operations need to satisfy the following properties:
+  \begin{itemize}
+  \item \texttt{`/`-is-`} states that converting a variable first to a ``variable-or-term'' and
+    then further to a term is the same as converting it directly to a
+    term using the variable constructor \texttt{`\_}.
+  \item \texttt{`/id-injective} and \texttt{id/`-injective} state that
+    \texttt{`/id} and \texttt{id/`} are injective. This property
+    follows easily, as both the identity function and the variable
+    constructor are injective.
+  \item \texttt{wk-id/`} characterizes the behaviour of the \texttt{wk} function.
+    Injecting a variable and then shifting it, is the same as injecting a shifted variable.
+  \end{itemize}
+
+  Building on the operations and axioms, we derive further operations:\footnote{
+    The definitions are included directly in the record module of the
+    Kit, so they are implicitly parameterized over a kit.
+  }
+  \begin{itemize}
+  \item 
+    We unify the definition of renamings \texttt{S₁ →ᵣ S₂} and
+    substitutions \texttt{S₁ →ₛ S₂}, which differ only in whether they
+    return variables or a terms:
+    \LibCode\KMap
+    We call a value of type \texttt{S₁ →ₖ S₂} a \emph{map}.
+  \item 
+    A map can be applied to a variable:
+    \LibCode\KAp
+  \item 
+    A map can be extended by a new variable or term:
+    \LibCode\KExt
+  \item 
+    A map can be shifted, as the Kit provides the necessary \texttt{wk} operation:
+    \LibCode\KWkm
+  \item 
+    A map can be lifted, which is the operation necessary to
+    move a map under a binder of sort \texttt{s}:
+    \LibCode\KLift
+  \item 
+    There is an identity map:
+    \LibCode\KId
+  \item 
+    There is a singleton map, which replaces \texttt{zero} and decreases
+    all other variables by one:
+    \LibCode\KSingle
+  \item 
+    There is a map which is equivalent to \texttt{wk}:
+    \LibCode\KWeaken
+  \end{itemize}
+
+  We also specify additional notation, which makes it less cumbersome
+  to explicitly pass a particular Kit to those definitions:
+  \LibCode\KKitNotation
+
+  The only task left now is to create the two instances for the kits
+  and define how a map can be applied to a term.
+  While the variable kit can be defined as expected, the term kit
+  raises a dependency problem: The \texttt{wk} operation for terms requires us
+  to already know how to apply a renaming to term.
+  To solve this problem, McBride came up with a clever solution:
+  Applying a map to a term, can be defined without even having any Kit-instances yet.
+  Afterwards we can first define the variable kit, and in the
+  definition of the term kit, we are then already allowed to apply renamings.
+  After the definition of the term-kit we are then allowed to apply
+  both renamings and substitutions as both kits defined.
+
+  \LibCode\KKitVar
+
   \subsection{Library}
-  \LibCode\KAp
-  \LibCode\KExt
-  \LibCode\KLift
-  \LibCode\KId
-  \LibCode\KSingle
-  \LibCode\KWeaken
+  \LibCode\KKitOpenInst
   \LibCode\KEq
@@
   % \LibCode\KIdLiftProof
-  \LibCode\KKitNotation
   \LibCode\KTraversal
-  \LibCode\KKitInstances
+  \LibCode\KKitVar
+  \LibCode\KKitTerm
   \LibCode\KKitOpen
@@
 
+  \section{Beyond System F}
+  \begin{itemize}
+  \item Binders with multiple variables are supported and even pattern matching can be encoded.
+  \item Heterogenous equality
+  \end{itemize}
+
   \section{Future Work}

### 2023-08-25 paper/paper.tex
Update.

@@
 \newcommand*\ExampleCode[1]{\begin{ExampleCode*}{#1}\end{ExampleCode*}}
+
+\newcommand*\ACode[1]{\AgdaFontStyle{#1}}
+\newcommand*\AField[1]{\AgdaField{#1}}
+\newcommand*\ACon[1]{\AgdaInductiveConstructor{#1}}
+\newcommand*\AKw[1]{\AgdaKeyword{#1}}
+\newcommand*\ADef[1]{\AgdaFunction{#1}}
 
@@
   \end{enumerate}
+  To avoid these drawbacks, we instead use a multi-sorted syntax.
 
   \subsection{Multi-Sorted Syntax}
-  To avoid these drawbacks, we define the syntax as a single
-  sort-indexed type of terms.
+  A multi-sorted syntax is defined by a single type of sort-indexed terms.
 
@@
   For example the parameter type \texttt{\_∋/⊢\_} will be instanciated
-  to \texttt{\_∋\_} for the variable kit, and to \texttt{\_⊢\_} for the
+  to \ACon{\_∋\_} for the variable kit, and to \AField{\_⊢\_} for the
   term kit.
@@
   \item 
-    \texttt{id/`} converts a variable \texttt{S ∋ s} into a
-    \texttt{S ∋/⊢ s}. For the variable kit, \texttt{S ∋/⊢ s} is instanciated to \texttt{S ∋ s}, so this
-    operation is the identity. For the term kit, \texttt{S ∋/⊢ s} is
-    instanciated to \texttt{S ⊢ s}, so this operation is the
-    variable constructor \texttt{`\_}.
+    \AField{id/`} converts a variable \ACode{S \ACon{∋} s} into a
+    \ACode{S \AField{∋/⊢} s}. For the variable kit, \ACode{S ∋/⊢ s} is instanciated to \ACode{S \ACon{∋} s}, so this
+    operation is the identity. For the term kit, \ACode{S ∋/⊢ s} is
+    instanciated to \ACode{S \AField{⊢} s}, so this operation is the
+    variable constructor \AField{`\_}.
   \item 
-    \texttt{`/id} converts a \texttt{S ∋/⊢ s} into a term \texttt{S ⊢ s}
-    and is dual to the \texttt{id/`} operation.
+    \AField{`/id} converts a \ACode{S ∋/⊢ s} into a term \ACode{S \AField{⊢} s}
+    and is dual to the \AField{id/`} operation.
   \item 
-    \texttt{wk} shifts the DeBruin-Indices in a \texttt{S ∋/⊢ s}.
-    The new, unused variable \texttt{zero} can assume any sort \texttt{s'}.
-    For variables \texttt{wk} is the successor \texttt{suc}.
-    For terms \texttt{wk} means applying a shifting renaming to the term.
+    \AField{wk} shifts the DeBruin-Indices in a \ACode{S ∋/⊢ s}.
+    The new, unused variable \ACon{zero} can assume any sort \ACode{s'}.
+    For variables \AField{wk} is the successor \ACon{suc}.
+    For terms \AField{wk} means applying a shifting renaming to the term.
   \end{itemize}
@@
   \begin{itemize}
-  \item \texttt{`/`-is-`} states that converting a variable first to a ``variable-or-term'' and
+  \item \AField{`/`-is-`} states that converting a variable first to a ``variable-or-term'' and
     then further to a term is the same as converting it directly to a
-    term using the variable constructor \texttt{`\_}.
-  \item \texttt{`/id-injective} and \texttt{id/`-injective} state that
-    \texttt{`/id} and \texttt{id/`} are injective. This property
+    term using the variable constructor \ACode{`\_}.
+  \item \AField{`/id-injective} and \AField{id/`-injective} state that
+    \AField{`/id} and \AField{id/`} are injective. This property
     follows easily, as both the identity function and the variable
     constructor are injective.
-  \item \texttt{wk-id/`} characterizes the behaviour of the \texttt{wk} function.
+  \item \AField{wk-id/`} characterizes the behaviour of the \AField{wk} function.
     Injecting a variable and then shifting it, is the same as injecting a shifted variable.
@@
   \item 
-    There is a map which is equivalent to \texttt{wk}:
+    There is a map which is equivalent to \texttt{wk} on terms:
     \LibCode\KWeaken

### 2023-08-25 paper/agdatex.py

@@ -21,7 +21,7 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
     prefixes = []
     mode = "none"
     stop_command_on_empty_line = False
-    def start_command(name):
+    def start_command(name, is_inline):
         global mode, tgt, prefixes
         for p in prefixes:
             name = p + name
@@ -31,7 +31,8 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
         elif mode == "command":
             print(f"ERROR: Line {line_num} starts a nested command:\n  {line}")
         mode = "command"
-        tgt += "\\newcommand*\\" + name + "{\\begin{code}\n"
+        opt = "[inline]" if is_inline else ""
+        tgt += "\\newcommand*\\" + name + "{\\begin{code}" + opt + "\n"
     def stop_command():
         global mode, tgt, prefixes
         tgt += "\\end{code}}\n"
@@ -43,9 +44,13 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
                 stop_command_on_empty_line = False
                 stop_command()
             l = l.split("!", 1)[1].strip()
+
+            is_inline = l[0] == "!"
+            if is_inline:
+                l = l[1:].strip()
             if "{" in l:
                 name = l.split(" ", 1)[0]
-                start_command(name)
+                start_command(name, is_inline)
             elif "}" in l:
                 stop_command()
             elif ">" in l:
@@ -54,7 +59,7 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
                 prefixes.pop()
             else:
                 name = l.split(" ", 1)[0]
-                start_command(name)
+                start_command(name, is_inline)
                 stop_command_on_empty_line = True
                 # print(f"ERROR: Line {line_num} contains invalid agdatex command:\n  {line}")
         elif line.strip() == "":

### 2023-08-25 paper/agdatex.py
Update.

@@ -3,10 +3,22 @@
 import os
 import sys
 from pathlib import Path
+from argparse import ArgumentParser
 import subprocess
 
-cache_dir = sys.argv[1]
-src_paths = [Path(x) for x in sys.argv[2:]]
+ap = ArgumentParser(
+    prog="agdatex",
+    description="Extract LaTeX-Macros from .agda files guided by comment annotations",
+)
+ap.add_argument("-c", "--cachedir", metavar="PATH", help="Cache directory")
+ap.add_argument("-r", "--root", metavar="PATH", help="Project root.")
+ap.add_argument("-g", "--git", help="Infer --root by searching for .git directory.")
+ap.add_argument("sources", metavar="SRC_PATH", nargs="+", help="Path to an annotated .agda-file.")
+
+args = ap.parse_args()
+
+cache_dir = args.cachedir
+src_paths = [Path(x) for x in args.sources]
 
 tgt_paths = [ p.parents[0] / (p.stem + ".lagda.tex") for p in src_paths ]
 bak_paths = [ p.parents[0] / (p.name + ".bak") for p in src_paths ]

### 2023-08-25 paper/agdatex.py
Update.

@@ -4,29 +4,107 @@ import os
 import sys
 from pathlib import Path
 from argparse import ArgumentParser
+import tempfile
 import subprocess
+import shutil
+
+
+# Parse Arguments
 
 ap = ArgumentParser(
     prog="agdatex",
     description="Extract LaTeX-Macros from .agda files guided by comment annotations",
 )
-ap.add_argument("-c", "--cachedir", metavar="PATH", help="Cache directory")
-ap.add_argument("-r", "--root", metavar="PATH", help="Project root.")
-ap.add_argument("-g", "--git", help="Infer --root by searching for .git directory.")
-ap.add_argument("sources", metavar="SRC_PATH", nargs="+", help="Path to an annotated .agda-file.")
+
+ap.add_argument("-o", "--outputdir", metavar="PATH", default="latex",
+                help="Output directory for agda's LaTeX backend. Forwarded as --latex-dir to agda. Default: 'latex'.")
+
+ap.add_argument("-e", "--exportfile", metavar="PATH",
+                help="This file will \\input all generated .tex files. Default: 'OUTPUTDIR/agda_all.tex'.")
+
+ap.add_argument("-t", "--tempdir", metavar="PATH",
+                help="Temporary directory to copy the project root to. Default: fresh system-dependent temporary dir.")
+
+ap.add_argument("-k", "--keeptempdir", metavar="PATH",
+                help="Keep temporary directory for debugging.")
+
+ap.add_argument("-r", "--root", metavar="PATH",
+                help="Project root. Default: Search for .git directory.")
+
+ap.add_argument("-i", "--index", metavar="PATH",
+                help="Write the list of generated macros to this file.")
+
+ap.add_argument("sources", metavar="SRC_PATH", nargs="+",
+                help="Path to an annotated .agda-file.")
 
 args = ap.parse_args()
 
-cache_dir = args.cachedir
-src_paths = [Path(x) for x in args.sources]
+
+# Resolve Project Root
+
+if args.root is not None:
+    root = Path(args.root)
+else:
+    # Search for .git directory.
+    root = Path(os.getcwd())
+    found_git_dir = False
+    while True:
+        for p in root.iterdir():
+            if p.is_dir() and p.name == ".git":
+                found_git_dir = True
[…]

### 2023-08-25 paper/agdatex.py
Update.

@@ -25,7 +25,7 @@ ap.add_argument("-e", "--exportfile", metavar="PATH",
 ap.add_argument("-t", "--tempdir", metavar="PATH",
                 help="Temporary directory to copy the project root to. Default: fresh system-dependent temporary dir.")
 
-ap.add_argument("-k", "--keeptempdir", metavar="PATH",
+ap.add_argument("-k", "--keeptempdir", action='store_true',
                 help="Keep temporary directory for debugging.")
 
 ap.add_argument("-r", "--root", metavar="PATH",
@@ -62,6 +62,7 @@ else:
 
 root = root.absolute()
 
+
 # Check if sources are relative to project root
 
 src_paths = []
@@ -76,14 +77,11 @@ for p in args.sources:
 # Create temporary directory
 
 if args.tempdir is not None:
-    tmp_dir = Path(args.tempdir)
-    tmp_dir.mkdir(exist_ok=True)
+    tmp_root = Path(args.tempdir)
+    tmp_root.mkdir(exist_ok=True)
 else:
-    tmp_dir_obj = tempfile.TemporaryDirectory(prefix="agdatex")
-    tmp_dir = Path(tmp_dir_obj.name)
-
-tmp_root = tmp_dir / "agda"
-tmp_root.mkdir(exist_ok=True)
+    tmp_root_obj = tempfile.TemporaryDirectory(prefix="agdatex")
+    tmp_root = Path(tmp_root_obj.name)
 
 
 # Copy project root to temporary directory
@@ -93,7 +91,7 @@ shutil.copytree(
     tmp_root,
     dirs_exist_ok=True,
     symlinks=True,
-    ignore=shutil.ignore_patterns(tmp_dir)
+    ignore=shutil.ignore_patterns(tmp_root)
 )
 
 
@@ -206,7 +204,8 @@ for tgt_path in tgt_paths:
     )
 
 
-# Create a agda-includes.tex file which includes all generated .tex-files.
+# Create an export .tex-file which imports all generated .tex-files.
+
 if args.exportfile is not None:
     export_file = Path(args.exportfile)
 else:
@@ -216,7 +215,6 @@ src_names = [ p.stem for p in src_paths ]
 s = ""
 for p in output_dir.glob("**/*.tex"):
     if p.stem in src_names:
-        print("in")
         s += "\\input{" + str(p.relative_to(output_dir)) + "}\n"
 with open(export_file, 'w') as f:
     f.write(s)

### 2023-08-26 paper/agdatex.py
Update.

@@ -20,7 +20,7 @@ ap.add_argument("-o", "--outputdir", metavar="PATH", default="latex",
                 help="Output directory for agda's LaTeX backend. Forwarded as --latex-dir to agda. Default: 'latex'.")
 
 ap.add_argument("-e", "--exportfile", metavar="PATH",
-                help="This file will \\input all generated .tex files. Default: 'OUTPUTDIR/agda_all.tex'.")
+                help="This file will \\input all generated .tex files. Both .tex and .sty are supported. Default: 'OUTPUTDIR/agda-generated.sty'.")
 
 ap.add_argument("-t", "--tempdir", metavar="PATH",
                 help="Temporary directory to copy the project root to. Default: fresh system-dependent temporary dir.")
@@ -209,10 +209,12 @@ for tgt_path in tgt_paths:
 if args.exportfile is not None:
     export_file = Path(args.exportfile)
 else:
-    export_file = output_dir / "agda_all.tex"
+    export_file = output_dir / "agda-generated.sty"
 
 src_names = [ p.stem for p in src_paths ]
 s = ""
+if export_file.suffix == ".sty":
+    s += "\ProvidesPackage{" + export_file.stem + "}\n"
 for p in output_dir.glob("**/*.tex"):
     if p.stem in src_names:
         s += "\\input{" + str(p.relative_to(output_dir)) + "}\n"

### 2023-08-26 paper/paper.tex
Update.

@@
 \usepackage{agda-unicode}
+\usepackage{agda-generated}
+\AgdaNoSpaceAroundCode{}
 
@@
 
-  % The following line will be replaced by make with an \input{} for each
-  % .lagda.tex file:
-
-  % SED MARKER
-
-  \AgdaNoSpaceAroundCode{}
-
   \section{Introduction}
@@
 
-  The term type \texttt{S~⊢~s} is then indexed by its sort \texttt{s} and the sorts of its free variables \texttt{S}.
-  For example, \texttt{[𝕥, 𝕥] ⊢ 𝕖} is the type of expressions (\texttt{𝕖}) with two free type-variables (\texttt{𝕥}).
+  The term type \ACode{S \ACon{⊢} s} is then indexed by its sort
+  \ACode{s} and the sorts of its free variables \ACode{S}.
+  For example, \ACon{[𝕥, 𝕥] ⊢ 𝕖} is the type of expressions (\ACon{𝕖}) with two free type-variables (\ACon{𝕥}).
   \newpage
   \AppCode\FSyntax
-  The notation \texttt{\_⊢\_} is often used for terms in
+  The notation \ACon{\_⊢\_} is often used for terms in
   intrinsically-typed languages. This is no accident: in effect, we
@@
   typing relation assures exactly that the syntactic categories are followed.
-  Sorts \texttt{s} correspond to types, and lists of sorts \texttt{S}
+  Sorts \ACode{s} correspond to types, and lists of sorts \ACode{S}
   correspond to type environments.
@@
   As it is typical in intrinsic typing, variables are represented as
-  typed (in our case sorted) DeBruijn indices \texttt{S~∋~s}, i.e.\
+  typed (in our case sorted) DeBruijn indices \ACode{S \ACon{∋} s}, i.e.\
   values of the usual proof-relevant list-membership relation:
@@
   Contrary to the usual intrinsic typing, we limit the environment
-  \texttt{S} to sorts with index \texttt{Var}.
+  \ACode{S} to sorts with index \ACon{Var}.
   This ensures that it is impossible to construct kind variables, as
-  it is not possible to extend \texttt{S} with the sort of kinds
-  \texttt{𝕜}.
+  it is not possible to extend \ACode{S} with the sort of kinds
+  \ACon{𝕜}.
 
@@
   The last field records that variable introduction
-  \texttt{`\_} is injective, which is trivially true for
+  \AField{`\_} is injective, which is trivially true for
   constructors. The instantiation for our System F syntax is
@@
 
-  A renaming \texttt{S₁ →ᵣ S₂} maps variables from \texttt{S₁} to
-  variables from \texttt{S₂}.
-  A substitution \texttt{S₁ →ₛ S₂} maps variables from \texttt{S₁} to
-  terms with free variables from \texttt{S₂}.
+  A renaming \ACode{S₁ \ADef{→ᵣ} S₂} maps variables from \ACode{S₁} to
+  variables from \ACode{S₂}.
+  A substitution \ACode{S₁ \ADef{→ₛ} S₂} maps variables from \ACode{S₁} to
+  terms with free variables from \ACode{S₂}.
 
@@
   A problem with mechanizing languages with DeBruijn-indices is that
-  the operation of applying a substitution \texttt{σ} to a term
-  \texttt{t} cannot be defined directly by structural recursion on
-  \texttt{t}.
-  If \texttt{t} is a term like \texttt{λx e},
-  where the subterm \texttt{e} is under a binder, then the variables
-  in \texttt{e} are shifted, i.e.\ the variable \texttt{zero} refers
-  to the variable from the binder, whereas a variable \texttt{suc x}
-  refers to the variable \texttt{x} from outside of the binder.
-  As the terms contained in \texttt{σ} may themselves contain free variables,
+  the operation of applying a substitution \ACode{σ} to a term
+  \ACode{t} cannot be defined directly by structural recursion on
+  \ACode{t}.
+  If \ACode{t} is a term like \ACode{\ACon{λx} e},
+  where the subterm \ACode{e} is under a binder, then the variables
+  in \ACode{e} are shifted, i.e.\ the variable \ACon{zero} refers
+  to the variable from the binder, whereas a variable \ACode{\ACon{suc} x}
+  refers to the variable \ACode{x} from outside of the binder.
+  As the terms contained in \ACode{σ} may themselves contain free variables,
   they need to be shifted, too, before they can be substituted
-  for variables in \texttt{e}. 
+  for variables in \ACode{e}. 
   Shifting the variables in a term is itself a substitution, which maps
-  each variable \texttt{x} to the term \texttt{` (suc x)}.
+  each variable \ACode{x} to the term \ACode{\ACon{`} (\ACon{suc} x)}.
   However, applying this shifting substitution to the terms in
-  \texttt{σ} is not structurally recursive, as the terms in \texttt{σ}
-  are not subterms of \texttt{λx e}.
+  \ACode{σ} is not structurally recursive, as the terms in \ACode{σ}
+  are not subterms of \ACode{\ACon{λx} e}.
 
@@
   In contrast to general substitutions, renamings can be easily
-  shifted: they only contain variables \texttt{x}, which can be
-  shifted by simply taking their successor \texttt{suc x}.
+  shifted: they only contain variables \ACode{x}, which can be
+  shifted by simply taking their successor \ACode{\ACon{suc} x}.
   As the shifting substitution is a renaming, we can then define substitution,
-  where we apply the shifting \emph{renaming} to the terms in \texttt{σ} when
+  where we apply the shifting \emph{renaming} to the terms in \ACode{σ} when
   the substitution needs to go under a binder.
@@
 
-  % A \texttt{Kit} is a structure that captures the commonalities
+  % A \ACode{Kit} is a structure that captures the commonalities
   % between variables and terms:
-  A \texttt{Kit} is a structure that allows to abstract over whether
+  A \ACode{Kit} is a structure that allows to abstract over whether
   something is a term or a variable. The intention is to instanciate
@@
   As we intend to have exactly two kit instances, we choose names of
-  the form \texttt{x/y}, where \texttt{x} is the name we choose for the
-  variable instance, and \texttt{y} is the name we choose for the term
+  the form \ACode{x/y}, where \ACode{x} is the name we choose for the
+  variable instance, and \ACode{y} is the name we choose for the term
   instance.
-  For example the parameter type \texttt{\_∋/⊢\_} will be instanciated
+  For example the parameter type \ACode{\_∋/⊢\_} will be instanciated
   to \ACon{\_∋\_} for the variable kit, and to \AField{\_⊢\_} for the
@@
     then further to a term is the same as converting it directly to a
-    term using the variable constructor \ACode{`\_}.
+    term using the variable constructor \AField{`\_}.
   \item \AField{`/id-injective} and \AField{id/`-injective} state that
@@
   \item 
-    We unify the definition of renamings \texttt{S₁ →ᵣ S₂} and
-    substitutions \texttt{S₁ →ₛ S₂}, which differ only in whether they
+    We unify the definition of renamings \ACode{S₁ \ADef{→ᵣ} S₂} and
+    substitutions \ACode{S₁ \ADef{→ₛ} S₂}, which differ only in whether they
     return variables or a terms:
     \LibCode\KMap
-    We call a value of type \texttt{S₁ →ₖ S₂} a \emph{map}.
+    We call a value of type \ACode{S₁ \ADef{→ₖ} S₂} a \emph{map}.
   \item 
@@
   \item 
-    A map can be shifted, as the Kit provides the necessary \texttt{wk} operation:
+    A map can be shifted, as the Kit provides the necessary \AField{wk} operation:
     \LibCode\KWkm
@@
     A map can be lifted, which is the operation necessary to
-    move a map under a binder of sort \texttt{s}:
+    move a map under a binder of sort \ACode{s}:
     \LibCode\KLift
@@
   \item 
-    There is a singleton map, which replaces \texttt{zero} and decreases
+    There is a singleton map, which replaces \ACon{zero} and decreases
     all other variables by one:
@@
   \item 
-    There is a map which is equivalent to \texttt{wk} on terms:
+    There is a map which is equivalent to \AField{wk} on terms:
     \LibCode\KWeaken
@@
   While the variable kit can be defined as expected, the term kit
-  raises a dependency problem: The \texttt{wk} operation for terms requires us
+  raises a dependency problem: The \AField{wk} operation for terms requires us
   to already know how to apply a renaming to term.

### 2023-08-26 paper/paper.tex
Update.

@@
 
-\newcommand*\ACode[1]{\AgdaFontStyle{#1}}
+% \setlength\fboxsep{0pt}
+% \usepackage{realboxes}
+% \definecolor{mygray}{rgb}{0.95,0.95,0.95}
+% \newcommand*\ACode[1]{\Colorbox{mygray}{\AgdaFontStyle{#1}}}
+
+% \newcommand*\ACode[1]{\AgdaFontStyle{#1}}
+
+% \newcommand*\ACode[1]{\texttt{#1}}
+
+\definecolor{mygray}{rgb}{0.4,0.4,0.4}
+\newcommand*\ACode[1]{\AgdaFontStyle{\textcolor{mygray}{#1}}}
+
 \newcommand*\AField[1]{\AgdaField{#1}}
@@
   \ACode{s} and the sorts of its free variables \ACode{S}.
-  For example, \ACon{[𝕥, 𝕥] ⊢ 𝕖} is the type of expressions (\ACon{𝕖}) with two free type-variables (\ACon{𝕥}).
+  For example, \ACode{\ACon{[𝕥, 𝕥] ⊢ 𝕖}} is the type of expressions
+  (\ACode{\ACon{𝕖}}) with two free type-variables (\ACode{\ACon{𝕥}}).
   \newpage
   \AppCode\FSyntax
-  The notation \ACon{\_⊢\_} is often used for terms in
+  The notation \ACode{\ACon{\_⊢\_}} is often used for terms in
   intrinsically-typed languages. This is no accident: in effect, we
@@
   Contrary to the usual intrinsic typing, we limit the environment
-  \ACode{S} to sorts with index \ACon{Var}.
+  \ACode{S} to sorts with index \ACode{\ACon{Var}}.
   This ensures that it is impossible to construct kind variables, as
   it is not possible to extend \ACode{S} with the sort of kinds
-  \ACon{𝕜}.
+  \ACode{\ACon{𝕜}}.
 
@@
   The last field records that variable introduction
-  \AField{`\_} is injective, which is trivially true for
+  \ACode{\AField{`\_}} is injective, which is trivially true for
   constructors. The instantiation for our System F syntax is
@@
   where the subterm \ACode{e} is under a binder, then the variables
-  in \ACode{e} are shifted, i.e.\ the variable \ACon{zero} refers
+  in \ACode{e} are shifted, i.e.\ the variable \ACode{\ACon{zero}} refers
   to the variable from the binder, whereas a variable \ACode{\ACon{suc} x}
@@
   For example the parameter type \ACode{\_∋/⊢\_} will be instanciated
-  to \ACon{\_∋\_} for the variable kit, and to \AField{\_⊢\_} for the
+  to \ACode{\ACon{\_∋\_}} for the variable kit, and to \ACode{\AField{\_⊢\_}} for the
   term kit.
@@
   \item 
-    \AField{id/`} converts a variable \ACode{S \ACon{∋} s} into a
+    \ACode{\AField{id/`}} converts a variable \ACode{S \ACon{∋} s} into a
     \ACode{S \AField{∋/⊢} s}. For the variable kit, \ACode{S ∋/⊢ s} is instanciated to \ACode{S \ACon{∋} s}, so this
@@
     instanciated to \ACode{S \AField{⊢} s}, so this operation is the
-    variable constructor \AField{`\_}.
+    variable constructor \ACode{\AField{`\_}}.
   \item 
-    \AField{`/id} converts a \ACode{S ∋/⊢ s} into a term \ACode{S \AField{⊢} s}
-    and is dual to the \AField{id/`} operation.
+    \ACode{\AField{`/id}} converts a \ACode{S ∋/⊢ s} into a term \ACode{S \AField{⊢} s}
+    and is dual to the \ACode{\AField{id/`}} operation.
   \item 
-    \AField{wk} shifts the DeBruin-Indices in a \ACode{S ∋/⊢ s}.
-    The new, unused variable \ACon{zero} can assume any sort \ACode{s'}.
-    For variables \AField{wk} is the successor \ACon{suc}.
-    For terms \AField{wk} means applying a shifting renaming to the term.
+    \ACode{\AField{wk}} shifts the DeBruin-Indices in a \ACode{S ∋/⊢ s}.
+    The new, unused variable \ACode{\ACon{zero}} can assume any sort \ACode{s'}.
+    For variables \ACode{\AField{wk}} is the successor \ACode{\ACon{suc}}.
+    For terms \ACode{\AField{wk}} means applying a shifting renaming to the term.
   \end{itemize}
@@
   \begin{itemize}
-  \item \AField{`/`-is-`} states that converting a variable first to a ``variable-or-term'' and
+  \item \ACode{\AField{`/`-is-`}} states that converting a variable first to a ``variable-or-term'' and
     then further to a term is the same as converting it directly to a
-    term using the variable constructor \AField{`\_}.
-  \item \AField{`/id-injective} and \AField{id/`-injective} state that
-    \AField{`/id} and \AField{id/`} are injective. This property
+    term using the variable constructor \ACode{\AField{`\_}}.
+  \item \ACode{\AField{`/id-injective}} and \ACode{\AField{id/`-injective}} state that
+    \ACode{\AField{`/id}} and \ACode{\AField{id/`}} are injective. This property
     follows easily, as both the identity function and the variable
     constructor are injective.
-  \item \AField{wk-id/`} characterizes the behaviour of the \AField{wk} function.
+  \item \ACode{\AField{wk-id/`}} characterizes the behaviour of the \ACode{\AField{wk}} function.
     Injecting a variable and then shifting it, is the same as injecting a shifted variable.
@@
     substitutions \ACode{S₁ \ADef{→ₛ} S₂}, which differ only in whether they
-    return variables or a terms:
+    return variables or terms:
     \LibCode\KMap
@@
   \item 
-    A map can be shifted, as the Kit provides the necessary \AField{wk} operation:
+    A map can be shifted, as the Kit provides the necessary \ACode{\AField{wk}} operation:
     \LibCode\KWkm
@@
   \item 
-    There is a singleton map, which replaces \ACon{zero} and decreases
+    There is a singleton map, which replaces \ACode{\ACon{zero}} and decreases
     all other variables by one:
@@
   \item 
-    There is a map which is equivalent to \AField{wk} on terms:
+    There is a map which is equivalent to \ACode{\AField{wk}} on terms:
     \LibCode\KWeaken
+  \item
+    Maps have an extensional equality:
+    \LibCode\KEq
+  \item
+    For simplicity we postulate functional extensionality:\footnote{
+      The actual implementation does \emph{not} use any postulates.
+      Appendix TODO explains the workaround.
+    }
+    \LibCode\KFunExt
+  \item
+    Lifting the identity map is a (larger) identity map:
+    \LibCode\KIdLift
+    % \LibCode\KIdLiftProof
   \end{itemize}
 
-  We also specify additional notation, which makes it less cumbersome
+  We also specify notation, which makes it less cumbersome
   to explicitly pass a particular Kit to those definitions:
-  \LibCode\KKitNotation
+  \begin{itemize}
+  \item \ACode{S \ADef{∋/⊢[} K \ADef{]} s} is \ACode{S ∋/⊢ s} for some \ACode{\ACon{Kit}} \ACode{K}; and
+  \item \ACode{S \ADef{–[} K \ADef{]→} s} is \ACode{S \ADef{→ₖ} s} for some \ACode{\ACon{Kit}} \ACode{K}.
+  \end{itemize}
 
-  The only task left now is to create the two instances for the kits
-  and define how a map can be applied to a term.
-  While the variable kit can be defined as expected, the term kit
-  raises a dependency problem: The \AField{wk} operation for terms requires us
-  to already know how to apply a renaming to term.
-  To solve this problem, McBride came up with a clever solution:
-  Applying a map to a term, can be defined without even having any Kit-instances yet.
-  Afterwards we can first define the variable kit, and in the
-  definition of the term kit, we are then already allowed to apply renamings.
-  After the definition of the term-kit we are then allowed to apply
-  both renamings and substitutions as both kits defined.
+  The operation of applying a map to a term depends on the concrete
+  structure of a term, so we define a structure for it:
+  \LibCode\KTraversal
+  The fields of this structure are:
+  \begin{itemize}
+  \item
+    \ACode{t \AField{⋯} ϕ} applies the map \ACode{ϕ} (a renaming or
+    substitution) to the term \ACode{t}.
+  \item
+    \ACode{\AField{⋯-var}} states that applying a map \ACode{ϕ} to a variable
+    term \ACode{\AField{`} x}, is the same as applying \ACode{ϕ} to
+    the variable \ACode{x}, and then converting the result from \ACode{S₂ ∋/⊢ S}
+    to \ACode{S₂ ⊢ S} via \ACode{\AField{id/`}}.
+  \item
+    \ACode{\AField{⋯-id}} states that applying the identity map
+    \ACode{id} to a term does not change the term.
+  \end{itemize}
 
+  Now we are finally ready to define the actual kit instances.
+  The variable kit definition is straightforward:
   \LibCode\KKitVar
 
-  \subsection{Library}
-  \LibCode\KKitOpenInst
-  \LibCode\KEq
-  \LibCode\KFunExt
-  \LibCode\KIdLift
-  % \LibCode\KIdLiftProof
-  \LibCode\KTraversal
-  \LibCode\KKitVar
+  The term kit requires both the variable kit and the
+  \ACode{\ACon{Traversal}} to be defined, because shifting a term \ACode{t} with \ACode{\AField{wk}}
+  means applying the shifting renaming to the term.
+  Hence, we define the term kit in the record module of \ACode{\ACon{Traversal}}:
   \LibCode\KKitTerm
-  \LibCode\KKitOpen
-  \subsection{System F}
+
+  % The only task left now is to create the two instances for the kits
+  % and define how a map can be applied to a term.
+  % While the variable kit can be defined as expected, the term kit
+  % raises a dependency problem: The \ACode{\AField{wk}} operation for terms requires us
+  % to already know how to apply a renaming to term.
+  % To solve this problem, McBride came up with a clever solution:
+  % Applying a map to a term, can be defined without even having any Kit-instances yet.
+  % Afterwards we can first define the variable kit, and in the
+  % definition of the term kit, we are then already allowed to apply renamings.
+  % After the definition of the term-kit we are then allowed to apply
+  % both renamings and substitutions as both kits defined.
+
+  Note that all definitions so far are library code. To make them
+  available for our System F example, all we need to do is
+  to create an instance of \ACode{\ACon{Traversal}}.
+
+  We start with the operation of applying a map to a term:
   \AppCode\FTraversalOp
+  The interesting cases are those with variables and binders:
+  \begin{itemize}
+  \item 
+    In the variable case \ACode{(\ACon{`} x) \ADef{⋯} ϕ}, we first apply the map
+    \ACode{ϕ} to the variable \ACode{x}. If \ACode{ϕ} is a renaming,
+    we get back a variable and need to apply the variable constructor
+    \ACode{\ACon{`\_}} to it. If \ACode{ϕ} is a substitution,
+    we get back a term that we can use directly.
+    Hence, the use of \ACode{`/id}.
+  \item 
+    In cases where the operation needs to go under a binder, like
+    \ACode{(\ACon{λx} e) \ADef{⋯} ϕ}, we lift the map using \AField{\_↑\_} before
+    we apply it to the subterm, to account for the bound variable.
+  \end{itemize}
+
   \AppCode\FTraversalId
@@
 
+  \ExampleCode\FExampleTrav
+
+  % \subsection{Unused}
+  % \LibCode\KKitOpenInst
+  % \LibCode\KKitOpen
+
   \section{Map Composition}
-  \subsection{Library}
+  An operation we did not cover in the previous section is the
+  composition of substitutions.
+  This operation and its properties are relevant in numerous places, e.g.\
+  TODO, which are also relevant for System F.
+
+  As we defined substitution in terms of renamings, we now have to
+  consider all four compositions between renamings and substitutions,
+  and their properties. The problem is again that they are dependent
+  on each other, i.e.\ composition between two substitutions relies on
+  the compositions between renamings and substitutions, which rely on
+  compositions between two renamings.
+
+  McBride et al.\cite{DBLP:journals/jar/BentonHKM12} addressed this
+  issue by duplicating the definitions and using tactics to reduce
+  boilerplate in proofs.
+
+  In contrast we define a structure similar to Kits, which allows us
+  to abstract over all four compositions and use the same trick as
+  before to eliminate the dependencies.
+  This has the benefit that a user of our library has to write less
+  boilerplate and also that we gain more expressiveness, which helps
+  us later to unify type preservation of all substitution and renaming
+  operations into a single lemma.
+
+  We first define a \ACode{\ACon{WkKit}}, which captures properties
+  between maps and map application, which we could not include in the
+  \ACode{\ACon{Kit}} definition, as we already require
+  \ACode{\ACon{Kit}}s to define a Traversal. We then create exactly
+  two instances of the \ACode{\ACon{WkKit}} for the variable and term
+  kit, respectively. Both the structure and the instances are part of
+  the library and do not require application code.
+
+  Using the \ACode{\ACon{WkKit}}, we then define
+  the \ACode{\ACon{ComposeKit}} and \ACode{\ACon{ComposeTraversal}}
+  structures, which serve analogous purposes to \ACode{\ACon{Kit}}
+  and \ACode{\ACon{Traversal}} from before:
+  the library provides a \ACode{\ACon{ComposeKit}} for each of the
+  four compositions, whereas the \ACode{\ACon{ComposeTraversal}}
+  structure needs to be instanciated by the application with
+  lemmas parameterized over arbitrary \ACode{\ACon{ComposeKit}}s.
+  This gives the user a composition operation
+  \ACode{\AField{\_·ₖ\_}} between arbitrary maps and corresponding
+  lemmas, instead of multiple definitions and lemmas about concrete
+  maps, enabling us to reason on a more abstract level and to avoid
+  further boilerplate.
+
+  \subsection{WkKits}
+
   \LibCode\KWkKit
   \LibCode\KWkKitInstances
+
+  \subsection{Composition}
+
   \LibCode\KComposeKit
@@
   % \LibCode\KDistLiftSingleTraverseProof
+
   \subsection{System F}

### 2023-08-29 paper/paper.tex
Update.

@@
   \subsection{A Structure for Multi-Sorted Syntax}
-  The multi-sorted syntax makes it easy to capture general syntax definitions
-  in a structure:
+  The multi-sorted syntax makes it easy to define a structure for
+  syntaxes of arbitrary object-languages, i.e.\ syntaxes with
+  arbitrary amounts of syntactic categories and variables:
   \LibCode\KTerms

### 2023-08-29 paper/paper.tex
Update.

@@
   arbitrary amounts of syntactic categories and variables:
-  \LibCode\KTerms
+  \LibCode\KSyntax
   The first three fields record the definitions of sorts, terms, and variable introduction.
@@
   straightforward:
-  \AppCode\FTerms
+  \AppCode\FSyntaxInst
 
@@
     The new, unused variable \ACode{\ACon{zero}} can assume any sort \ACode{s'}.
-    For variables \ACode{\AField{wk}} is the successor \ACode{\ACon{suc}}.
-    For terms \ACode{\AField{wk}} means applying a shifting renaming to the term.
+    For variables, \ACode{\AField{wk}} is the successor \ACode{\ACon{suc}}.
+    For terms, \ACode{\AField{wk}} means applying a shifting renaming to the term.
   \end{itemize}
@@
 
-  We first define a \ACode{\ACon{WkKit}}, which captures properties
-  between maps and map application, which we could not include in the
-  \ACode{\ACon{Kit}} definition, as we already require
-  \ACode{\ACon{Kit}}s to define a Traversal. We then create exactly
-  two instances of the \ACode{\ACon{WkKit}} for the variable and term
-  kit, respectively. Both the structure and the instances are part of
-  the library and do not require application code.
+  \subsection{WkKits}
+
+  In the definition of \ACode{\ACon{Kit}} we required the axiom
+  \ACode{\AField{wk-id/`}}, which captured the interaction between
+  \ACode{\AField{wk}} and \ACode{\AField{id/`}}:
+  \ExampleCode{\KWkId}
+  Injecting a variable with \ACode{\AField{id/`}} and then shifting it,
+  is the same as injecting a shifted variable.
+
+  Now that we have defined the \ACode{\ACon{Traversal}} structure, we
+  can formulate a similar interaction between
+  \ACode{\AField{wk}} and \ACode{\AField{`/id}}:
+  \ExampleCode{\KIdWk}
+  Shifting a variable/term \ACode{x/t} with \ACode{\AField{wk}} and then injecting
+  it to a term, is the same as injecting \ACode{x/t} to a term and then applying
+  a shifting renaming.
+
+  As this lemma depends on map application \ACode{\ADef{\_⋯\_}} being
+  already defined, we cannot add it to the original \ACode{\ACon{Kit}}
+  structure, but need to define a new structure for it, which we call
+  a \ACode{\ACon{WkKit}}:
+  \LibCode\KWkKit
+
+  We then create exactly two instances of the \ACode{\ACon{WkKit}} for
+  the variable and term kit, respectively. 
+  \LibCode\KWkKitInstances
+
+  Both the structure and the instances are part of the library and do
+  not require application code.
+
+  \subsection{Composition}
 
@@
 
-  \subsection{WkKits}
-
-  \LibCode\KWkKit
-  \LibCode\KWkKitInstances
-
-  \subsection{Composition}
+  \ExampleCode\KExFourComps
+  \ExampleCode\KExTwoComps
+  \ExampleCode\KComposition
 

### 2023-08-30 paper/paper.tex
Update.

@@
     \LibCode\KMap
-    We call a value of type \ACode{S₁ \ADef{→ₖ} S₂} a \emph{map}.
+    We call a value of type \ACode{S₁ \ADef{→ₖ} S₂} a \emph{map} and
+    use the meta-variable \ACode{ϕ} for it.
   \item 
@@
 
-  Now we are finally ready to define the actual kit instances.
+  Finally, we define the actual kit instances.
   The variable kit definition is straightforward:
@@
   \section{Map Composition}
-  An operation we did not cover in the previous section is the
-  composition of substitutions.
-  This operation and its properties are relevant in numerous places, e.g.\
-  TODO, which are also relevant for System F.
+  In this section we extend our framework with the composition of
+  arbitrary maps.
 
-  As we defined substitution in terms of renamings, we now have to
-  consider all four compositions between renamings and substitutions,
-  and their properties. The problem is again that they are dependent
-  on each other, i.e.\ composition between two substitutions relies on
-  the compositions between renamings and substitutions, which rely on
-  compositions between two renamings.
+  Composition of substitutions is important for most language
+  formalizations, as it allows to reason about the meaning of applying
+  multiple substitutions to a term. This appears for example in the proof of
+  substitution-preserves-typing for System F, where in the
+  case of type-application \ACode{(e \ACon{∙} t) \ADef{⋯} σ} we need to reason about
+  \ACode{(e \ADef{⋯} \ADef{⦅} t \ADef{⦆}) \ADef{⋯} σ}, which is equivalent to
+  \ACode{e \ADef{⋯} (\ADef{⦅} t \ADef{⦆} \ADef{·ₖ} σ)},
+  where \ACode{\_\ADef{·ₖ}\_} is backward composition.
+
+  As we defined substitution in terms of renamings, we need to
+  consider all four compositions between renamings and substitutions.
+  While the composition operations are defined independently of each other,
+  the \ACode{\AField{⋯-assoc}} lemma for two substitutions, depends on the
+  \ACode{\AField{⋯-assoc}} lemmas between a renaming and a substitution, which in
+  turn depend on the \ACode{\AField{⋯-assoc}} lemma between two renamings.
 
@@
 
-  In contrast we define a structure similar to Kits, which allows us
-  to abstract over all four compositions and use the same trick as
-  before to eliminate the dependencies.
+  In contrast we define structures similar to \ACode{\ACon{Kit}} and
+  \ACode{\ACon{Traversal}}, which allows us to abstract over all four
+  compositions and use the same trick as before to eliminate the
+  dependencies.
   This has the benefit that a user of our library has to write less
-  boilerplate and also that we gain more expressiveness, which helps
-  us later to unify type preservation of all substitution and renaming
-  operations into a single lemma.
+  boilerplate, but more importantly it allows to prove lemmas about the
+  composition of arbitrary maps, which enables us to later unify type
+  preservation of all substitution and renaming operations into a single
+  lemma.
 
@@
   \begin{itemize}
-  \item Binders with multiple variables are supported and even pattern matching can be encoded.
-  \item Heterogenous equality
+  \item
+    Binders with multiple variables are supported and even pattern
+    matching can be encoded.
+  \item
+    Heterogenous equality
   \end{itemize}
@@
   \section{Future Work}
+  \begin{itemize}
+  \item
+    Enabling more object languages by allowing subterms with less
+    restrictions on the form of \ACode{S}.
+  \end{itemize}
 

### 2023-08-30 paper/paper.tex
Update.

@@
   boilerplate, but more importantly it allows to prove lemmas about the
-  composition of arbitrary maps, which enables us to later unify type
+  composition of arbitrary maps, which enable us later to unify type
   preservation of all substitution and renaming operations into a single
   lemma.
+
+  \subsection{An Examination of Composition}
+  With the infrastructure we build in the previous sections, we can
+  define the four compositions as followed:
+  \ExampleCode\KExFourComps
+  The definitions reveal an interesting asymetry, in the sense that
+  the first argument of composition guides the implementation: if we
+  compose a renaming with any map, then we first apply renaming to the
+  the variable, which yields another variable, so we apply the second
+  map to that \emph{variable}.
+  However, if we compose a substitution with any map, then we first
+  apply the substitution to that variable, which yields a term, so we
+  need to apply the second map to that \emph{term}.
+
+  With the machinery, we built so far, we can unify the first two and
+  last two compositions as followed:
+  \ExampleCode\KExTwoComps
+
+  But to unify those two compositions, we need a new abstraction.
+
+  \ExampleCode\KComposition
 
@@
   further boilerplate.
-
-  \ExampleCode\KExFourComps
-  \ExampleCode\KExTwoComps
-  \ExampleCode\KComposition
 

### 2023-08-30 paper/paper.tex
Update.

@@
 
-  But to unify those two compositions, we need a new abstraction.
-
+  But to unify those two compositions, we need a new abstraction:
   \ExampleCode\KComposition
+  The Kit-abstraction from before allowed us to abstract over what we
+  apply, i.e.\ a renaming or a substitution, but not about what we apply it
+  to, i.e.\ a variable or a term. For the latter we still have two
+  distinct operations \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}.
+  A \ACode{\ACon{ComposeKit} K₁ K₂ K₁⊔K₂} fills exactly that gap by providing
+  the operation \ACode{\ADef{\_\&/⋯\_}} which applies a renaming/substitution (\ACode{K₂})
+  to a variable/term (\ACode{K₁}) and produces a variable/term (\ACode{K₁⊔K₂}).
 

### 2023-08-30 paper/paper.tex
Update.

@@
   A \ACode{\ACon{ComposeKit} K₁ K₂ K₁⊔K₂} fills exactly that gap by providing
-  the operation \ACode{\ADef{\_\&/⋯\_}} which applies a renaming/substitution (\ACode{K₂})
+  the operation \ACode{\AField{\_\&/⋯\_}} which applies a renaming/substitution (\ACode{K₂})
   to a variable/term (\ACode{K₁}) and produces a variable/term (\ACode{K₁⊔K₂}).
+  We want to have a ComposeKit for each \ACode{K₁} and \ACode{K₂},
+  where the output Kit \ACode{K₁⊔K₂}, is defined as a function of
+  \ACode{K₁} and \ACode{K₂}, similar to functional dependencies in Haskell.
+  If \ACode{K₁} and \ACode{K₂} are both the variable kit \ACode{Kᵣ},
+  then \ACode{K₁⊔K₂} is also the variable kit \ACode{Kᵣ}. But if either
+  \ACode{K₁} or \ACode{K₂} is a term kit \ACode{Kₛ}, then also
+  \ACode{K₁⊔K₂} is the term kit \ACode{Kₛ}. This is captured in the
+  notation \ACode{K₁⊔K₂}, referring to the partial order generated by
+  \ACode{Kᵣ < Kₛ}.
+
+  \ACode{\ACon{ComposeKit}} and \ACode{\ACon{\_\&/⋯\_}} are analogous to
+  \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
+  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-assoc}} are analogous to
+  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{\_⋯\_}}.
 

### 2023-09-08 paper/paper.tex
Update.

--- 2023-08-30 paper/paper.tex
@@
 
-  As we intend to have exactly two kit instances, we choose names of
+  As we intend to have exactly two \ACode{\ACon{Kit}} instances, we choose names of
   the form \ACode{x/y}, where \ACode{x} is the name we choose for the
@@
   The term kit requires both the variable kit and the
-  \ACode{\ACon{Traversal}} to be defined, because shifting a term \ACode{t} with \ACode{\AField{wk}}
+  \ACode{\ACon{Traversal}} to be defined, because shifting a term with \ACode{\AField{wk}}
   means applying the shifting renaming to the term.
@@
     \ACode{(\ACon{λx} e) \ADef{⋯} ϕ}, we lift the map using \AField{\_↑\_} before
-    we apply it to the subterm, to account for the bound variable.
+    we apply it to the subterm to account for the bound variable.
   \end{itemize}
@@
   \subsection{An Examination of Composition}
-  With the infrastructure we build in the previous sections, we can
-  define the four compositions as followed:
+  To motivate our abstraction for composition, we first look
+  at the four compositions individually:
   \ExampleCode\KExFourComps
-  The definitions reveal an interesting asymetry, in the sense that
-  the first argument of composition guides the implementation: if we
-  compose a renaming with any map, then we first apply renaming to the
-  the variable, which yields another variable, so we apply the second
-  map to that \emph{variable}.
-  However, if we compose a substitution with any map, then we first
-  apply the substitution to that variable, which yields a term, so we
-  need to apply the second map to that \emph{term}.
-
-  With the machinery, we built so far, we can unify the first two and
-  last two compositions as followed:
+  The definitions reveal two interesting properties:
+  \begin{enumerate}
+  \item
+    If we compose two maps \ACode{ϕ₁} and \ACode{ϕ₂}, then the
+    resulting map is a renaming, iff both \ACode{ϕ₁} and \ACode{ϕ₂} are
+    renamings.
+    In other words, 
+    if \ACode{ϕ₁} is a \ACode{K₁}-map and \ACode{ϕ₂} is a \ACode{K₂}-map,
+    then the result is a \ACode{(K₁ ⊔ K₂)}-map, where \ACode{⊔} refers to the
+    lattice for \ACode{\{ \ADef{Kᵣ} , \ADef{Kₛ} \}}
+    generated by \ACode{\ADef{Kᵣ} $\leq$ \ADef{Kₛ}}.
+  \item
+    All four compositions first apply \ACode{ϕ₁} to \ACode{x},
+    and then apply \ACode{ϕ₂} to the result.
+    If \ACode{ϕ₁} is a renaming, this result is another variable, but
+    if \ACode{ϕ₁} is a substitution, this result is a term.
+  \end{enumerate}
+  With the \ACode{\ACon{Kit}} abstraction, we can easily abstract over
+  \ACode{ϕ₂} being a renaming or a substitution:
   \ExampleCode\KExTwoComps
 
-  But to unify those two compositions, we need a new abstraction.
+  But to abstract over \ACode{ϕ₁}, the \ACode{\ACon{Kit}} abstraction
+  is not sufficient: while it allows us to abstract over what we are applying, i.e.
+  a renaming or a substitution, it does not allow us to
+  abstract over what we are applying it to, i.e. a variable or a term.
+  For the latter we have two distinct operations \ACode{\ADef{\_\&\_}}
+  and \ACode{\ADef{\_⋯\_}}.
 
+  To fill this gap, we introduce a new abstraction that we call a
+  \ACode{\ACon{ComposeKit}},
+  which provides an operation \ACode{\AField{\_\&/⋯\_}} that unifies
+  \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}, allowing us to
+  to define a general composition in the spirit of:
   \ExampleCode\KComposition
+
+  \ACode{\ACon{ComposeKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
+  \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
+  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-assoc}} are analogous to
+  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{\_⋯\_}}.
 

### 2023-09-15 paper/paper.tex
Update.

@@
 % \documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[sigplan,10pt]{acmart}
+\documentclass[sigplan,10pt, anonymous]{acmart}
 
@@
   \begin{abstract}
-    Variable binding and substitution are a long-standing nuisance for mechanized
-    metatheory of programming languages. A common approach is to represent
-    variables as DeBruijn Indices, where substitution is defined in terms of
-    variable renaming. If a language has $n$ different kinds of variables, and
-    $m$ different kinds of terms, we potentially need to define both
-    renaming and substitution for all $m\cdot n$ combinations and prove the
-    theory of those operations both in isolation and with respect to each other.
+    Formalizing a typed programming language in a proof assistant
+    requires to choose representations for variables and typing.
+    Variables are often represented as de Bruijn indices, where
+    substitution is usually defined in terms of renamings to allow for
+    proofs by structural induction.
+    Typing can be represented extrinsically by defining untyped terms and a typing
+    relation, or intrinsically by combining syntax and typing into a single
+    definition of well-typed terms.
+    For extrinsic typing, there is again a choice between extrinsic
+    scoping, where terms and the notion of free variables are defined
+    separately, and intrinsic scoping, where terms are indexed by their
+    free variables.
 
-    % McBride introduced kits to define both substitution and renaming by a
-    % single term traversal for intrinsically typed terms.\cite{mcbride2005kits}
+    This paper describes an Agda framework for formalizing programming
+    languages with extrinsic typing, intrinsic scoping, and de Bruijn
+    Indices for variables.
+    The framework supports object languages with arbitrary many
+    variable sorts and dependencies, making it suitable for
+    polymorphic languages and dependent types.
+    Given an Agda definition of syntax and typing, the framework derives
+    substitution operations and lemmas for untyped terms, and provides an
+    abstraction to prove type preservation of these operations with
+    just a single lemma.
+    The key insights behind the framework are the use of multi-sorted syntax
+    definitions, which enable parallel substitutions that replace
+    all variables of all sorts simultaneously, and
+    abstractions that unify the definitions, compositions, and typings
+    of multi-sorted renamings and substitutions.
+    Case studies have been conducted to prove subject reduction for
+    System F with subtyping, dependently typed lambda calculus, and lambda
+    calculus with pattern matching.
+  \end{abstract}
 
-    We propose a technique to derive the implementation and metatheory of
-    renaming and substitution for all $m\cdot n$ combinations of variables and terms
-    from a single term traversal and three simple lemmas for extrinsically typed
-    but intrinsically scoped terms.
-    We demonstrate this technique by formalizing type soundness of System F with
-    very parallel substitutions, which replace both type and term variables at
-    once, enabling a single substitution-preserves-typing lemma.
-    We provide a well-documented implementation of the technique as an Agda
-    library.
+  % \begin{abstract}
+  %   Formalizing a typed programming language in a proof assistant
+  %   requires to choose between an extrinsic and an intrinsic
+  %   representation of typing.
+  %   \emph{Extrinsic} typing means to separately define a set of
+  %   untyped terms and a typing relation.
+  %   \emph{Intrinsic} typing means to directly define a set of well-typed terms,
+  %   incorporating the typing relation into the syntax definition.
 
-    As the technique is based on extrinsic typing, the metatheory of renaming
-    and substitution is decoupled from type preservation.
-    This has the downside, that type preservation of renaming and substitution
-    are not for free, but also the upside, that the library can be directly
-    instantiated for object languages with non-standard typing relations like
-    dependent types, type state, or resource environments.
-  \end{abstract}
+  %   One consequence of intrinsic typing is that operations on terms
+  %   are also proofs that the operations preserve typing.
+  %   This often leads to less code, as shared structure between
+  %   definitions and preservation lemmas does not need to be repeated,
+  %   but has the drawback that specification and proofs are
+  %   intertwined, making it more difficult to check if a specification is
+  %   correct.
+  % \end{abstract}
+
+  % \begin{abstract}
+  %   Formalizing a typed programming language in a proof assistant
+  %   requires to choose between an extrinsic and intrinsic representation of typing.
+  %   For extrinsic typing, 
+  %   There are two approaches to formalizing the syntax of typed object languages
+  %   in proof assistants. The extrinsic approach defines
+  %   object terms and typing relation separately, whereas the
+  %   instrinsic approach only defines well-typed terms.
+
+  %   This paper describes an Agda framework for deriving substitution
+  %   operations and lemmas for syntaxes with extrinsic typing,
+  %   intrinsic scoping, and DeBruijn Indices as variable representation.
+
+  %   This paper describes an Agda framework for formalizing programming
+  %   languages with extrinsic typing, intrinsic scoping, and DeBruijn
+  %   Indices as variable representation. The key feature of the
+  %   framework 
+  %   all sorts of object variables.
+  %   Agda's reflection mechanism is used to derive 
+  %   substitution properties for a
+  %   large class of syntaxes and only requires user code for to prove that the
+
+  %   framework is to deal with multiple sorts of variables in a uniform way
+  % \end{abstract}
+
+  % \begin{abstract}
+  %   Variable binding and substitution are central concepts in the formalization
+  %   of programming languages in proof assistants. 
+  %   % Many different representations of variables have been developed with different
+  %   % trade-offs on proof simplicity, clarity, and reduction of boilerplate.
+  %   A common approach is to represent variables as DeBruijn Indices
+  %   and to define substitution in terms of variable renamings for the sake
+  %   of structural induction.
+  % \end{abstract}
+
+  % \begin{abstract}
+  %   Variable binding and substitution are a long-standing nuisance for mechanized
+  %   metatheory of programming languages. A common approach is to represent
+  %   variables as DeBruijn Indices, where substitution is defined in terms of
+  %   variable renaming. If a language has $n$ different kinds of variables, and
+  %   $m$ different kinds of terms, we potentially need to define both
+  %   renaming and substitution for all $m\cdot n$ combinations and prove the
+  %   theory of those operations both in isolation and with respect to each other.
+
+  %   % McBride introduced kits to define both substitution and renaming by a
+  %   % single term traversal for intrinsically typed terms.\cite{mcbride2005kits}
+
+  %   We propose a technique to derive the implementation and metatheory of
+  %   renaming and substitution for all $m\cdot n$ combinations of variables and terms
+  %   from a single term traversal and three simple lemmas for extrinsically typed
+  %   but intrinsically scoped terms.
+  %   We demonstrate this technique by formalizing type soundness of System F with
+  %   very parallel substitutions, which replace both type and term variables at
+  %   once, enabling a single substitution-preserves-typing lemma.
+  %   We provide a well-documented implementation of the technique as an Agda
+  %   library.
+
+  %   As the technique is based on extrinsic typing, the metatheory of renaming
+  %   and substitution is decoupled from type preservation.
+  %   This has the downside, that type preservation of renaming and substitution
+  %   are not for free, but also the upside, that the library can be directly
+  %   instantiated for object languages with non-standard typing relations like
+  %   dependent types, type state, or resource environments.
+  % \end{abstract}
 
@@
   \section{Introduction}
-  Formalizing metatheory of programming languages in a proof assistant 
-  requires to deal with variables, binders and substitution.
+  In recent years intrinsic typing\cite{TODO} has become rather popular.
+  A consequence of intrinsic typing is that operations on terms
+  are also proofs that the operations preserve typing.
+  This often has the benefit of more concise code, as shared structure
+  between definitions and preservation lemmas does not need to be
+  repeated.
+  However, intrinsic typing also has three major drawbacks:
+  \begin{enumerate}
+  \item
+    It is more difficult to judge whether a specification is correct,
+    because there is no clear separation between specification and proofs.
+    To understand the specification one has to additionally
+    understand the proofs, which can introduce significant cognitive overhead.
+    % For example specifying a small step semantics is directly intertwined
+    % with the proof of subject reduction.
+  \item
+    It is less uniform for object languages with multiple variable
+    sorts, e.g. type- and expression-variables. In such a scenario,
+    types are usually indexed by kinds and expressions by kinds and types.
+    This requires not only a type-in-type and expression-in-expression
+    substitution, but also a type-in-expression substitution, which causes
+    interactions between different types of substitutions introducing
+    additional complexity.
+    The duplication increases even more if one additionally introduces
+    kind-variables. Then a total of 6 different substitutions and
+    their interactions is required.
+    Unifying these substitutions to reduce repetition is difficult
+    with intrinsic typing due to the interdependence of definitions.
+  \item
+    It is less suitable for automation, because type preservation is baked
+    into the syntax definition and type preservation can become arbitrary complex
+    depending on the typing relation. This can prevent substitition
+    definition and lemmas from being derived, where otherwise it would be possible.
+  \end{enumerate}
+  With extrinsic typing the major drawback is the duplication between
+  definitions and type preservation lemmas, which we found more
+  suitable to be eliminated by automation.
 
-  In most languages those constructs behave uniformly,
-  i.e. substitution is only concerned with variables and binders and
-  behaves homomorphic on all other syntax constructors.
-
-  Intrinsic typing
-
-  \begin{enumerate}
-  \item DeBruijn
-  \item Intrinsic vs Extrinsic
-  \item Reduction under binder
-  \end{enumerate}
+  To combat those drawbacks, we went in the direction of intrinsically
+  scoped syntax in our search for a general, clean, and automated framework
+  for metatheory.
 
@@
 
-  Contrary to the usual intrinsic typing, we limit the environment
-  \ACode{S} to sorts with index \ACode{\ACon{Var}}.
-  This ensures that it is impossible to construct kind variables, as
-  it is not possible to extend \ACode{S} with the sort of kinds
-  \ACode{\ACon{𝕜}}.
+  Note that there is no straightforward way to construct a
+  multi-sorted syntax with intrinsic typing: in a direct translation,
+  the type of terms \ACode{\ACon{\_⊢\_}} would be indexed by itself, which
+  most proof assistants forbid to avoid breaking logical consistency.
+  We conjecture that an encoding with universe codes is possible, as
+  it has been done for intrinsically typed languages with dependent
+  types, but those constructions come with their own drawbacks and complexity.
+
+  % Contrary to the usual intrinsic typing, we limit the environment
+  % \ACode{S} to sorts with index \ACode{\ACon{Var}}.
+  % This ensures that it is impossible to construct kind variables, as
+  % it is not possible to extend \ACode{S} with the sort of kinds
+  % \ACode{\ACon{𝕜}}.
 
@@
   \section{Map Composition}
-  In this section we extend our framework with the composition of
-  arbitrary maps.
+
+  % - Reason about multiple applications in terms of a single application
+  % - e.g. wk-cancels-⦅⦆ or dist-↑-⦅⦆, often important for induction hypothesis
+  % - Mcbride uses tactics
+  % - We introduce an abstraction similar to kits
+  % - ComposeKit has primitives for composition, ComposeTraversal has the property
+
+  %%%%%%%%%%
+
+  In this section, we extend our framework with a single composition
+  operator for arbitrary maps.
+
+  Composition \AField{\_·ₖ\_} equips our framework with the power to
+  reason about multiple maps \ACode{ϕ₁} and \ACode{ϕ₂} being applied
+  to a term in sequence, in terms of a single application:
+
+  The core property of composition is that applying two maps
+  \ACode{ϕ₁} and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same
+  as applying their composition \ACode{(ϕ₁ \AField{·ₖ} ϕ₂)} to \ACode{t}, i.e.
+  $$
+  \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
+  $$
+
+  %%%%%%%%%%
+
+  In this section, we extend our framework with a single composition
+  operator for arbitrary maps.
+
+  Composition \AField{\_·ₖ\_} equips our framework with the power to
+  reason about multiple maps \ACode{ϕ₁} and \ACode{ϕ₂} being applied
+  to a term in sequence:
+
+  \ACode{\AField{⋯-assoc} t ϕ₁ ϕ₂} \ \ : \ \ \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
+
+  For many lemmas, e.g. substitution-preserves-typing in System F,
+  it is necessary to apply an equality like \ACode{t ⋯ σ₁ ⋯ σ₂ ≡ t ⋯ σ₂' ⋯ σ₁'}
+  where the induction hypothesis requires \ACode{t' ⋯ σ₁'}.
+
+  %%%%%%%%%%
+
+  In this section, we extend our framework with a single composition
+  operator for arbitrary maps.
 
@@
   which provides an operation \ACode{\AField{\_\&/⋯\_}} that unifies
-  \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}, allowing us to
+  \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}, allowing us
   to define a general composition in the spirit of:
@@
   \section{Types \& Contexts}
-  \subsection{Library}
+  % To define typing contexts and a typing relation, we need to know
+  % which sorts of terms play the role of types for other sorts of terms.
+  % When modeling a typed language with multi-sorted syntax, both terms
+  % and types are represented as terms, but with different sorts.
+  % To define typing contexts and a typing relation, we need to record for each term
+  % from 
+  % describe terms, and ocan be mapped to 
+  % In a typed language,
+
+  In the context of multi-sorted syntax, the notation of a type can be
+  described as a mapping between sorts.
+  For System F, the expression sort \ACode{\ACon{𝕖}} maps to the type sort
+  \ACode{\ACon{𝕥}} and the sort \ACode{\ACon{𝕥}} maps to the kind sort
+  \ACode{\ACon{𝕜}}.
+  To teach our framework about types, we have to instantiate the following structure:
   \LibCode\KTypeSorts
+  For System F the instanciation is
+  \AppCode\FTypes
+  For a nicer interface, we define a type analogously to a term:
   \LibCode\KTypes
+
+  \subsection{Typing Context}
+  A typing context is then a function which maps variables to types.
+  As we want our framework to support dependent types, the types returned by
+  a typing context are allowed to use the domain of the 
+  To be able to deal with dependent types, the result types are only allowed
+  to refer to the variables that come before 
   \LibCode\KContextHelper
@@
   \LibCode\KContextLookup
-  \subsection{System F}
-  \AppCode\FTypes
+
+  \LibCode\KVariableTyping
 
   \section{Typing}
-  % TODO move to typing section
-  The only drawback we found, is that types can now also be indexed
-  with free expression variables, even though types can never contain
-  free variables as assured by the variable constructor. However, this
-  only causes minor issues.
+  Equipped with a notion of types and type context we are ready to define
+  the multi-sorted typing relation for System F, incorporating both
+  typing and kinding:
+  \AppCode\FTyping
+  The interesting cases are:
+  \begin{itemize}
+  \item
+    the variable typing constructor covers both expression- and
+    type-variables, analogously to the variable term constructor;
+  \item
+    the typing of the lambda abstraction \ACode{\ACon{⊢λ}} weakens the
+    type of the codomain \ACode{t₂}.
+    This is necessary, because with multi-sorted syntax, types could
+    also depend on expressions, so the sub typing derivation accounts for
+    another variable which is not used by the type.
+  \item
+    the kinding constructor \ACode{\ACon{⊢τ}} states that all types
+    have kind \ACode{\ACon{★}}. This is sufficient for System F as types
+    are automatically well-scoped due to intrinsic scoping.
+  \end{itemize}
+  To teach the framework about typing, we create a structure analogously to
+  \ACode{\ACon{Syntax}}:
+  \LibCode\KTyping
+  and instanciate it for System F:
+  \AppCode\FTypingInst
 
+  Next we define \ACode{\ACon{TypingKit}}s, which are analog to
+  \ACode{\ACon{Kit}}s, but abstract about variable and term
+  \emph{typing}, instead of variables and terms, themselves.
+
+  \LibCode\KTypingKit
   \subsection{Library}
-  \LibCode\KVariableTyping
-  \LibCode\KTyping
-  \LibCode\KTypingKit
   \LibCode\KMapTyping
@@
   \subsection{System F}
-  \AppCode\FTyping
-  \AppCode\FTypingInst
   \AppCode\FPreserve
@@
   \section{Semantics}
-  % \AppCode\FValues
+  We define a standard small-step operational semantics for System F
+  with full reduction:
   \AppCode\FReduction
@@
   \section{Subject Reduction}
+  Proving subject reduction is now straightforward:
   \AppCode\FSubjectReduction
   \AppCode\FSubjectReductionProofInteresting
+  We only listed the cases that do not follow directly from the
+  induction hypothesis:
+  \begin{itemize}
+  \item The beta-reduction rule for type-application, \ACode{\ACon{β-Λ}},
+    is a direct consequence of substitution preserves typing.
+    \ACode{⊢e₂} is the typing for \ACode{e₂}.
+    \ACode{⊢⦅ ⊢e₂ ⦆} is the substitution-typing for the singleton substitution 
+    \ACode{\AField{⦅ e₂ ⦆}}.
+    \ACode{\AField{⊢e₁ ⊢⋯ ⊢⦅ ⊢e₂ ⦆}} states that applying the singleton substitution
+    to \ACode{e₁} preserves typing.
+  \item The beta-reduction rule for value-application is almost the same,
+    but requires us to addionaly apply an equality to the type index.
+  \end{itemize}
+
   % \FSubjectReductionProofRest
@@
   \section{Reflection \& Generics}
+  We used Agda's reflection mechanism\cite{TODO} to derive
+  instanciations related to all structures for untyped substitition,
+  i.e. \ACode{\ACon{Traversal}} and \ACode{\ACon{ComposeTraversal}}.
+
+  To gain insight into the class of object languages supported by our
+  reflection algorithm, we modeled a class of object languages using a
+  generic programming technique and instanciated the structures
+  generically. The class contains all object languages which:
+  \begin{enumerate}
+  \item Have a variable constructor of type TODO;
+  \item Uses subterms only directly (e.g. not in lists);
+  \item Subterms only extend the scope-context.
+  \end{enumerate}
+
+  Our reflection algorithm derives proofs with exactly the same
+  structure as the generic proof, giving high confidence that it covers
+  the same class of languages.
+
+  The formal definition of the generic class can be found in
+  Appendix~TODO and the actual implementation.
+
+  We were able to describe many languages in this shape, including
+  session types and type state.
+
+  We believe restriction 2 is purely technically and can be lifted
+  with a more sophisticated reflection algorithm, as parameterized
+  datastructures can also be inlined into the syntax definition as terms
+  of a new sort.
+
+  We believe restriction 3 can be lifted partially to allow for
+  example the encoding of linear types that requires context
+  splitting. This has already be done for intrinsical kits\cite{TODO}.
 

### 2023-09-15 paper/agdatex.py

@@ -109,21 +109,28 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
     prefixes = []
     mode = "none"
     stop_command_on_empty_line = False
+    opt = ""
     def start_command(name, is_inline):
-        global mode, tgt, prefixes
+        global mode, tgt, prefixes, opt
         for p in prefixes:
             name = p + name
         commands.append(name)
         if mode == "hide":
             tgt += "\\end{code}\n"
         elif mode == "command":
-            print(f"ERROR: Line {line_num} starts a nested command:\n  {line}")
+            print(f"ERROR: Line {line_num+1} starts a nested command:\n  {line}")
         mode = "command"
         opt = "[inline]" if is_inline else ""
-        tgt += "\\newcommand*\\" + name + "{\\begin{code}" + opt + "\n"
+        tgt += "\\newcommand*\\" + name + "{"
+        tgt += "\\begin{AgdaAlign}"
+        tgt += "\\begin{AgdaSuppressSpace}"
+        tgt += "\\begin{code}" + opt + "\n"
     def stop_command():
         global mode, tgt, prefixes
-        tgt += "\\end{code}}\n"
+        tgt += "\\end{code}"
+        tgt += "\\end{AgdaSuppressSpace}"
+        tgt += "\\end{AgdaAlign}"
+        tgt += "}\n"
         mode = "none"
     for line_num, line in enumerate(src.splitlines()):
         l = line.strip()
@@ -145,11 +152,26 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
                 prefixes.append(l.split(" ", 1)[0])
             elif "<" in l:
                 prefixes.pop()
+            elif "[" in l:
+                if mode == "command":
+                    mode = "hide"
+                    tgt += "\\end{code}\n"
+                    tgt += "\\begin{code}[hide]\n"
+                else:
+                    print(f"ERROR: Line {line_num+1} starts a hiding block outside of a command:\n  {line}")
+                    sys.exit(1)
+            elif "]" in l:
+                if mode == "hide":
+                    mode = "command"
+                    tgt += "\\end{code}\n"
+                    tgt += "\\begin{code}" + opt + "\n"
+                else:
+                    print(f"ERROR: Line {line_num+1} stops a hiding block outside of hiding mode:\n  {line}")
             else:
                 name = l.split(" ", 1)[0]
                 start_command(name, is_inline)
                 stop_command_on_empty_line = True
[…]

### 2023-09-15 paper/paper.tex
Update.

@@
   \emph{typing}, instead of variables and terms, themselves.
-
   \LibCode\KTypingKit
-  \subsection{Library}
+  Building on the fields of a \ACode{\ACon{TypingKit}}, we define further operations:
+  \begin{itemize}
+  \item
+    Using the variable/term typing, we can define a renaming/substitution typing:
     \LibCode\KMapTyping
+  \item 
+    Lifting a map preserves its typing:
     \LibCode\KLiftTyping
     % \LibCode\KLiftTypingProof
+  \item 
+    If a variable/term has a typing, then so does its singleton renaming/substitution:
     \LibCode\KSingleTyping
     % \LibCode\KSingleTypingProof
-  \LibCode\KTypingNotation
+  \end{itemize}
+  We also specify notation, which makes it less cumbersome
+  to explicitly pass a particular Kit to those definitions:
+  \begin{itemize}
+  \item \ACode{Γ₂ \ADef{∋*/⊢*[} TK \ADef{]} ϕ \ADef{∶} Γ₁} is
+    \ACode{Γ₂ \ADef{∋*/⊢*} ϕ \ADef{∶} Γ₁} for some \ACode{\ACon{TypingKit}} \ACode{TK}.
+  \end{itemize}
+  % \LibCode\KTypingNotation
+
+  We then define a \ACode{\ACon{TypingTraversal}} analogously to
+  \ACode{\ACon{Traversal}}, but instead of defining the application of
+  renamings/substitutions, it defines the type preservation of applying
+  renamings/substitutions:
   \LibCode\KTypingTraversal

### 2023-09-15 paper/paper.tex
Update.

@@
   \end{itemize}
-  We also specify notation, which makes it less cumbersome
-  to explicitly pass a particular Kit to those definitions:
-  \begin{itemize}
-  \item \ACode{Γ₂ \ADef{∋*/⊢*[} TK \ADef{]} ϕ \ADef{∶} Γ₁} is
-    \ACode{Γ₂ \ADef{∋*/⊢*} ϕ \ADef{∶} Γ₁} for some \ACode{\ACon{TypingKit}} \ACode{TK}.
-  \end{itemize}
+  To make it easier to explicitly pass typing kits,
+  we define the notation \ACode{Γ₂ \ADef{∋*/⊢*[} TK \ADef{]} ϕ \ADef{∶} Γ₁}
+  to stand for \ACode{Γ₂ \ADef{∋*/⊢*} ϕ \ADef{∶} Γ₁} given some
+  \ACode{\ACon{TypingKit}} \ACode{TK}.
   % \LibCode\KTypingNotation
@@
   \ACode{\ACon{Traversal}}, but instead of defining the application of
-  renamings/substitutions, it defines the type preservation of applying
-  renamings/substitutions:
+  maps, it defines the type preservation of applying maps:
   \LibCode\KTypingTraversal
+  Given a variable/term \ACode{e} with typing \ACode{⊢e} and
+  a renaming/substitution \ACode{ϕ} with typing \ACode{⊢ϕ},
+  the term \ACode{⊢e \AField{⊢⋯} ⊢ϕ} is a typing for \ACode{e \AField{⋯} ϕ}.
+
+  Given a \ACode{\ACon{TypingTraversal}}, we can easily instanciate the
+  \ACode{\ACon{TypingKits}} for variables and terms:
   \LibCode\KTypingInstances
   % \LibCode\KTypingTraversalNotation
-  \subsection{System F}
+
+  We are then ready to create a \ACode{\ACon{TypingTraversal}}
+  instance for System F:
   \AppCode\FPreserve

### 2023-09-16 paper/paper.tex
Update.

@@
   \section{Introduction}
+  \label{sec:introduction}
   In recent years intrinsic typing\cite{TODO} has become rather popular.
@@
   \subsection{Structure}
+  \label{sec:introduction:structure}
   The rest of this paper will introduce our framework using System F
@@
   Code of examples is written in yellow boxes.
-  Code of the System F formalization is written without boxes. This is
-  the only code a user of our framework has to write.
-
-  In the paper we state all definitions and lemmas, but omit some proofs
-  in the interest of space. However the omitted proofs exclusively
-  depend on lemmas and definitions stated in this paper. The full code
-  from the paper can be found at TODO.
-
-  In the paper we only show the part of the framework necessary for
-  the System F soundness proof. The actual implementation is more
-  general and complete and can be found at TODO.
+  Code of the System F formalization is written without boxes. The
+  latter is the only code a user of our framework has to write.
+
+  In this paper we present a simplified version of the actual
+  framework, focusing on the core concepts. We present all necessary
+  definitions and lemmas, but omit some proofs for the sake of
+  space. The omitted proofs can be found in Appendix~{TODO} and the
+  supplementary material. The full framework can be found at TODO.
+
+  The rest of this paper is structured as followed:
+  Section~\ref{sec:syntax} introduces the multi-sorted syntax and compares it to the
+  more common unsorted syntax.
+  Section~\ref{sec:maps} introduces multi-sorted substitutions and renamings, and
+  an abstraction to unify them.
+  Section~\ref{sec:composition} introduces composition of multi-sorted
+  substitions and renamings, and an abstraction to unify all four
+  compositions.
+  Section~\ref{sec:types} shows how to introduce the notion of types
+  for multi-sorted syntax and how to define typing contexts.
+  Section~\ref{sec:typing} shows how to define a multi-sorted typing relation,
+  and an abstraction to unify type preservattion lemmas for renamings and substitutions.
+  Section~\ref{sec:semantics} introduces a small-step operational semantics for System F.
+  Section~\ref{sec:subjectreduction} proves the subject reduction lemma for System F.
+  Section~\ref{sec:reflection} describes the class of object languages covered by
+  our reflection algorithm.
+  Section~\ref{sec:beyond} describes case studies of object languages beyond System F.
+  Section~\ref{sec:future} describes future work.
+  Section~\ref{sec:related} discusses related work.
+  Section~\ref{sec:conclusion} concludes.
 
   \subsection{Contributions}
+  \label{sec:introduction:contributions}
   \begin{enumerate}
   \item
-    a multi-sorted approach to extrinsic typing;
-  \item
-    a multi-sorted extension of McBride's kits to unify renamings and substitutions and their metatheory;
-  \item
-    an abstraction for composition and it's metatheory, unifying the four combinations of renamings and substitutions;
-  \item
-    an abstraction for typing relations, unifying type preservation of renaming and substitution into a single lemma;
-  \item
-    an Agda-implementation featuring derived term-kits, representation
-    independence for substitutions and typing contexts, heterogeneous
-    equality, independence of functional extensionality, case studies
-    on System F, dependent lambda calculus, pattern matching, linear
-    types.
+    a novel approach for formalizing intrinsically-scoped syntax with
+    multiple variable sorts as a special kind of intrinsically-typed
+    syntax, we call \emph{multi-sorted syntax};
+  \item
+    an adaptation of McBride's kit-abstraction for unifying renamings
+    and substitutions to multi-sorted syntax;
+  \item
+    a novel abstraction for composition and it's metatheory, unifying
+    the four compositions between renamings and substitutions;
+  \item
+    a novel abstraction for typing relations, unifying type
+    preservation of renaming and substitution;
+  \item
+    a formalized specification of a large class of object languages
+    for which substitution and lemmas can be derived generically.
+  \item
+    an implementation of the concepts as framework in Agda featuring
+    a reflection algorithm, representation independence for
+    substitutions and typing contexts, heterogeneous equality between
+    renamings and substitutions, and free of postulated axioms.
+  \item
+    three case studies using our approach to prove subject reduction
+    for System~F with subtyping, a dependently-typed lambda calculus, and
+    pattern matching.
   \end{enumerate}
 
-  \newpage
-
   \section{Syntax}
+  \label{sec:syntax}
   \subsection{Single-Sorted Syntax}
+  \label{sec:syntax:single}
   The following shows a typical intrinsically-scoped syntax of System F:
@@
 
-  \texttt{Type}s are indexed by the number of free type variables \texttt{n}.
-  \texttt{Expr}essions are additionally indexed by the number of free expression variables \texttt{m}.
-  Variables \texttt{`\_} are represented as DeBruijn-indices, where \texttt{Fin n} is the type of \texttt{n} elements.
+  \ACode{\ACon{Type}}s are indexed by the number of free type
+  variables \ACode{n}.
+  \ACode{\ACon{Expr}}essions are additionally indexed by the number of
+  free expression variables \ACode{m}.
+  Variables \ACode{\ACon{`\_}} are represented as DeBruijn-indices, where
+  \ACode{\ACon{Fin} n} is the type of \ACode{n} elements.
 
@@
   \begin{enumerate}
-  \item the syntactic categories (\texttt{Kind}, \texttt{Type}, and \texttt{Expr})
+  \item
+    the syntactic categories (\ACode{\ACon{Kind}},
+    \ACode{\ACon{Type}}, and \ACode{\ACon{Expr}})
     have different types, which makes it difficult to treat them uniformly; and
@@
     substitutions, e.g.\ adding kind variables would lead to a total of 6
-    different substitutions.
+    different substitutions and corresponding interaction lemmas.
   \end{enumerate}
@@
   \subsection{Multi-Sorted Syntax}
+  \label{sec:syntax:multi}
   A multi-sorted syntax is defined by a single type of sort-indexed terms.
@@
   (\ACode{\ACon{𝕖}}) with two free type-variables (\ACode{\ACon{𝕥}}).
-  \newpage
   \AppCode\FSyntax
@@
   \subsection{A Structure for Multi-Sorted Syntax}
+  \label{sec:syntax:structure}
   The multi-sorted syntax makes it easy to define a structure for
@@
   \section{Renamings \& Substitutions}
+  \label{sec:maps}
   \subsection{Multi-Sorted Renamings \& Substitutions}
+  \label{sec:maps:example}
   Working with a sort-indexed syntax allows us to define renamings and substitutions
@@
   \subsection{The Need for Renamings}
+  \label{sec:maps:renamings}
   A problem with mechanizing languages with DeBruijn-indices is that
@@
   \subsection{Unifying Renamings \& Substitutions}
+  \label{sec:maps:kits}
   To avoid the duplication between renamings and substitutions,
@@
   \section{Map Composition}
+  \label{sec:composition}
 
@@
   \subsection{An Examination of Composition}
+  \label{sec:composition:example}
   To motivate our abstraction for composition, we first look
@@
   \subsection{WkKits}
+  \label{sec:composition:wkkit}
 
@@
   \subsection{Composition}
+  \label{sec:composition:ckit}
 
@@
   \section{Types \& Contexts}
+  \label{sec:types}
   % To define typing contexts and a typing relation, we need to know
@@
   \section{Typing}
+  \label{sec:typing}
   Equipped with a notion of types and type context we are ready to define
@@
 
+  As the \ACode{\ACon{Traversal}} and  \ACode{\ACon{ComposeTraversal}} instances
+  can be generally derived via reflection,
+  the \ACode{\AField{\_⊢⋯\_}} lemma is in fact the only lemma that
+  has to been proved manually for subject reduction.
+  From this single lemma follows type preservation for all operations
+  of both renamings and substitutions for all sorts of variables
+  simultaneously.
+
   \section{Semantics}
+  \label{sec:semantics}
   We define a standard small-step operational semantics for System F
@@
   \section{Subject Reduction}
+  \label{sec:subjectreduction}
   Proving subject reduction is now straightforward:
@@
   \section{Reflection \& Generics}
+  \label{sec:reflection}
   We used Agda's reflection mechanism\cite{TODO} to derive
@@
   \section{Beyond System F}
+  \label{sec:beyond}
   \begin{itemize}
@@
   \section{Future Work}
+  \label{sec:future}
   \begin{itemize}
@@
     restrictions on the form of \ACode{S}.
+  \item
+    Providing abstractions not just for type-preservation, but also
+    for different kinds of semantics, e.g. confluence or lemmas related
+    to denotational semantics.
   \end{itemize}
@@
   \section{Related Work}
+  \label{sec:related}
   Autosubst \cite{DBLP:conf/cpp/StarkSK19, DBLP:conf/itp/SchaferTS15}.
@@
   \section{Conclusion}
+  \label{sec:conclusion}
 

### 2023-09-16 paper/paper.tex
Update.

@@
 \begin{document}
+
   \begin{abstract}
@@
 
-  % \begin{abstract}
-  %   Formalizing a typed programming language in a proof assistant
-  %   requires to choose between an extrinsic and an intrinsic
-  %   representation of typing.
-  %   \emph{Extrinsic} typing means to separately define a set of
-  %   untyped terms and a typing relation.
-  %   \emph{Intrinsic} typing means to directly define a set of well-typed terms,
-  %   incorporating the typing relation into the syntax definition.
-
-  %   One consequence of intrinsic typing is that operations on terms
-  %   are also proofs that the operations preserve typing.
-  %   This often leads to less code, as shared structure between
-  %   definitions and preservation lemmas does not need to be repeated,
-  %   but has the drawback that specification and proofs are
-  %   intertwined, making it more difficult to check if a specification is
-  %   correct.
-  % \end{abstract}
-
-  % \begin{abstract}
-  %   Formalizing a typed programming language in a proof assistant
-  %   requires to choose between an extrinsic and intrinsic representation of typing.
-  %   For extrinsic typing, 
-  %   There are two approaches to formalizing the syntax of typed object languages
-  %   in proof assistants. The extrinsic approach defines
-  %   object terms and typing relation separately, whereas the
-  %   instrinsic approach only defines well-typed terms.
-
-  %   This paper describes an Agda framework for deriving substitution
-  %   operations and lemmas for syntaxes with extrinsic typing,
-  %   intrinsic scoping, and DeBruijn Indices as variable representation.
-
-  %   This paper describes an Agda framework for formalizing programming
-  %   languages with extrinsic typing, intrinsic scoping, and DeBruijn
-  %   Indices as variable representation. The key feature of the
-  %   framework 
-  %   all sorts of object variables.
-  %   Agda's reflection mechanism is used to derive 
-  %   substitution properties for a
-  %   large class of syntaxes and only requires user code for to prove that the
-
-  %   framework is to deal with multiple sorts of variables in a uniform way
-  % \end{abstract}
-
-  % \begin{abstract}
-  %   Variable binding and substitution are central concepts in the formalization
-  %   of programming languages in proof assistants. 
-  %   % Many different representations of variables have been developed with different
-  %   % trade-offs on proof simplicity, clarity, and reduction of boilerplate.
-  %   A common approach is to represent variables as DeBruijn Indices
-  %   and to define substitution in terms of variable renamings for the sake
-  %   of structural induction.
-  % \end{abstract}
-
-  % \begin{abstract}
-  %   Variable binding and substitution are a long-standing nuisance for mechanized
-  %   metatheory of programming languages. A common approach is to represent
-  %   variables as DeBruijn Indices, where substitution is defined in terms of
-  %   variable renaming. If a language has $n$ different kinds of variables, and
-  %   $m$ different kinds of terms, we potentially need to define both
-  %   renaming and substitution for all $m\cdot n$ combinations and prove the
-  %   theory of those operations both in isolation and with respect to each other.
-
-  %   % McBride introduced kits to define both substitution and renaming by a
-  %   % single term traversal for intrinsically typed terms.\cite{mcbride2005kits}
-
-  %   We propose a technique to derive the implementation and metatheory of
-  %   renaming and substitution for all $m\cdot n$ combinations of variables and terms
-  %   from a single term traversal and three simple lemmas for extrinsically typed
-  %   but intrinsically scoped terms.
-  %   We demonstrate this technique by formalizing type soundness of System F with
-  %   very parallel substitutions, which replace both type and term variables at
-  %   once, enabling a single substitution-preserves-typing lemma.
-  %   We provide a well-documented implementation of the technique as an Agda
-  %   library.
-
-  %   As the technique is based on extrinsic typing, the metatheory of renaming
-  %   and substitution is decoupled from type preservation.
-  %   This has the downside, that type preservation of renaming and substitution
-  %   are not for free, but also the upside, that the library can be directly
-  %   instantiated for object languages with non-standard typing relations like
-  %   dependent types, type state, or resource environments.
-  % \end{abstract}
-
   \maketitle
@@
   \label{sec:introduction}
-  In recent years intrinsic typing\cite{TODO} has become rather popular.
-  A consequence of intrinsic typing is that operations on terms
-  are also proofs that the operations preserve typing.
-  This often has the benefit of more concise code, as shared structure
-  between definitions and preservation lemmas does not need to be
-  repeated.
+
+  Formalizing a typed programming language in a proof assistant
+  requires to choose between an extrinsic and intrinsic representation
+  of typing.
+
+  With \emph{intrinsic typing}, the definitions of syntax and typing
+  are merged into a single definition of well-typed terms.
+  Consequently, operations on terms have to map
+  well-typed terms to well-typed terms proving that they preserve typing.
+  For example, in an intrinsic setting, the definition of substitution
+  also acts as a proof that substitution preserves typing, and the
+  definition of a small-step semantics also acts as a proof for subject
+  reduction.
+  When working without library support, this often has the benefit of
+  more concise code, as shared structure between definitions and
+  preservation lemmas does not need to be repeated.
   However, intrinsic typing also has three major drawbacks:
@@
   \label{sec:introduction:structure}
+
   The rest of this paper will introduce our framework using System F

### 2023-09-16 paper/paper.tex
Update.

@@
   well-typed terms to well-typed terms proving that they preserve typing.
-  For example, in an intrinsic setting, the definition of substitution
+  For example, the definition of substitution
   also acts as a proof that substitution preserves typing, and the
@@
   reduction.
-  When working without library support, this often has the benefit of
-  more concise code, as shared structure between definitions and
-  preservation lemmas does not need to be repeated.
+  When working without library support, intrinsic typing often has the
+  benefit of more concise code, as shared structure between definitions
+  and preservation lemmas does not need to be repeated.
+
+  With \emph{extrinsic typing}, the definitions of syntax and typing
+  are kept separately: the syntax is defined first, and then typing is
+  defined as a relation. Consequently, operations on terms and
+  their type preservation lemmas are also separate entities.
+
+  However, when working with library support, this is less of a benefit.
+  Most operations, like substitution, have a generic definition and
+  metatheory across languages, which can be derived automatically via
+  reflection or generic programming techniques.
+  Only by adding a typing-relation
+
   However, intrinsic typing also has three major drawbacks:

### 2023-09-16 paper/paper.tex
Update.

@@
   Consequently, operations on terms have to map
-  well-typed terms to well-typed terms proving that they preserve typing.
+  well-typed terms to well-typed terms proving that the
+  operations preserve typing.
   For example, the definition of substitution
@@
   reduction.
-  When working without library support, intrinsic typing often has the
-  benefit of more concise code, as shared structure between definitions
-  and preservation lemmas does not need to be repeated.
+  This often has the benefit of more concise code, as shared structure
+  between definitions and preservation lemmas does not need to be
+  repeated.
 
@@
   their type preservation lemmas are also separate entities.
+  While extrinsic typing cannot eliminate the shared structure between
+  definitions and preservation lemmas, it can automated by reflection
+  or generic programming techniques, leaving only the preservation lemmas
+  to be proven by the programmer. While typing can be arbitrary complex,
+  the definition and metatheory of untyped substitution behaves the same
+  across languages.
 

### 2023-09-16 paper/paper.tex
Update.

@@
   are merged into a single definition of well-typed terms.
-  Consequently, operations on terms have to map
-  well-typed terms to well-typed terms proving that the
-  operations preserve typing.
+  Consequently, the definition of an operation on terms also acts
+  as a type preservation lemma for that operation, as it has to
+  map well-typed terms to well-typed terms.
   For example, the definition of substitution
@@
   their type preservation lemmas are also separate entities.
+  However, hope is not lost for extrinsic typing to keep up with the
+  conciseness of intrinsic typing: 
   While extrinsic typing cannot eliminate the shared structure between

### 2023-09-16 paper/paper.tex
Update.

@@
   our reflection algorithm.
-  Section~\ref{sec:beyond} describes case studies of object languages beyond System F.
+  Section~\ref{sec:casestudies} describes case studies of object languages beyond System F.
   Section~\ref{sec:future} describes future work.
@@
 
-  \section{Beyond System F}
-  \label{sec:beyond}
+  \section{Case Studies}
+  \label{sec:casestudies}
   \begin{itemize}
   \item
-    Binders with multiple variables are supported and even pattern
-    matching can be encoded.
-  \item
-    Heterogenous equality
+    For lambda calculus with dependent function types, the framework
+    just works out of the box for both deriving and instanciating the
+    \ACode{\ACon{TypingTraversal}}. Additionally, having the
+    \ACode{\ACon{ComposeKit}} abstraction turned out to be useful
+    to prove confluence, allowing to unify certain lemmas, which
+    otherwise need to be proved for both renamings and substitutions,
+    e.g. that for two terms \ACode{t} and \ACode{t'}
+    with \ACode{t \ACon{↪} t'} it follows that \ACode{(t \AField{⋯} ϕ)
+    \ACon{↪} (t' \AField{⋯} ϕ)} for any map \ACode{ϕ}.
+  \item
+    For System F with subtyping the main challenges are how to represent
+    subtyping constraints and how to deal with the fact that
+    substitution preserves typing is not generally true, as type variables
+    have subtyping bounds that need to be respected.
+    Instead of defining typing contexts and type preservation lemmas independent
+    of our framework, we found a way that still allows us to benefit
+    from the typing machinery.
+    When binding a type variable with a subtyping constraint \ACode{α <: t},
+    we first bind the type variable as \ACode{α : ★}, and then bind the constraint
+    with a dummy name \ACode{\_ ∶ (α <: t)}. This description is
+    similar to first-class constraints, but restricted enough to be
+    isomorphic to the original formalization.
+    With that encoding substitution preserves typing is generally true again, enabling
+    the use of the framework's typing contexts and \ACode{\ACon{TypingKit}} machinery.
+    Replacing a type variable \ACode{α <: t} with a type \ACode{t'}
+    which is not a subtype of \ACode{t'}, the resulting term is still
+    well-typed, but in a context with an insatisfiable constraint.
+  \item
+    Object language with pattern matching pose an interesting problem.
+    A clause \ACode{p ⇒ e} in a match-expression consists of a pattern
+    \ACode{p} and a continuation expression \ACode{e}.
+    The pattern \ACode{p} can bind arbitrary many variables in subpatterns,
+    which are then available as free variables in \ACode{e}.
+    % The challenge is to track the sorts of variables bound by a pattern,
+    % to make them available to \ACode{e} by extending \ACode{e}'s sort
+    % context correspondingly.
+    To track the sorts of variables bound by a pattern, one can define
+    the sort of patterns to contain the sorts of the variables that it binds:
+    % 𝕡 : List (Sort Var) → Sort NoVar
+    The syntax for variable patterns and clauses can then be specified as:
+    % varpat : S ⊢ 𝕡 [ 𝕖 ]  -- pattern binding an expression variable
+    % pairpat : S ⊢ 𝕡 S₁ → S ⊢ 𝕡 S₂ → S ⊢ 𝕡 (S₁ ++ S₂)
+    % clause : S ⊢ 𝕡 S' → (S' ++ S) ⊢ 𝕖 → S ⊢ 𝕔
   \end{itemize}

### 2023-09-16 paper/paper.tex
Update.

@@
   definitions and lemmas, but omit some proofs for the sake of
-  space. The omitted proofs can be found in Appendix~{TODO} and the
+  space. The omitted proofs can be found in Appendix~\ref{sec:proofs} and the
   supplementary material. The full framework can be found at TODO.
@@
   The formal definition of the generic class can be found in
-  Appendix~TODO and the actual implementation.
+  Appendix~\ref{sec:generic-class} and the actual implementation.
 
@@
   \appendix
+  \onecolumn
+
+  \section{Omitted Proofs}
+  \label{sec:proofs}
+
+  \subsection{Proofs for Section~\ref{sec:maps}}
+  \LibCode\KIdLift
+  \LibCode\KIdLiftProof
+
+  % \AppCode\FTraversalId
+  % \AppCode\FTraversalIdProofInteresting
+  % \LibCode\FTraversalIdProofRest
+
+  \subsection{Proofs for Section~\ref{sec:composition}}
+  \LibCode\KComposeKitAp
+  \LibCode\KComposeKitApProof
+
+  \LibCode\KDistLiftCompose
+  \LibCode\KDistLiftComposeProof
+
+  \LibCode\KCommLiftWeaken
+  \LibCode\KCommLiftWeakenProof
+
+  \LibCode\KCommLiftWeakenTraverse
+  \LibCode\KCommLiftWeakenTraverseProof
+
+  \LibCode\KWeakenCancelsSingle
+  \LibCode\KWeakenCancelsSingleProof
+
+  \LibCode\KWeakenCancelsSingleTraverse
+  \LibCode\KWeakenCancelsSingleTraverseProof
+
+  \LibCode\KDistLiftSingle
+  \LibCode\KDistLiftSingleProof
+
+  \LibCode\KDistLiftSingleTraverse
+  \LibCode\KDistLiftSingleTraverseProof
+
+  % \AppCode\FAssoc
+  % \AppCode\FAssocProofInteresting
+  % \AppCode\FAssocProofRest
+
+  \subsection{Proofs for Section~\ref{sec:typing}}
+  \LibCode\KLiftTyping
+  \LibCode\KLiftTypingProof
+
+  \LibCode\KSingleTyping
+  \LibCode\KSingleTypingProof
+
+  % \AppCode\FSubjectReduction
+  % \AppCode\FSubjectReductionProofInteresting
+
+  \section{Generic Class of Object Languages}
+  \label{sec:generic-class}
+
 \end{document}

### 2023-09-17 paper/paper.tex
Update.

@@
 
+  Formalizing programming languages in proof assistants quickly gets
+  repetitive. Almost every programming language supports variables
+  with static binding, and hence requires numerous definitions and
+  lemmas related to variable substitution.
+
+  Additionally, repetition can also occur within a single formalization.
+  This can be seen with polymorphic languages, where multiple sorts
+  of variables are present.
+  Consider for example System F, which supports both expression- and
+  type-variables.
+  With a naive approach, the whole substitution machinery needs to be
+  duplicated three times! We need to substitute expression-variables in
+  expressions, type-variables in types, but we also need to substitute
+  type-variables in expressions.
+  Even worse, having two substitutions acting on expressions, requires
+  to also prove lemmas about their interactions.
+  If we would additionally introduce kind-variables, we would end up
+  with a total of six duplications of the substition machinery and
+  corresponding interaction lemmas!
+
+  Further repetition can be introduced by the choice of variable
+  representation. For example, for de Bruijn indices, substitution is
+  usually defined in terms of renamings to allow for structural
+  induction.
+  With a naive approach, this would again double the amount of
+  substitution machinery, as all definitions and lemmas need to be
+  made first for renamings and then again for substitutions.
+
+  Even more repetition is introduced by having a typed language,
+  where for each substitution (and renaming) operation, we need to prove
+  that the operation preserves typing, again doubling the repetition.
+
+  % In summary, to have enough substitution machinery to prove subject
+  % reduction for a language with expression-, type-, and
+  % kind-variables one can end up with a total of 24
+  % repetitions for each definition related to substitution, not counting
+  % the interaction lemmas between different substitutions that are
+  % acting on the same syntactic categories!
+  In summary, formalizing a language with expression-, type-, and
+  kind-variables with a naive approach can end up with a total of 24
+  repetitions for each definition related to substitution, not counting
+  the interaction lemmas between different substitutions that are
+  acting on the same syntactic categories!
+
+  In our framework we address those repetitions as followed:
+  \begin{itemize}
+  \item
+    Instead of requiring a separate substitution for each sort of variable,
+    we define a multi-sorted substitution which replaces all variables
+    of all sorts simultaneously and can be applied to terms of all syntactic categories.
+    This not only unifies many different definitions, removing redundancy on the way,
+    but also eliminates the need for interaction lemmas.
+  \item
+    We use de Bruijn indices, but unify renamings and typings using the Kit abstraction.
+    We further developed a Kit-like abstraction to also unify
+    compositions of renamings and substitutions.
+  \end{itemize}
+
+  A close examination of the formalizations reveals that variable
+  substitution behaves uniformly across languages, in the sense that
+  the behavior of substitution is completely determined by the positions
+  of variable bindings and variable uses.
+
   Formalizing a typed programming language in a proof assistant
@@
   Consequently, the definition of an operation on terms also acts
-  as a type preservation lemma for that operation, as it has to
+  as a proof of type preservation for that operation, as it has to
   map well-typed terms to well-typed terms.
@@
   their type preservation lemmas are also separate entities.
-  However, hope is not lost for extrinsic typing to keep up with the
-  conciseness of intrinsic typing: 
-  While extrinsic typing cannot eliminate the shared structure between
-  definitions and preservation lemmas, it can automated by reflection
-  or generic programming techniques, leaving only the preservation lemmas
-  to be proven by the programmer. While typing can be arbitrary complex,
-  the definition and metatheory of untyped substitution behaves the same
-  across languages.
-
-  However, when working with library support, this is less of a benefit.
-  Most operations, like substitution, have a generic definition and
-  metatheory across languages, which can be derived automatically via
-  reflection or generic programming techniques.
-  Only by adding a typing-relation
-
-  However, intrinsic typing also has three major drawbacks:
-  \begin{enumerate}
-  \item
-    It is more difficult to judge whether a specification is correct,
-    because there is no clear separation between specification and proofs.
-    To understand the specification one has to additionally
-    understand the proofs, which can introduce significant cognitive overhead.
-    % For example specifying a small step semantics is directly intertwined
-    % with the proof of subject reduction.
-  \item
-    It is less uniform for object languages with multiple variable
-    sorts, e.g. type- and expression-variables. In such a scenario,
-    types are usually indexed by kinds and expressions by kinds and types.
-    This requires not only a type-in-type and expression-in-expression
-    substitution, but also a type-in-expression substitution, which causes
-    interactions between different types of substitutions introducing
-    additional complexity.
-    The duplication increases even more if one additionally introduces
-    kind-variables. Then a total of 6 different substitutions and
-    their interactions is required.
-    Unifying these substitutions to reduce repetition is difficult
-    with intrinsic typing due to the interdependence of definitions.
-  \item
-    It is less suitable for automation, because type preservation is baked
-    into the syntax definition and type preservation can become arbitrary complex
-    depending on the typing relation. This can prevent substitition
-    definition and lemmas from being derived, where otherwise it would be possible.
-  \end{enumerate}
-  With extrinsic typing the major drawback is the duplication between
-  definitions and type preservation lemmas, which we found more
-  suitable to be eliminated by automation.
-
-  To combat those drawbacks, we went in the direction of intrinsically
-  scoped syntax in our search for a general, clean, and automated framework
-  for metatheory.
+  This has the benefit that specification and proofs are kept strictly separate,
+  allowing a human to judge whether a specification is correct without also
+  having to understand the proofs.
+  % To understand the specification of intrinsic typing one has to
+  % additionally understand the proofs, which can introduce significant
+  % cognitive overhead.
+
+  For our substition framework, we choose to work with extrinsic
+  typing for a different reason.
+
+  % For substitution, this separation brings an opportunity for automation:
+  % free from the arbitrary complexity of typing relations,
+  % the definitions and lemmas of substitution for untyped terms
+  % behaves uniformly across object languages, and can be derived
+  % completely.
+
+  % While extrinsic typing cannot eliminate the shared structure between
+  % definitions and preservation lemmas, it can automated by reflection
+  % or generic programming techniques, leaving only the preservation lemmas
+  % to be proven by the programmer. While typing can be arbitrary complex,
+  % the definition and metatheory of untyped substitution behaves the same
+  % across languages.
+
+  % However, when working with library support, this is less of a benefit.
+  % Most operations, like substitution, have a generic definition and
+  % metatheory across languages, which can be derived automatically via
+  % reflection or generic programming techniques.
+  % Only by adding a typing-relation
+
+  % However, intrinsic typing also has three major drawbacks:
+  % \begin{enumerate}
+  % \item
+  %   It is more difficult to judge whether a specification is correct,
+  %   because there is no clear separation between specification and proofs.
+  %   To understand the specification one has to additionally
+  %   understand the proofs, which can introduce significant cognitive overhead.
+  %   % For example specifying a small step semantics is directly intertwined
+  %   % with the proof of subject reduction.
+  % \item
+  %   It is less uniform for object languages with multiple variable
+  %   sorts, e.g. type- and expression-variables. In such a scenario,
+  %   types are usually indexed by kinds and expressions by kinds and types.
+  %   This requires not only a type-in-type and expression-in-expression
+  %   substitution, but also a type-in-expression substitution, which causes
+  %   interactions between different types of substitutions introducing
+  %   additional complexity.
+  %   The duplication increases even more if one additionally introduces
+  %   kind-variables. Then a total of 6 different substitutions and
+  %   their interactions is required.
+  %   Unifying these substitutions to reduce repetition is difficult
+  %   with intrinsic typing due to the interdependence of definitions.
+  % \item
+  %   It is less suitable for automation, because type preservation is baked
+  %   into the syntax definition and type preservation can become arbitrary complex
+  %   depending on the typing relation. This can prevent substitition
+  %   definition and lemmas from being derived, where otherwise it would be possible.
+  % \end{enumerate}
+  % With extrinsic typing the major drawback is the duplication between
+  % definitions and type preservation lemmas, which we found more
+  % suitable to be eliminated by automation.
+
+  % To combat those drawbacks, we went in the direction of intrinsically
+  % scoped syntax in our search for a general, clean, and automated framework
+  % for metatheory.
 
@@
   definitions and lemmas, but omit some proofs for the sake of
-  space. The omitted proofs can be found in Appendix~\ref{sec:proofs} and the
-  supplementary material. The full framework can be found at TODO.
+  space. The omitted proofs can be found in Appendix~\ref{sec:proofs}
+  and in the supplementary material. The full framework can be found at
+  TODO.
 
@@
   \label{sec:syntax}
-  \subsection{Single-Sorted Syntax}
+  \subsection{Unsorted Syntax}
   \label{sec:syntax:single}
@@
   \label{sec:related}
+
+  The kit abstraction for unifying renamings and substitutions
+  appeared first in an unpublished article by
+  McBride\cite{unpublished:mcbride2005kits}, and then later in
+  Benton et al.\cite{DBLP:journals/jar/BentonHKM12}. 
+  Wood and Atkey\cite{DBLP:journals/corr/abs-2005-02247} propose an
+  extension to kits that supports substructural types via resource
+  vectors.
+  In all three cases they are formulated for intrinsic typing.
+
+  \emph{Autosubst}\cite{DBLP:conf/itp/SchaferTS15}
+  is a Coq framework, which derives parallel substitution definitions and
+  lemmas for languages from annotated Coq syntax definitions using
+  extrinsic typing, extrinsic scoping, and de Bruijn indices.
+  The framework is implemented in Coq's tactic language LTAC and comes
+  with a decision procedure for all assumption-free, equational
+  substitution-lemmas.
+  The implementation of Autosubst deals with multiple variable sorts
+  by generating multiple substitutions and corresponding interaction lemmas.
+  Autosubst does not provide any machinery for type preservation and
+  does not unify renaming and substitution.
+  % Assumes fun-ext.
+
+  \emph{Autosubst 2}\cite{DBLP:conf/cpp/StarkSK19} is a standalone
+  code generator, which translates second-order HOAS specifications into
+  mutual inductive term sorts. Compared to Autosubst 1, it features
+  mutually recursive object languages, intrinsically scoping, and
+  vectorized substitutions.
+  They also define and prove lemmas for all four compositions.
+  The generated syntax consists of multiple term types, indexed by
+  different amounts of free variables as shown in Section~\ref{sec:syntax}.
+  % Does not assume fun-ext.
+
+  Allais et al.\cite{DBLP:conf/cpp/Allais0MM17} propose a powerful
+  abstraction for denotational semantics and proving semantic fusion
+  lemmas.
+  In later work\cite{DBLP:journals/pacmpl/AllaisA0MM18}, they use
+  generic programming to instanciate this abstraction for a class of
+  object languages comparable to ours.
+  They demonstrate how both renaming and substitutions can be
+  described as semantics and how the four composition laws follow from
+  their generic fusion laws.
+  They do not show how to work with multiple variable sorts, how to
+  abstract over the compositions, and how to abstract over typing in
+  an extrinsically typed scenario.
+
   Autosubst \cite{DBLP:conf/cpp/StarkSK19, DBLP:conf/itp/SchaferTS15}.
@@
 
+  Abstract binding trees
+
   \section{Conclusion}

### 2023-09-18 paper/paper.tex
Update.

@@
   Section~\ref{sec:casestudies} describes case studies of object languages beyond System F.
+  Section~\ref{sec:related} discusses related work.
   Section~\ref{sec:future} describes future work.
-  Section~\ref{sec:related} discusses related work.
   Section~\ref{sec:conclusion} concludes.
@@
 
-  \section{Future Work}
-  \label{sec:future}
-  \begin{itemize}
-  \item
-    Enabling more object languages by allowing subterms with less
-    restrictions on the form of \ACode{S}.
-  \item
-    Providing abstractions not just for type-preservation, but also
-    for different kinds of semantics, e.g. confluence or lemmas related
-    to denotational semantics.
-  \end{itemize}
-
   \section{Related Work}
@@
   The kit abstraction for unifying renamings and substitutions
-  appeared first in an unpublished article by
+  appeared first in an unpublished manuscript by
   McBride\cite{unpublished:mcbride2005kits}, and then later in
@@
   does not unify renaming and substitution.
-  % Assumes fun-ext.
 
@@
   different amounts of free variables as shown in Section~\ref{sec:syntax}.
-  % Does not assume fun-ext.
 
   Allais et al.\cite{DBLP:conf/cpp/Allais0MM17} propose a powerful
-  abstraction for denotational semantics and proving semantic fusion
-  lemmas.
+  abstraction for denotational semantics and semantic fusion lemmas.
   In later work\cite{DBLP:journals/pacmpl/AllaisA0MM18}, they use
@@
   object languages comparable to ours.
-  They demonstrate how both renaming and substitutions can be
-  described as semantics and how the four composition laws follow from
-  their generic fusion laws.
+  They demonstrate how both renamings and substitutions can be
+  described as semantics and how the four composition lemmas follow from
+  their generic fusion lemma.
+  They do cover an abstraction to unify renamings and substitutions.
   They do not show how to work with multiple variable sorts, how to
@@
   an extrinsically typed scenario.
-
-  Autosubst \cite{DBLP:conf/cpp/StarkSK19, DBLP:conf/itp/SchaferTS15}.
-  Kits \cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}.
-  Kits linear \cite{DBLP:journals/corr/abs-2005-02247}.
-  Universe \cite{DBLP:journals/pacmpl/AllaisA0MM18}.
+  However, we believe that their framework is general enough such that
+  our concepts can also be transferred to their framework.
 
   Abstract binding trees
+
+  \section{Future Work}
+  \label{sec:future}
+  \begin{itemize}
+  \item
+    Enabling more object languages by allowing subterms with less
+    restrictions on the form of \ACode{S}, allowing recursive
+    occurences in container types like lists and finite functions, and
+    adding support for substructural types as in
+    \cite{DBLP:journals/corr/abs-2005-02247}.
+  \item
+    Incorporating a decision procedure for equational
+    substitution-lemmas based on $\sigma$-calculus as in Autosubst
+    \cite{DBLP:conf/itp/SchaferTS15, DBLP:conf/cpp/StarkSK19}.
+  \item
+    Developing abstractions and derived lemmas beyond substition and
+    type preservation, e.g. for confluence or denotational semantics
+    as in \cite{DBLP:journals/pacmpl/AllaisA0MM18}.
+  \end{itemize}
 

### 2023-09-18 paper/paper.tex
Update.

@@
   duplicated three times! We need to substitute expression-variables in
-  expressions, type-variables in types, but we also need to substitute
-  type-variables in expressions.
+  expressions, type-variables in types, but also type-variables in
+  expressions.
   Even worse, having two substitutions acting on expressions, requires
@@
 
-  Even more repetition is introduced by having a typed language,
-  where for each substitution (and renaming) operation, we need to prove
-  that the operation preserves typing, again doubling the repetition.
+  Similarly, for a typed language formalized via extrinsic typing we
+  need type preservation lemmas for each substitution and renaming
+  operation, which again doubles the repetition.
 
@@
   % acting on the same syntactic categories!
+  If one is not careful, a formalization with three sorts of variables
+  can easily end up with 24 definitions for each operation related
+  to substitution, which usually includes at least a substitution type, and
+  operations for application, extension, weakening, lifting, singleton
+  substitution, extensional equality, and an identity substitution.
+
   In summary, formalizing a language with expression-, type-, and
@@
       The actual implementation does \emph{not} use any postulates.
-      Appendix TODO explains the workaround.
+      The workaround is explained in Appendix~\ref{sec:fun-ext}.
     }
@@
   \LibCode\KTypingTraversal
-  Given a variable/term \ACode{e} with typing \ACode{⊢e} and
+  Given a term \ACode{e} with typing \ACode{⊢e} and
   a renaming/substitution \ACode{ϕ} with typing \ACode{⊢ϕ},
@@
   appeared first in an unpublished manuscript by
-  McBride\cite{unpublished:mcbride2005kits}, and then later in
+  McBride\cite{unpublished:mcbride2005kits}, and later in
   Benton et al.\cite{DBLP:journals/jar/BentonHKM12}. 
@@
   \item
+    Extending the class of typing relations covered by
+    the \ACode{\ACon{TypingTraversal}} abstraction and identifying
+    a subclass for which it can be derived.
+  \item
     Incorporating a decision procedure for equational
@@
 
+  \section{System F with Reflection}
+  \label{sec:reflection-example}
+
+  \section{Map Equality Without Functional Extensionality}
+  \label{sec:fun-ext}
+
+  \section{Features of the Actual Implementation}
+  \label{sec:features}
+
 \end{document}

### 2023-09-18 paper/paper.tex
Update.

@@
   need type preservation lemmas for each substitution and renaming
-  operation, which again doubles the repetition.
+  operation, which again doubles the number of definitions.
 
@@
   % acting on the same syntactic categories!
-  If one is not careful, a formalization with three sorts of variables
-  can easily end up with 24 definitions for each operation related
-  to substitution, which usually includes at least a substitution type, and
-  operations for application, extension, weakening, lifting, singleton
-  substitution, extensional equality, and an identity substitution.
-
-  In summary, formalizing a language with expression-, type-, and
-  kind-variables with a naive approach can end up with a total of 24
-  repetitions for each definition related to substitution, not counting
-  the interaction lemmas between different substitutions that are
-  acting on the same syntactic categories!
-
-  In our framework we address those repetitions as followed:
-  \begin{itemize}
-  \item
-    Instead of requiring a separate substitution for each sort of variable,
-    we define a multi-sorted substitution which replaces all variables
-    of all sorts simultaneously and can be applied to terms of all syntactic categories.
-    This not only unifies many different definitions, removing redundancy on the way,
-    but also eliminates the need for interaction lemmas.
-  \item
-    We use de Bruijn indices, but unify renamings and typings using the Kit abstraction.
-    We further developed a Kit-like abstraction to also unify
-    compositions of renamings and substitutions.
-  \end{itemize}
-
-  A close examination of the formalizations reveals that variable
-  substitution behaves uniformly across languages, in the sense that
-  the behavior of substitution is completely determined by the positions
-  of variable bindings and variable uses.
-
-  Formalizing a typed programming language in a proof assistant
-  requires to choose between an extrinsic and intrinsic representation
-  of typing.
-
-  With \emph{intrinsic typing}, the definitions of syntax and typing
-  are merged into a single definition of well-typed terms.
-  Consequently, the definition of an operation on terms also acts
-  as a proof of type preservation for that operation, as it has to
-  map well-typed terms to well-typed terms.
-  For example, the definition of substitution
-  also acts as a proof that substitution preserves typing, and the
-  definition of a small-step semantics also acts as a proof for subject
-  reduction.
-  This often has the benefit of more concise code, as shared structure
-  between definitions and preservation lemmas does not need to be
-  repeated.
-
-  With \emph{extrinsic typing}, the definitions of syntax and typing
-  are kept separately: the syntax is defined first, and then typing is
-  defined as a relation. Consequently, operations on terms and
-  their type preservation lemmas are also separate entities.
-  This has the benefit that specification and proofs are kept strictly separate,
-  allowing a human to judge whether a specification is correct without also
-  having to understand the proofs.
-  % To understand the specification of intrinsic typing one has to
-  % additionally understand the proofs, which can introduce significant
-  % cognitive overhead.
-
-  For our substition framework, we choose to work with extrinsic
-  typing for a different reason.
-
-  % For substitution, this separation brings an opportunity for automation:
-  % free from the arbitrary complexity of typing relations,
-  % the definitions and lemmas of substitution for untyped terms
-  % behaves uniformly across object languages, and can be derived
-  % completely.
-
-  % While extrinsic typing cannot eliminate the shared structure between
-  % definitions and preservation lemmas, it can automated by reflection
-  % or generic programming techniques, leaving only the preservation lemmas
-  % to be proven by the programmer. While typing can be arbitrary complex,
-  % the definition and metatheory of untyped substitution behaves the same
-  % across languages.
-
-  % However, when working with library support, this is less of a benefit.
-  % Most operations, like substitution, have a generic definition and
-  % metatheory across languages, which can be derived automatically via
-  % reflection or generic programming techniques.
-  % Only by adding a typing-relation
-
-  % However, intrinsic typing also has three major drawbacks:
-  % \begin{enumerate}
+  If one is not careful, a formalization of a language with type and
+  kind polymorphism can easily end up with 24 slightly changed copies of
+  the whole substitution machinery, which for de Bruijn indices tends to
+  include at least a type for parallel substitution, and operations for
+  application, extension, weakening, lifting, singleton substitution,
+  extensional equality, and an identity substitution - and this is not
+  counting the slightly changed copies of interaction lemmas!
+
+  Clearly, this situation is in need of automation.
+  Our framework approaches this problem using a combination of
+  abstractions and reflection in the context of extrinsically typed,
+  intrinsically scoped syntax with de Bruijn indices.
+  The user can write a language specification using regular Agda
+  definitions (no generic encodings) and our framework derives
+  definitions and lemmas for untyped substitution, and provides an
+  abstraction to prove type preservation for all substitution operations
+  using a single lemma. For System F, our framework allows to prove
+  subject reduction with only a single handwritten lemma for type
+  preservation.
+  Compared to many other approaches, we do not derive substitutions
+  for each of the variable sorts and syntactic categories, but instead
+  use a novel approach for defining syntax, which directly supports
+  substitutions that replace all variable sorts in parallel and can be
+  applied to all syntactic categories.
+  By further unifying renamings and substitutions, we gain the ability
+  to talk abstractly about any kind of renaming or substitution that
+  can occur in the formalization.
+  Together with abstractions to unify renamings and substitutions,
+  this allows us to create an abstraction that is general enough to
+  prove type preservation for all sub
+
+  % Choosing an extrinsic typing approach might seem counter intuitive,
+  % as separating
+
+  % In our framework we address those repetitions as followed:
+  % \begin{itemize}
   % \item
-  %   It is more difficult to judge whether a specification is correct,
-  %   because there is no clear separation between specification and proofs.
-  %   To understand the specification one has to additionally
-  %   understand the proofs, which can introduce significant cognitive overhead.
-  %   % For example specifying a small step semantics is directly intertwined
-  %   % with the proof of subject reduction.
+  %   Instead of requiring a separate substitution for each sort of variable,
+  %   we define a multi-sorted substitution which replaces all variables
+  %   of all sorts simultaneously and can be applied to terms of all syntactic categories.
+  %   This not only unifies many different definitions, removing redundancy on the way,
+  %   but also eliminates the need for interaction lemmas.
   % \item
-  %   It is less uniform for object languages with multiple variable
-  %   sorts, e.g. type- and expression-variables. In such a scenario,
-  %   types are usually indexed by kinds and expressions by kinds and types.
-  %   This requires not only a type-in-type and expression-in-expression
-  %   substitution, but also a type-in-expression substitution, which causes
-  %   interactions between different types of substitutions introducing
-  %   additional complexity.
-  %   The duplication increases even more if one additionally introduces
-  %   kind-variables. Then a total of 6 different substitutions and
-  %   their interactions is required.
-  %   Unifying these substitutions to reduce repetition is difficult
-  %   with intrinsic typing due to the interdependence of definitions.
-  % \item
-  %   It is less suitable for automation, because type preservation is baked
-  %   into the syntax definition and type preservation can become arbitrary complex
-  %   depending on the typing relation. This can prevent substitition
-  %   definition and lemmas from being derived, where otherwise it would be possible.
-  % \end{enumerate}
-  % With extrinsic typing the major drawback is the duplication between
-  % definitions and type preservation lemmas, which we found more
-  % suitable to be eliminated by automation.
-
-  % To combat those drawbacks, we went in the direction of intrinsically
-  % scoped syntax in our search for a general, clean, and automated framework
-  % for metatheory.
+  %   We use de Bruijn indices, but unify renamings and typings using the Kit abstraction.
+  %   We further developed a Kit-like abstraction to also unify
+  %   compositions of renamings and substitutions.
+  % \end{itemize}
+
+  % A close examination of the formalizations reveals that variable
+  % substitution behaves uniformly across languages, in the sense that
+  % the behavior of substitution is completely determined by the positions
+  % of variable bindings and variable uses.
+
+  % Formalizing a typed programming language in a proof assistant
+  % requires to choose between an extrinsic and intrinsic representation
+  % of typing.
+
+  % With \emph{intrinsic typing}, the definitions of syntax and typing
+  % are merged into a single definition of well-typed terms.
+  % Consequently, the definition of an operation on terms also acts
+  % as a proof of type preservation for that operation, as it has to
+  % map well-typed terms to well-typed terms.
+  % For example, the definition of substitution
+  % also acts as a proof that substitution preserves typing, and the
+  % definition of a small-step semantics also acts as a proof for subject
+  % reduction.
+  % This often has the benefit of more concise code, as shared structure
+  % between definitions and preservation lemmas does not need to be
+  % repeated.
+
+  % With \emph{extrinsic typing}, the definitions of syntax and typing
+  % are kept separately: the syntax is defined first, and then typing is
+  % defined as a relation. Consequently, operations on terms and
+  % their type preservation lemmas are also separate entities.
+  % This has the benefit that specification and proofs are kept strictly separate,
+  % allowing a human to judge whether a specification is correct without also
+  % having to understand the proofs.
+  % % To understand the specification of intrinsic typing one has to
+  % % additionally understand the proofs, which can introduce significant
+  % % cognitive overhead.
+
+  % For our substition framework, we choose to work with extrinsic
+  % typing for a different reason.
+
+  % % For substitution, this separation brings an opportunity for automation:
+  % % free from the arbitrary complexity of typing relations,
+  % % the definitions and lemmas of substitution for untyped terms
+  % % behaves uniformly across object languages, and can be derived
+  % % completely.
+
+  % % While extrinsic typing cannot eliminate the shared structure between
+  % % definitions and preservation lemmas, it can automated by reflection
+  % % or generic programming techniques, leaving only the preservation lemmas
+  % % to be proven by the programmer. While typing can be arbitrary complex,
+  % % the definition and metatheory of untyped substitution behaves the same
+  % % across languages.
+
+  % % However, when working with library support, this is less of a benefit.
+  % % Most operations, like substitution, have a generic definition and
+  % % metatheory across languages, which can be derived automatically via
+  % % reflection or generic programming techniques.
+  % % Only by adding a typing-relation
+
+  % % However, intrinsic typing also has three major drawbacks:
+  % % \begin{enumerate}
+  % % \item
+  % %   It is more difficult to judge whether a specification is correct,
+  % %   because there is no clear separation between specification and proofs.
+  % %   To understand the specification one has to additionally
+  % %   understand the proofs, which can introduce significant cognitive overhead.
+  % %   % For example specifying a small step semantics is directly intertwined
+  % %   % with the proof of subject reduction.
+  % % \item
+  % %   It is less uniform for object languages with multiple variable
+  % %   sorts, e.g. type- and expression-variables. In such a scenario,
+  % %   types are usually indexed by kinds and expressions by kinds and types.
+  % %   This requires not only a type-in-type and expression-in-expression
+  % %   substitution, but also a type-in-expression substitution, which causes
+  % %   interactions between different types of substitutions introducing
+  % %   additional complexity.
+  % %   The duplication increases even more if one additionally introduces
+  % %   kind-variables. Then a total of 6 different substitutions and
+  % %   their interactions is required.
+  % %   Unifying these substitutions to reduce repetition is difficult
+  % %   with intrinsic typing due to the interdependence of definitions.
+  % % \item
+  % %   It is less suitable for automation, because type preservation is baked
+  % %   into the syntax definition and type preservation can become arbitrary complex
+  % %   depending on the typing relation. This can prevent substitition
+  % %   definition and lemmas from being derived, where otherwise it would be possible.
+  % % \end{enumerate}
+  % % With extrinsic typing the major drawback is the duplication between
+  % % definitions and type preservation lemmas, which we found more
+  % % suitable to be eliminated by automation.
+
+  % % To combat those drawbacks, we went in the direction of intrinsically
+  % % scoped syntax in our search for a general, clean, and automated framework
+  % % for metatheory.
 
@@
 
+  Frameworks for programming languages metatheory can be broadly
+  categorized by representation of variable bindings, representation
+  of typing and scoping, and requirements on the meta-language.
+
+  Variable bindings have been represented as de Bruijn indices, de
+  Bruijn levels, names, locally nameless, locally named, shifted names,
+  nominal techniques, (parameterized) higher order abstract syntax.
+
+  Typing and scoping can be both extrinsic, both intrinsic, or
+  extrinsically typed and intrinsically scoped.
+
+  As the amount of related work is rather large, we focus on work that
+  is more closely related to ours and refer to other papers for the
+  broader picture. In particular we cover abstractions for renamings
+  and substitutions, comparable extrinsically typed approaches, but also
+  intrinsically typed approaches that could be used as a basis for
+  multi-sorted syntax, which as mentioned before is intrinsically typed
+  in disguise.
+
+  \subsection{Unifying Renamings \& Substitution}
+
   The kit abstraction for unifying renamings and substitutions
@@
   vectors.
-  In all three cases they are formulated for intrinsic typing.
+  In all three cases kits are formulated for intrinsic typing
+  and scenarios with polymorphism are not considered.
+
+  \subsection{Extrinsically Typed Approaches}
 
@@
   different amounts of free variables as shown in Section~\ref{sec:syntax}.
+
+  Abstract binding trees
+
+  \subsection{Intrinsically Typed Approaches}
 
@@
 
-  Abstract binding trees
-
   \section{Future Work}

### 2023-09-18 paper/paper.tex
Update.

@@
   is more closely related to ours and refer to other papers for the
-  broader picture. In particular we cover abstractions for renamings
+  broader picture. In particular, we discuss abstractions for renamings
   and substitutions, comparable extrinsically typed approaches, but also

### 2023-09-18 paper/paper.tex
Update.

@@
 
-  Further repetition can be introduced by the choice of variable
+  Further repetition can occur due to the choice of variable
   representation. For example, for de Bruijn indices, substitution is
@@
   preservation.
+
   Compared to many other approaches, we do not derive substitutions
@@
   can occur in the formalization.
+  Having unified the untyped substitution machinery, we build further
+  abstractions to similarly unify typing and type preservation lemmas.
+  Our reflection algorithm instanciates those abstractions
+
   Together with abstractions to unify renamings and substitutions,
@@
 
-  Frameworks for programming languages metatheory can be broadly
-  categorized by representation of variable bindings, representation
-  of typing and scoping, and requirements on the meta-language.
+  Frameworks for programming language metatheory can be broadly
+  categorized by representation of variable bindings and
+  representation of typing and scoping.
 
@@
   by generating multiple substitutions and corresponding interaction lemmas.
-  Autosubst does not provide any machinery for type preservation and
-  does not unify renaming and substitution.
 
@@
   different amounts of free variables as shown in Section~\ref{sec:syntax}.
+
+  Both works do not provide machinery for type preservation and
+  do not unify renaming and substitution (although the latter is listed as
+  future work). We believe it should be possible to translate their
+  decision procedure from vectorized substitutions to the setting of
+  our multi-sorted substitutions.
 
@@
 
+  This work is orthogonal but compatible to our approach.
+  We can instantiate the system using our syntax type system,
+  but then the semantics abstraction would work with terms
+  which are untyped with respect to the actual type system.
+  We think it is worthwile to explore if their semantics abstraction
+  can also be generalized to multi-sorted typing relations.
+
   \section{Future Work}

### 2023-09-19 paper/paper.tex
Update.

@@
   expressions.
-  Even worse, having two substitutions acting on expressions, requires
+  Even worse, having two substitutions acting on expressions requires
   to also prove lemmas about their interactions.
@@
   counting the slightly changed copies of interaction lemmas!
-
   Clearly, this situation is in need of automation.
+
   Our framework approaches this problem using a combination of
@@
   can occur in the formalization.
-  Having unified the untyped substitution machinery, we build further
-  abstractions to similarly unify typing and type preservation lemmas.
-  Our reflection algorithm instanciates those abstractions
-
-  Together with abstractions to unify renamings and substitutions,
-  this allows us to create an abstraction that is general enough to
-  prove type preservation for all sub
+  This generality is key to then define abstractions for typing and
+  type preservation on the same level of generality, allowing to prove
+  type preservation for renamings and substitutions of all variable
+  sorts and syntactic categories in a single lemma for many typing
+  relations, including those of our case studies.
+
+  % TODO: avoiding interaction lemmas
+
+  % By using extrinsic typing, we clearly separate the definitions and
+  % lemmas of untyped substitutions from typing and type preservation.
+  % This is useful, because the former can be derived for virtually any
+  % object language, while the latter cannot, as typing relations can
+  % be arbitrarily complex.
+
+  % This does not only separate specification from proofs, making it
+  % easier for a human to verify if a specification is correct without
+  % understanding the proofs, but also separates things which can be
+  % derived for virtually any object language, from things which cannot.
+
+  % This allows untyped substitution to be derived for virtually all
+  % object languages, while still allowing for general abstractions
+  % for type preservation to be built on top, which cannot be proved
+  % automatically in general, as typing relations can become arbitrarily
+  % complex.
+
+  % The class of supported typing relations is currently limited to 
+  % a subclass of the classical tertiary relations between typing
+  % context, term, and type, but strong enough work for our three case studies
+  % of System F with subtyping, lambda calculus with dependent function
+  % types, and pattern matching.
 
@@
   space. The omitted proofs can be found in Appendix~\ref{sec:proofs}
-  and in the supplementary material. The full framework can be found at
-  TODO.
+  and in the supplementary material.
+  The full framework is available online.\footnote{See supplementary material. In the final version here will be a link to github.}
 
@@
   and an abstraction to unify type preservattion lemmas for renamings and substitutions.
-  Section~\ref{sec:semantics} introduces a small-step operational semantics for System F.
-  Section~\ref{sec:subjectreduction} proves the subject reduction lemma for System F.
+  Section~\ref{sec:subjectreduction} introduces a small-step
+  operational semantics for System F and proves subject reduction.
   Section~\ref{sec:reflection} describes the class of object languages covered by
   our reflection algorithm.
-  Section~\ref{sec:casestudies} describes case studies of object languages beyond System F.
+  Section~\ref{sec:casestudies} discusses our case studies.
   Section~\ref{sec:related} discusses related work.
-  Section~\ref{sec:future} describes future work.
+  Section~\ref{sec:future} discusses future work.
   Section~\ref{sec:conclusion} concludes.
@@
   \item
-    an adaptation of McBride's kit-abstraction for unifying renamings
-    and substitutions to multi-sorted syntax;
+    an adaptation of McBride's
+    kit-abstraction\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
+    for unifying renamings and substitutions to multi-sorted syntax;
   \item
@@
   \item
-    an implementation of the concepts as framework in Agda featuring
+    an implementation as an Agda framework featuring
     a reflection algorithm, representation independence for
     substitutions and typing contexts, heterogeneous equality between
-    renamings and substitutions, and free of postulated axioms.
+    renamings and substitutions, and absence of postulated axioms.
   \item
@@
 
-  \section{Semantics}
-  \label{sec:semantics}
+  \section{Subject Reduction}
+  \label{sec:subjectreduction}
+
   We define a standard small-step operational semantics for System F
@@
 
-  \section{Subject Reduction}
-  \label{sec:subjectreduction}
   Proving subject reduction is now straightforward:

### 2023-09-19 paper/paper.tex
Update.

@@
     Consequently, interaction lemmas between those substitutions are required.
-    In general, this leads to a combinatory explosion of
-    substitutions, e.g.\ adding kind variables would lead to a total of 6
-    different substitutions and corresponding interaction lemmas.
+    % In general, this leads to a combinatory explosion of
+    % substitutions, e.g.\ adding kind variables would lead to a total of 6
+    % different substitutions and corresponding interaction lemmas.
   \end{enumerate}
@@
   it has been done for intrinsically typed languages with dependent
-  types, but those constructions come with their own drawbacks and complexity.
+  types\cite{TODO}, but those constructions come with additional
+  complexity and are less suitable to automation.
 
@@
   \label{sec:syntax:structure}
-  The multi-sorted syntax makes it easy to define a structure for
-  syntaxes of arbitrary object-languages, i.e.\ syntaxes with
-  arbitrary amounts of syntactic categories and variables:
+  The regularity of the multi-sorted syntax makes it easy to define a
+  structure for syntaxes of arbitrary object-languages, i.e.\ syntaxes
+  with arbitrary amounts of syntactic categories and variables:
   \LibCode\KSyntax
@@
 
-  Building on the operations and axioms, we derive further operations:\footnote{
+  Building on the operations and axioms of a \ACode{\ACon{Kit}}, we
+  derive further operations:\footnote{
     The definitions are included directly in the record module of the
@@
   \item 
-    A map can be shifted, as the Kit provides the necessary \ACode{\AField{wk}} operation:
+    A map can be shifted, as the \ACode{\ACon{Kit}} provides the
+    necessary \ACode{\AField{wk}} operation:
     \LibCode\KWkm
@@
   We also specify notation, which makes it less cumbersome
-  to explicitly pass a particular Kit to those definitions:
+  to explicitly pass a particular \ACode{\ACon{Kit}} to those definitions:
   \begin{itemize}
@@
   \LibCode\KTraversal
-  The fields of this structure are:
+  The fields of this structure have the following meaning:
   \begin{itemize}
@@
     \ACode{\AField{⋯-var}} states that applying a map \ACode{ϕ} to a variable
-    term \ACode{\AField{`} x}, is the same as applying \ACode{ϕ} to
-    the variable \ACode{x}, and then converting the result from \ACode{S₂ ∋/⊢ S}
-    to \ACode{S₂ ⊢ S} via \ACode{\AField{id/`}}.
+    term\\ \ACode{\AField{`} x}, is the same as applying \ACode{ϕ} to
+    the variable \ACode{x}, and then converting the result to a term
+    via \ACode{\AField{id/`}}.
   \item
@@
     we get back a term that we can use directly.
-    Hence, the use of \ACode{`/id}.
+    This is exactly what \ACode{\AField{`/id}} does.
   \item 
@@
 
+  Then we prove that applying an identity map does not change the term:
   \AppCode\FTraversalId
   \AppCode\FTraversalIdProofInteresting
+  The interesting cases are:
+  \begin{itemize}
+  \item
+    the variable case \ACode{\ACon{`} x}, where we need to combine the
+    \ACon{\AField{id/`}} from the identity map, with the \ACon{\AField{`/id}}
+    from the traversal operation using \ACon{\AField{`/`-is-`}}.
+  \item
+    the lambda abstraction case \ACode{\ACon{λx} e}, where the
+    traversal lifts the identity over its binder. Here
+    \ACon{\AField{id↑-id}} is used to state that this lifted identity
+    map is still the identity.
+  \item
+    all other cases are analogous and hence omitted.
+  \end{itemize}
+
+  Finally, we instanciate the \ACode{\ACon{Traversal}} structure:
   % \LibCode\FTraversalIdProofRest
@@
 
-  \ExampleCode\FExampleTrav
+  % \ExampleCode\FExampleTrav
 

### 2023-09-19 paper/paper.tex
Update.

@@
 
+  % TODO note that its derived
+
+  % \subsection{Unused}
   % \ExampleCode\FExampleTrav
-
-  % \subsection{Unused}
   % \LibCode\KKitOpenInst
@@
 
-  Composition \AField{\_·ₖ\_} equips our framework with the power to
-  reason about multiple maps \ACode{ϕ₁} and \ACode{ϕ₂} being applied
-  to a term in sequence, in terms of a single application:
-
   The core property of composition is that applying two maps
@@
   $$
-  \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
+  \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
   $$
-
-  %%%%%%%%%%
-
-  In this section, we extend our framework with a single composition
-  operator for arbitrary maps.
-
-  Composition \AField{\_·ₖ\_} equips our framework with the power to
-  reason about multiple maps \ACode{ϕ₁} and \ACode{ϕ₂} being applied
-  to a term in sequence:
-
-  \ACode{\AField{⋯-assoc} t ϕ₁ ϕ₂} \ \ : \ \ \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
-
-  For many lemmas, e.g. substitution-preserves-typing in System F,
-  it is necessary to apply an equality like \ACode{t ⋯ σ₁ ⋯ σ₂ ≡ t ⋯ σ₂' ⋯ σ₁'}
-  where the induction hypothesis requires \ACode{t' ⋯ σ₁'}.
-
-  %%%%%%%%%%
-
-  In this section, we extend our framework with a single composition
-  operator for arbitrary maps.
-
-  Composition of substitutions is important for most language
-  formalizations, as it allows to reason about the meaning of applying
-  multiple substitutions to a term. This appears for example in the proof of
-  substitution-preserves-typing for System F, where in the
-  case of type-application \ACode{(e \ACon{∙} t) \ADef{⋯} σ} we need to reason about
-  \ACode{(e \ADef{⋯} \ADef{⦅} t \ADef{⦆}) \ADef{⋯} σ}, which is equivalent to
-  \ACode{e \ADef{⋯} (\ADef{⦅} t \ADef{⦆} \ADef{·ₖ} σ)},
-  where \ACode{\_\ADef{·ₖ}\_} is backward composition.
+  We call this property \ACode{\AField{⋯-fusion}}.
+  This property equips our framework with the power to
+  reason about the application of multiple maps in terms of
+  the application of a single map.
+  As such it forms the basis for all lemmas involving multiple maps,
+  e.g.\ that applying a weakening and then a singleton
+  substitution cancel each other out.
+
+  % For example, we can prove that first applying a weakening
+  % \ACode{\ADef{weaken} s}, and then applying a singleton substitution
+  % \ACode{\ADef{⦅} t \ADef{⦆}} cancel each other cancel each other out
+  % by proving that
+  % \ACode{(\ADef{weaken} s \AField{·ₖ} \ADef{⦅} t \ADef{⦆}) \ADef{$\sim$} id}.
+  % As such it forms the basis for proving that a shifting cancels a
+  % singleton substitution, or that
+
+  % For many lemmas, e.g. substitution-preserves-typing in System F,
+  % it is necessary to apply an equality like \ACode{t ⋯ σ₁ ⋯ σ₂ ≡ t ⋯ σ₂' ⋯ σ₁'}
+  % where the induction hypothesis requires \ACode{t' ⋯ σ₁'}.
+
+  % Composition of substitutions is important for most language
+  % formalizations, as it allows to reason about the meaning of applying
+  % multiple substitutions to a term. This appears for example in the proof of
+  % substitution-preserves-typing for System F, where in the
+  % case of type-application \ACode{(e \ACon{∙} t) \ADef{⋯} σ} we need to reason about
+  % \ACode{(e \ADef{⋯} \ADef{⦅} t \ADef{⦆}) \ADef{⋯} σ}, which is equivalent to
+  % \ACode{e \ADef{⋯} (\ADef{⦅} t \ADef{⦆} \ADef{·ₖ} σ)},
+  % where \ACode{\_\ADef{·ₖ}\_} is backward composition.
 
@@
   consider all four compositions between renamings and substitutions.
-  While the composition operations are defined independently of each other,
-  the \ACode{\AField{⋯-assoc}} lemma for two substitutions, depends on the
-  \ACode{\AField{⋯-assoc}} lemmas between a renaming and a substitution, which in
-  turn depend on the \ACode{\AField{⋯-assoc}} lemma between two renamings.
-
-  McBride et al.\cite{DBLP:journals/jar/BentonHKM12} addressed this
+  While the composition operations can be defined independently of each other,
+  the \ACode{\AField{⋯-fusion}} lemma for two substitutions, depends on the
+  \ACode{\AField{⋯-fusion}} lemmas between a renaming and a substitution, which in
+  turn depend on the \ACode{\AField{⋯-fusion}} lemma between two renamings.
+
+  Benton et al.\cite{DBLP:journals/jar/BentonHKM12} addresses this
   issue by duplicating the definitions and using tactics to reduce
   boilerplate in proofs.
-
-  In contrast we define structures similar to \ACode{\ACon{Kit}} and
-  \ACode{\ACon{Traversal}}, which allows us to abstract over all four
+  In contrast, we define structures similar to \ACode{\ACon{Kit}} and
+  \ACode{\ACon{Traversal}}, which allow us to abstract over all four
   compositions and use the same trick as before to eliminate the
   dependencies.
-  This has the benefit that a user of our library has to write less
-  boilerplate, but more importantly it allows to prove lemmas about the
-  composition of arbitrary maps, which enable us later to unify type
+  This has the benefit, that we can prove lemmas about
+  the composition of arbitrary maps, which enable us later to unify type
   preservation of all substitution and renaming operations into a single
   lemma.
+  % TODO: clarify
 
@@
   \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
-  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-assoc}} are analogous to
+  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-fusion}} are analogous to
   \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{\_⋯\_}}.
@@
   \subsection{System F}
-  \AppCode\FAssoc
-  \AppCode\FAssocProofInteresting
-  % \AppCode\FAssocProofRest
+  \AppCode\FFusion
+  \AppCode\FFusionProofInteresting
+  % \AppCode\FFusionProofRest
   \AppCode\FComposeTraversal
@@
 
-  % \AppCode\FAssoc
-  % \AppCode\FAssocProofInteresting
-  % \AppCode\FAssocProofRest
+  % \AppCode\FFusion
+  % \AppCode\FFusionProofInteresting
+  % \AppCode\FFusionProofRest
 

### 2023-09-19 paper/paper.tex
Update.

@@
   which provides an operation \ACode{\AField{\_\&/⋯\_}} that unifies
-  \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}, allowing us
-  to define a general composition in the spirit of:
+  \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}. This allows us to
+  define a general composition in the spirit of:
   \ExampleCode\KComposition
-
-  \ACode{\ACon{ComposeKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
-  \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
-  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-fusion}} are analogous to
-  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{\_⋯\_}}.
+  We then define a \ACode{\ACon{ComposeTraversal}}, which consists of the
+  \ACode{\AField{⋯-fusion}} lemma for general maps.
+
+  Just like a \ACode{\ACon{Traversal}} allowed us to bootstrap the two
+  \ACode{\ACon{Kit}} instances, the \ACode{\ACon{ComposeTraversal}} allows
+  us to bootstrap the four \ACode{\ACon{ComposeKit}} instances.
+
+  % \ACode{\ACon{ComposeKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
+  % \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
+  % \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-fusion}} are analogous to
+  % \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{\_⋯\_}}.
 
@@
   \label{sec:composition:wkkit}
+
+  Before we define \ACode{\ACon{ComposeKit}} and
+  \ACode{\ACon{ComposeTraversal}}, we need one more ingredient.
 
@@
   a shifting renaming.
+  Instead of explaining \ACode{\AField{wk}} through its action on variables,
+  this lemma explains \ACode{\AField{wk}} through its action on terms.
 
@@
   \LibCode\KWkKitInstances
+
+  % Note that we have exactly two instances for \ACode{\ACon{Kit}}s
+  % and both of them have a \ACode{\ACon{WkKit}} instance. Consequently,
+  % whenever we abstract over a \ACode{\ACon{Kit}}, we can also safely assume
+  % that it has a \ACode{\ACon{WkKit}} instance, without restricting the
+  % lemmas usage.
 

### 2023-09-19 paper/paper.tex
Update.

@@
 
+  \subsection{Extension Kits}
+  \label{sec:maps:wkkit}
+
+  % Before we define \ACode{\ACon{ComposeKit}} and
+  % \ACode{\ACon{ComposeTraversal}}, we need one more ingredient.
+
+  As we defined the \ACode{\ACon{Kit}} structure before the
+  \ACode{\ACon{Traversal}} structure, the fields of
+  \ACode{\ACon{Kit}} could not use the \ACode{\ACon{\_⋯\_}} function in
+  their types.
+
+  This prevented us to define a useful lemma, which also needs to
+  be instanciated separately for variables and terms: the counterpiece
+  to \ACode{\AField{wk-id/`}}, which we call \ACode{\AField{wk-`/id}}.
+
+  Recall that \ACode{\AField{wk-id/`}} explains the
+  \ACode{\AField{wk}} function of an arbitrary \ACode{\ACon{Kit}} by how it acts on
+  \emph{variables}:
+  \ExampleCode{\KWkId}
+
+  Similarly, the \ACode{\AField{wk-`/id}} lemma
+  explains the \ACode{\AField{wk}} function of an
+  arbitrary \ACode{\ACon{Kit}} by how it acts on \emph{terms}:
+  \ExampleCode{\KIdWk}
+
+  This lemma is useful for proving extensional equalities between maps
+  involving weakening, where \ACode{\AField{`/id-injective}} allows to
+  add \ACode{\AField{`/id}} on both sides of the equation, such that
+  \AField{wk-`/id} can be used to make further progress.
+
+  To prove \AField{wk-`/id}, we define a structure
+  \ACode{\ACon{WkKit}}, which extends a regular \ACode{\ACon{Kit}} with
+  the \AField{wk-`/id} lemma:
+  \LibCode\KWkKit
+
+  We then create exactly two instances of the \ACode{\ACon{WkKit}} for
+  the variable and term \ACode{\ACon{Kit}}, respectively. 
+  \LibCode\KWkKitInstances
+
+  As the variable and term \ACode{\ACon{Kit}}s are the only two \ACode{\ACon{Kit}}s,
+  and both have \ACode{\ACon{WkKit}} instances, it is always safe to assume that
+  a \ACode{\ACon{Kit}} also supports the \ACode{\ACon{WkKit}} extension.
+
+  % Note that we have exactly two instances for \ACode{\ACon{Kit}}s
+  % and both of them have a \ACode{\ACon{WkKit}} instance. Consequently,
+  % whenever we abstract over a \ACode{\ACon{Kit}}, we can also safely assume
+  % that it has a \ACode{\ACon{WkKit}} instance, without restricting the
+  % lemmas usage.
+
+  % Both the structure and the instances are part of the library and do
+  % not require application code.
+
   \section{Map Composition}
@@
 
-  \subsection{WkKits}
-  \label{sec:composition:wkkit}
-
-  Before we define \ACode{\ACon{ComposeKit}} and
-  \ACode{\ACon{ComposeTraversal}}, we need one more ingredient.
-
-  In the definition of \ACode{\ACon{Kit}} we required the axiom
-  \ACode{\AField{wk-id/`}}, which captured the interaction between
-  \ACode{\AField{wk}} and \ACode{\AField{id/`}}:
-  \ExampleCode{\KWkId}
-  Injecting a variable with \ACode{\AField{id/`}} and then shifting it,
-  is the same as injecting a shifted variable.
-
-  Now that we have defined the \ACode{\ACon{Traversal}} structure, we
-  can formulate a similar interaction between
-  \ACode{\AField{wk}} and \ACode{\AField{`/id}}:
-  \ExampleCode{\KIdWk}
-  Shifting a variable/term \ACode{x/t} with \ACode{\AField{wk}} and then injecting
-  it to a term, is the same as injecting \ACode{x/t} to a term and then applying
-  a shifting renaming.
-  Instead of explaining \ACode{\AField{wk}} through its action on variables,
-  this lemma explains \ACode{\AField{wk}} through its action on terms.
-
-  As this lemma depends on map application \ACode{\ADef{\_⋯\_}} being
-  already defined, we cannot add it to the original \ACode{\ACon{Kit}}
-  structure, but need to define a new structure for it, which we call
-  a \ACode{\ACon{WkKit}}:
-  \LibCode\KWkKit
-
-  We then create exactly two instances of the \ACode{\ACon{WkKit}} for
-  the variable and term kit, respectively. 
-  \LibCode\KWkKitInstances
-
-  % Note that we have exactly two instances for \ACode{\ACon{Kit}}s
-  % and both of them have a \ACode{\ACon{WkKit}} instance. Consequently,
-  % whenever we abstract over a \ACode{\ACon{Kit}}, we can also safely assume
-  % that it has a \ACode{\ACon{WkKit}} instance, without restricting the
-  % lemmas usage.
-
-  Both the structure and the instances are part of the library and do
-  not require application code.
-
-  \subsection{Composition}
+  \subsection{An Abstraction for Composition}
   \label{sec:composition:ckit}

### 2023-09-19 paper/paper.tex
Update.

@@
 
-  In this section, we extend our framework with a single composition
-  operator for arbitrary maps.
-
-  The core property of composition is that applying two maps
-  \ACode{ϕ₁} and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same
-  as applying their composition \ACode{(ϕ₁ \AField{·ₖ} ϕ₂)} to \ACode{t}, i.e.
+  In this section, we extend our framework with an abstraction that
+  allows the composition of arbitrary maps.
+
+  The core property of composition, which we call
+  \ACode{\AField{⋯-fusion}}, states that applying two maps \ACode{ϕ₁}
+  and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same as
+  applying their composition \ACode{(ϕ₁ \AField{·ₖ} ϕ₂)} to \ACode{t},
+  i.e.
   $$
@@
   $$
-  We call this property \ACode{\AField{⋯-fusion}}.
   This property equips our framework with the power to
@@
     renamings.
-    In other words, 
+    In other words:
     if \ACode{ϕ₁} is a \ACode{K₁}-map and \ACode{ϕ₂} is a \ACode{K₂}-map,
@@
     and then apply \ACode{ϕ₂} to the result.
-    If \ACode{ϕ₁} is a renaming, this result is another variable, but
-    if \ACode{ϕ₁} is a substitution, this result is a term.
+    If \ACode{ϕ₁} is a renaming, then this result is another variable, but
+    if \ACode{ϕ₁} is a substitution, then this result is a term.
   \end{enumerate}
@@
   \LibCode\KComposeKit
+
+  Similar as for the \ACode{\ACon{Kit}}, we derive further definitions
+  from the the axioms of a \ACode{\ACon{ComposeKit}}:
+  \begin{itemize}
+  \item
+    We define map composition as motivated by the previous subsection:
     \LibCode\KComposition
+  \item
+    We show that the \ACode{\AField{\_\&/⋯\_}} operator behaves like
+    \ACode{\AField{\_\&\_}} when applied to variables:
     \LibCode\KComposeKitAp
     % \LibCode\KComposeKitApProof
+
+    % The \ACode{\AField{`/id}} on the left side is from \ACode{K₁⊔K₂}, whereas
+    % the \ACode{\AField{`/id}} on the right side is from \ACode{K₂}.
+  \item
+    We show that lifting \ACode{\AField{\_↑\_}} distributes over
+    composition:
     \LibCode\KDistLiftCompose
     % \LibCode\KDistLiftComposeProof
-  \LibCode\KComposeKitNotation
+  \end{itemize}
+  To make it easier to explicitly specify a \ACode{\ACon{ComposeKit}},
+  we define the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}
+  to stand for \ACode{ϕ₁ \ADef{·ₖ} ϕ₂} for some
+  \ACode{\ACon{ComposeKit}} \ACode{C}.
+  % \LibCode\KComposeKitNotation
+
+  We then define the \ACode{\ACon{ComposeTraversal}} structure,
+  which provides the \ACode{\ACon{⋯-fusion}} for the compositions
+  described by arbitrary \ACode{\ACon{ComposeKit}}s:
   \LibCode\KComposeTraversal
+
+  Similar as for the \ACode{\ACon{Traversal}}, we derive further
+  definitions from the the axiom of a \ACode{\ACon{ComposeTraversal}}:
+  \begin{itemize}
+  \item 
+    A map \ACode{ϕ} followed by a weakening is equivalent to a
+    weakening followed by \ACode{ϕ} that has been lifted over the weakened
+    variable:
     \LibCode\KCommLiftWeaken
     % \LibCode\KCommLiftWeakenProof
+    By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the same
+    statement for multiple map applications:
     \LibCode\KCommLiftWeakenTraverse
     % \LibCode\KCommLiftWeakenTraverseProof
+  \item 
+    A weakening followed by a singleton substitution cancel each other out:
+    \LibCode\KWeakenCancelsSingle
+    % \LibCode\KWeakenCancelsSingleProof
+    By applying \ACode{\AField{⋯-fusion}} and \ACode{\AField{⋯-id}} to
+    this lemma, we get the same statement for multiple map applications:
+    \LibCode\KWeakenCancelsSingleTraverse
+    % \LibCode\KWeakenCancelsSingleTraverseProof
+  \item 
+    A singleton substitution followed by some map \ACode{ϕ} is the same
+    as \ACode{ϕ} lifted over the variable of the singleton substitution
+    followed by the singleton substitution with \ACode{ϕ} applied to its term:
+    \LibCode\KDistLiftSingle
+    % \LibCode\KDistLiftSingleProof
+    By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the
+    same statement for multiple map applications:
+    \LibCode\KWeakenCancelsSingleTraverse
+    \LibCode\KDistLiftSingleTraverse
+    % \LibCode\KDistLiftSingleTraverseProof
+  \item 
+    Given the \ACode{\ADef{⋯-↑-wk}} lemma, we can define the four
+    \ACode{\ACon{ComposeKit}} instances:
     \LibCode\KComposeKitInstances
     \LibCode\KComposeKitInstancesConcrete
-  \LibCode\KWeakenCancelsSingle
-  % \LibCode\KWeakenCancelsSingleProof
-  \LibCode\KWeakenCancelsSingleTraverse
-  % \LibCode\KWeakenCancelsSingleTraverseProof
-  \LibCode\KDistLiftSingle
-  % \LibCode\KDistLiftSingleProof
-  \LibCode\KDistLiftSingleTraverse
-  % \LibCode\KDistLiftSingleTraverseProof
-
-  \subsection{System F}
+  \end{itemize}
+
+  Finally, we instantiate the \ACode{\ACon{ComposeTraversal}} for
+  System F to make the composition machinery available:
   \AppCode\FFusion

### 2023-09-19 paper/paper.tex
Update.

@@
     as \ACode{ϕ} lifted over the variable of the singleton substitution
-    followed by the singleton substitution with \ACode{ϕ} applied to its term:
+    followed by the singleton substitution with \ACode{ϕ} applied to
+    its content:
     \LibCode\KDistLiftSingle

### 2023-09-19 paper/paper.tex
Update.

@@
   A \ACode{Kit} is a structure that allows to abstract over whether
-  something is a term or a variable. The intention is to instanciate
+  something is a term or a variable. The intention is to instantiate
   this structure exactly twice (once for variables and once for terms),
@@
   instance.
-  For example the parameter type \ACode{\_∋/⊢\_} will be instanciated
+  For example the parameter type \ACode{\_∋/⊢\_} will be instantiated
   to \ACode{\ACon{\_∋\_}} for the variable kit, and to \ACode{\AField{\_⊢\_}} for the
@@
     \ACode{\AField{id/`}} converts a variable \ACode{S \ACon{∋} s} into a
-    \ACode{S \AField{∋/⊢} s}. For the variable kit, \ACode{S ∋/⊢ s} is instanciated to \ACode{S \ACon{∋} s}, so this
+    \ACode{S \AField{∋/⊢} s}. For the variable kit, \ACode{S ∋/⊢ s} is instantiated to \ACode{S \ACon{∋} s}, so this
     operation is the identity. For the term kit, \ACode{S ∋/⊢ s} is
-    instanciated to \ACode{S \AField{⊢} s}, so this operation is the
+    instantiated to \ACode{S \AField{⊢} s}, so this operation is the
     variable constructor \ACode{\AField{`\_}}.
@@
 
-  Note that all definitions so far are library code. To make them
-  available for our System F example, all we need to do is
-  to create an instance of \ACode{\ACon{Traversal}}.
-
-  We start with the operation of applying a map to a term:
+  \subsection{Instantiation for System F}
+  In this subsection, we show how to instantiate the
+  \ACode{\ACon{Traversal}} abstraction for System F.
+  In practice, this is done by our reflection algorithm automatically,
+  but it can be instructive to see what happens under the hood.
+
+  First, we define the operation of applying a map to a term:
   \AppCode\FTraversalOp
@@
 
-  Finally, we instanciate the \ACode{\ACon{Traversal}} structure:
+  Finally, we instantiate the \ACode{\ACon{Traversal}} structure:
   % \LibCode\FTraversalIdProofRest
@@
   This prevented us to define a useful lemma, which also needs to
-  be instanciated separately for variables and terms: the counterpiece
+  be instantiated separately for variables and terms: the counterpiece
   to \ACode{\AField{wk-id/`}}, which we call \ACode{\AField{wk-`/id}}.
@@
   four compositions, whereas the \ACode{\ACon{ComposeTraversal}}
-  structure needs to be instanciated by the application with
+  structure needs to be instantiated by the application with
   lemmas parameterized over arbitrary \ACode{\ACon{ComposeKit}}s.
@@
     same statement for multiple map applications:
-    \LibCode\KWeakenCancelsSingleTraverse
     \LibCode\KDistLiftSingleTraverse
@@
 
-  Finally, we instantiate the \ACode{\ACon{ComposeTraversal}} for
-  System F to make the composition machinery available:
+  \subsection{Instantiation for System F}
+  In this subsection, we show how to instantiate the
+  \ACode{\ACon{ComposeTraversal}} abstraction for System F.
+  In practice, this is done by our reflection algorithm automatically,
+  but it can be instructive to see what happens under the hood.
+
   \AppCode\FFusion
   \AppCode\FFusionProofInteresting
+  The interesting cases are:
+  \begin{itemize}
+  \item 
+  \end{itemize}
+
   % \AppCode\FFusionProofRest
@@
   \LibCode\KTypeSorts
-  For System F the instanciation is
+  For System F the instantiation is
   \AppCode\FTypes
@@
   \LibCode\KTyping
-  and instanciate it for System F:
+  and instantiate it for System F:
   \AppCode\FTypingInst
@@
 
-  Given a \ACode{\ACon{TypingTraversal}}, we can easily instanciate the
+  Given a \ACode{\ACon{TypingTraversal}}, we can easily instantiate the
   \ACode{\ACon{TypingKits}} for variables and terms:
@@
 
-  We are then ready to create a \ACode{\ACon{TypingTraversal}}
-  instance for System F:
+  \subsection{Instantiation for System F}
+  In this subsection, we show how to instantiate the
+  \ACode{\ACon{TypingTraversal}} abstraction for System F.
+  This is the only structure that is \emph{not} instantiated
+  automatically via reflection, as typing relations can be arbitrary
+  complex in general.
+
   \AppCode\FPreserve
+  The interesting parts of the proof are:
+  \begin{itemize}
+  \item 
+  \end{itemize}
+
   \AppCode\FTypingTraversal
 
-  As the \ACode{\ACon{Traversal}} and  \ACode{\ACon{ComposeTraversal}} instances
-  can be generally derived via reflection,
-  the \ACode{\AField{\_⊢⋯\_}} lemma is in fact the only lemma that
-  has to been proved manually for subject reduction.
-  From this single lemma follows type preservation for all operations
-  of both renamings and substitutions for all sorts of variables
-  simultaneously.
+  % As the \ACode{\ACon{Traversal}} and  \ACode{\ACon{ComposeTraversal}} instances
+  % can be generally derived via reflection,
+  % the \ACode{\AField{\_⊢⋯\_}} lemma is in fact the only lemma that
+  % has to been proved manually for subject reduction.
+  % From this single lemma follows type preservation for all operations
+  % of both renamings and substitutions for all sorts of variables
+  % simultaneously.
 
@@
   We used Agda's reflection mechanism\cite{TODO} to derive
-  instanciations related to all structures for untyped substitition,
+  instantiations related to all structures for untyped substitition,
   i.e. \ACode{\ACon{Traversal}} and \ACode{\ACon{ComposeTraversal}}.
@@
   reflection algorithm, we modeled a class of object languages using a
-  generic programming technique and instanciated the structures
+  generic programming technique and instantiated the structures
   generically. The class contains all object languages which:
@@
     For lambda calculus with dependent function types, the framework
-    just works out of the box for both deriving and instanciating the
+    just works out of the box for both deriving and instantiating the
     \ACode{\ACon{TypingTraversal}}. Additionally, having the
@@
   In later work\cite{DBLP:journals/pacmpl/AllaisA0MM18}, they use
-  generic programming to instanciate this abstraction for a class of
+  generic programming to instantiate this abstraction for a class of
   object languages comparable to ours.

### 2023-09-19 paper/paper.tex
Update.

@@
     lattice for \ACode{\{ \ADef{Kᵣ} , \ADef{Kₛ} \}}
-    generated by \ACode{\ADef{Kᵣ} $\leq$ \ADef{Kₛ}}.
+    generated by \ACode{\ADef{Kᵣ} < \ADef{Kₛ}}.
   \item

### 2023-09-19 paper/paper.tex
Update.

@@
 
-  Benton et al.\cite{DBLP:journals/jar/BentonHKM12} addresses this
+  Previous work on kits\cite{DBLP:journals/jar/BentonHKM12} addresses this
   issue by duplicating the definitions and using tactics to reduce
@@
   \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}. This allows us to
-  define a general composition in the spirit of:
+  define a general composition, which informally can be described as:
   \ExampleCode\KComposition
@@
   \ACode{\AField{⋯-fusion}} lemma for general maps.
-
   Just like a \ACode{\ACon{Traversal}} allowed us to bootstrap the two
@@
   \label{sec:composition:ckit}
-
-  Using the \ACode{\ACon{WkKit}}, we then define
-  the \ACode{\ACon{ComposeKit}} and \ACode{\ACon{ComposeTraversal}}
-  structures, which serve analogous purposes to \ACode{\ACon{Kit}}
-  and \ACode{\ACon{Traversal}} from before:
-  the library provides a \ACode{\ACon{ComposeKit}} for each of the
-  four compositions, whereas the \ACode{\ACon{ComposeTraversal}}
-  structure needs to be instantiated by the application with
-  lemmas parameterized over arbitrary \ACode{\ACon{ComposeKit}}s.
-  This gives the user a composition operation
-  \ACode{\AField{\_·ₖ\_}} between arbitrary maps and corresponding
-  lemmas, instead of multiple definitions and lemmas about concrete
-  maps, enabling us to reason on a more abstract level and to avoid
-  further boilerplate.
 
@@
   \item
+    variables, where we need to use the \ACode{\AField{\&/⋯-⋯}} lemma
+    provided by the \ACode{\ACon{ComposeKit}}.
+  \item
+    bindings, where the traversal operation \ACode{\AField{\_⋯\_}}
+    needs to lift the map via \ACode{\AField{\_↑\_}}, requiring us to
+    distribute the lifting over the composition using
+    \ACode{\AField{dist-↑-·}}.
   \end{itemize}

### 2023-09-20 paper/paper.tex
Update.

@@
     all variables of all sorts simultaneously, and
-    abstractions that unify the definitions, compositions, and typings
-    of multi-sorted renamings and substitutions.
+    abstractions that unify the definitions, compositions, typings,
+    and type preservation lemmas of multi-sorted renamings and substitutions.
     Case studies have been conducted to prove subject reduction for
@@
   Section~\ref{sec:typing} shows how to define a multi-sorted typing relation,
-  and an abstraction to unify type preservattion lemmas for renamings and substitutions.
+  and an abstraction to unify type preservation lemmas for renamings and substitutions.
   Section~\ref{sec:subjectreduction} introduces a small-step
@@
     a formalized specification of a large class of object languages
-    for which substitution and lemmas can be derived generically.
+    for which untyped substitution and lemmas can be derived generically.
   \item
@@
   \item
-    three case studies using our approach to prove subject reduction
+    three case studies in using our approach to prove subject reduction
     for System~F with subtyping, a dependently-typed lambda calculus, and
@@
     constructor are injective.
-  \item \ACode{\AField{wk-id/`}} characterizes the behaviour of the \ACode{\AField{wk}} function.
+  \item \ACode{\AField{wk-id/`}} characterizes the behaviour of the
+    \ACode{\AField{wk}} function by how it acts on variables.
     Injecting a variable and then shifting it, is the same as injecting a shifted variable.
@@
   \ACode{\ACon{Traversal}} structure, the fields of
-  \ACode{\ACon{Kit}} could not use the \ACode{\ACon{\_⋯\_}} function in
+  \ACode{\ACon{Kit}} could not use the \ACode{\AField{\_⋯\_}} function in
   their types.
@@
   This property equips our framework with the power to
-  reason about the application of multiple maps in terms of
+  reason about the application of multiple maps by reasoning about
   the application of a single map.
@@
   dependencies.
-  This has the benefit, that we can prove lemmas about
-  the composition of arbitrary maps, which enable us later to unify type
-  preservation of all substitution and renaming operations into a single
-  lemma.
-  % TODO: clarify
+  While composition can be dervied, this has the benefit, that
+  we've gained a new level of generality, which enables us in the next
+  sections to unify type preservation of all substitution and renaming
+  operations into one lemma.
 
@@
 
+  The \ACode{\ACon{ComposeKit}} abstraction provides the
+  \ACode{\AField{\_\&/⋯\_}} operation, necessary for defining
+  composition, and two lemmas about its behavior:
   \LibCode\KComposeKit
+  The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
+  (according to \ACode{K₁}) and a renaming or substitution
+  (according to \ACode{K₂}) and performs the map traversal returning
+  a variable or term (according to \ACode{K₁⊔K₂}).
+  We later define four instances for \ACode{\ACon{ComposeKit}}, where
+  \ACode{K₁} and \ACode{K₂} can each be either the variable or term kit.
+  The third parameter \ACode{K₁⊔K₂} can be seen as a functional
+  dependency and is completely determined by the choice of \ACode{K₁}
+  and \ACode{K₂}.
+
+  The \ACode{\AField{\&/⋯-⋯}} lemma describes the behavior of
+  \ACode{\AField{\_\&/⋯\_}} in terms of \ACode{\AField{\_⋯\_}}, allowing
+  subsequent lemmas to make use of the lemmas we previously proved
+  about \ACode{\AField{\_⋯\_}}.
+
+  The \ACode{\AField{\&/⋯-wk-↑}} lemma states that applying a map and then
+  weakening is the same as weakening first and then lifting the map over
+  the variable introduced by the weakening.
 
   Similar as for the \ACode{\ACon{Kit}}, we derive further definitions
-  from the the axioms of a \ACode{\ACon{ComposeKit}}:
+  from the axioms of a \ACode{\ACon{ComposeKit}}:
   \begin{itemize}
@@
   We then define the \ACode{\ACon{ComposeTraversal}} structure,
-  which provides the \ACode{\ACon{⋯-fusion}} for the compositions
+  which provides the \ACode{\AField{⋯-fusion}} lemma for the compositions
   described by arbitrary \ACode{\ACon{ComposeKit}}s:
@@
   Similar as for the \ACode{\ACon{Traversal}}, we derive further
-  definitions from the the axiom of a \ACode{\ACon{ComposeTraversal}}:
+  definitions from the axiom of a \ACode{\ACon{ComposeTraversal}}:
   \begin{itemize}
@@
 
-  In the context of multi-sorted syntax, the notation of a type can be
+  In the context of multi-sorted syntax, the notion of a type can be
   described as a mapping between sorts.
@@
   instantiations related to all structures for untyped substitition,
-  i.e. \ACode{\ACon{Traversal}} and \ACode{\ACon{ComposeTraversal}}.
+  i.e. \ACode{\ACon{Syntax}}, \ACode{\ACon{Traversal}} and
+  \ACode{\ACon{ComposeTraversal}}.
 
   To gain insight into the class of object languages supported by our
-  reflection algorithm, we modeled a class of object languages using a
-  generic programming technique and instantiated the structures
-  generically. The class contains all object languages which:
+  reflection algorithm, we have formalized a class of object languages
+  using a generic programming technique and instantiated the structures
+  generically. Informally, this class contains all object languages which
   \begin{enumerate}
-  \item Have a variable constructor of type TODO;
-  \item Uses subterms only directly (e.g. not in lists);
-  \item Subterms only extend the scope-context.
+  \item have a variable constructor of type TODO;
+  \item use subterms only directly (e.g. not in lists);
+  \item only extend the scope-context of subterms.
   \end{enumerate}
@@
   example the encoding of linear types that requires context
-  splitting. This has already be done for intrinsical kits\cite{TODO}.
+  splitting.
 

### 2023-09-20 paper/paper.tex
Update.

@@
   \ExampleCode\KComposition
-  We then define a \ACode{\ACon{ComposeTraversal}}, which consists of the
-  \ACode{\AField{⋯-fusion}} lemma for general maps.
-  Just like a \ACode{\ACon{Traversal}} allowed us to bootstrap the two
-  \ACode{\ACon{Kit}} instances, the \ACode{\ACon{ComposeTraversal}} allows
-  us to bootstrap the four \ACode{\ACon{ComposeKit}} instances.
 
@@
 
-  The \ACode{\ACon{ComposeKit}} abstraction provides the
-  \ACode{\AField{\_\&/⋯\_}} operation, necessary for defining
-  composition, and two lemmas about its behavior:
+  In this subsection, we define the \ACode{\ACon{ComposeKit}}
+  abstraction, which provides the \ACode{\AField{\_\&/⋯\_}} operation
+  that is necessary for defining composition, and the
+  \ACode{\ACon{ComposeTraversal}}, which provides the
+  \ACode{\AField{⋯-fusion}} lemma for general maps.
+
+  Similarly as it was the case for \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}},
+  the \ACode{\ACon{ComposeTraversal}} needs to be instantiated for
+  abstract \ACode{\ACon{ComposeKit}}s, which then provides the actual
+  four \ACode{\ACon{ComposeKit}} instances. This allows us to break
+  the dependencies between the different \ACode{\AField{⋯-fusion}} lemmas,
+  similar to how \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}} allowed
+  us to break the dependency between traversals with renamings and substitutions.
+
+  For this purpose, the \ACode{\ACon{ComposeKit}} abstraction, needs
+  to satisfy two requirements:
+  it needs to provide us with enough machinery to define composition, but also
+  it needs to provide us with enough axioms that we are able to instanciate
+  the \ACode{\ACon{ComposeTraversal}}. It is defined as followed:
   \LibCode\KComposeKit
@@
   a variable or term (according to \ACode{K₁⊔K₂}).
-  We later define four instances for \ACode{\ACon{ComposeKit}}, where
-  \ACode{K₁} and \ACode{K₂} can each be either the variable or term kit.
   The third parameter \ACode{K₁⊔K₂} can be seen as a functional
@@
   \item 
-    Given the \ACode{\ADef{⋯-↑-wk}} lemma, we can define the four
-    \ACode{\ACon{ComposeKit}} instances:
+    We define \ACode{\ACon{ComposeKit}} instances for each of the four
+    combinations between renamings and substitutions:
     \LibCode\KComposeKitInstances
+    Interestingly, the \ACode{\ADef{Cₛ}} instance has another
+    \ACode{\ADef{ComposeKit}} parameter.
+    This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
+    which yields the \ACode{\ADef{ComposeKit}} between substitutions
+    and renamings, and then use this resulting \ACode{\ADef{ComposeKit}}
+    as an argument for the case with two substitutions.
+    As shown below, Agda's instance resolution is up for the task:
+
+    % To get the \ACode{\ADef{ComposeKit}} between substitutions and
+    % renamings, the argument \ACode{\ADef{Cᵣ}} is used.
+    % The resulting \ACode{\ADef{ComposeKit}} is then again used as an argument
+    % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{ComposeKit}}
+    % between two substitutions. As shown below, Agda's instance
+    % resolution is up for the task:
     \LibCode\KComposeKitInstancesConcrete
+
   \end{itemize}
@@
   In practice, this is done by our reflection algorithm automatically,
-  but it can be instructive to see what happens under the hood.
+  but it can be instructive to see, as it motivates the axioms of the
+  \ACode{\ACon{ComposeKit}}.
 

### 2023-09-20 paper/paper.tex
Update.

@@
   \ACode{\ACon{𝕜}}.
-  To teach our framework about types, we have to instantiate the following structure:
+  The following structure is used to teach our framework about types:
   \LibCode\KTypeSorts
@@
   \AppCode\FTypes
-  For a nicer interface, we define a type analogously to a term:
+  There are two things to discuss:
+  \begin{enumerate}
+  \item
+    The \ACode{\AField{↑ᵗ}} function maps a sort of arbitrary sort type,
+    to a sort of a potentially different sort type, which is expressed
+    by the use of an existential.
+    For System F we require this generality, as the sort
+    \ACode{\ACon{𝕥}} can have variables, whereas its corresponding type
+    sort \ACode{\ACon{𝕜}} cannot.
+  \item
+    Some sorts, like \ACode{\ACon{𝕜}}, do not have corresponding type
+    sorts, but we still need to provide one, as \ACode{\AField{↑ᵗ}} is a
+    total function.
+    For such sorts, we can simply use arbitrary sort types, as the
+    formalization will never use them.
+  \end{enumerate}
+  To hide the existential, we define \ACode{S \ACon{∶⊢} s}, which represents
+  the type for a term \ACode{S \ACon{⊢} s}.
   \LibCode\KTypes
@@
   \subsection{Typing Context}
-  A typing context is then a function which maps variables to types.
-  As we want our framework to support dependent types, the types returned by
-  a typing context are allowed to use the domain of the 
-  To be able to deal with dependent types, the result types are only allowed
-  to refer to the variables that come before 
-  \LibCode\KContextHelper
+  Equipped with a notion of types, we define a typing context as a
+  function that maps variables to types of the same sort.
+
+  As we want our framework to support dependent types, we allow a type returned
+  by the typing context to use all variables bound previously in the typing context.
+  This is achieved by the \ACode{drop-∈} function, which cuts of all
+  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
+
+  % \LibCode\KContextHelper
   \LibCode\KContexts
+  To use a type returned from a typing context, it needs to be
+  weakened appropriately through the \ACode{\ADef{wk-telescope}} function:
   \LibCode\KContextLookup
-
+  We then define a variable typing as followed:
   \LibCode\KVariableTyping
@@
   \label{sec:typing}
-  Equipped with a notion of types and type context we are ready to define
+  Equipped with a notion of types and type context, we are ready to define
   the multi-sorted typing relation for System F, incorporating both
@@
   \item
-    the variable typing constructor covers both expression- and
+    the variable typing constructor, which covers both expression- and
     type-variables, analogously to the variable term constructor;
   \item
-    the typing of the lambda abstraction \ACode{\ACon{⊢λ}} weakens the
-    type of the codomain \ACode{t₂}.
-    This is necessary, because with multi-sorted syntax, types could
-    also depend on expressions, so the sub typing derivation accounts for
-    another variable which is not used by the type.
+    the typing \ACode{\ACon{⊢λ}} of the lambda abstraction, which
+    weakens the type of the codomain \ACode{t₂}.
+    This is necessary, because multi-sorted syntax allows types
+    to depend on expressions, so the typing derivation for \ACode{e}
+    has to account for a variable, which is not used by the type.
   \item
@@
   \LibCode\KTyping
-  and instantiate it for System F:
+  The instantiation for System F is straightforward:
   \AppCode\FTypingInst
 
-  Next we define \ACode{\ACon{TypingKit}}s, which are analog to
-  \ACode{\ACon{Kit}}s, but abstract about variable and term
-  \emph{typing}, instead of variables and terms, themselves.
+  \subsection{An Abstraction for Type Preservation}
+  By now, the reader probably knows what comes next: we build an
+  abstraction to unify type preservation for renamings and substitutions,
+  eliminating the dependencies by yet another type of kits.
+
+  We start with \ACode{\ACon{TypingKit}}s, which abstract over
+  variable and term \emph{typing}, and then define a
+  \ACode{\ACon{TypingTraversal}}, which provides substitution
+  preserves typing for all \ACode{\ACon{TypingKit}}s.
   \LibCode\KTypingKit
+  The operations are completely analog to those of a \ACode{\ACon{Kit}}, but talk
+  about typing instead of terms.
   Building on the fields of a \ACode{\ACon{TypingKit}}, we define further operations:
@@
     \LibCode\KMapTyping
+    \ACode{Γ₂ \ADef{∋*/⊢*} ϕ \ADef{∶} Γ₁} states that \ACode{ϕ} is a map that takes
+    terms in \ACode{Γ₁} to terms in \ACode{Γ₂}.
   \item 
@@
   the term \ACode{⊢e \AField{⊢⋯} ⊢ϕ} is a typing for \ACode{e \AField{⋯} ϕ}.
+  The large amount of kit-parameters does not impose any restriction,
+  as both our \ACode{\ACon{Kit}}s support the \ACode{\ACon{WkKit}}
+  extension and can be composed arbitrarily. They only make the previously
+  built machinery available to the proof of \ACode{\AField{\_⊢⋯\_}}.
 
@@
   \item
+    There is a strong similarity to the instantiation of
+    map traversal \ACode{\AField{\_⋯\_}}. Where \ACode{\AField{\_⋯\_}}
+    used \ACode{\AField{`/id}} or \ACode{\AField{\_↑\_}}, we now use their
+    preservation lemmas \ACode{\AField{⊢`/id}} or \ACode{\AField{\_∋↑/⊢↑\_}}.
+  \item
+    The lambda typing constructor \ACode{\ACon{⊢λ}} weakens the type \ACode{t₂} to
+    shield it from expression-substitution, requiring us to use the
+    \ACode{\AField{⋯-↑-wk}} lemma.
+  \item
+    The type application constructor \ACode{\ACon{⊢∙}} substitutes \ACode{t₁} into \ACode{t₂},
+    requiring us to use the \ACode{\AField{dist-↑-⦅⦆}} lemma.
   \end{itemize}
-
+  Finally, we instantiate the \ACode{\ACon{TypingTraversal}} structure:
   \AppCode\FTypingTraversal
@@
 
-  We define a standard small-step operational semantics for System F
-  with full reduction:
+  To finish our running example, we define a standard small-step
+  operational semantics for System F with full reduction and prove subject reduction:
   \AppCode\FReduction
-
-  Proving subject reduction is now straightforward:
+  \vspace{0mm}
   \AppCode\FSubjectReduction
   \AppCode\FSubjectReductionProofInteresting
-  We only listed the cases that do not follow directly from the
-  induction hypothesis:
+  We only listed the cases for $\beta$-reduction, as the other cases
+  follow directly from the induction hypothesis:
   \begin{itemize}
@@
     is a direct consequence of substitution preserves typing.
-    \ACode{⊢e₂} is the typing for \ACode{e₂}.
-    \ACode{⊢⦅ ⊢e₂ ⦆} is the substitution-typing for the singleton substitution 
-    \ACode{\AField{⦅ e₂ ⦆}}.
-    \ACode{\AField{⊢e₁ ⊢⋯ ⊢⦅ ⊢e₂ ⦆}} states that applying the singleton substitution
-    to \ACode{e₁} preserves typing.
   \item The beta-reduction rule for value-application is almost the same,
-    but requires us to addionaly apply an equality to the type index.
+    but requires us to additionally show that the type is not affected by the
+    expression substitution. This is a direct consequence of the weakening
+    used in \ACode{\ACon{⊢λ}}.
   \end{itemize}
@@
   \label{sec:reflection}
-  We used Agda's reflection mechanism\cite{TODO} to derive
+  We used Agda's reflection mechanism to derive
   instantiations related to all structures for untyped substitition,
@@
   using a generic programming technique and instantiated the structures
-  generically. Informally, this class contains all object languages which
+  generically. Informally, this class contains all object languages described
+  via multi-sorted syntax, which
   \begin{enumerate}
-  \item have a variable constructor of type TODO;
+  \item have a variable constructor of type
+    \ACode{S \ACon{∋} s \AKw{→} S \ACon{⊢} s};
   \item use subterms only directly (e.g. not in lists);
-  \item only extend the scope-context of subterms.
+  \item only extend the scope-context of subterms, but never modify it.
   \end{enumerate}
@@
 
-  We were able to describe many languages in this shape, including
-  session types and type state.
-
-  We believe restriction 2 is purely technically and can be lifted
+  We believe restriction 2 is purely technical and can be lifted
   with a more sophisticated reflection algorithm, as parameterized
@@
   \label{sec:casestudies}
+  Using our full implementation of the framework, we proved subject
+  reduction for the following object languages:
   \begin{itemize}
@@
     just works out of the box for both deriving and instantiating the
-    \ACode{\ACon{TypingTraversal}}. Additionally, having the
-    \ACode{\ACon{ComposeKit}} abstraction turned out to be useful
-    to prove confluence, allowing to unify certain lemmas, which
-    otherwise need to be proved for both renamings and substitutions,
-    e.g. that for two terms \ACode{t} and \ACode{t'}
-    with \ACode{t \ACon{↪} t'} it follows that \ACode{(t \AField{⋯} ϕ)
-    \ACon{↪} (t' \AField{⋯} ϕ)} for any map \ACode{ϕ}.
+    \ACode{\ACon{TypingTraversal}}.
+    Furthermore, the \ACode{\ACon{ComposeKit}} abstraction turned out
+    to be useful to prove confluence, allowing to unify certain
+    lemmas, which otherwise need to be proved for both renamings and
+    substitutions.
+    % e.g. that for two terms \ACode{t} and \ACode{t'}
+    % with \ACode{t \ACon{↪} t'} it follows that \ACode{(t \AField{⋯} ϕ)
+    % \ACon{↪} (t' \AField{⋯} ϕ)} for any map \ACode{ϕ}.
   \item
@@
     have subtyping bounds that need to be respected.
-    Instead of defining typing contexts and type preservation lemmas independent
-    of our framework, we found a way that still allows us to benefit
-    from the typing machinery.
-    When binding a type variable with a subtyping constraint \ACode{α <: t},
-    we first bind the type variable as \ACode{α : ★}, and then bind the constraint
-    with a dummy name \ACode{\_ ∶ (α <: t)}. This description is
+    While it would be possible to use our framework only to derive
+    untyped substitution and define typing contexts and type preservation
+    lemmas by hand, we found an encoding, which still allows us to use the
+    the \ACode{\ACon{TypingKit}} abstraction.
+    Instead of binding a type variable with a subtyping constraint
+    \ACode{α <: t}, we first bind the type variable as \ACode{α : ★}, and
+    then bind the constraint as \ACode{c ∶ (α <: t)}. This description is
     similar to first-class constraints, but restricted enough to be
-    isomorphic to the original formalization.
-    With that encoding substitution preserves typing is generally true again, enabling
-    the use of the framework's typing contexts and \ACode{\ACon{TypingKit}} machinery.
-    Replacing a type variable \ACode{α <: t} with a type \ACode{t'}
-    which is not a subtype of \ACode{t'}, the resulting term is still
-    well-typed, but in a context with an insatisfiable constraint.
-  \item
-    Object language with pattern matching pose an interesting problem.
-    A clause \ACode{p ⇒ e} in a match-expression consists of a pattern
-    \ACode{p} and a continuation expression \ACode{e}.
-    The pattern \ACode{p} can bind arbitrary many variables in subpatterns,
-    which are then available as free variables in \ACode{e}.
-    % The challenge is to track the sorts of variables bound by a pattern,
-    % to make them available to \ACode{e} by extending \ACode{e}'s sort
-    % context correspondingly.
-    To track the sorts of variables bound by a pattern, one can define
-    the sort of patterns to contain the sorts of the variables that it binds:
-    % 𝕡 : List (Sort Var) → Sort NoVar
-    The syntax for variable patterns and clauses can then be specified as:
-    % varpat : S ⊢ 𝕡 [ 𝕖 ]  -- pattern binding an expression variable
-    % pairpat : S ⊢ 𝕡 S₁ → S ⊢ 𝕡 S₂ → S ⊢ 𝕡 (S₁ ++ S₂)
-    % clause : S ⊢ 𝕡 S' → (S' ++ S) ⊢ 𝕖 → S ⊢ 𝕔
+    isomorphic to the original formalization, as constraint variables
+    cannot be accesses by the user.
+    With that encoding substitution preserves typing is generally true
+    again, enabling the use of the framework's typing contexts and
+    \ACode{\ACon{TypingKit}} machinery.
+    When replacing a type variable \ACode{α <: t} with a type \ACode{t'},
+    which is not a subtype of \ACode{t'}, then the resulting term is still
+    well-typed, but in a context with an unsatisfiable constraint.
+  \item
+    Object languages with pattern matching can be encoded by adding
+    the sorts of the variables bound by a pattern to the pattern sort
+    \ACode{\ACon{𝕡}} itself:
+    \begin{quote}
+    \ACode{\ACon{𝕡} \AKw{:} \ACon{List (Sort Var)} \AKw{→} \ACon{Sort NoVar}} 
+    \end{quote}
+    This allows the clause of a \ACode{\ACon{match}} expression to
+    read the bound variables from the pattern sort and make them
+    available as free variables in its body expression.
   \end{itemize}

### 2023-09-20 paper/paper.tex
Update.

@@
   extrinsic typing, extrinsic scoping, and de Bruijn indices.
-  The framework is implemented in Coq's tactic language LTAC and comes
+  The framework is implemented in Coq's tactic language Ltac and comes
   with a decision procedure for all assumption-free, equational

### 2023-09-20 paper/paper.tex
Update.

@@
   relations, including those of our case studies.
-
-  % TODO: avoiding interaction lemmas
 
@@
   Section~\ref{sec:related} discusses related work.
-  Section~\ref{sec:future} discusses future work.
-  Section~\ref{sec:conclusion} concludes.
+  Section~\ref{sec:future} concludes with future work.
 
@@
   \AppCode\FTraversal
-
-  % TODO note that its derived
 
@@
 
-  Frameworks for programming language metatheory can be broadly
-  categorized by representation of variable bindings and
-  representation of typing and scoping.
-
-  Variable bindings have been represented as de Bruijn indices, de
-  Bruijn levels, names, locally nameless, locally named, shifted names,
-  nominal techniques, (parameterized) higher order abstract syntax.
-
-  Typing and scoping can be both extrinsic, both intrinsic, or
-  extrinsically typed and intrinsically scoped.
+  % Frameworks for programming language metatheory can be broadly
+  % categorized by representation of variable bindings and
+  % representation of typing and scoping.
+
+  % Variable bindings have been represented as de Bruijn indices, de
+  % Bruijn levels, names, locally nameless, locally named, shifted names,
+  % nominal techniques, (parameterized) higher order abstract syntax.
+
+  % Typing and scoping can be both extrinsic, both intrinsic, or
+  % extrinsically typed and intrinsically scoped.
 
   As the amount of related work is rather large, we focus on work that
-  is more closely related to ours and refer to other papers for the
-  broader picture. In particular, we discuss abstractions for renamings
+  is closely related to ours and refer to other papers for the
+  broader
+  picture\cite{DBLP:conf/cpp/StarkSK19,DBLP:journals/pacmpl/AllaisA0MM18}. In
+  particular, we discuss abstractions for renamings
   and substitutions, comparable extrinsically typed approaches, but also
@@
   In all three cases kits are formulated for intrinsic typing
-  and scenarios with polymorphism are not considered.
+  and scenarios with polymorphism are not yet considered.
 
@@
   vectorized substitutions.
-  They also define and prove lemmas for all four compositions.
-  The generated syntax consists of multiple term types, indexed by
-  different amounts of free variables as shown in Section~\ref{sec:syntax}.
-
-  Both works do not provide machinery for type preservation and
-  do not unify renaming and substitution (although the latter is listed as
-  future work). We believe it should be possible to translate their
-  decision procedure from vectorized substitutions to the setting of
-  our multi-sorted substitutions.
-
-  Abstract binding trees
+  Compared to our work, the syntax they generate takes the form of
+  what we described as \emph{unsorted syntax} in
+  Section~\ref{sec:syntax}, i.e. different syntactic categories are
+  described by different types with different amount of indices for
+  variable counts. To eliminate the need for interaction lemmas, they
+  define the notion of \emph{vectorized substitution}, which combines
+  the individual substitutions by putting them in a vector.
+
+  Both works do not provide machinery for type preservation and do not
+  unify renaming and substitution (although the latter is listed as
+  future work). We believe their amazing work of creating a decision
+  procedure for vectorized substitutions should also translate to our
+  setting with multi-sorted substitutions.
+
+  % Abstract binding trees
 
@@
   They demonstrate how both renamings and substitutions can be
-  described as semantics and how the four composition lemmas follow from
-  their generic fusion lemma.
-  They do cover an abstraction to unify renamings and substitutions.
-  They do not show how to work with multiple variable sorts, how to
-  abstract over the compositions, and how to abstract over typing in
-  an extrinsically typed scenario.
-  However, we believe that their framework is general enough such that
-  our concepts can also be transferred to their framework.
-
-  This work is orthogonal but compatible to our approach.
-  We can instantiate the system using our syntax type system,
-  but then the semantics abstraction would work with terms
-  which are untyped with respect to the actual type system.
+  described as semantics, how the four composition lemmas follow from
+  their generic fusion lemma, and also provide an abstraction to unify
+  renamings and substitutions.
+  They show how to use their framework for both intrinsic and
+  extrinsic typing, but have not yet explored the realm of polymorphism.
+
+  This work is orthogonal, but compatible to our approach.
+  We can instantiate their system using the hidden intrinsic type
+  system of our multi-sorted syntax, but unfortunately then their
+  semantics abstraction would only work with terms, which are untyped
+  with respect to the actual extrinsic type system.
   We think it is worthwile to explore if their semantics abstraction
@@
   \label{sec:future}
+  We want to focus our future work in the following directions:
   \begin{itemize}
@@
   \end{itemize}
-
-  \section{Conclusion}
-  \label{sec:conclusion}
+  \vspace{10mm}
 

### 2023-12-14 paper/paper.tex
Update.

@@
   substitution machinery, as all definitions and lemmas need to be
-  made first for renamings and then again for substitutions.
+  written first for renamings and then again for substitutions.
 
@@
   most proof assistants forbid to avoid breaking logical consistency.
-  We conjecture that an encoding with universe codes is possible, as
-  it has been done for intrinsically typed languages with dependent
-  types\cite{TODO}, but those constructions come with additional
-  complexity and are less suitable to automation.
+  We conjecture that an encoding might be possible, as it has been
+  done for intrinsically typed languages with dependent
+  types\cite{DBLP:conf/popl/AltenkirchK16}, but those constructions come
+  with additional complexity and are less suitable to automation.
 
@@
       The actual implementation does \emph{not} use any postulates.
-      The workaround is explained in Appendix~\ref{sec:fun-ext}.
+      % The workaround is explained in Appendix~\ref{sec:fun-ext}.
     }
@@
   Similarly as it was the case for \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}},
-  the \ACode{\ACon{ComposeTraversal}} needs to be instantiated for
-  abstract \ACode{\ACon{ComposeKit}}s, which then provides the actual
-  four \ACode{\ACon{ComposeKit}} instances. This allows us to break
+  the lemma of the \ACode{\ACon{ComposeTraversal}} needs to be proved for
+  an abstract \ACode{\ACon{ComposeKit}} and in return defines the
+  four concrete \ACode{\ACon{ComposeKit}} instances for us. This allows us to break
   the dependencies between the different \ACode{\AField{⋯-fusion}} lemmas,
@@
   We then define the \ACode{\ACon{ComposeTraversal}} structure,
-  which provides the \ACode{\AField{⋯-fusion}} lemma for the compositions
-  described by arbitrary \ACode{\ACon{ComposeKit}}s:
+  which provides the \ACode{\AField{⋯-fusion}} lemma for arbitrary
+  \ACode{\ACon{ComposeKit}}s:
   \LibCode\KComposeTraversal
@@
   \item 
-    A singleton substitution followed by some map \ACode{ϕ} is the same
-    as \ACode{ϕ} lifted over the variable of the singleton substitution
-    followed by the singleton substitution with \ACode{ϕ} applied to
-    its content:
+    A singleton map can be swapped with any map \ACode{ϕ}:
     \LibCode\KDistLiftSingle
@@
   % \AppCode\FFusionProofRest
-  \AppCode\FComposeTraversal
+  % \AppCode\FComposeTraversal
 
@@
   \label{sec:typing}
-  Equipped with a notion of types and type context, we are ready to define
+  Now that we have a notion of types and type context, we are ready to define
   the multi-sorted typing relation for System F, incorporating both
@@
     have kind \ACode{\ACon{★}}. This is sufficient for System F as types
-    are automatically well-scoped due to intrinsic scoping.
+    are automatically well-typed due to intrinsic scoping.
   \end{itemize}
@@
   variable and term \emph{typing}, and then define a
-  \ACode{\ACon{TypingTraversal}}, which provides substitution
-  preserves typing for all \ACode{\ACon{TypingKit}}s.
+  \ACode{\ACon{TypingTraversal}}, which provides
+  substitution-preserves-typing for all \ACode{\ACon{TypingKit}}s.
   \LibCode\KTypingKit
-  The operations are completely analog to those of a \ACode{\ACon{Kit}}, but talk
+  The operations are completely analogous to those of a \ACode{\ACon{Kit}}, but talk
   about typing instead of terms.
@@
   \AppCode\FSubjectReductionProofInteresting
-  We only listed the cases for $\beta$-reduction, as the other cases
+  We only list the cases for $\beta$-reduction, as the other cases
   follow directly from the induction hypothesis:
   \begin{itemize}
-  \item The beta-reduction rule for type-application, \ACode{\ACon{β-Λ}},
-    is a direct consequence of substitution preserves typing.
-  \item The beta-reduction rule for value-application is almost the same,
+  \item The $\beta$-reduction rule for type-application, \ACode{\ACon{β-Λ}},
+    is a direct consequence of substitution-preserves-typing.
+  \item The $\beta$-reduction rule for value-application is almost the same,
     but requires us to additionally show that the type is not affected by the
-    expression substitution. This is a direct consequence of the weakening
-    used in \ACode{\ACon{⊢λ}}.
+    expression substitution.
+    % This is a direct consequence of the weakening used in \ACode{\ACon{⊢λ}}.
   \end{itemize}
@@
   The formal definition of the generic class can be found in
-  Appendix~\ref{sec:generic-class} and the actual implementation.
+  the supplementary material.
+  % Appendix~\ref{sec:generic-class} and the actual implementation.
 
@@
     When replacing a type variable \ACode{α <: t} with a type \ACode{t'},
-    which is not a subtype of \ACode{t'}, then the resulting term is still
+    which is not a subtype of \ACode{t}, then the resulting term is still
     well-typed, but in a context with an unsatisfiable constraint.
@@
   \item
-    Incorporating a decision procedure for equational
-    substitution-lemmas based on $\sigma$-calculus as in Autosubst
+    Developing a decision procedure for equational
+    substitution lemmas based on $\sigma$-calculus as in Autosubst
     \cite{DBLP:conf/itp/SchaferTS15, DBLP:conf/cpp/StarkSK19}.
@@
   \end{itemize}
-  \vspace{10mm}
 
@@
 
-  \section{Generic Class of Object Languages}
-  \label{sec:generic-class}
-
-  \section{System F with Reflection}
-  \label{sec:reflection-example}
-
-  \section{Map Equality Without Functional Extensionality}
-  \label{sec:fun-ext}
-
-  \section{Features of the Actual Implementation}
-  \label{sec:features}
+  % \section{Generic Class of Object Languages}
+  % \label{sec:generic-class}
+
+  % \section{System F with Reflection}
+  % \label{sec:reflection-example}
+
+  % \section{Map Equality Without Functional Extensionality}
+  % \label{sec:fun-ext}
+
+  % \section{Features of the Actual Implementation}
+  % \label{sec:features}
 

### 2024-03-07 paper/agdatex.py
Started to refactor paper for ITP 2024

@@ -34,6 +34,12 @@ ap.add_argument("-r", "--root", metavar="PATH",
 ap.add_argument("-i", "--index", metavar="PATH",
                 help="Write the list of generated macros to this file.")
 
+ap.add_argument("-v", "--verbose", action='store_true',
+                     help="Turn on verbosity")
+
+ap.add_argument("-n", "--noop", action='store_true',
+                    help="No action: dry run")
+
 ap.add_argument("sources", metavar="SRC_PATH", nargs="+",
                 help="Path to an annotated .agda-file.")
 
@@ -62,6 +68,8 @@ else:
 
 root = root.absolute()
 
+if args.verbose:
+    print (f"VERBOSE: root = {root}")
 
 # Check if sources are relative to project root
 
@@ -73,6 +81,8 @@ for p in args.sources:
         print(f"ERROR: Source path '{p}' is not relative to root '{root}'.")
         sys.exit(1)
 
+if args.verbose:
+    print (f"VERBOSE: src_paths = {src_paths}")
 
 # Create temporary directory
 
@@ -83,6 +93,11 @@ else:
     tmp_root_obj = tempfile.TemporaryDirectory(prefix="agdatex")
     tmp_root = Path(tmp_root_obj.name)
 
+if args.verbose:
+    print (f"VERBOSE: tmp_root = {tmp_root}")
+
+if args.noop:
+    sys.exit("[no harm done]")
 
 # Copy project root to temporary directory
 
@@ -132,6 +147,15 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
         tgt += "\\end{AgdaAlign}"
         tgt += "}\n"
         mode = "none"
+    def latexname(name: str) -> bool:
+        return name.isascii() and name.isalpha()
+    def asciify(name: str) -> str:
+        return ''.join( (c if latexname(c) else 'X') for c in name )
+    def checked_latex_ident(name: str, line_num: int) -> str:
+        if not latexname(name):
+            print(f"WARNING: Illegal characters in '{name}' in line {line_num} replaced by X")
+            name = asciify(name)
+        return name
     for line_num, line in enumerate(src.splitlines()):
         l = line.strip()
         if l.startswith("--!") or l.startswith("-- !"):
@@ -144,12 +168,13 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
             if is_inline:
                 l = l[1:].strip()
             if "{" in l:
[…]

### 2024-03-07 paper/paper.tex
Refactored paper for ITP 2024

@@
+% PAGE LIMIT: 16
+
 % \documentclass[sigplan, screen]{acmart}
@@
 % \documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[sigplan,10pt, anonymous]{acmart}
+% \documentclass[sigplan,10pt, anonymous]{acmart}
+\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
 
@@
 
-\bibliographystyle{ACM-Reference-Format}
+% \bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{plainurl}% the mandatory bibstyle
 
 \title{Abstractions for Multi-Sorted Substitutions}
-
-\author{Hannes Saffrich}
-\orcid{0000-0002-1825-0097}                 %% \orcid is optional
-\affiliation{
-  % \department{Department1}                %% \department is recommended
-  \institution{University of Freiburg}      %% \institution is required
-  \country{Germany}                         %% \country is recommended
-}
-\email{saffrich@informatik.uni-freiburg.de} %% \email is recommended
+\author{Hannes Saffrich}{University of Freiburg, Germany}{saffrich@informatik.uni-freiburg.de}{https://orcid.org/0009-0004-7014-754X}{}
+\authorrunning{H. Saffrich}
+\Copyright{CC-BY}
+\keywords{Agda, Metatheory, Framework}
+% \relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
+%\relatedversiondetails[linktext={opt.
+text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g.
+Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional
+% \acknowledgements{I want to thank \dots}%optional
+
+%\nolinenumbers %uncomment to disable line numbering
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{John Q. Open and Joan R. Access}
+\EventNoEds{2}
+\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
+\EventShortTitle{CVIT 2016}
+\EventAcronym{CVIT}
+\EventYear{2016}
+\EventDate{December 24--27, 2016}
+\EventLocation{Little Whinging, United Kingdom}
+\EventLogo{}
+\SeriesVolume{42}
+\ArticleNo{23}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \author{Hannes Saffrich}
+% \orcid{0000-0002-1825-0097}                 %% \orcid is optional
+% \affiliation{
+%   % \department{Department1}                %% \department is recommended
+%   \institution{University of Freiburg}      %% \institution is required
+%   \country{Germany}                         %% \country is recommended
+% }
+% \email{saffrich@informatik.uni-freiburg.de} %% \email is recommended
 
@@
 % \email{thiemann@informatik.uni-freiburg.de} %% \email is recommended
+
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10003752.10003790.10011740</concept_id>
+       <concept_desc>Theory of computation~Type theory</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10011007.10011006.10011039.10011040</concept_id>
+       <concept_desc>Software and its engineering~Syntax</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10003752.10003790.10002990</concept_id>
+       <concept_desc>Theory of computation~Logic and verification</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Theory of computation~Type theory}
+\ccsdesc[500]{Software and its engineering~Syntax}
+\ccsdesc[500]{Theory of computation~Logic and verification}
 
@@
 \begin{document}
+
+  \maketitle
 
@@
   \end{abstract}
-
-  \maketitle
 
@@
   and in the supplementary material.
-  The full framework is available online.\footnote{See supplementary material. In the final version here will be a link to github.}
+  The full framework is available online.\footnote{See supplementary material. In the final version, this text will be replaced by a link to github.}
 
@@
 
-  For example, the substitution, which replaces the term variable 0 with the
-  identity function and the type variable 1 with the bottom type, is defined
-  as followed:
-  \ExampleCode\FExampleSub
+  % For example, the substitution, which replaces the term variable 0 with the
+  % identity function and the type variable 1 with the bottom type, is defined
+  % as followed:
+  % \ExampleCode\FExampleSub
 
@@
 
-  \subsection{The Need for Renamings}
-  \label{sec:maps:renamings}
-  A problem with mechanizing languages with DeBruijn-indices is that
-  the operation of applying a substitution \ACode{σ} to a term
-  \ACode{t} cannot be defined directly by structural recursion on
-  \ACode{t}.
-  If \ACode{t} is a term like \ACode{\ACon{λx} e},
-  where the subterm \ACode{e} is under a binder, then the variables
-  in \ACode{e} are shifted, i.e.\ the variable \ACode{\ACon{zero}} refers
-  to the variable from the binder, whereas a variable \ACode{\ACon{suc} x}
-  refers to the variable \ACode{x} from outside of the binder.
-  As the terms contained in \ACode{σ} may themselves contain free variables,
-  they need to be shifted, too, before they can be substituted
-  for variables in \ACode{e}. 
-  Shifting the variables in a term is itself a substitution, which maps
-  each variable \ACode{x} to the term \ACode{\ACon{`} (\ACon{suc} x)}.
-  However, applying this shifting substitution to the terms in
-  \ACode{σ} is not structurally recursive, as the terms in \ACode{σ}
-  are not subterms of \ACode{\ACon{λx} e}.
-
-  The usual workaround is to first define renamings,
-  i.e.\ substitutions that replace variables with variables.
-  In contrast to general substitutions, renamings can be easily
-  shifted: they only contain variables \ACode{x}, which can be
-  shifted by simply taking their successor \ACode{\ACon{suc} x}.
-  As the shifting substitution is a renaming, we can then define substitution,
-  where we apply the shifting \emph{renaming} to the terms in \ACode{σ} when
-  the substitution needs to go under a binder.
-
-  % The usual workaround is to define renamings first and use them
-  % to weaken the terms in substitutions when going under binders.
-  While this keeps lemmas simple (requiring only structural induction),
-  it duplicates the amount of work as all operations and lemmas have
-  to be defined and proved for both renamings and substitutions.
+  % \subsection{The Need for Renamings}
+  % \label{sec:maps:renamings}
+  % A problem with mechanizing languages with DeBruijn-indices is that
+  % the operation of applying a substitution \ACode{σ} to a term
+  % \ACode{t} cannot be defined directly by structural recursion on
+  % \ACode{t}.
+  % If \ACode{t} is a term like \ACode{\ACon{λx} e},
+  % where the subterm \ACode{e} is under a binder, then the variables
+  % in \ACode{e} are shifted, i.e.\ the variable \ACode{\ACon{zero}} refers
+  % to the variable from the binder, whereas a variable \ACode{\ACon{suc} x}
+  % refers to the variable \ACode{x} from outside of the binder.
+  % As the terms contained in \ACode{σ} may themselves contain free variables,
+  % they need to be shifted, too, before they can be substituted
+  % for variables in \ACode{e}. 
+  % Shifting the variables in a term is itself a substitution, which maps
+  % each variable \ACode{x} to the term \ACode{\ACon{`} (\ACon{suc} x)}.
+  % However, applying this shifting substitution to the terms in
+  % \ACode{σ} is not structurally recursive, as the terms in \ACode{σ}
+  % are not subterms of \ACode{\ACon{λx} e}.
+
+  % The usual workaround is to first define renamings,
+  % i.e.\ substitutions that replace variables with variables.
+  % In contrast to general substitutions, renamings can be easily
+  % shifted: they only contain variables \ACode{x}, which can be
+  % shifted by simply taking their successor \ACode{\ACon{suc} x}.
+  % As the shifting substitution is a renaming, we can then define substitution,
+  % where we apply the shifting \emph{renaming} to the terms in \ACode{σ} when
+  % the substitution needs to go under a binder.
+
+  % % The usual workaround is to define renamings first and use them
+  % % to weaken the terms in substitutions when going under binders.
+  % While this keeps lemmas simple (requiring only structural induction),
+  % it duplicates the amount of work as all operations and lemmas have
+  % to be defined and proved for both renamings and substitutions.
 
@@
   Building on the operations and axioms of a \ACode{\ACon{Kit}}, we
-  derive further operations:\footnote{
+  define the usual operations for renamings and substitutions:\footnote{
     The definitions are included directly in the record module of the
@@
     \ACode{\AField{⋯-var}} states that applying a map \ACode{ϕ} to a variable
-    term\\ \ACode{\AField{`} x}, is the same as applying \ACode{ϕ} to
+    term \ACode{\AField{`} x}, is the same as applying \ACode{ϕ} to
     the variable \ACode{x}, and then converting the result to a term
@@
   \AppCode\FTyping
+
+  % \begin{minipage}[b]{0.5\linewidth}
+  % \AppCode\FTypingI
+  % \end{minipage}
+  % \begin{minipage}[b]{0.5\linewidth}
+  % \AppCode\FTypingII
+  % \end{minipage}
+
   The interesting cases are:
@@
 
-  \section{Subject Reduction}
-  \label{sec:subjectreduction}
-
-  To finish our running example, we define a standard small-step
-  operational semantics for System F with full reduction and prove subject reduction:
-  \AppCode\FReduction
-  \vspace{0mm}
-  \AppCode\FSubjectReduction
-  \AppCode\FSubjectReductionProofInteresting
-  We only list the cases for $\beta$-reduction, as the other cases
-  follow directly from the induction hypothesis:
-  \begin{itemize}
-  \item The $\beta$-reduction rule for type-application, \ACode{\ACon{β-Λ}},
-    is a direct consequence of substitution-preserves-typing.
-  \item The $\beta$-reduction rule for value-application is almost the same,
-    but requires us to additionally show that the type is not affected by the
-    expression substitution.
-    % This is a direct consequence of the weakening used in \ACode{\ACon{⊢λ}}.
-  \end{itemize}
+  % \section{Subject Reduction}
+  % \label{sec:subjectreduction}
+
+  % To finish our running example, we define a standard small-step
+  % operational semantics for System F with full reduction and prove subject reduction:
+  % \AppCode\FReduction
+  % \vspace{0mm}
+  % \AppCode\FSubjectReduction
+  % \AppCode\FSubjectReductionProofInteresting
+  % We only list the cases for $\beta$-reduction, as the other cases
+  % follow directly from the induction hypothesis:
+  % \begin{itemize}
+  % \item The $\beta$-reduction rule for type-application, \ACode{\ACon{β-Λ}},
+  %   is a direct consequence of substitution-preserves-typing.
+  % \item The $\beta$-reduction rule for value-application is almost the same,
+  %   but requires us to additionally show that the type is not affected by the
+  %   expression substitution.
+  %   % This is a direct consequence of the weakening used in \ACode{\ACon{⊢λ}}.
+  % \end{itemize}
 

### 2024-03-11 paper/paper.tex
Update.

@@
 % \documentclass[sigplan,10pt, anonymous]{acmart}
-\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
+\documentclass[a4paper, UKenglish, cleveref, autoref, thm-restate]{lipics-v2021}
 
@@
 \authorrunning{H. Saffrich}
-\Copyright{CC-BY}
+\Copyright{Hannes Saffrich}
 \keywords{Agda, Metatheory, Framework}
@@
   and an abstraction to unify type preservation lemmas for renamings and substitutions.
-  Section~\ref{sec:subjectreduction} introduces a small-step
-  operational semantics for System F and proves subject reduction.
+  % Section~\ref{sec:subjectreduction} introduces a small-step
+  % operational semantics for System F and proves subject reduction.
   Section~\ref{sec:reflection} describes the class of object languages covered by
@@
   most proof assistants forbid to avoid breaking logical consistency.
-  We conjecture that an encoding might be possible, as it has been
-  done for intrinsically typed languages with dependent
-  types\cite{DBLP:conf/popl/AltenkirchK16}, but those constructions come
-  with additional complexity and are less suitable to automation.
+  % We conjecture that an encoding might be possible, as it has been
+  % done for intrinsically typed languages with dependent
+  % types\cite{DBLP:conf/popl/AltenkirchK16}, but those constructions come
+  % with additional complexity and are less suitable to automation.
 
@@
     \ACode{\AField{`/id}} converts a \ACode{S ∋/⊢ s} into a term \ACode{S \AField{⊢} s}
-    and is dual to the \ACode{\AField{id/`}} operation.
+    and is analogous to the \ACode{\AField{id/`}} operation.
   \item 
@@
   \item 
-    A map can be lifted, which is the operation necessary to
+    A map can be lifted, which allows to
     move a map under a binder of sort \ACode{s}:
@@
   \begin{itemize}
-  \item \ACode{S \ADef{∋/⊢[} K \ADef{]} s} is \ACode{S ∋/⊢ s} for some \ACode{\ACon{Kit}} \ACode{K}; and
-  \item \ACode{S \ADef{–[} K \ADef{]→} s} is \ACode{S \ADef{→ₖ} s} for some \ACode{\ACon{Kit}} \ACode{K}.
+  \item
+     we write \ACode{(S \ADef{∋/⊢[} K \ADef{]} s)} for the \ACode{(S ∋/⊢ s)}
+     of some specific \ACode{\ACon{Kit}} \ACode{K}; and
+   \item
+     we write \ACode{(S \ADef{–[} K \ADef{]→} s)} for the \ACode{(S \ADef{→ₖ} s)}
+     of some specific \ACode{\ACon{Kit}} \ACode{K}.
   \end{itemize}

### 2024-03-11 paper/paper.tex
Update.

@@
 
-  We then create exactly two instances of the \ACode{\ACon{WkKit}} for
-  the variable and term \ACode{\ACon{Kit}}, respectively. 
+  We then create two instances of the \ACode{\ACon{WkKit}} for
+  the variable and term \ACode{\ACon{Kit}}, respectively:
   \LibCode\KWkKitInstances
@@
   allows the composition of arbitrary maps.
-
   The core property of composition, which we call
@@
   and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same as
-  applying their composition \ACode{(ϕ₁ \AField{·ₖ} ϕ₂)} to \ACode{t},
+  applying their composition \ACode{(ϕ₁~\AField{·ₖ}~ϕ₂)} to \ACode{t},
   i.e.
@@
   dependencies.
-  While composition can be dervied, this has the benefit, that
-  we've gained a new level of generality, which enables us in the next
+  % TODO
+  While composition can be derived, our approach has the benefit, that
+  we have gained a new level of generality, which enables us in the next
   sections to unify type preservation of all substitution and renaming
@@
   to satisfy two requirements:
-  it needs to provide us with enough machinery to define composition, but also
-  it needs to provide us with enough axioms that we are able to instanciate
+  it needs to provide us with enough machinery to define composition, but
+  it also needs to provide us with enough axioms that we are able to instanciate
   the \ACode{\ACon{ComposeTraversal}}. It is defined as followed:

### 2024-03-11 paper/paper.tex
Update.

@@
 
-  % Before we define \ACode{\ACon{ComposeKit}} and
-  % \ACode{\ACon{ComposeTraversal}}, we need one more ingredient.
+  % Before we define \ACode{\ACon{CKit}} and
+  % \ACode{\ACon{CTraversal}}, we need one more ingredient.
 
@@
   % - We introduce an abstraction similar to kits
-  % - ComposeKit has primitives for composition, ComposeTraversal has the property
+  % - CKit has primitives for composition, CTraversal has the property
 
@@
   To fill this gap, we introduce a new abstraction that we call a
-  \ACode{\ACon{ComposeKit}},
+  \ACode{\ACon{CKit}},
   which provides an operation \ACode{\AField{\_\&/⋯\_}} that unifies
@@
 
-  % \ACode{\ACon{ComposeKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
+  % \ACode{\ACon{CKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
   % \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
-  % \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-fusion}} are analogous to
-  % \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{\_⋯\_}}.
+  % \ACode{\ACon{CTraversal}} and \ACode{\AField{⋯-fusion}} are analogous to
+  % \ACode{\ACon{CTraversal}} and \ACode{\AField{\_⋯\_}}.
 
@@
 
-  In this subsection, we define the \ACode{\ACon{ComposeKit}}
+  In this subsection, we define the \ACode{\ACon{CKit}}
   abstraction, which provides the \ACode{\AField{\_\&/⋯\_}} operation
   that is necessary for defining composition, and the
-  \ACode{\ACon{ComposeTraversal}}, which provides the
+  \ACode{\ACon{CTraversal}}, which provides the
   \ACode{\AField{⋯-fusion}} lemma for general maps.
@@
   Similarly as it was the case for \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}},
-  the lemma of the \ACode{\ACon{ComposeTraversal}} needs to be proved for
-  an abstract \ACode{\ACon{ComposeKit}} and in return defines the
-  four concrete \ACode{\ACon{ComposeKit}} instances for us. This allows us to break
+  the lemma of the \ACode{\ACon{CTraversal}} needs to be proved for
+  an abstract \ACode{\ACon{CKit}} and in return defines the
+  four concrete \ACode{\ACon{CKit}} instances for us. This allows us to break
   the dependencies between the different \ACode{\AField{⋯-fusion}} lemmas,
@@
 
-  For this purpose, the \ACode{\ACon{ComposeKit}} abstraction, needs
+  For this purpose, the \ACode{\ACon{CKit}} abstraction, needs
   to satisfy two requirements:
@@
   it also needs to provide us with enough axioms that we are able to instanciate
-  the \ACode{\ACon{ComposeTraversal}}. It is defined as followed:
-  \LibCode\KComposeKit
+  the \ACode{\ACon{CTraversal}}. It is defined as followed:
+  \LibCode\KCKit
   The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
@@
   Similar as for the \ACode{\ACon{Kit}}, we derive further definitions
-  from the axioms of a \ACode{\ACon{ComposeKit}}:
+  from the axioms of a \ACode{\ACon{CKit}}:
   \begin{itemize}
@@
     \ACode{\AField{\_\&\_}} when applied to variables:
-    \LibCode\KComposeKitAp
-    % \LibCode\KComposeKitApProof
+    \LibCode\KCKitAp
+    % \LibCode\KCKitApProof
 
@@
   \end{itemize}
-  To make it easier to explicitly specify a \ACode{\ACon{ComposeKit}},
+  To make it easier to explicitly specify a \ACode{\ACon{CKit}},
   we define the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}
   to stand for \ACode{ϕ₁ \ADef{·ₖ} ϕ₂} for some
-  \ACode{\ACon{ComposeKit}} \ACode{C}.
-  % \LibCode\KComposeKitNotation
-
-  We then define the \ACode{\ACon{ComposeTraversal}} structure,
+  \ACode{\ACon{CKit}} \ACode{C}.
+  % \LibCode\KCKitNotation
+
+  We then define the \ACode{\ACon{CTraversal}} structure,
   which provides the \ACode{\AField{⋯-fusion}} lemma for arbitrary
-  \ACode{\ACon{ComposeKit}}s:
-  \LibCode\KComposeTraversal
+  \ACode{\ACon{CKit}}s:
+  \LibCode\KCTraversal
 
   Similar as for the \ACode{\ACon{Traversal}}, we derive further
-  definitions from the axiom of a \ACode{\ACon{ComposeTraversal}}:
+  definitions from the axiom of a \ACode{\ACon{CTraversal}}:
   \begin{itemize}
@@
   \item 
-    We define \ACode{\ACon{ComposeKit}} instances for each of the four
+    We define \ACode{\ACon{CKit}} instances for each of the four
     combinations between renamings and substitutions:
-    \LibCode\KComposeKitInstances
+    \LibCode\KCKitInstances
     Interestingly, the \ACode{\ADef{Cₛ}} instance has another
-    \ACode{\ADef{ComposeKit}} parameter.
+    \ACode{\ADef{CKit}} parameter.
     This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
-    which yields the \ACode{\ADef{ComposeKit}} between substitutions
-    and renamings, and then use this resulting \ACode{\ADef{ComposeKit}}
+    which yields the \ACode{\ADef{CKit}} between substitutions
+    and renamings, and then use this resulting \ACode{\ADef{CKit}}
     as an argument for the case with two substitutions.
@@
 
-    % To get the \ACode{\ADef{ComposeKit}} between substitutions and
+    % To get the \ACode{\ADef{CKit}} between substitutions and
     % renamings, the argument \ACode{\ADef{Cᵣ}} is used.
-    % The resulting \ACode{\ADef{ComposeKit}} is then again used as an argument
-    % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{ComposeKit}}
+    % The resulting \ACode{\ADef{CKit}} is then again used as an argument
+    % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{CKit}}
     % between two substitutions. As shown below, Agda's instance
     % resolution is up for the task:
-    \LibCode\KComposeKitInstancesConcrete
+    \LibCode\KCKitInstancesConcrete
 
@@
   In this subsection, we show how to instantiate the
-  \ACode{\ACon{ComposeTraversal}} abstraction for System F.
+  \ACode{\ACon{CTraversal}} abstraction for System F.
   In practice, this is done by our reflection algorithm automatically,
   but it can be instructive to see, as it motivates the axioms of the
-  \ACode{\ACon{ComposeKit}}.
+  \ACode{\ACon{CKit}}.
 
@@
     variables, where we need to use the \ACode{\AField{\&/⋯-⋯}} lemma
-    provided by the \ACode{\ACon{ComposeKit}}.
+    provided by the \ACode{\ACon{CKit}}.
   \item
@@
   % \AppCode\FFusionProofRest
-  % \AppCode\FComposeTraversal
+  % \AppCode\FCTraversal
 
@@
 
-  We start with \ACode{\ACon{TypingKit}}s, which abstract over
+  We start with \ACode{\ACon{TKit}}s, which abstract over
   variable and term \emph{typing}, and then define a
-  \ACode{\ACon{TypingTraversal}}, which provides
-  substitution-preserves-typing for all \ACode{\ACon{TypingKit}}s.
-  \LibCode\KTypingKit
+  \ACode{\ACon{TTraversal}}, which provides
+  substitution-preserves-typing for all \ACode{\ACon{TKit}}s.
+  \LibCode\KTKit
   The operations are completely analogous to those of a \ACode{\ACon{Kit}}, but talk
   about typing instead of terms.
-  Building on the fields of a \ACode{\ACon{TypingKit}}, we define further operations:
+  Building on the fields of a \ACode{\ACon{TKit}}, we define further operations:
   \begin{itemize}
@@
   to stand for \ACode{Γ₂ \ADef{∋*/⊢*} ϕ \ADef{∶} Γ₁} given some
-  \ACode{\ACon{TypingKit}} \ACode{TK}.
+  \ACode{\ACon{TKit}} \ACode{TK}.
   % \LibCode\KTypingNotation
 
-  We then define a \ACode{\ACon{TypingTraversal}} analogously to
+  We then define a \ACode{\ACon{TTraversal}} analogously to
   \ACode{\ACon{Traversal}}, but instead of defining the application of
   maps, it defines the type preservation of applying maps:
-  \LibCode\KTypingTraversal
+  \LibCode\KTTraversal
   Given a term \ACode{e} with typing \ACode{⊢e} and
@@
 
-  Given a \ACode{\ACon{TypingTraversal}}, we can easily instantiate the
-  \ACode{\ACon{TypingKits}} for variables and terms:
+  Given a \ACode{\ACon{TTraversal}}, we can easily instantiate the
+  \ACode{\ACon{TKits}} for variables and terms:
   \LibCode\KTypingInstances
-  % \LibCode\KTypingTraversalNotation
+  % \LibCode\KTTraversalNotation
 
@@
   In this subsection, we show how to instantiate the
-  \ACode{\ACon{TypingTraversal}} abstraction for System F.
+  \ACode{\ACon{TTraversal}} abstraction for System F.
   This is the only structure that is \emph{not} instantiated
@@
     The type application constructor \ACode{\ACon{⊢∙}} substitutes \ACode{t₁} into \ACode{t₂},
-    requiring us to use the \ACode{\AField{dist-↑-⦅⦆}} lemma.
+    requiring us to use \ACode{\AField{dist-↑-⦅⦆}}.
   \end{itemize}
-  Finally, we instantiate the \ACode{\ACon{TypingTraversal}} structure:
-  \AppCode\FTypingTraversal
-
-  % As the \ACode{\ACon{Traversal}} and  \ACode{\ACon{ComposeTraversal}} instances
+  Finally, we instantiate the \ACode{\ACon{TTraversal}} structure:
+  \AppCode\FTTraversal
+
+  % As the \ACode{\ACon{Traversal}} and  \ACode{\ACon{CTraversal}} instances
   % can be generally derived via reflection,
@@
   i.e. \ACode{\ACon{Syntax}}, \ACode{\ACon{Traversal}} and
-  \ACode{\ACon{ComposeTraversal}}.
+  \ACode{\ACon{CTraversal}}.
 
@@
     just works out of the box for both deriving and instantiating the
-    \ACode{\ACon{TypingTraversal}}.
-    Furthermore, the \ACode{\ACon{ComposeKit}} abstraction turned out
+    \ACode{\ACon{TTraversal}}.
+    Furthermore, the \ACode{\ACon{CKit}} abstraction turned out
     to be useful to prove confluence, allowing to unify certain
@@
     lemmas by hand, we found an encoding, which still allows us to use the
-    the \ACode{\ACon{TypingKit}} abstraction.
+    the \ACode{\ACon{TKit}} abstraction.
     Instead of binding a type variable with a subtyping constraint
@@
     again, enabling the use of the framework's typing contexts and
-    \ACode{\ACon{TypingKit}} machinery.
+    \ACode{\ACon{TKit}} machinery.
     When replacing a type variable \ACode{α <: t} with a type \ACode{t'},
@@
     Extending the class of typing relations covered by
-    the \ACode{\ACon{TypingTraversal}} abstraction and identifying
+    the \ACode{\ACon{TTraversal}} abstraction and identifying
     a subclass for which it can be derived.
@@
   \subsection{Proofs for Section~\ref{sec:composition}}
-  \LibCode\KComposeKitAp
-  \LibCode\KComposeKitApProof
+  \LibCode\KCKitAp
+  \LibCode\KCKitApProof
 

### 2024-03-11 paper/agdatex.py
Update.

@@ -7,7 +7,8 @@ from argparse import ArgumentParser
 import tempfile
 import subprocess
 import shutil
-
+import json
+import hashlib
 
 # Parse Arguments
 
@@ -39,6 +40,9 @@ ap.add_argument("-v", "--verbose", action='store_true',
 
 ap.add_argument("-n", "--noop", action='store_true',
                     help="No action: dry run")
+
+ap.add_argument ("-c", "--clear", action='store_true',
+                     help="Clear cashes to force rebuild all")
 
 ap.add_argument("sources", metavar="SRC_PATH", nargs="+",
                 help="Path to an annotated .agda-file.")
@@ -96,6 +100,29 @@ else:
 if args.verbose:
     print (f"VERBOSE: tmp_root = {tmp_root}")
 
+old_hashes = dict()
+if args.clear:
+    if args.verbose:
+        print(f"VERBOSE: ignoring cached hashes")
+else:
+    try:
+        with open(".agdatex-hashes.json") as digest:
+            old_hashes = json.load(digest)
+    except (OSError, json.decoder.JSONDecodeError) as e:
+        if args.verbose:
+            print (f"VERBOSE: cannot read hashes ({e})")
+
+if args.verbose:
+    print(f"VERBOSE: {old_hashes=}")
+
+current_hashes = dict()
+for src_path in src_paths:
+    with open (root / src_path, "rb") as fp:
+        current_hashes[str(src_path)] = hashlib.file_digest(fp, "sha256").hexdigest()
+
+if args.verbose:
+    print(f"VERBOSE: {current_hashes=}")
+
 if args.noop:
     sys.exit("[no harm done]")
 
@@ -127,7 +154,7 @@ for src_path, tgt_path in zip(src_paths, tgt_paths):
     opt = ""
     def start_command(name, is_inline):
         global mode, tgt, prefixes, opt
-        for p in prefixes:
+        for p in reversed(prefixes):
             name = p + name
         commands.append(name)
         if mode == "hide":
@@ -243,13 +270,24 @@ output_dir = Path(args.outputdir)
 output_dir.mkdir(exist_ok=True)
 
 print("Running agda with latex backend...")
-for tgt_path in tgt_paths:
+for src_path, tgt_path in zip(src_paths, tgt_paths):
+    if current_hashes[str(src_path)] == old_hashes.get(str(src_path), ''):
+        print(f"  Skipping {tgt_path} (no change)...")
[…]

### 2024-03-11 paper/paper.tex
Update.

@@
   Section~\ref{sec:related} discusses related work.
-  Section~\ref{sec:future} concludes with future work.
+  Section~\ref{sec:conclusion} concludes with future work.
 
@@
 
-  \section{Future Work}
-  \label{sec:future}
+  \subsection{Fibrational}
+  \cite{DBLP:conf/rta/LicataSR17}
+
+  \subsection{Pure Type Systems}
+  \cite{DBLP:conf/esop/BarendregtH90, berarditowards, terlouw1989een}
+
+  \section{Conclusion}
+  \label{sec:conclusion}
   We want to focus our future work in the following directions:

### 2024-03-11 paper/paper.tex
Update.

@@
     \LibCode\KAp
-  \item 
-    A map can be extended by a new variable or term:
-    \LibCode\KExt
-  \item 
-    A map can be shifted, as the \ACode{\ACon{Kit}} provides the
-    necessary \ACode{\AField{wk}} operation:
-    \LibCode\KWkm
+  % \item 
+  %   A map can be extended by a new variable or term:
+  %   \LibCode\KExt
+  % \item 
+  %   A map can be shifted, as the \ACode{\ACon{Kit}} provides the
+  %   necessary \ACode{\AField{wk}} operation:
+  %   \LibCode\KWkm
   \item 
@@
     move a map under a binder of sort \ACode{s}:
-    \LibCode\KLift
-  \item 
-    There is an identity map:
-    \LibCode\KId
+    % \LibCode\KLift
+    \LibCode\KLiftAlt
   \item 
@@
     all other variables by one:
-    \LibCode\KSingle
+    \LibCode\KSingleAlt
   \item 
     There is a map which is equivalent to \ACode{\AField{wk}} on terms:
-    \LibCode\KWeaken
+    \LibCode\KWeakenAlt
   \item
@@
   \item 
+    There is an identity map:
+    \LibCode\KId
+  \item
     Lifting the identity map is a (larger) identity map:
@@
   \end{itemize}
-
-  We also specify notation, which makes it less cumbersome
+  We also specify notation 
   to explicitly pass a particular \ACode{\ACon{Kit}} to those definitions:
+  % We also specify notation, which makes it less cumbersome
+  % to explicitly pass a particular \ACode{\ACon{Kit}} to those definitions:
   \begin{itemize}
@@
     % resolution is up for the task:
-    \LibCode\KCKitInstancesConcrete
+    \begin{minipage}[t]{0.5\linewidth}
+    \LibCode\KCKitInstancesConcreteI
+    \end{minipage}
+    \begin{minipage}[t]{0.5\linewidth}
+    \LibCode\KCKitInstancesConcreteII
+    \end{minipage}
 

### 2024-03-11 paper/paper.tex
Update.

@@
 
+  an implementation as an Agda framework featuring
+  a reflection algorithm, representation independence for
+  substitutions and typing contexts, heterogeneous equality between
+  renamings and substitutions, and absence of postulated axioms.
+  \begin{code}
+fooo
+  \end{code}
+
   \ACode{\ACon{Type}}s are indexed by the number of free type
@@
   applying their composition \ACode{(ϕ₁~\AField{·ₖ}~ϕ₂)} to \ACode{t},
-  i.e.
-  $$
-  \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
-  $$
+  i.e. \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
+  % $$
+  % \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
+  % $$
   This property equips our framework with the power to
@@
   \label{sec:composition:example}
-  To motivate our abstraction for composition, we first look
+  To motivate our abstraction, we first look
   at the four compositions individually:
-  \ExampleCode\KExFourComps
+
+  \noindent
+  \begin{minipage}[t]{0.60\linewidth}
+  \ExampleCode\KExFourCompsI
+  \end{minipage}
+  \begin{minipage}[t]{0.40\linewidth}
+  \ExampleCode\KExFourCompsII
+  \end{minipage}
+
   The definitions reveal two interesting properties:
@@
   % \LibCode\KContextHelper
+  \noindent
+  \begin{minipage}{1.0\linewidth}
+    \begin{minipage}{0.5\linewidth}
       \LibCode\KContexts
+      \vspace{3mm}
+    \end{minipage}
+    \begin{minipage}{0.5\linewidth}
+      \LibCode\KContextExt
+    \end{minipage}
+  \end{minipage}
+
   To use a type returned from a typing context, it needs to be
   weakened appropriately through the \ACode{\ADef{wk-telescope}} function:
-  \LibCode\KContextLookup
+
+  \noindent
+  \begin{minipage}{0.5\linewidth}
+    \LibCode\KContextLookupI
+  \end{minipage}
+  \begin{minipage}{0.5\linewidth}
+    \LibCode\KContextLookupII
+  \end{minipage}
+  % \LibCode\KContextLookup
+
   We then define a variable typing as followed:

### 2024-03-11 paper/paper.tex
Update.

@@
 
-  an implementation as an Agda framework featuring
-  a reflection algorithm, representation independence for
-  substitutions and typing contexts, heterogeneous equality between
-  renamings and substitutions, and absence of postulated axioms.
-  \begin{code}
-fooo
-  \end{code}
-
   \ACode{\ACon{Type}}s are indexed by the number of free type
@@
   \label{sec:syntax:structure}
+  % The regularity of the multi-sorted syntax makes it easy to define a
+  % structure for syntaxes of arbitrary object-languages, i.e.\ syntaxes
+  % with arbitrary amounts of syntactic categories and variables:
   The regularity of the multi-sorted syntax makes it easy to define a
-  structure for syntaxes of arbitrary object-languages, i.e.\ syntaxes
-  with arbitrary amounts of syntactic categories and variables:
+  structure for arbitrary syntaxes, i.e. with arbitrarily many syntactic categories and variable types:
   \LibCode\KSyntax
@@
   McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
-  introduced an abstraction called \emph{kits}.\footnote{
-    While kits were originally formulated for intrinsically-typed languages,
-    they nicely extend to our extrinsically-typed multi-sorted syntax, whose
-    sort-indexing is a form of intrinsic typing.
-  }
-
+  introduced the \emph{Kit} abstraction.%
+  % \footnote{
+  %   While kits were originally formulated for intrinsically-typed languages,
+  %   they nicely extend to our extrinsically-typed multi-sorted syntax, whose
+  %   sort-indexing is a form of intrinsic typing.
+  % }
   % A \ACode{Kit} is a structure that captures the commonalities
@@
 
-  A kit provides the following operations:
+  A kit consists of the following components:
   \begin{itemize}
@@
     For terms, \ACode{\AField{wk}} means applying a shifting renaming to the term.
-  \end{itemize}
-
-  The operations need to satisfy the following properties:
-  \begin{itemize}
   \item \ACode{\AField{`/`-is-`}} states that converting a variable first to a ``variable-or-term'' and
@@
   \item \ACode{\AField{`/id-injective}} and \ACode{\AField{id/`-injective}} state that
-    \ACode{\AField{`/id}} and \ACode{\AField{id/`}} are injective. This property
-    follows easily, as both the identity function and the variable
-    constructor are injective.
+    \ACode{\AField{`/id}} and \ACode{\AField{id/`}} are injective.
+    % This property
+    % follows easily, as both the identity function and the variable
+    % constructor are injective.
   \item \ACode{\AField{wk-id/`}} characterizes the behaviour of the
-    \ACode{\AField{wk}} function by how it acts on variables.
-    Injecting a variable and then shifting it, is the same as injecting a shifted variable.
+    \ACode{\AField{wk}} function by how it acts on variables:
+    injecting a variable and then shifting it, is the same as injecting a shifted variable.
   \end{itemize}
 
-  Building on the operations and axioms of a \ACode{\ACon{Kit}}, we
-  define the usual operations for renamings and substitutions:\footnote{
+  \begin{figure}
+    \label{fig:map-ops}
+    \centering
+    \begin{minipage}[t]{0.5\linewidth}
+    \LibCode\KMap
+    \LibCode\KAp
+    \LibCode\KLiftAlt
+    \LibCode\KSingleAlt
+    \end{minipage}
+    \begin{minipage}[t]{0.4\linewidth}
+    \LibCode\KWeakenAlt
+    \LibCode\KEq
+    \LibCode\KFunExt
+    \LibCode\KId
+    \LibCode\KIdLift
+    \end{minipage}
+    \caption{Map Operations}
+  \end{figure}
+  Figure~\ref{fig:map-ops} shows the usual operations for renamings and substitutions
+  generalized to maps.\footnote{
     The definitions are included directly in the record module of the
@@
   }
-  \begin{itemize}
-  \item 
-    We unify the definition of renamings \ACode{S₁ \ADef{→ᵣ} S₂} and
-    substitutions \ACode{S₁ \ADef{→ₛ} S₂}, which differ only in whether they
-    return variables or terms:
-    \LibCode\KMap
+  The type \ACode{S₁ \ADef{→ₖ} S₂} unifies
+    renamings \ACode{S₁ \ADef{→ᵣ} S₂} and
+    substitutions \ACode{S₁ \ADef{→ₛ} S₂}.
     We call a value of type \ACode{S₁ \ADef{→ₖ} S₂} a \emph{map} and
     use the meta-variable \ACode{ϕ} for it.
-  \item 
-    A map can be applied to a variable:
-    \LibCode\KAp
-  % \item 
-  %   A map can be extended by a new variable or term:
-  %   \LibCode\KExt
-  % \item 
-  %   A map can be shifted, as the \ACode{\ACon{Kit}} provides the
-  %   necessary \ACode{\AField{wk}} operation:
-  %   \LibCode\KWkm
-  \item 
-    A map can be lifted, which allows to
-    move a map under a binder of sort \ACode{s}:
-    % \LibCode\KLift
-    \LibCode\KLiftAlt
-  \item 
-    There is a singleton map, which replaces \ACode{\ACon{zero}} and decreases
-    all other variables by one:
-    \LibCode\KSingleAlt
-  \item 
-    There is a map which is equivalent to \ACode{\AField{wk}} on terms:
-    \LibCode\KWeakenAlt
-  \item
-    Maps have an extensional equality:
-    \LibCode\KEq
-  \item
-    For simplicity we postulate functional extensionality:\footnote{
+  The operation \ACode{ϕ \& x} applies a map to a variable.
+  The operation \ACode{ϕ ↑ s} lifts a map under a binder of sort \ACode{s}.
+  The operation \ACode{⦅ x/t ⦆} constructs a singleton map that replaces
+  \ACode{\ACon{zero}} with \ACode{x/t} and decreases all other variables by one.
+  The \ACode{\AField{weaken s}} map increases all variables by one.
+  The \ACode{ϕ₁ $\sim$ ϕ₂} type expresses extensional equality of maps.
+    For simplicity, we postulate functional extensionality \ACode{\~-ext}.\footnote{
       The actual implementation does \emph{not} use any postulates.
@@
     }
-    \LibCode\KFunExt
-  \item 
-    There is an identity map:
-    \LibCode\KId
-  \item
-    Lifting the identity map is a (larger) identity map:
-    \LibCode\KIdLift
-    % \LibCode\KIdLiftProof
-  \end{itemize}
-  We also specify notation 
-  to explicitly pass a particular \ACode{\ACon{Kit}} to those definitions:
+  There is an identity map \ACode{id}.
+  The lemma \ACode{id↑-id} states that a lifted identity map is again an identity map.
+
+  To make it easier to talk about a specific kit, we introduce the following notations:
+  % We also specify notation 
+  % to explicitly pass a particular \ACode{\ACon{Kit}} to those definitions:
   % We also specify notation, which makes it less cumbersome
@@
    \item
-     we write \ACode{(S \ADef{–[} K \ADef{]→} s)} for the \ACode{(S \ADef{→ₖ} s)}
+     we write \ACode{(S₁ \ADef{–[} K \ADef{]→} S₂)} for the \ACode{(S₁ \ADef{→ₖ} S₂)}
      of some specific \ACode{\ACon{Kit}} \ACode{K}.
@@
     we get back a variable and need to apply the variable constructor
-    \ACode{\ACon{`\_}} to it. If \ACode{ϕ} is a substitution,
+    \ACode{\ACon{`\_}}. If \ACode{ϕ} is a substitution,
     we get back a term that we can use directly.
@@
   \AppCode\FTraversalIdProofInteresting
-  The interesting cases are:
+  We only display and discuss the interesting cases:
   \begin{itemize}
   \item
-    the variable case \ACode{\ACon{`} x}, where we need to combine the
-    \ACon{\AField{id/`}} from the identity map, with the \ACon{\AField{`/id}}
-    from the traversal operation using \ACon{\AField{`/`-is-`}}.
-  \item
-    the lambda abstraction case \ACode{\ACon{λx} e}, where the
-    traversal lifts the identity over its binder. Here
-    \ACon{\AField{id↑-id}} is used to state that this lifted identity
-    map is still the identity.
-  \item
-    all other cases are analogous and hence omitted.
+    in the variable case \ACode{\ACon{`} x}, the
+    \ACon{\AField{id/`}} from the identity map meets the \ACon{\AField{`/id}}
+    from the traversal operation, so we use \ACon{\AField{`/`-is-`}}.
+  \item
+    in the lambda abstraction case \ACode{\ACon{λx} e}, the
+    traversal lifts the identity under its binder. Here
+    we need to use \ACon{\AField{id↑-id}} to show that a lifted identity
+    map is again an identity map.
   \end{itemize}
@@
   their types.
-
   This prevented us to define a useful lemma, which also needs to
@@
   \ACode{ϕ₂} being a renaming or a substitution:
-  \ExampleCode\KExTwoComps
+
+  \noindent
+  \begin{minipage}[t]{0.60\linewidth}
+  \ExampleCode\KExTwoCompsI
+  \end{minipage}
+  \begin{minipage}[t]{0.40\linewidth}
+  \ExampleCode\KExTwoCompsII
+  \end{minipage}
 

### 2024-03-11 paper/paper.tex
Update.

@@
   \ACode{\ACon{Traversal}} structure, the fields of
-  \ACode{\ACon{Kit}} could not use the \ACode{\AField{\_⋯\_}} function in
+  \ACode{\ACon{Kit}} could not use map application \ACode{\AField{\_⋯\_}} in
   their types.
-  This prevented us to define a useful lemma, which also needs to
-  be instantiated separately for variables and terms: the counterpiece
-  to \ACode{\AField{wk-id/`}}, which we call \ACode{\AField{wk-`/id}}.
-
-  Recall that \ACode{\AField{wk-id/`}} explains the
-  \ACode{\AField{wk}} function of an arbitrary \ACode{\ACon{Kit}} by how it acts on
-  \emph{variables}:
-  \ExampleCode{\KWkId}
-
-  Similarly, the \ACode{\AField{wk-`/id}} lemma
-  explains the \ACode{\AField{wk}} function of an
-  arbitrary \ACode{\ACon{Kit}} by how it acts on \emph{terms}:
-  \ExampleCode{\KIdWk}
+  This prevented us to include another useful axiom into the
+  \ACode{\ACon{Kit}} structure. As this axiom also needs to be proved
+  separately for variables and terms, we define a new structure for
+  it which extends a \ACode{\ACon{Kit}}:
+  \LibCode\KWkKit
+
+  The \ACode{\AField{wk-`/id}} axiom
+  explains the \ACode{\AField{wk}} function by how it acts on \emph{terms}.
+  It is the counterpiece to the \ACode{\ACon{Kit}} axiom
+  \ACode{\AField{wk-id/`}}, which explains the \ACode{\AField{wk}}
+  function by how it acts on \emph{variables}.
 
@@
 
-  To prove \AField{wk-`/id}, we define a structure
-  \ACode{\ACon{WkKit}}, which extends a regular \ACode{\ACon{Kit}} with
-  the \AField{wk-`/id} lemma:
-  \LibCode\KWkKit
-
-  We then create two instances of the \ACode{\ACon{WkKit}} for
-  the variable and term \ACode{\ACon{Kit}}, respectively:
+  The instantiations of the \ACode{\ACon{WkKit}} are straightforward:
   \LibCode\KWkKitInstances
@@
   a \ACode{\ACon{Kit}} also supports the \ACode{\ACon{WkKit}} extension.
-
-  % Note that we have exactly two instances for \ACode{\ACon{Kit}}s
-  % and both of them have a \ACode{\ACon{WkKit}} instance. Consequently,
-  % whenever we abstract over a \ACode{\ACon{Kit}}, we can also safely assume
-  % that it has a \ACode{\ACon{WkKit}} instance, without restricting the
-  % lemmas usage.
-
-  % Both the structure and the instances are part of the library and do
-  % not require application code.
 
@@
   \noindent
-  \begin{minipage}[t]{0.60\linewidth}
+  \begin{minipage}[t]{0.65\linewidth}
   \ExampleCode\KExFourCompsI
   \end{minipage}
-  \begin{minipage}[t]{0.40\linewidth}
+  \begin{minipage}[t]{0.35\linewidth}
   \ExampleCode\KExFourCompsII
@@
   \noindent
-  \begin{minipage}[t]{0.60\linewidth}
+  \begin{minipage}[t]{0.65\linewidth}
   \ExampleCode\KExTwoCompsI
   \end{minipage}
-  \begin{minipage}[t]{0.40\linewidth}
+  \begin{minipage}[t]{0.35\linewidth}
   \ExampleCode\KExTwoCompsII
@@
   it needs to provide us with enough machinery to define composition, but
-  it also needs to provide us with enough axioms that we are able to instanciate
+  it also needs to provide us with enough axioms that we are able to instantiate
   the \ACode{\ACon{CTraversal}}. It is defined as followed:

### 2024-03-11 paper/paper.tex
Update.

@@
     right=2.5pt,%
+    before skip=5pt,%
+    after skip=5pt,%
     boxsep=0pt%
-  ]\vspace{-0.2\baselineskip}%
+  ]
 }{%
-  \vspace{-1\baselineskip}%
   \end{tcolorbox}%
@@
     right=2.5pt,%
+    before skip=5pt,%
+    after skip=5pt,%
     boxsep=0pt%
@@
   \noindent
-  \begin{minipage}{0.5\linewidth}
+  \begin{minipage}{0.55\linewidth}
     \LibCode\KContextLookupI
   \end{minipage}
-  \begin{minipage}{0.5\linewidth}
+  \begin{minipage}{0.45\linewidth}
     \LibCode\KContextLookupII
@@
   \LibCode\KTypingInstances
+
+  % \noindent
+  % \begin{minipage}[t]{0.5\linewidth}
+  % \LibCode\KTypingInstancesI
+  % \end{minipage}
+  % \begin{minipage}[t]{0.5\linewidth}
+  % \LibCode\KTypingInstancesII
+  % \end{minipage}
+
   % \LibCode\KTTraversalNotation
@@
   \AppCode\FPreserve
+  The type of \ACode{\ADef{\_⊢⋯\_}} is the same as in the record definition and hence omitted.
   The interesting parts of the proof are:
@@
   \end{itemize}
-  Finally, we instantiate the \ACode{\ACon{TTraversal}} structure:
-  \AppCode\FTTraversal
+  % Finally, we instantiate the \ACode{\ACon{TTraversal}} structure:
+  % \AppCode\FTTraversal
 

### 2024-03-11 paper/paper.tex
Update.

@@
   The third parameter \ACode{K₁⊔K₂} can be seen as a functional
-  dependency and is completely determined by the choice of \ACode{K₁}
-  and \ACode{K₂}.
+  dependency\cite{DBLP:conf/esop/Jones00} and is completely determined
+  by the choice of \ACode{K₁} and \ACode{K₂}.
 

### 2024-03-11 paper/paper.tex
Update.

@@
     \LibCode\KComposition
-  \item
-    We show that the \ACode{\AField{\_\&/⋯\_}} operator behaves like
-    \ACode{\AField{\_\&\_}} when applied to variables:
-    \LibCode\KCKitAp
-    % \LibCode\KCKitApProof
-
-    % The \ACode{\AField{`/id}} on the left side is from \ACode{K₁⊔K₂}, whereas
-    % the \ACode{\AField{`/id}} on the right side is from \ACode{K₂}.
+  % \item
+  %   We show that the \ACode{\AField{\_\&/⋯\_}} operator behaves like
+  %   \ACode{\AField{\_\&\_}} when applied to variables:
+  %   \LibCode\KCKitAp
+  %   % \LibCode\KCKitApProof
+
+  %   % The \ACode{\AField{`/id}} on the left side is from \ACode{K₁⊔K₂}, whereas
+  %   % the \ACode{\AField{`/id}} on the right side is from \ACode{K₂}.
   \item
@@
 
-  We then define the \ACode{\ACon{CTraversal}} structure,
+  We then define a structure,
   which provides the \ACode{\AField{⋯-fusion}} lemma for arbitrary
@@
   Similar as for the \ACode{\ACon{Traversal}}, we derive further
-  definitions from the axiom of a \ACode{\ACon{CTraversal}}:
+  definitions from the axioms of a \ACode{\ACon{CTraversal}}:
   \begin{itemize}

### 2024-03-11 paper/paper.tex
Update.

@@
   \begin{figure}
-    \label{fig:map-ops}
     \centering
@@
     \caption{Map Operations}
+    \label{fig:map-ops}
   \end{figure}
@@
   The \ACode{ϕ₁ $\sim$ ϕ₂} type expresses extensional equality of maps.
-    For simplicity, we postulate functional extensionality \ACode{\~-ext}.\footnote{
+    For simplicity, we postulate functional extensionality \ACode{$\sim$-ext}.\footnote{
       The actual implementation does \emph{not} use any postulates.
@@
     variable:
-    \LibCode\KCommLiftWeaken
-    % \LibCode\KCommLiftWeakenProof
-    By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the same
-    statement for multiple map applications:
     \LibCode\KCommLiftWeakenTraverse
-    % \LibCode\KCommLiftWeakenTraverseProof
+
+    % A map \ACode{ϕ} followed by a weakening is equivalent to a
+    % weakening followed by \ACode{ϕ} that has been lifted over the weakened
+    % variable:
+    % \LibCode\KCommLiftWeaken
+    % % \LibCode\KCommLiftWeakenProof
+    % By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the same
+    % statement for multiple map applications:
+    % \LibCode\KCommLiftWeakenTraverse
+    % % \LibCode\KCommLiftWeakenTraverseProof
   \item 
     A weakening followed by a singleton substitution cancel each other out:
-    \LibCode\KWeakenCancelsSingle
-    % \LibCode\KWeakenCancelsSingleProof
-    By applying \ACode{\AField{⋯-fusion}} and \ACode{\AField{⋯-id}} to
-    this lemma, we get the same statement for multiple map applications:
     \LibCode\KWeakenCancelsSingleTraverse
-    % \LibCode\KWeakenCancelsSingleTraverseProof
+
+    % A weakening followed by a singleton substitution cancel each other out:
+    % \LibCode\KWeakenCancelsSingle
+    % % \LibCode\KWeakenCancelsSingleProof
+    % By applying \ACode{\AField{⋯-fusion}} and \ACode{\AField{⋯-id}} to
+    % this lemma, we get the same statement for multiple map applications:
+    % \LibCode\KWeakenCancelsSingleTraverse
+    % % \LibCode\KWeakenCancelsSingleTraverseProof
   \item 
     A singleton map can be swapped with any map \ACode{ϕ}:
-    \LibCode\KDistLiftSingle
-    % \LibCode\KDistLiftSingleProof
-    By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the
-    same statement for multiple map applications:
     \LibCode\KDistLiftSingleTraverse
-    % \LibCode\KDistLiftSingleTraverseProof
+    % A singleton map can be swapped with any map \ACode{ϕ}:
+    % \LibCode\KDistLiftSingle
+    % % \LibCode\KDistLiftSingleProof
+    % By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the
+    % same statement for multiple map applications:
+    % \LibCode\KDistLiftSingleTraverse
+    % % \LibCode\KDistLiftSingleTraverseProof
   \item 
@@
   \noindent
-  \begin{minipage}{0.55\linewidth}
+  \begin{minipage}[t]{1.0\linewidth}
+    \begin{minipage}[t]{0.55\linewidth}
       \LibCode\KContextLookupI
     \end{minipage}
-  \begin{minipage}{0.45\linewidth}
+    \begin{minipage}[t]{0.45\linewidth}
       \LibCode\KContextLookupII
+    \end{minipage}
   \end{minipage}

### 2024-03-13 paper/paper.tex
Update.

@@
   \LibCode\KSyntax
-  The first three fields record the definitions of sorts, terms, and variable introduction.
+  \noindent The first three fields record the definitions of sorts, terms, and variable introduction.
   The last field records that variable introduction
@@
 
-  A renaming \ACode{S₁ \ADef{→ᵣ} S₂} maps variables from \ACode{S₁} to
+  \noindent A renaming \ACode{S₁ \ADef{→ᵣ} S₂} maps variables from \ACode{S₁} to
   variables from \ACode{S₂}.
@@
   \end{figure}
-  Figure~\ref{fig:map-ops} shows the usual operations for renamings and substitutions
-  generalized to maps.\footnote{
+  Figure~\ref{fig:map-ops} shows the usual operations for renamings and substitutions.\footnote{
     The definitions are included directly in the record module of the
@@
   \LibCode\KTraversal
-  The fields of this structure have the following meaning:
+  \noindent The fields of this structure have the following meaning:
   \begin{itemize}

### 2024-03-13 paper/paper.tex
Update.

@@
   \LibCode\KWkKit
-
-  The \ACode{\AField{wk-`/id}} axiom
+  \noindent The \ACode{\AField{wk-`/id}} axiom
   explains the \ACode{\AField{wk}} function by how it acts on \emph{terms}.
@@
   function by how it acts on \emph{variables}.
-
+  %
   This lemma is useful for proving extensional equalities between maps
@@
   \AField{wk-`/id} can be used to make further progress.
-
+  %
   The instantiations of the \ACode{\ACon{WkKit}} are straightforward:
@@
   dependencies.
-  % TODO
-  While composition can be derived, our approach has the benefit, that
-  we have gained a new level of generality, which enables us in the next
-  sections to unify type preservation of all substitution and renaming
-  operations into one lemma.
+  With the help of a general map composition, we can prove lemmas
+  about the interactions of general maps, which is crucial for proving a
+  type preservation lemma for general map application instead of individual
+  lemmas for renamings and substitutions.
 
@@
     and then apply \ACode{ϕ₂} to the result.
-    If \ACode{ϕ₁} is a renaming, then this result is another variable, but
-    if \ACode{ϕ₁} is a substitution, then this result is a term.
+    If \ACode{ϕ₁} is a renaming, this result is another variable, but
+    if \ACode{ϕ₁} is a substitution, this result is a term.
   \end{enumerate}
@@
 
-  But to abstract over \ACode{ϕ₁}, the \ACode{\ACon{Kit}} abstraction
+  \noindent But to abstract over \ACode{ϕ₁}, the \ACode{\ACon{Kit}} abstraction
   is not sufficient: while it allows us to abstract over what we are applying, i.e.
@@
   To fill this gap, we introduce a new abstraction that we call a
-  \ACode{\ACon{CKit}},
+  \emph{compose kit} (\ACode{\ACon{CKit}}),
   which provides an operation \ACode{\AField{\_\&/⋯\_}} that unifies
   \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}. This allows us to
-  define a general composition, which informally can be described as:
+  define a general composition as follows:
   \ExampleCode\KComposition
@@
   \ACode{\AField{\_\&/⋯\_}} in terms of \ACode{\AField{\_⋯\_}}, allowing
-  subsequent lemmas to make use of the lemmas we previously proved
-  about \ACode{\AField{\_⋯\_}}.
+  subsequent lemmas to make use of the lemmas that we have already proved
+  for \ACode{\AField{\_⋯\_}}.
 

### 2024-03-13 paper/paper.tex
Update.

@@
   allows the composition of arbitrary maps.
-  The core property of composition, which we call
-  \ACode{\AField{⋯-fusion}}, states that applying two maps \ACode{ϕ₁}
+  The core property of composition is the
+  \ACode{\AField{fusion}} lemma, which states that applying two maps \ACode{ϕ₁}
   and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same as
@@
   % $$
-  This property equips our framework with the power to
+  This property gives our framework the ability to
   reason about the application of multiple maps by reasoning about
@@
   While the composition operations can be defined independently of each other,
-  the \ACode{\AField{⋯-fusion}} lemma for two substitutions, depends on the
-  \ACode{\AField{⋯-fusion}} lemmas between a renaming and a substitution, which in
-  turn depend on the \ACode{\AField{⋯-fusion}} lemma between two renamings.
+  the \ACode{\AField{fusion}} lemma for two substitutions, depends on the
+  \ACode{\AField{fusion}} lemmas between a renaming and a substitution, which in
+  turn depend on the \ACode{\AField{fusion}} lemma between two renamings.
 
@@
   % \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
-  % \ACode{\ACon{CTraversal}} and \ACode{\AField{⋯-fusion}} are analogous to
+  % \ACode{\ACon{CTraversal}} and \ACode{\AField{fusion}} are analogous to
   % \ACode{\ACon{CTraversal}} and \ACode{\AField{\_⋯\_}}.
@@
   \ACode{\ACon{CTraversal}}, which provides the
-  \ACode{\AField{⋯-fusion}} lemma for general maps.
+  \ACode{\AField{fusion}} lemma for general maps.
 
@@
   four concrete \ACode{\ACon{CKit}} instances for us. This allows us to break
-  the dependencies between the different \ACode{\AField{⋯-fusion}} lemmas,
+  the dependencies between the different \ACode{\AField{fusion}} lemmas,
   similar to how \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}} allowed
@@
   We then define a structure,
-  which provides the \ACode{\AField{⋯-fusion}} lemma for arbitrary
+  which provides the \ACode{\AField{fusion}} lemma for arbitrary
   \ACode{\ACon{CKit}}s:
@@
     % % \LibCode\KCommLiftWeakenProof
-    % By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the same
+    % By applying \ACode{\AField{fusion}} to this lemma, we get the same
     % statement for multiple map applications:
@@
     % % \LibCode\KWeakenCancelsSingleProof
-    % By applying \ACode{\AField{⋯-fusion}} and \ACode{\AField{⋯-id}} to
+    % By applying \ACode{\AField{fusion}} and \ACode{\AField{⋯-id}} to
     % this lemma, we get the same statement for multiple map applications:
@@
     % % \LibCode\KDistLiftSingleProof
-    % By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the
+    % By applying \ACode{\AField{fusion}} to this lemma, we get the
     % same statement for multiple map applications:

### 2024-03-13 paper/paper.tex
Update.

@@
 
+  A compose kit \ACode{\ACon{CKit}~K₁~K₂~K₁⊔K₂}
+  describes the operations necessary for definining the composition of a
+  map from \ACode{K₁} with a map from \ACode{K₂} that results in a map from
+  \ACode{K₁⊔K₂}:
+  \LibCode\KCKit
+  \noindent The third parameter \ACode{K₁⊔K₂} can be seen as a functional
+  dependency\cite{DBLP:conf/esop/Jones00} and is determined
+  by the choice of \ACode{K₁} and \ACode{K₂}.
+  The fields of a compose kit have the following meaning:
+  \begin{itemize}
+  \item
+    The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
+    \ACode{x/t} (according to \ACode{K₁}) and a renaming or substitution
+    \ACode{ϕ} (according to \ACode{K₂}) and applies \ACode{ϕ} to \ACode{x/t} resulting in
+    a variable or term (according to \ACode{K₁⊔K₂}).
+    From this operation we derive map composition \ACode{\ADef{\_·ₖ\_}} as
+    shown in the previous subsection.
+    % TODO: use the notation def to the first time we use it?
+    We use the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}, if we need to
+    make it explicit that we are using the composition of some
+    specific \ACode{\ACon{CKit}} \ACode{C}.
+  \item 
+    The \ACode{\AField{\&/⋯-⋯}} lemma describes the behavior of
+    \ACode{\AField{\_\&/⋯\_}} in terms of \ACode{\AField{\_⋯\_}}, allowing
+    subsequent lemmas to make use of the lemmas that we have already proved
+    for \ACode{\AField{\_⋯\_}}.
+  \item 
+    The \ACode{\AField{\&/⋯-wk-↑}} lemma states that applying a map and then
+    weakening is the same as weakening first and then lifting the map over
+    the variable introduced by the weakening.
+    From this lemma, we can derive that lifting distributes over composition:
+    \LibCode\KDistLiftCompose
+    % \LibCode\KDistLiftComposeProof
+  \end{itemize}
+
+  A \ACode{\ACon{CTraversal}} provides a generic \ACode{\AField{fusion}} lemma
+  that works for any composition described by a \ACode{\ACon{CKit}}:
+  \LibCode\KCTraversal
+
+  Given a \ACode{\ACon{CTraversal}}, we can prove the lemmas about interactions
+  of multiple maps, which are required for substitution preserves typing:
+  \begin{itemize}
+  \item 
+    A map \ACode{ϕ} followed by a weakening is equivalent to a
+    weakening followed by \ACode{ϕ} that has been lifted over the weakened
+    variable:
+    \LibCode\KCommLiftWeakenTraverse
+
+    % A map \ACode{ϕ} followed by a weakening is equivalent to a
+    % weakening followed by \ACode{ϕ} that has been lifted over the weakened
+    % variable:
+    % \LibCode\KCommLiftWeaken
+    % % \LibCode\KCommLiftWeakenProof
+    % By applying \ACode{\AField{fusion}} to this lemma, we get the same
+    % statement for multiple map applications:
+    % \LibCode\KCommLiftWeakenTraverse
+    % % \LibCode\KCommLiftWeakenTraverseProof
+  \item 
+    A weakening followed by a singleton substitution cancel each other out:
+    \LibCode\KWeakenCancelsSingleTraverse
+
+    % A weakening followed by a singleton substitution cancel each other out:
+    % \LibCode\KWeakenCancelsSingle
+    % % \LibCode\KWeakenCancelsSingleProof
+    % By applying \ACode{\AField{fusion}} and \ACode{\AField{⋯-id}} to
+    % this lemma, we get the same statement for multiple map applications:
+    % \LibCode\KWeakenCancelsSingleTraverse
+    % % \LibCode\KWeakenCancelsSingleTraverseProof
+  \item 
+    A singleton map can be swapped with any map \ACode{ϕ}:
+    \LibCode\KDistLiftSingleTraverse
+    % A singleton map can be swapped with any map \ACode{ϕ}:
+    % \LibCode\KDistLiftSingle
+    % % \LibCode\KDistLiftSingleProof
+    % By applying \ACode{\AField{fusion}} to this lemma, we get the
+    % same statement for multiple map applications:
+    % \LibCode\KDistLiftSingleTraverse
+    % % \LibCode\KDistLiftSingleTraverseProof
+  \end{itemize}
+  We define \ACode{\ACon{CKit}} instances for each of the four
+  combinations between renamings and substitutions:
+  \LibCode\KCKitInstances
+  Interestingly, the \ACode{\ADef{Cₛ}} instance has another
+  \ACode{\ADef{CKit}} parameter.
+  This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
+  which yields the \ACode{\ADef{CKit}} between substitutions
+  and renamings, and then use this resulting \ACode{\ADef{CKit}}
+  as an argument for the case with two substitutions.
+  As shown below, Agda's instance resolution is up for the task:
+
+  % To get the \ACode{\ADef{CKit}} between substitutions and
+  % renamings, the argument \ACode{\ADef{Cᵣ}} is used.
+  % The resulting \ACode{\ADef{CKit}} is then again used as an argument
+  % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{CKit}}
+  % between two substitutions. As shown below, Agda's instance
+  % resolution is up for the task:
+  \noindent
+  \begin{minipage}[t]{0.5\linewidth}
+  \LibCode\KCKitInstancesConcreteI
+  \end{minipage}
+  \begin{minipage}[t]{0.5\linewidth}
+  \LibCode\KCKitInstancesConcreteII
+  \end{minipage}
+
+  % \LibCode\KCKitNotation
+
+  \subsection{An Abstraction for Composition}
+  \label{sec:composition:ckit}
+
   In this subsection, we define the \ACode{\ACon{CKit}}
@@
   The third parameter \ACode{K₁⊔K₂} can be seen as a functional
-  dependency\cite{DBLP:conf/esop/Jones00} and is completely determined
+  dependency\cite{DBLP:conf/esop/Jones00} and is determined
   by the choice of \ACode{K₁} and \ACode{K₂}.

### 2024-03-13 paper/paper.tex
Update.

@@
 
-  Given a \ACode{\ACon{CTraversal}}, we can prove the lemmas about interactions
-  of multiple maps, which are required for substitution preserves typing:
+  \noindent Given a \ACode{\ACon{CTraversal}}, we can prove the usual lemmas about interactions
+  of multiple maps:
   \begin{itemize}
@@
   \item 
-    A weakening followed by a singleton substitution cancel each other out:
+    A weakening followed by a singleton substitution act as an identity map:
     \LibCode\KWeakenCancelsSingleTraverse
@@
   \AppCode\FFusionProofInteresting
-  The interesting cases are:
+  We only show the interesting cases, which are:
   \begin{itemize}
@@
     variables, where we need to use the \ACode{\AField{\&/⋯-⋯}} lemma
-    provided by the \ACode{\ACon{CKit}}.
+    provided by the \ACode{\ACon{CKit}}; and
   \item
@@
   For System F, the expression sort \ACode{\ACon{𝕖}} maps to the type sort
-  \ACode{\ACon{𝕥}} and the sort \ACode{\ACon{𝕥}} maps to the kind sort
+  \ACode{\ACon{𝕥}}, and the sort \ACode{\ACon{𝕥}} maps to the kind sort
   \ACode{\ACon{𝕜}}.
@@
   by the typing context to use all variables bound previously in the typing context.
-  This is achieved by the \ACode{drop-∈} function, which cuts of all
+  This is achieved by the \ACode{drop-∈} function, which cuts off all
   sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.

### 2024-03-13 paper/paper.tex
Update.

@@
     shown in the previous subsection.
-    % TODO: use the notation def to the first time we use it?
-    We use the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}, if we need to
-    make it explicit that we are using the composition of some
-    specific \ACode{\ACon{CKit}} \ACode{C}.
+    % We use the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}, if we need to
+    % make it explicit that we are using the composition of some
+    % specific \ACode{\ACon{CKit}} \ACode{C}.
   \item 
@@
   \end{itemize}
-  To make it easier to explicitly specify a \ACode{\ACon{CKit}},
-  we define the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}
-  to stand for \ACode{ϕ₁ \ADef{·ₖ} ϕ₂} for some
-  \ACode{\ACon{CKit}} \ACode{C}.
+  % To make it easier to explicitly specify a \ACode{\ACon{CKit}},
+  % we define the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}
+  % to stand for \ACode{ϕ₁ \ADef{·ₖ} ϕ₂} for some
+  % \ACode{\ACon{CKit}} \ACode{C}.
   % \LibCode\KCKitNotation
@@
     have kind \ACode{\ACon{★}}. This is sufficient for System F as types
-    are automatically well-typed due to intrinsic scoping.
+    are automatically well-kinded due to intrinsic scoping.
   \end{itemize}
@@
     \LibCode\KMapTyping
-    \ACode{Γ₂ \ADef{∋*/⊢*} ϕ \ADef{∶} Γ₁} states that \ACode{ϕ} is a map that takes
+    \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒ₖ} Γ₂} states that \ACode{ϕ} is a map that takes
     terms in \ACode{Γ₁} to terms in \ACode{Γ₂}.
@@
   \end{itemize}
-  To make it easier to explicitly pass typing kits,
-  we define the notation \ACode{Γ₂ \ADef{∋*/⊢*[} TK \ADef{]} ϕ \ADef{∶} Γ₁}
-  to stand for \ACode{Γ₂ \ADef{∋*/⊢*} ϕ \ADef{∶} Γ₁} given some
-  \ACode{\ACon{TKit}} \ACode{TK}.
+  % To make it easier to explicitly pass typing kits,
+  % we define the notation \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒[} TK \ADef{]} Γ₂}
+  % to stand for \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒ₖ} Γ₂} of some
+  % \ACode{\ACon{TKit}} \ACode{TK}.
   % \LibCode\KTypingNotation

### 2024-03-13 paper/paper.tex
Update.

@@
 
-  \subsection{An Abstraction for Composition}
-  \label{sec:composition:ckit}
-
-  In this subsection, we define the \ACode{\ACon{CKit}}
-  abstraction, which provides the \ACode{\AField{\_\&/⋯\_}} operation
-  that is necessary for defining composition, and the
-  \ACode{\ACon{CTraversal}}, which provides the
-  \ACode{\AField{fusion}} lemma for general maps.
-
-  Similarly as it was the case for \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}},
-  the lemma of the \ACode{\ACon{CTraversal}} needs to be proved for
-  an abstract \ACode{\ACon{CKit}} and in return defines the
-  four concrete \ACode{\ACon{CKit}} instances for us. This allows us to break
-  the dependencies between the different \ACode{\AField{fusion}} lemmas,
-  similar to how \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}} allowed
-  us to break the dependency between traversals with renamings and substitutions.
-
-  For this purpose, the \ACode{\ACon{CKit}} abstraction, needs
-  to satisfy two requirements:
-  it needs to provide us with enough machinery to define composition, but
-  it also needs to provide us with enough axioms that we are able to instantiate
-  the \ACode{\ACon{CTraversal}}. It is defined as followed:
-  \LibCode\KCKit
-  The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
-  (according to \ACode{K₁}) and a renaming or substitution
-  (according to \ACode{K₂}) and performs the map traversal returning
-  a variable or term (according to \ACode{K₁⊔K₂}).
-  The third parameter \ACode{K₁⊔K₂} can be seen as a functional
-  dependency\cite{DBLP:conf/esop/Jones00} and is determined
-  by the choice of \ACode{K₁} and \ACode{K₂}.
-
-  The \ACode{\AField{\&/⋯-⋯}} lemma describes the behavior of
-  \ACode{\AField{\_\&/⋯\_}} in terms of \ACode{\AField{\_⋯\_}}, allowing
-  subsequent lemmas to make use of the lemmas that we have already proved
-  for \ACode{\AField{\_⋯\_}}.
-
-  The \ACode{\AField{\&/⋯-wk-↑}} lemma states that applying a map and then
-  weakening is the same as weakening first and then lifting the map over
-  the variable introduced by the weakening.
-
-  Similar as for the \ACode{\ACon{Kit}}, we derive further definitions
-  from the axioms of a \ACode{\ACon{CKit}}:
-  \begin{itemize}
-  \item
-    We define map composition as motivated by the previous subsection:
-    \LibCode\KComposition
+  % \subsection{An Abstraction for Composition}
+  % \label{sec:composition:ckit}
+
+  % In this subsection, we define the \ACode{\ACon{CKit}}
+  % abstraction, which provides the \ACode{\AField{\_\&/⋯\_}} operation
+  % that is necessary for defining composition, and the
+  % \ACode{\ACon{CTraversal}}, which provides the
+  % \ACode{\AField{fusion}} lemma for general maps.
+
+  % Similarly as it was the case for \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}},
+  % the lemma of the \ACode{\ACon{CTraversal}} needs to be proved for
+  % an abstract \ACode{\ACon{CKit}} and in return defines the
+  % four concrete \ACode{\ACon{CKit}} instances for us. This allows us to break
+  % the dependencies between the different \ACode{\AField{fusion}} lemmas,
+  % similar to how \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}} allowed
+  % us to break the dependency between traversals with renamings and substitutions.
+
+  % For this purpose, the \ACode{\ACon{CKit}} abstraction, needs
+  % to satisfy two requirements:
+  % it needs to provide us with enough machinery to define composition, but
+  % it also needs to provide us with enough axioms that we are able to instantiate
+  % the \ACode{\ACon{CTraversal}}. It is defined as followed:
+  % \LibCode\KCKit
+  % The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
+  % (according to \ACode{K₁}) and a renaming or substitution
+  % (according to \ACode{K₂}) and performs the map traversal returning
+  % a variable or term (according to \ACode{K₁⊔K₂}).
+  % The third parameter \ACode{K₁⊔K₂} can be seen as a functional
+  % dependency\cite{DBLP:conf/esop/Jones00} and is determined
+  % by the choice of \ACode{K₁} and \ACode{K₂}.
+
+  % The \ACode{\AField{\&/⋯-⋯}} lemma describes the behavior of
+  % \ACode{\AField{\_\&/⋯\_}} in terms of \ACode{\AField{\_⋯\_}}, allowing
+  % subsequent lemmas to make use of the lemmas that we have already proved
+  % for \ACode{\AField{\_⋯\_}}.
+
+  % The \ACode{\AField{\&/⋯-wk-↑}} lemma states that applying a map and then
+  % weakening is the same as weakening first and then lifting the map over
+  % the variable introduced by the weakening.
+
+  % Similar as for the \ACode{\ACon{Kit}}, we derive further definitions
+  % from the axioms of a \ACode{\ACon{CKit}}:
+  % \begin{itemize}
   % \item
-  %   We show that the \ACode{\AField{\_\&/⋯\_}} operator behaves like
-  %   \ACode{\AField{\_\&\_}} when applied to variables:
-  %   \LibCode\KCKitAp
-  %   % \LibCode\KCKitApProof
-
-  %   % The \ACode{\AField{`/id}} on the left side is from \ACode{K₁⊔K₂}, whereas
-  %   % the \ACode{\AField{`/id}} on the right side is from \ACode{K₂}.
-  \item
-    We show that lifting \ACode{\AField{\_↑\_}} distributes over
-    composition:
-    \LibCode\KDistLiftCompose
-    % \LibCode\KDistLiftComposeProof
-  \end{itemize}
-  % To make it easier to explicitly specify a \ACode{\ACon{CKit}},
-  % we define the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}
-  % to stand for \ACode{ϕ₁ \ADef{·ₖ} ϕ₂} for some
-  % \ACode{\ACon{CKit}} \ACode{C}.
-  % \LibCode\KCKitNotation
-
-  We then define a structure,
-  which provides the \ACode{\AField{fusion}} lemma for arbitrary
-  \ACode{\ACon{CKit}}s:
-  \LibCode\KCTraversal
-
-  Similar as for the \ACode{\ACon{Traversal}}, we derive further
-  definitions from the axioms of a \ACode{\ACon{CTraversal}}:
-  \begin{itemize}
-  \item 
-    A map \ACode{ϕ} followed by a weakening is equivalent to a
-    weakening followed by \ACode{ϕ} that has been lifted over the weakened
-    variable:
-    \LibCode\KCommLiftWeakenTraverse
-
+  %   We define map composition as motivated by the previous subsection:
+  %   \LibCode\KComposition
+  % % \item
+  % %   We show that the \ACode{\AField{\_\&/⋯\_}} operator behaves like
+  % %   \ACode{\AField{\_\&\_}} when applied to variables:
+  % %   \LibCode\KCKitAp
+  % %   % \LibCode\KCKitApProof
+
+  % %   % The \ACode{\AField{`/id}} on the left side is from \ACode{K₁⊔K₂}, whereas
+  % %   % the \ACode{\AField{`/id}} on the right side is from \ACode{K₂}.
+  % \item
+  %   We show that lifting \ACode{\AField{\_↑\_}} distributes over
+  %   composition:
+  %   \LibCode\KDistLiftCompose
+  %   % \LibCode\KDistLiftComposeProof
+  % \end{itemize}
+  % % To make it easier to explicitly specify a \ACode{\ACon{CKit}},
+  % % we define the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}
+  % % to stand for \ACode{ϕ₁ \ADef{·ₖ} ϕ₂} for some
+  % % \ACode{\ACon{CKit}} \ACode{C}.
+  % % \LibCode\KCKitNotation
+
+  % We then define a structure,
+  % which provides the \ACode{\AField{fusion}} lemma for arbitrary
+  % \ACode{\ACon{CKit}}s:
+  % \LibCode\KCTraversal
+
+  % Similar as for the \ACode{\ACon{Traversal}}, we derive further
+  % definitions from the axioms of a \ACode{\ACon{CTraversal}}:
+  % \begin{itemize}
+  % \item 
   %   A map \ACode{ϕ} followed by a weakening is equivalent to a
@@
   %   variable:
-    % \LibCode\KCommLiftWeaken
-    % % \LibCode\KCommLiftWeakenProof
-    % By applying \ACode{\AField{fusion}} to this lemma, we get the same
-    % statement for multiple map applications:
   %   \LibCode\KCommLiftWeakenTraverse
-    % % \LibCode\KCommLiftWeakenTraverseProof
-  \item 
-    A weakening followed by a singleton substitution cancel each other out:
-    \LibCode\KWeakenCancelsSingleTraverse
-
+
+  %   % A map \ACode{ϕ} followed by a weakening is equivalent to a
+  %   % weakening followed by \ACode{ϕ} that has been lifted over the weakened
+  %   % variable:
+  %   % \LibCode\KCommLiftWeaken
+  %   % % \LibCode\KCommLiftWeakenProof
+  %   % By applying \ACode{\AField{fusion}} to this lemma, we get the same
+  %   % statement for multiple map applications:
+  %   % \LibCode\KCommLiftWeakenTraverse
+  %   % % \LibCode\KCommLiftWeakenTraverseProof
+  % \item 
   %   A weakening followed by a singleton substitution cancel each other out:
-    % \LibCode\KWeakenCancelsSingle
-    % % \LibCode\KWeakenCancelsSingleProof
-    % By applying \ACode{\AField{fusion}} and \ACode{\AField{⋯-id}} to
-    % this lemma, we get the same statement for multiple map applications:
   %   \LibCode\KWeakenCancelsSingleTraverse
-    % % \LibCode\KWeakenCancelsSingleTraverseProof
-  \item 
-    A singleton map can be swapped with any map \ACode{ϕ}:
-    \LibCode\KDistLiftSingleTraverse
+
+  %   % A weakening followed by a singleton substitution cancel each other out:
+  %   % \LibCode\KWeakenCancelsSingle
+  %   % % \LibCode\KWeakenCancelsSingleProof
+  %   % By applying \ACode{\AField{fusion}} and \ACode{\AField{⋯-id}} to
+  %   % this lemma, we get the same statement for multiple map applications:
+  %   % \LibCode\KWeakenCancelsSingleTraverse
+  %   % % \LibCode\KWeakenCancelsSingleTraverseProof
+  % \item 
   %   A singleton map can be swapped with any map \ACode{ϕ}:
-    % \LibCode\KDistLiftSingle
-    % % \LibCode\KDistLiftSingleProof
-    % By applying \ACode{\AField{fusion}} to this lemma, we get the
-    % same statement for multiple map applications:
   %   \LibCode\KDistLiftSingleTraverse
-    % % \LibCode\KDistLiftSingleTraverseProof
-  \item 
-    We define \ACode{\ACon{CKit}} instances for each of the four
-    combinations between renamings and substitutions:
-    \LibCode\KCKitInstances
-    Interestingly, the \ACode{\ADef{Cₛ}} instance has another
-    \ACode{\ADef{CKit}} parameter.
-    This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
-    which yields the \ACode{\ADef{CKit}} between substitutions
-    and renamings, and then use this resulting \ACode{\ADef{CKit}}
-    as an argument for the case with two substitutions.
-    As shown below, Agda's instance resolution is up for the task:
-
-    % To get the \ACode{\ADef{CKit}} between substitutions and
-    % renamings, the argument \ACode{\ADef{Cᵣ}} is used.
-    % The resulting \ACode{\ADef{CKit}} is then again used as an argument
-    % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{CKit}}
-    % between two substitutions. As shown below, Agda's instance
-    % resolution is up for the task:
-    \begin{minipage}[t]{0.5\linewidth}
-    \LibCode\KCKitInstancesConcreteI
-    \end{minipage}
-    \begin{minipage}[t]{0.5\linewidth}
-    \LibCode\KCKitInstancesConcreteII
-    \end{minipage}
-
-  \end{itemize}
+  %   % A singleton map can be swapped with any map \ACode{ϕ}:
+  %   % \LibCode\KDistLiftSingle
+  %   % % \LibCode\KDistLiftSingleProof
+  %   % By applying \ACode{\AField{fusion}} to this lemma, we get the
+  %   % same statement for multiple map applications:
+  %   % \LibCode\KDistLiftSingleTraverse
+  %   % % \LibCode\KDistLiftSingleTraverseProof
+  % \item 
+  %   We define \ACode{\ACon{CKit}} instances for each of the four
+  %   combinations between renamings and substitutions:
+  %   \LibCode\KCKitInstances
+  %   Interestingly, the \ACode{\ADef{Cₛ}} instance has another
+  %   \ACode{\ADef{CKit}} parameter.
+  %   This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
+  %   which yields the \ACode{\ADef{CKit}} between substitutions
+  %   and renamings, and then use this resulting \ACode{\ADef{CKit}}
+  %   as an argument for the case with two substitutions.
+  %   As shown below, Agda's instance resolution is up for the task:
+
+  %   % To get the \ACode{\ADef{CKit}} between substitutions and
+  %   % renamings, the argument \ACode{\ADef{Cᵣ}} is used.
+  %   % The resulting \ACode{\ADef{CKit}} is then again used as an argument
+  %   % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{CKit}}
+  %   % between two substitutions. As shown below, Agda's instance
+  %   % resolution is up for the task:
+  %   \begin{minipage}[t]{0.5\linewidth}
+  %   \LibCode\KCKitInstancesConcreteI
+  %   \end{minipage}
+  %   \begin{minipage}[t]{0.5\linewidth}
+  %   \LibCode\KCKitInstancesConcreteII
+  %   \end{minipage}
+
+  % \end{itemize}
 
@@
     used \ACode{\AField{`/id}} or \ACode{\AField{\_↑\_}}, we now use their
-    preservation lemmas \ACode{\AField{⊢`/id}} or \ACode{\AField{\_∋↑/⊢↑\_}}.
+    preservation lemmas \ACode{\AField{⊢`/id}} or \ACode{\AField{\_⊢↑\_}}.
   \item

### 2024-03-14 paper/paper.tex
Update.

@@
 
-  A \ACode{\ACon{CTraversal}} provides a generic \ACode{\AField{fusion}} lemma
-  that works for any composition described by a \ACode{\ACon{CKit}}:
+  A \ACode{\ACon{CTraversal}} provides a \ACode{\AField{fusion}} lemma
+  that works for the composition of any \ACode{\ACon{CKit}}:
   \LibCode\KCTraversal
@@
   \end{itemize}
-  We define \ACode{\ACon{CKit}} instances for each of the four
-  combinations between renamings and substitutions:
+  Similarly, as it was the case for the \ACode{\ADef{Kit}} and \ACode{\ADef{Traversal}}
+  structures, the idea is that we instantiate the \ACode{\ADef{CTraversal}} for
+  our object language, and in return the framework defines the concrete \ACode{\ADef{CKit}}
+  instances for us:
   \LibCode\KCKitInstances
@@
   \LibCode\KTypeSorts
-  For System F the instantiation is
+  For System F, the instantiation is
   \AppCode\FTypes
@@
     For such sorts, we can simply use arbitrary sort types, as the
-    formalization will never use them.
+    formalization will have no typing rules that use them.
   \end{enumerate}
@@
   \begin{minipage}[t]{1.0\linewidth}
+    \begin{minipage}[t]{0.45\linewidth}
+      \LibCode\KContextLookupII
+    \end{minipage}
     \begin{minipage}[t]{0.55\linewidth}
       \LibCode\KContextLookupI
-    \end{minipage}
-    \begin{minipage}[t]{0.45\linewidth}
-      \LibCode\KContextLookupII
     \end{minipage}

### 2024-03-14 paper/paper.tex
Update.

@@
   sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
+  Contexts \ACode{Γ} can be extended by a type \ACode{t} via \ACode{t \ADef{∷ᵗ} Γ}.
 
@@
   Now that we have a notion of types and type context, we are ready to define
-  the multi-sorted typing relation for System F, incorporating both
+  the multi-sorted typing relation for System F, which describes both
   typing and kinding:
@@
   \item
-    the typing \ACode{\ACon{⊢λ}} of the lambda abstraction, which
-    weakens the type of the codomain \ACode{t₂}.
+    the typing \ACode{\ACon{⊢λ}} of a lambda abstraction, which
+    weakens codomain type \ACode{t₂}.
     This is necessary, because multi-sorted syntax allows types
@@
   \ACode{\ACon{Traversal}}, but instead of defining the application of
-  maps, it defines the type preservation of applying maps:
+  % maps, it defines that the application of typed maps preserves term types:
+  maps, it defines that the application of a typed map to a typed term yields a typed term:
   \LibCode\KTTraversal
@@
 
-  We believe restriction 3 can be lifted partially to allow for
-  example the encoding of linear types that requires context
-  splitting.
+  % We believe restriction 3 can be lifted partially to allow for
+  % example the encoding of linear types that requires context
+  % splitting.
 
@@
   \label{sec:conclusion}
-  We want to focus our future work in the following directions:
-  \begin{itemize}
-  \item
-    Enabling more object languages by allowing subterms with less
-    restrictions on the form of \ACode{S}, allowing recursive
-    occurences in container types like lists and finite functions, and
-    adding support for substructural types as in
-    \cite{DBLP:journals/corr/abs-2005-02247}.
-  \item
-    Extending the class of typing relations covered by
-    the \ACode{\ACon{TTraversal}} abstraction and identifying
-    a subclass for which it can be derived.
-  \item
-    Developing a decision procedure for equational
-    substitution lemmas based on $\sigma$-calculus as in Autosubst
-    \cite{DBLP:conf/itp/SchaferTS15, DBLP:conf/cpp/StarkSK19}.
-  \item
-    Developing abstractions and derived lemmas beyond substition and
-    type preservation, e.g. for confluence or denotational semantics
-    as in \cite{DBLP:journals/pacmpl/AllaisA0MM18}.
-  \end{itemize}
+  We have presented a framework for deriving untyped substitution and its properties, and
+  and abstraction for proving type preservation of substitution with a single lemma.
+
+  Using extrinsic typing allowed us to unify the renamings and
+  substitutions for variables of arbitrary many syntactic categories.
+
+  Furthermore, extrinsic typing allowed us treat untyped substitution
+  and its properties separately from typing and type preservation.
+  This has the benefit that the former can be derived automatically for virtually
+  any object language, while the latter is more domain specific and
+  can be handled independently.
+
+  We believe our framework is a viable alternative to intrinsic
+  typing, as the duplication is automated away.
+
+  We believe our framework is particularly useful for polymorphic languages,
+  where intrinsic typing requires necessarily to deal with multiple substitutions
+  and their interactions. Furthermore, in more complex scenarios, intrinsic typing
+  can easily lead to subst-hell, which is avoided with extrinsic typing.
+
+  We belive the core limitation of our framework are the typing kits, which
+  require a classical ternary typing relation. However, this problem is hard
+  to avoid as typing relations can become arbitrarily complicated.
+  To adapt the framework to more complicated type systems, only the typing
+  kits and typing traversals need to be extended, which we have successfully
+  done for linear typing ala Wood and Atkey\cite{DBLP:journals/corr/abs-2005-02247} and
+  general substructural typing ala Licata et al\cite{DBLP:conf/rta/LicataSR17}.
 

### 2024-03-14 paper/paper.tex
Update.

@@
   kind polymorphism can easily end up with 24 slightly changed copies of
-  the whole substitution machinery, which for de Bruijn indices tends to
-  include at least a type for parallel substitution, and operations for
-  application, extension, weakening, lifting, singleton substitution,
-  extensional equality, and an identity substitution - and this is not
-  counting the slightly changed copies of interaction lemmas!
+  the whole substitution machinery!
   Clearly, this situation is in need of automation.
 
-  Our framework approaches this problem using a combination of
+  Our framework approaches this problem by using a combination of
   abstractions and reflection in the context of extrinsically typed,
@@
   abstraction to prove type preservation for all substitution operations
-  using a single lemma. For System F, our framework allows to prove
+  with only a single lemma. For System F, our framework allows to prove
   subject reduction with only a single handwritten lemma for type
@@
   The rest of this paper will introduce our framework using System F
-  as a running example for a type soundness proof.
+  as a running example for a substitution-preserves-typing proof.
 
@@
   and in the supplementary material.
-  The full framework is available online.\footnote{See supplementary material. In the final version, this text will be replaced by a link to github.}
-
-  The rest of this paper is structured as followed:
+  The full framework is available on Github: \textcolor{blue}{\url{https://github.com/m0rphism/kitty}}
+  % The full framework is available online.\footnote{See supplementary material. In the final version, this text will be replaced by a link to github.}
+
+  The rest of this paper is structured as follows:
   Section~\ref{sec:syntax} introduces the multi-sorted syntax and compares it to the
@@
   compositions.
-  Section~\ref{sec:types} shows how to introduce the notion of types
-  for multi-sorted syntax and how to define typing contexts.
+  Section~\ref{sec:types} shows how to define types
+  and typing contexts for multi-sorted syntax.
   Section~\ref{sec:typing} shows how to define a multi-sorted typing relation,
@@
   Section~\ref{sec:related} discusses related work.
-  Section~\ref{sec:conclusion} concludes with future work.
+  Section~\ref{sec:conclusion} concludes.
 
@@
     syntax, we call \emph{multi-sorted syntax};
-  \item
-    an adaptation of McBride's
-    kit-abstraction\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
-    for unifying renamings and substitutions to multi-sorted syntax;
+  % \item
+  %   an adaptation of McBride's
+  %   kit-abstraction\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
+  %   for unifying renamings and substitutions to multi-sorted syntax;
   \item
@@
   \item
-    a novel abstraction for typing relations, unifying type
+    a novel abstraction for typing, unifying type
     preservation of renaming and substitution;

### 2024-03-14 paper/paper.tex
Update.

@@
 
-  \subsection{Fibrational}
-  \cite{DBLP:conf/rta/LicataSR17}
+  % \subsection{Fibrational}
+  % \cite{DBLP:conf/rta/LicataSR17}
 
   \subsection{Pure Type Systems}
-  \cite{DBLP:conf/esop/BarendregtH90, berarditowards, terlouw1989een}
+  Pure Type Systems\cite{DBLP:conf/esop/BarendregtH90, berarditowards, terlouw1989een}
+  describe a class of typed lambda calculi parameterized over a set of
+  sorts, dependencies between sorts, and rules for quantification.
+
+  While pure type systems may seem very similar to multi-sorted
+  syntax, they are actually quite different:
+  \begin{itemize}
+  \item 
+    In multi-sorted syntax, sorts describe syntactic categories of terms. Terms
+    of different sorts are kept syntactically different, e.g. the set of expressions
+    \ACode{S ⊢ 𝕖} and types \ACode{S ⊢ 𝕥} in our System F example.
+  \item 
+    In pure type systems, sorts are universes, e.g. as
+    in Agda a type has itself type \ACode{Set} or a kind has type \ACode{Set₁}.
+    Terms which have different sorts as types, do still belong to the
+    same syntactic category.
+  \end{itemize}
+
+  We can use our framework to model pure type systems, but the sorts of pure type systems
+  do not correspond to the sorts of the multi-sorted syntax. The multi-sorted syntax
+  has just a single sort that describes all terms, as it is common for
+  dependently typed languages.
 

### 2024-03-14 paper/paper.tex
Update.

@@
 
-  The rest of this paper will introduce our framework using System F
+  The rest of this paper introduces our framework using System F
   as a running example for a substitution-preserves-typing proof.
@@
   definitions and lemmas, but omit some proofs for the sake of
-  space. The omitted proofs can be found in Appendix~\ref{sec:proofs}
-  and in the supplementary material.
-  The full framework is available on Github: \textcolor{blue}{\url{https://github.com/m0rphism/kitty}}
+  space. The omitted proofs can be found in our supplementary material,
+  which includes the simplified framework (504 lines of code) and the
+  System F formalization (228 lines of code).
+  The full framework including the case studies is available on Github: \textcolor{blue}{\url{https://github.com/m0rphism/kitty}}
   % The full framework is available online.\footnote{See supplementary material. In the final version, this text will be replaced by a link to github.}
@@
   free expression variables \ACode{m}.
-  Variables \ACode{\ACon{`\_}} are represented as DeBruijn-indices, where
+  Variables \ACode{\ACon{`\_}} are represented as de Bruijn indices, where
   \ACode{\ACon{Fin} n} is the type of \ACode{n} elements.
@@
 
-  The term type \ACode{S \ACon{⊢} s} is then indexed by its sort
+  The term type \ACode{S \ACon{⊢} s} is indexed by its sort
   \ACode{s} and the sorts of its free variables \ACode{S}.
@@
   As it is typical in intrinsic typing, variables are represented as
-  typed (in our case sorted) DeBruijn indices \ACode{S \ACon{∋} s}, i.e.\
+  typed (in our case sorted) de Bruijn indices \ACode{S \ACon{∋} s}, i.e.\
   values of the usual proof-relevant list-membership relation:
@@
   The regularity of the multi-sorted syntax makes it easy to define a
-  structure for arbitrary syntaxes, i.e. with arbitrarily many syntactic categories and variable types:
+  structure for arbitrary syntaxes, i.e.\ syntaxes with arbitrarily
+  many syntactic categories and variable types:
   \LibCode\KSyntax
@@
   % \label{sec:maps:renamings}
-  % A problem with mechanizing languages with DeBruijn-indices is that
+  % A problem with mechanizing languages with de bruijn indices is that
   % the operation of applying a substitution \ACode{σ} to a term
@@
   McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
-  introduced the \emph{Kit} abstraction.%
+  introduced the \emph{kit} abstraction.
   % \footnote{
@@
   % between variables and terms:
-  A \ACode{Kit} is a structure that allows to abstract over whether
+  A kit is a structure that allows to abstract over whether
   something is a term or a variable. The intention is to instantiate
@@
   \item 
-    \ACode{\AField{wk}} shifts the DeBruin-Indices in a \ACode{S ∋/⊢ s}.
+    \ACode{\AField{wk}} shifts the de Bruin indices in a variable or term.
     The new, unused variable \ACode{\ACon{zero}} can assume any sort \ACode{s'}.
@@
   \end{figure}
-  Figure~\ref{fig:map-ops} shows the usual operations for renamings and substitutions.\footnote{
+  Figure~\ref{fig:map-ops} shows the usual operations for renamings and substitutions.
   The definitions are included directly in the record module of the
   Kit, so they are implicitly parameterized over a kit.
-  }
   The type \ACode{S₁ \ADef{→ₖ} S₂} unifies
@@
 
-  Then we prove that applying an identity map does not change the term:
+  We then prove that applying an identity map does not change the term:
   \AppCode\FTraversalId
@@
     \ACon{\AField{id/`}} from the identity map meets the \ACon{\AField{`/id}}
-    from the traversal operation, so we use \ACon{\AField{`/`-is-`}}.
+    from the traversal operation, so we need to use \ACon{\AField{`/`-is-`}}.
   \item
@@
   applying their composition \ACode{(ϕ₁~\AField{·ₖ}~ϕ₂)} to \ACode{t},
-  i.e. \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
+  i.e.\ \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
   % $$
@@
   the \ACode{\AField{fusion}} lemma for two substitutions, depends on the
-  \ACode{\AField{fusion}} lemmas between a renaming and a substitution, which in
-  turn depend on the \ACode{\AField{fusion}} lemma between two renamings.
+  \ACode{\AField{fusion}} lemmas for a renaming and a substitution, which in
+  turn depend on the \ACode{\AField{fusion}} lemma for two renamings.
 
@@
   a renaming or a substitution, it does not allow us to
-  abstract over what we are applying it to, i.e. a variable or a term.
+  abstract over what we are applying it to, i.e.\ a variable or a term.
   For the latter we have two distinct operations \ACode{\ADef{\_\&\_}}
@@
   our object language, and in return the framework defines the concrete \ACode{\ADef{CKit}}
-  instances for us:
+  instances for us. Hence, we define them in the record module of \ACode{\ADef{CTraversal}}:
   \LibCode\KCKitInstances
-  Interestingly, the \ACode{\ADef{Cₛ}} instance has another
-  \ACode{\ADef{CKit}} parameter.
-  This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
-  which yields the \ACode{\ADef{CKit}} between substitutions
-  and renamings, and then use this resulting \ACode{\ADef{CKit}}
-  as an argument for the case with two substitutions.
-  As shown below, Agda's instance resolution is up for the task:
+  \noindent
+  \ACode{\ADef{Cᵣ}} is the compose kit between a renaming and another kit \ACode{K}.
+  \ACode{\ADef{Cₛ}} is the compose kit between a substitution and another kit \ACode{K},
+  and requires that we already know how to compose a \ACode{K}-map with a renaming.
+  The following verifies that \ACode{\ADef{Cᵣ}} and \ACode{\ADef{Cₛ}} indeed get us all four compositions:
+
+  \noindent
+  \begin{minipage}[t]{0.244\linewidth}
+  \ExampleCode\KCKitInstancesConcreteI
+  \end{minipage}
+  \begin{minipage}[t]{0.244\linewidth}
+  \ExampleCode\KCKitInstancesConcreteII
+  \end{minipage}
+  \begin{minipage}[t]{0.244\linewidth}
+  \ExampleCode\KCKitInstancesConcreteIII
+  \end{minipage}
+  \begin{minipage}[t]{0.244\linewidth}
+  \ExampleCode\KCKitInstancesConcreteIV
+  \end{minipage}
 
@@
   % resolution is up for the task:
-  \noindent
-  \begin{minipage}[t]{0.5\linewidth}
-  \LibCode\KCKitInstancesConcreteI
-  \end{minipage}
-  \begin{minipage}[t]{0.5\linewidth}
-  \LibCode\KCKitInstancesConcreteII
-  \end{minipage}
+
+  % \noindent
+  % \begin{minipage}[t]{0.5\linewidth}
+  % \LibCode\KCKitInstancesConcreteI
+  % \end{minipage}
+  % \begin{minipage}[t]{0.5\linewidth}
+  % \LibCode\KCKitInstancesConcreteII
+  % \end{minipage}
 
@@
   instantiations related to all structures for untyped substitition,
-  i.e. \ACode{\ACon{Syntax}}, \ACode{\ACon{Traversal}} and
+  i.e.\ \ACode{\ACon{Syntax}}, \ACode{\ACon{Traversal}} and
   \ACode{\ACon{CTraversal}}.
@@
   what we described as \emph{unsorted syntax} in
-  Section~\ref{sec:syntax}, i.e. different syntactic categories are
+  Section~\ref{sec:syntax}, i.e.\ different syntactic categories are
   described by different types with different amount of indices for
@@
   \item 
-    In pure type systems, sorts are universes, e.g. as
-    in Agda a type has itself type \ACode{Set} or a kind has type \ACode{Set₁}.
+    In pure type systems, sorts are universe types, e.g.\ like
+    \ACode{Set} and \ACode{Set₁} in Agda.
     Terms which have different sorts as types, do still belong to the

### 2024-03-14 paper/paper.tex
Update.

@@
   Section~\ref{sec:types} shows how to define types
-  and typing contexts for multi-sorted syntax.
+  and type contexts for multi-sorted syntax.
   Section~\ref{sec:typing} shows how to define a multi-sorted typing relation,
@@
     a reflection algorithm, representation independence for
-    substitutions and typing contexts, heterogeneous equality between
+    substitutions and type contexts, heterogeneous equality between
     renamings and substitutions, and absence of postulated axioms.
@@
   \label{sec:types}
-  % To define typing contexts and a typing relation, we need to know
+  % To define type contexts and a typing relation, we need to know
   % which sorts of terms play the role of types for other sorts of terms.
@@
   % and types are represented as terms, but with different sorts.
-  % To define typing contexts and a typing relation, we need to record for each term
+  % To define type contexts and a typing relation, we need to record for each term
   % from 
@@
 
-  \subsection{Typing Context}
-  Equipped with a notion of types, we define a typing context as a
+  \subsection{Type Contexts}
+  Equipped with a notion of types, we define a type context as a
   function that maps variables to types of the same sort.
@@
   As we want our framework to support dependent types, we allow a type returned
-  by the typing context to use all variables bound previously in the typing context.
+  by the type context to use all variables bound previously in the type context.
   This is achieved by the \ACode{drop-∈} function, which cuts off all
@@
 
-  To use a type returned from a typing context, it needs to be
+  To use a type returned from a type context, it needs to be
   weakened appropriately through the \ACode{\ADef{wk-telescope}} function:
@@
   \item
-    the variable typing constructor, which covers both expression- and
+    the variable typing constructor \ACode{\ACon{⊢`}}, which covers both expression- and
     type-variables, analogously to the variable term constructor;
@@
     the typing \ACode{\ACon{⊢λ}} of a lambda abstraction, which
-    weakens codomain type \ACode{t₂}.
+    weakens the codomain type \ACode{t₂}.
     This is necessary, because multi-sorted syntax allows types
     to depend on expressions, so the typing derivation for \ACode{e}
-    has to account for a variable, which is not used by the type.
+    has to account for a variable, which is not used by the type; and
   \item

### 2024-03-14 paper/paper.tex
Update.

@@
   compositions.
-  Section~\ref{sec:types} shows how to define types
-  and type contexts for multi-sorted syntax.
-  Section~\ref{sec:typing} shows how to define a multi-sorted typing relation,
-  and an abstraction to unify type preservation lemmas for renamings and substitutions.
+  Section~\ref{sec:types} shows how to define types,
+  type contexts, and typing relations,
+  and presents an abstraction for unifying type preservation lemmas for renamings and substitutions.
   % Section~\ref{sec:subjectreduction} introduces a small-step
@@
 
-  \section{Types \& Contexts}
+  \section{Types \& Typing}
   \label{sec:types}
+  \subsection{Types}
   % To define type contexts and a typing relation, we need to know
@@
   \subsection{Type Contexts}
-  Equipped with a notion of types, we define a type context as a
-  function that maps variables to types of the same sort.
-
-  As we want our framework to support dependent types, we allow a type returned
-  by the type context to use all variables bound previously in the type context.
-  This is achieved by the \ACode{drop-∈} function, which cuts off all
-  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
-  Contexts \ACode{Γ} can be extended by a type \ACode{t} via \ACode{t \ADef{∷ᵗ} Γ}.
-
+  \begin{figure}
+    \centering
     % \LibCode\KContextHelper
-  \noindent
     \begin{minipage}{1.0\linewidth}
-    \begin{minipage}{0.5\linewidth}
+      \begin{minipage}{0.57\linewidth}
         \LibCode\KContexts
-      \vspace{3mm}
+        \LibCode\KContextLookupI
+        \LibCode\KVariableTyping
       \end{minipage}
-    \begin{minipage}{0.5\linewidth}
+      \begin{minipage}{0.43\linewidth}
         \LibCode\KContextExt
+        \LibCode\KContextLookupII
+        \vspace{10.1mm}
       \end{minipage}
     \end{minipage}
-
-  To use a type returned from a type context, it needs to be
-  weakened appropriately through the \ACode{\ADef{wk-telescope}} function:
-
-  \noindent
-  \begin{minipage}[t]{1.0\linewidth}
-    \begin{minipage}[t]{0.45\linewidth}
-      \LibCode\KContextLookupII
-    \end{minipage}
-    \begin{minipage}[t]{0.55\linewidth}
-      \LibCode\KContextLookupI
-    \end{minipage}
-  \end{minipage}
-  % \LibCode\KContextLookup
-
-  We then define a variable typing as followed:
-  \LibCode\KVariableTyping
-
-  \section{Typing}
-  \label{sec:typing}
-  Now that we have a notion of types and type context, we are ready to define
+    \caption{Type Contexts}
+    \label{fig:type-contexts}
+  \end{figure}
+  Equipped with a notion of types, we are ready to define type contexts as shown
+  in Figure~\ref{fig:type-contexts}.
+  The type \ACode{\ADef{Ctx} S} describes type contexts with bindings for variables from \ACode{S}
+  as functions that map variables to types of the same sort.
+  As we want our framework to support dependent types, we allow a type
+  in the context to use all variables bound previously in the context.
+  This is achieved with \ACode{\ADef{drop-∈} x S}, which cuts off all
+  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
+  The \ACode{\ADef{\_∷ᵗ\_}} function extends a context with a new binding.
+  The \ACode{\ADef{lookup}} function looks up the type of a variable in a context
+  and weakens it appropriately.
+  Finally, a variable typing \ACode{Γ \ADef{∋} x \ADef{∶} t} is defined in terms of \ACode{\ADef{lookup}}.
+
+  \subsection{Typing}
+  \label{sec:types:typing}
+  Now that we have a notion of types and type contexts, we are ready to define
   the multi-sorted typing relation for System F, which describes both

### 2024-03-14 paper/paper.tex
Update.

@@
   in Figure~\ref{fig:type-contexts}.
-  The type \ACode{\ADef{Ctx} S} describes type contexts with bindings for variables from \ACode{S}
+  \ACode{\ADef{Ctx} S} describes type contexts with bindings for variables from \ACode{S}
   as functions that map variables to types of the same sort.

### 2024-03-14 paper/paper.tex
Update.

@@
   and weakens it appropriately.
-  Finally, a variable typing \ACode{Γ \ADef{∋} x \ADef{∶} t} is defined in terms of \ACode{\ADef{lookup}}.
+  Finally, a variable typing \ACode{Γ \ADef{∋} x \ADef{∶} t}
+  states that \ACode{\ADef{lookup} Γ x} is \ACode{t}.
 
@@
   \item
-    the variable typing constructor \ACode{\ACon{⊢`}}, which covers both expression- and
+    the variable rule \ACode{\ACon{⊢`}}, which covers both expression- and
     type-variables, analogously to the variable term constructor;
   \item
-    the typing \ACode{\ACon{⊢λ}} of a lambda abstraction, which
+    the lambda rule \ACode{\ACon{⊢λ}}, which
     weakens the codomain type \ACode{t₂}.
@@
   \item
-    the kinding constructor \ACode{\ACon{⊢τ}} states that all types
+    the kinding rule \ACode{\ACon{⊢τ}} states that all types
     have kind \ACode{\ACon{★}}. This is sufficient for System F as types
@@
   \LibCode\KTKit
-  The operations are completely analogous to those of a \ACode{\ACon{Kit}}, but talk
-  about typing instead of terms.
+  The fields of a \ACode{\ACon{TKit}} are completely analogous to
+  those of a \ACode{\ACon{Kit}}, but talk about typing instead of terms.
   Building on the fields of a \ACode{\ACon{TKit}}, we define further operations:

### 2024-03-15 paper/paper.tex
Update.

@@
   \subsection{Type Contexts}
-  \begin{figure}
-    \centering
-    % \LibCode\KContextHelper
-    \begin{minipage}{1.0\linewidth}
-      \begin{minipage}{0.57\linewidth}
+  Equipped with a notion of types, we are ready to define type contexts.
+  As we want our framework to support dependent types, we allow a type
+  in the context to use all variables bound previously in the context:
   \LibCode\KContexts
-        \LibCode\KContextLookupI
+
+  When looking up the type of a variable, we need to weaken it for each
+  binding that comes after the variable:
+  \LibCode\KContextLookup
+
+  % Finally, we define a variable typing:
+  Finally, a variable typing \ACode{Γ \ADef{∋} x \ADef{∶} t}
+  states that looking up \ACode{x} in \ACode{Γ} yields \ACode{t}:
   \LibCode\KVariableTyping
-      \end{minipage}
-      \begin{minipage}{0.43\linewidth}
-        \LibCode\KContextExt
-        \LibCode\KContextLookupII
-        \vspace{10.1mm}
-      \end{minipage}
-    \end{minipage}
-    \caption{Type Contexts}
-    \label{fig:type-contexts}
-  \end{figure}
-  Equipped with a notion of types, we are ready to define type contexts as shown
-  in Figure~\ref{fig:type-contexts}.
-  \ACode{\ADef{Ctx} S} describes type contexts with bindings for variables from \ACode{S}
-  as functions that map variables to types of the same sort.
-  As we want our framework to support dependent types, we allow a type
-  in the context to use all variables bound previously in the context.
-  This is achieved with \ACode{\ADef{drop-∈} x S}, which cuts off all
-  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
-  The \ACode{\ADef{\_∷ᵗ\_}} function extends a context with a new binding.
-  The \ACode{\ADef{lookup}} function looks up the type of a variable in a context
-  and weakens it appropriately.
-  Finally, a variable typing \ACode{Γ \ADef{∋} x \ADef{∶} t}
-  states that \ACode{\ADef{lookup} Γ x} is \ACode{t}.
 

### 2024-03-15 paper/paper.tex
Update.

@@
   \LibCode\KTKit
-  The fields of a \ACode{\ACon{TKit}} are completely analogous to
-  those of a \ACode{\ACon{Kit}}, but talk about typing instead of terms.
-  Building on the fields of a \ACode{\ACon{TKit}}, we define further operations:
+  The first field abstracts over variable and term typing.
+  The other fields express typings for the fields of a \ACode{\ACon{Kit}}.
+  Building on the fields of a \ACode{\ACon{TKit}}, we define map
+  typing and type preservation for map lifting and the singleton map
+  in the record module of \ACode{\ACon{TKit}}:
   \begin{itemize}
@@
     \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒ₖ} Γ₂} states that \ACode{ϕ} is a map that takes
-    terms in \ACode{Γ₁} to terms in \ACode{Γ₂}.
+    terms from \ACode{Γ₁} to terms in \ACode{Γ₂}.
   \item 
@@
   the term \ACode{⊢e \AField{⊢⋯} ⊢ϕ} is a typing for \ACode{e \AField{⋯} ϕ}.
-  The large amount of kit-parameters does not impose any restriction,
-  as both our \ACode{\ACon{Kit}}s support the \ACode{\ACon{WkKit}}
-  extension and can be composed arbitrarily. They only make the previously
-  built machinery available to the proof of \ACode{\AField{\_⊢⋯\_}}.
-
-  Given a \ACode{\ACon{TTraversal}}, we can easily instantiate the
-  \ACode{\ACon{TKits}} for variables and terms:
+
+  As before, we define the \ACode{\ACon{TKit}} instances in the record module
+  of \ACode{\ACon{TTraversal}}:
   \LibCode\KTypingInstances
+
+  The large amount of kit-parameters of \ACode{\AField{\_⊢⋯\_}} does
+  not impose any restriction, as both our \ACode{\ACon{Kit}}s support
+  the \ACode{\ACon{WkKit}} extension and can be composed
+  arbitrarily. Agda's instance resolution allows us to easily
+  instantiate a concrete substitution-preserves-typing lemma:
+  \LibCode\KTPresS
 

### 2024-03-15 paper/paper.tex
Update.

@@
   space. The omitted proofs can be found in our supplementary material,
-  which includes the simplified framework (504 lines of code) and the
-  System F formalization (228 lines of code).
-  The full framework including the case studies is available on Github: \textcolor{blue}{\url{https://github.com/m0rphism/kitty}}
+  which includes the simplified framework (368 lines of code) and the
+  System F formalization (221 lines of code).
+  The full framework including the case studies is available on Github:
+  \textcolor{blue}{\url{https://github.com/m0rphism/kitty}}
   % The full framework is available online.\footnote{See supplementary material. In the final version, this text will be replaced by a link to github.}

### 2024-03-16 paper/paper.tex
Update.

@@
     The lambda typing constructor \ACode{\ACon{⊢λ}} weakens the type \ACode{t₂} to
-    shield it from expression-substitution, requiring us to use the
-    \ACode{\AField{⋯-↑-wk}} lemma.
+    shield it from expression-substitution, requiring us to use
+    \ACode{\AField{⋯-↑-wk}} to move the map under the weakening.
   \item
     The type application constructor \ACode{\ACon{⊢∙}} substitutes \ACode{t₁} into \ACode{t₂},
-    requiring us to use \ACode{\AField{dist-↑-⦅⦆}}.
+    requiring us to use \ACode{\AField{dist-↑-⦅⦆}} to move the map
+    under the singleton substitution.
   \end{itemize}
@@
   \label{sec:reflection}
-  We used Agda's reflection mechanism to derive
+  We use Agda's reflection mechanism to derive
   instantiations related to all structures for untyped substitition,
@@
   To gain insight into the class of object languages supported by our
-  reflection algorithm, we have formalized a class of object languages
-  using a generic programming technique and instantiated the structures
-  generically. Informally, this class contains all object languages described
-  via multi-sorted syntax, which
+  reflection algorithm, we have instantiated the structures for a generic syntax
+  similar to the one in Allais et al.\cite{DBLP:journals/pacmpl/AllaisA0MM18}.
+  Our reflection algorithm derives proofs with the same
+  structure as the generic proofs, giving high confidence that it covers
+  the same class of languages.
+
+  Informally, all objects languages with multi-sorted syntax are supported that
   \begin{enumerate}
   \item have a variable constructor of type
-    \ACode{S \ACon{∋} s \AKw{→} S \ACon{⊢} s};
-  \item use subterms only directly (e.g. not in lists);
-  \item only extend the scope-context of subterms, but never modify it.
+    \ACode{\AKw{∀} \{S s\} \AKw{→} S \ACon{∋} s \AKw{→} S \ACon{⊢} s};
+  \item use subterms only directly (e.g. not in lists); and
+  \item only extend the scope-context of subterms, but never modify it otherwise.
   \end{enumerate}
 
-  Our reflection algorithm derives proofs with exactly the same
-  structure as the generic proof, giving high confidence that it covers
-  the same class of languages.
-
-  The formal definition of the generic class can be found in
+  % Appendix~\ref{sec:generic-class} and the actual implementation.
+
+  We believe restriction 2 is purely technical and can be lifted
+  with a more sophisticated reflection algorithm. In the current system,
+  this restriction can be worked around by inlining the data structure
+  constructors into the syntax definition as terms of a new sort.
+
+  The formal definition of the generic syntax can be found in
   the supplementary material.
-  % Appendix~\ref{sec:generic-class} and the actual implementation.
-
-  We believe restriction 2 is purely technical and can be lifted
-  with a more sophisticated reflection algorithm, as parameterized
-  datastructures can also be inlined into the syntax definition as terms
-  of a new sort.
 

### 2024-03-18 paper/paper.tex
Update.

@@
     For lambda calculus with dependent function types, the framework
-    just works out of the box for both deriving and instantiating the
-    \ACode{\ACon{TTraversal}}.
-    Furthermore, the \ACode{\ACon{CKit}} abstraction turned out
-    to be useful to prove confluence, allowing to unify certain
-    lemmas, which otherwise need to be proved for both renamings and
+    works out of the box for both deriving untyped substitution
+    and instantiating the \ACode{\ACon{TTraversal}}.
+    In the proof of confluence, the \ACode{\ACon{CKit}} abstraction
+    allowed us to unify lemmas about the reduction of renamings and
     substitutions.
-    % e.g. that for two terms \ACode{t} and \ACode{t'}
-    % with \ACode{t \ACon{↪} t'} it follows that \ACode{(t \AField{⋯} ϕ)
-    % \ACon{↪} (t' \AField{⋯} ϕ)} for any map \ACode{ϕ}.
   \item
@@
     subtyping constraints and how to deal with the fact that
-    substitution preserves typing is not generally true, as type variables
+    substitution-preserves-typing is not generally true, as type variables
     have subtyping bounds that need to be respected.
@@
     untyped substitution and define typing contexts and type preservation
-    lemmas by hand, we found an encoding, which still allows us to use the
-    the \ACode{\ACon{TKit}} abstraction.
+    lemmas by hand, we instead used an encoding that allows us to use the
+    the \ACode{\ACon{TKit}} abstraction directly.
     Instead of binding a type variable with a subtyping constraint
@@
     cannot be accesses by the user.
-    With that encoding substitution preserves typing is generally true
-    again, enabling the use of the framework's typing contexts and
-    \ACode{\ACon{TKit}} machinery.
-    When replacing a type variable \ACode{α <: t} with a type \ACode{t'},
-    which is not a subtype of \ACode{t}, then the resulting term is still
-    well-typed, but in a context with an unsatisfiable constraint.
+    With this encoding, substitution-preserves-typing is generally true
+    again: replacing a type variable \ACode{α <: t} with a type \ACode{t'},
+    which is not a subtype of \ACode{t}, results in a term that is still
+    well-typed, but in a context with an unsatisfiable constraint \ACode{t' <: t}.
   \item
@@
   broader
-  picture\cite{DBLP:conf/cpp/StarkSK19,DBLP:journals/pacmpl/AllaisA0MM18}. In
-  particular, we discuss abstractions for renamings
-  and substitutions, comparable extrinsically typed approaches, but also
-  intrinsically typed approaches that could be used as a basis for
-  multi-sorted syntax, which as mentioned before is intrinsically typed
-  in disguise.
+  picture\cite{DBLP:conf/cpp/StarkSK19,DBLP:journals/pacmpl/AllaisA0MM18}.
+  % In
+  % particular, we discuss abstractions for renamings
+  % and substitutions, comparable extrinsically typed approaches, but also
+  % intrinsically typed approaches that could be used as a basis for
+  % multi-sorted syntax, which as mentioned before is intrinsically typed
+  % in disguise.
+
+  \subsection{Variable Binding}
+  There is a plethora of different methods for representing
+  variable binders in the literature:
+  de Bruijn\cite{scholar:de1972lambda},
+  co de Bruijn\cite{DBLP:journals/corr/abs-1807-04085},
+  locally nameless\cite{DBLP:journals/jar/Chargueraud12},
+  locally named\cite{DBLP:conf/tlca/McKinnaP93},
+  higher order abstract syntax\cite{DBLP:conf/pldi/PfenningE88} and
+  its parametric variant\cite{DBLP:conf/icfp/Chlipala08},
+  nominal logic\cite{DBLP:journals/tcs/UrbanPG04},
+  shifted names\cite{scholar:dolansyntax},
+  nameless painless\cite{DBLP:conf/icfp/Pouillard11}, and
+  scope graphs\cite{DBLP:journals/pacmpl/AntwerpenPRV18}.
+  Many of these representations have been studied in solutions to the POPLMark
+  challenges\cite{DBLP:conf/tphol/AydemirBFFPSVWWZ05, DBLP:journals/jfp/AbelAHPMSS19}.
 
@@
   In all three cases kits are formulated for intrinsic typing
-  and scenarios with polymorphism are not yet considered.
+  and scenarios with polymorphism are not considered.
 
@@
   the individual substitutions by putting them in a vector.
-
-  Both works do not provide machinery for type preservation and do not
-  unify renaming and substitution (although the latter is listed as
-  future work). We believe their amazing work of creating a decision
-  procedure for vectorized substitutions should also translate to our
-  setting with multi-sorted substitutions.
+  We believe their great work of creating a decision procedure for
+  vectorized substitutions should also translate to our setting with
+  multi-sorted substitutions.
+
+  \emph{Needle and Knot}\cite{DBLP:conf/esop/KeuchelWS16} is a code
+  generator for unscoped syntax with de Bruijn indices. They generate
+  substitution and interaction lemmas for single-pointed substitution
+  for languages with multiple variable sorts and binders that bind
+  lists of variables.
+
+  All of the above work does not provide machinery to model typing
+  and type preservation and does not unify renaming and substitution and
+  their compositions.
+  Hence, type preservation needs to be modeled manually and
+  individually for renamings and substitutions.
 
@@
   They show how to use their framework for both intrinsic and
-  extrinsic typing, but have not yet explored the realm of polymorphism.
-
-  This work is orthogonal, but compatible to our approach.
-  We can instantiate their system using the hidden intrinsic type
-  system of our multi-sorted syntax, but unfortunately then their
-  semantics abstraction would only work with terms, which are untyped
-  with respect to the actual extrinsic type system.
-  We think it is worthwile to explore if their semantics abstraction
-  can also be generalized to multi-sorted typing relations.
+  extrinsic typing, but are missing a story for polymorphism.
+
+  With only a slight modification to their framework, we can instantiate it
+  for multi-sorted syntax, enabling the definition of polymorphic languages.
+  However, as the intrinsic typing is then
+  used to ensure syntactic sorts, the semantic abstractions talk about untyped
+  terms, so typing and type preservation lemmas have to be modeled entirely manually.
+  We believe it would be worthwhile to explore how their semantics
+  abstractions can be lifted to typing relations similar to how our typing kits lift
+  regular kits from terms to typing relations.
 
@@
   sorts, dependencies between sorts, and rules for quantification.
-
   While pure type systems may seem very similar to multi-sorted
@@
     In pure type systems, sorts are universe types, e.g.\ like
-    \ACode{Set} and \ACode{Set₁} in Agda.
+    \ACode{Set} in Agda or \ACode{Prop} in Coq.
     Terms which have different sorts as types, do still belong to the
@@
 
-  We can use our framework to model pure type systems, but the sorts of pure type systems
-  do not correspond to the sorts of the multi-sorted syntax. The multi-sorted syntax
-  has just a single sort that describes all terms, as it is common for
-  dependently typed languages.
+  We can model pure type systems as a multi-sorted syntax with a single sort,
+  where the sorts of the pure type system occur as terms representing universe types.
+
+  % We can use our framework to model pure type systems, but the sorts of pure type systems
+  % do not correspond to the sorts of the multi-sorted syntax. The multi-sorted syntax
+  % has just a single sort that describes all terms, as it is common for
+  % dependently typed languages.
 
@@
   \label{sec:conclusion}
-  We have presented a framework for deriving untyped substitution and its properties, and
-  and abstraction for proving type preservation of substitution with a single lemma.
-
-  Using extrinsic typing allowed us to unify the renamings and
-  substitutions for variables of arbitrary many syntactic categories.
-
-  Furthermore, extrinsic typing allowed us treat untyped substitution
-  and its properties separately from typing and type preservation.
-  This has the benefit that the former can be derived automatically for virtually
-  any object language, while the latter is more domain specific and
-  can be handled independently.
-
-  We believe our framework is a viable alternative to intrinsic
-  typing, as the duplication is automated away.
-
-  We believe our framework is particularly useful for polymorphic languages,
-  where intrinsic typing requires necessarily to deal with multiple substitutions
-  and their interactions. Furthermore, in more complex scenarios, intrinsic typing
-  can easily lead to subst-hell, which is avoided with extrinsic typing.
-
-  We belive the core limitation of our framework are the typing kits, which
-  require a classical ternary typing relation. However, this problem is hard
-  to avoid as typing relations can become arbitrarily complicated.
-  To adapt the framework to more complicated type systems, only the typing
-  kits and typing traversals need to be extended, which we have successfully
-  done for linear typing ala Wood and Atkey\cite{DBLP:journals/corr/abs-2005-02247} and
+  % We have presented a framework for deriving untyped substitution and its properties, and
+  % and abstraction for proving type preservation of substitution with a single lemma.
+
+  % Using extrinsic typing allowed us to unify the renamings and
+  % substitutions for variables of arbitrary many syntactic categories.
+
+  % Furthermore, extrinsic typing allowed us treat untyped substitution
+  % and its properties separately from typing and type preservation.
+  % This has the benefit that the former can be derived automatically for virtually
+  % any object language, while the latter is more domain specific and
+  % can be handled independently.
+
+  % We believe our framework is a viable alternative to intrinsic
+  % typing, as the duplication is automated away.
+
+  % We believe our framework is particularly useful for polymorphic languages,
+  % where intrinsic typing requires necessarily to deal with multiple substitutions
+  % and their interactions. Furthermore, in more complex scenarios, intrinsic typing
+  % can easily lead to subst-hell, which is avoided with extrinsic typing.
+
+  % We belive the core limitation of our framework are the typing kits, which
+  % require a classical ternary typing relation. However, this problem is hard
+  % to avoid as typing relations can become arbitrarily complicated.
+  % To adapt the framework to more complicated type systems, only the typing
+  % kits and typing traversals need to be extended, which we have successfully
+  % done for linear typing ala Wood and Atkey\cite{DBLP:journals/corr/abs-2005-02247} and
+  % general substructural typing ala Licata et al\cite{DBLP:conf/rta/LicataSR17}.
+
+  % - Main use case: polymorphism / multiple syntactic categories / dependent types
+  % - Limitations and Flexibility with only extending typing kits
+  % - Comparison with intrinsic typing
+  %   - Deriving untyped substitution is on par with intrinsic
+  %   - Subst-Hell / Interaction lemmas unavoidable
+
+  While intrinsic typing allows to unify definitions with their type
+  preservation proofs, extrinsic typing allows to unify
+  substitutions across different syntactic categories, as we have demonstrated.
+  We believe this makes extrinsic typing more suitable for polymorphic languages:
+  with extrinsic typing, the definitions and lemmas for untyped
+  substitutions can be automated for virtually any object language,
+  whereas with intrinsic typing, it seems impossible to unify
+  substitutions across syntactic categories, because an intrinsically
+  typed multi-sorted syntax would require to index terms with terms,
+  which is the same problem that occurs with giving intrinsic types to a dependently typed language.
+
+  Our framework is able to automatically derive definitions and lemmas
+  about untyped substitution, and provides an abstraction for proving
+  type preservation of substitution with a single lemma (\ACode{\AField{\_⊢⋯\_}}).
+
+  Similarly to frameworks with intrinsic typing, the main limitation in our work is the shape of
+  typing relations: our typing kits can only model classical ternary typing relations.
+  To adapt our framework to more complicated typing relations, the machinery for untyped substitution
+  can be reused, but the typing kits and typing traversals need to be modified.
+  We found that this works surprisingly well in practice,
+  where we have already made extended typing kits to support
+  linear typing ala Wood and Atkey\cite{DBLP:journals/corr/abs-2005-02247} and
   general substructural typing ala Licata et al\cite{DBLP:conf/rta/LicataSR17}.
+  In both cases, the typing relation has a fourth component which models usage restrictions
+  on the type context.
+
+  % In our opinion, our framework makes a case for extrinsic typing for
+  % polymorphic, languages as the
 
@@
   \clearpage
-  \appendix
-  \onecolumn
-
-  \section{Omitted Proofs}
-  \label{sec:proofs}
-
-  \subsection{Proofs for Section~\ref{sec:maps}}
-  \LibCode\KIdLift
-  \LibCode\KIdLiftProof
-
-  % \AppCode\FTraversalId
-  % \AppCode\FTraversalIdProofInteresting
-  % \LibCode\FTraversalIdProofRest
-
-  \subsection{Proofs for Section~\ref{sec:composition}}
-  \LibCode\KCKitAp
-  \LibCode\KCKitApProof
-
-  \LibCode\KDistLiftCompose
-  \LibCode\KDistLiftComposeProof
-
-  \LibCode\KCommLiftWeaken
-  \LibCode\KCommLiftWeakenProof
-
-  \LibCode\KCommLiftWeakenTraverse
-  \LibCode\KCommLiftWeakenTraverseProof
-
-  \LibCode\KWeakenCancelsSingle
-  \LibCode\KWeakenCancelsSingleProof
-
-  \LibCode\KWeakenCancelsSingleTraverse
-  \LibCode\KWeakenCancelsSingleTraverseProof
-
-  \LibCode\KDistLiftSingle
-  \LibCode\KDistLiftSingleProof
-
-  \LibCode\KDistLiftSingleTraverse
-  \LibCode\KDistLiftSingleTraverseProof
-
-  % \AppCode\FFusion
-  % \AppCode\FFusionProofInteresting
-  % \AppCode\FFusionProofRest
-
-  \subsection{Proofs for Section~\ref{sec:typing}}
-  \LibCode\KLiftTyping
-  \LibCode\KLiftTypingProof
-
-  \LibCode\KSingleTyping
-  \LibCode\KSingleTypingProof
-
-  % \AppCode\FSubjectReduction
-  % \AppCode\FSubjectReductionProofInteresting
-
-  % \section{Generic Class of Object Languages}
-  % \label{sec:generic-class}
-
-  % \section{System F with Reflection}
-  % \label{sec:reflection-example}
-
-  % \section{Map Equality Without Functional Extensionality}
-  % \label{sec:fun-ext}
-
-  % \section{Features of the Actual Implementation}
-  % \label{sec:features}
+  % \appendix
+  % \onecolumn
+
+  % \section{Omitted Proofs}
+  % \label{sec:proofs}
+
+  % \subsection{Proofs for Section~\ref{sec:maps}}
+  % \LibCode\KIdLift
+  % \LibCode\KIdLiftProof
+
+  % % \AppCode\FTraversalId
+  % % \AppCode\FTraversalIdProofInteresting
+  % % \LibCode\FTraversalIdProofRest
+
+  % \subsection{Proofs for Section~\ref{sec:composition}}
+  % \LibCode\KCKitAp
+  % \LibCode\KCKitApProof
+
+  % \LibCode\KDistLiftCompose
+  % \LibCode\KDistLiftComposeProof
+
+  % \LibCode\KCommLiftWeaken
+  % \LibCode\KCommLiftWeakenProof
+
+  % \LibCode\KCommLiftWeakenTraverse
+  % \LibCode\KCommLiftWeakenTraverseProof
+
+  % \LibCode\KWeakenCancelsSingle
+  % \LibCode\KWeakenCancelsSingleProof
+
+  % \LibCode\KWeakenCancelsSingleTraverse
+  % \LibCode\KWeakenCancelsSingleTraverseProof
+
+  % \LibCode\KDistLiftSingle
+  % \LibCode\KDistLiftSingleProof
+
+  % \LibCode\KDistLiftSingleTraverse
+  % \LibCode\KDistLiftSingleTraverseProof
+
+  % % \AppCode\FFusion
+  % % \AppCode\FFusionProofInteresting
+  % % \AppCode\FFusionProofRest
+
+  % \subsection{Proofs for Section~\ref{sec:typing}}
+  % \LibCode\KLiftTyping
+  % \LibCode\KLiftTypingProof
+
+  % \LibCode\KSingleTyping
+  % \LibCode\KSingleTypingProof
+
+  % % \AppCode\FSubjectReduction
+  % % \AppCode\FSubjectReductionProofInteresting
+
+  % % \section{Generic Class of Object Languages}
+  % % \label{sec:generic-class}
+
+  % % \section{System F with Reflection}
+  % % \label{sec:reflection-example}
+
+  % % \section{Map Equality Without Functional Extensionality}
+  % % \label{sec:fun-ext}
+
+  % % \section{Features of the Actual Implementation}
+  % % \label{sec:features}
 

### 2024-03-18 paper/paper.tex
Update.

@@
   Code of the framework is displayed in gray boxes.
-  Code of examples is written in yellow boxes.
-  Code of the System F formalization is written without boxes. The
+  Code of examples is displayed in yellow boxes.
+  Code of the System F formalization is displayed without boxes. The
   latter is the only code a user of our framework has to write.
 
-  In this paper we present a simplified version of the actual
+  In this paper, we present a simplified version of the actual
   framework, focusing on the core concepts. We present all necessary
@@
   \item
-    a novel abstraction for composition and it's metatheory, unifying
+    a novel abstraction for composition and its metatheory, unifying
     the four compositions between renamings and substitutions;
@@
     type-in-type and expression-in-expression substitution, but also type-in-expression substitution.
-    Consequently, interaction lemmas between those substitutions are required.
+    Consequently, interaction lemmas between the substitutions are required.
     % In general, this leads to a combinatory explosion of
@@
     use the meta-variable \ACode{ϕ} for it.
-  The operation \ACode{ϕ \& x} applies a map to a variable.
-  The operation \ACode{ϕ ↑ s} lifts a map under a binder of sort \ACode{s}.
-  The operation \ACode{⦅ x/t ⦆} constructs a singleton map that replaces
+  The operation \ACode{ϕ \ADef{\&} x} applies a map to a variable.
+  The operation \ACode{ϕ \ADef{↑} s} lifts a map under a binder of sort \ACode{s}.
+  The operation \ACode{\ADef{⦅} x/t \ADef{⦆}} constructs a singleton map that replaces
   \ACode{\ACon{zero}} with \ACode{x/t} and decreases all other variables by one.
-  The \ACode{\AField{weaken s}} map increases all variables by one.
-  The \ACode{ϕ₁ $\sim$ ϕ₂} type expresses extensional equality of maps.
-    For simplicity, we postulate functional extensionality \ACode{$\sim$-ext}.\footnote{
+  The \ACode{\ADef{weaken}} map increases all variables by one.
+  The \ACode{ϕ₁ \ADef{$\sim$} ϕ₂} type expresses extensional equality of maps.
+    For simplicity, we postulate functional extensionality \ACode{\AField{$\sim$-ext}}.\footnote{
       The actual implementation does \emph{not} use any postulates.
@@
     }
-  There is an identity map \ACode{id}.
-  The lemma \ACode{id↑-id} states that a lifted identity map is again an identity map.
+  There is an identity map \ACode{\ADef{id}}.
+  The lemma \ACode{\ADef{id↑-id}} states that a lifted identity map is again an identity map.
 
@@
   \item
-     we write \ACode{(S \ADef{∋/⊢[} K \ADef{]} s)} for the \ACode{(S ∋/⊢ s)}
+    we write \ACode{S₁ \ADef{–[} K \ADef{]→} S₂} for the \ACode{S₁ \ADef{→ₖ} S₂}
+    of some specific \ACode{\ACon{Kit}} \ACode{K}.
+  \item
+    we write \ACode{S \ADef{∋/⊢[} K \ADef{]} s} for the \ACode{S ∋/⊢ s}
     of some specific \ACode{\ACon{Kit}} \ACode{K}; and
-   \item
-     we write \ACode{(S₁ \ADef{–[} K \ADef{]→} S₂)} for the \ACode{(S₁ \ADef{→ₖ} S₂)}
-     of some specific \ACode{\ACon{Kit}} \ACode{K}.
   \end{itemize}
@@
   The operation of applying a map to a term depends on the concrete
-  structure of a term, so we define a structure for it:
+  object language. It is captured by the following structure:
   \LibCode\KTraversal
@@
     In cases where the operation needs to go under a binder, like
-    \ACode{(\ACon{λx} e) \ADef{⋯} ϕ}, we lift the map using \AField{\_↑\_} before
-    we apply it to the subterm to account for the bound variable.
+    \ACode{(\ACon{λx} e) \ADef{⋯} ϕ}, we lift the map using
+    \AField{\_↑\_} to account for the bound variable before
+    we apply it to the subterm.
   \end{itemize}
@@
   describes the operations necessary for definining the composition of a
-  map from \ACode{K₁} with a map from \ACode{K₂} that results in a map from
-  \ACode{K₁⊔K₂}:
+  \ACode{K₁}-map with a \ACode{K₂}-map that results in a \ACode{K₁⊔K₂}-map:
   \LibCode\KCKit
@@
   our object language, and in return the framework defines the concrete \ACode{\ADef{CKit}}
-  instances for us. Hence, we define them in the record module of \ACode{\ADef{CTraversal}}:
+  instances for us. Hence, we define the \ACode{\ADef{CKit}} instances in the record module of \ACode{\ADef{CTraversal}}:
   \LibCode\KCKitInstances
@@
     There is a strong similarity to the instantiation of
-    map traversal \ACode{\AField{\_⋯\_}}. Where \ACode{\AField{\_⋯\_}}
-    used \ACode{\AField{`/id}} or \ACode{\AField{\_↑\_}}, we now use their
+    map traversal \ACode{\ADef{\_⋯\_}}: where \ACode{\ADef{\_⋯\_}}
+    used \ACode{\AField{`/id}} or \ACode{\AField{\_↑\_}}, our \ACode{\ADef{\_⊢⋯\_}} uses their
     preservation lemmas \ACode{\AField{⊢`/id}} or \ACode{\AField{\_⊢↑\_}}.
@@
     isomorphic to the original formalization, as constraint variables
-    cannot be accesses by the user.
+    cannot be accessed by the user.
     With this encoding, substitution-preserves-typing is generally true
@@
   \item
-    Object languages with pattern matching can be encoded by adding
+    % Object languages with pattern matching can be encoded by adding
+    % the sorts of the variables bound by a pattern to the pattern sort
+    % \ACode{\ACon{𝕡}} itself:
+    % \begin{quote}
+    % \ACode{\ACon{𝕡} \AKw{:} \ACon{List (Sort Var)} \AKw{→} \ACon{Sort NoVar}} 
+    % \end{quote}
+    % This allows the clause of a \ACode{\ACon{match}} expression to
+    % read the bound variables from the pattern sort and make them
+    % available as free variables in its body expression.
+    Object languages with pattern matching can be modeled by adding
     the sorts of the variables bound by a pattern to the pattern sort
-    \ACode{\ACon{𝕡}} itself:
+    \ACode{\ACon{𝕡}} itself. A pattern matching clause \ACode{p~\ACon{⇒}~e} can
+    then be defined as
     \begin{quote}
-    \ACode{\ACon{𝕡} \AKw{:} \ACon{List (Sort Var)} \AKw{→} \ACon{Sort NoVar}} 
+    \ACode{\ACon{\_⇒\_} \AKw{:} S \ACon{⊢} \ACon{𝕡} S' \AKw{→} \AKw{(}S' \ADef{++} S\AKw{)} \ACon{⊢} \ACon{𝕖} \AKw{→} S \ACon{⊢} \ACon{𝕔}} 
     \end{quote}
-    This allows the clause of a \ACode{\ACon{match}} expression to
-    read the bound variables from the pattern sort and make them
-    available as free variables in its body expression.
+    where \ACode{S'} describes the variables bound by the pattern and \ACode{\ACon{𝕔}} is the sort of a clause.
   \end{itemize}
@@
   There is a plethora of different methods for representing
-  variable binders in the literature:
-  de Bruijn\cite{scholar:de1972lambda},
-  co de Bruijn\cite{DBLP:journals/corr/abs-1807-04085},
+  variable binders:
+  de Bruijn indices\cite{scholar:de1972lambda},
+  co de Bruijn indices\cite{DBLP:journals/corr/abs-1807-04085},
   locally nameless\cite{DBLP:journals/jar/Chargueraud12},
@@
   However, as the intrinsic typing is then
-  used to ensure syntactic sorts, the semantic abstractions talk about untyped
+  used to describe syntactic categories, the semantic abstractions talk about untyped
   terms, so typing and type preservation lemmas have to be modeled entirely manually.
@@
 
+  We have presented an Agda framework, which automatically derives definitions and lemmas
+  for untyped substitution, and provides an abstraction for proving
+  type preservation of renaming and substitution for all syntactic
+  categories with a single lemma (\ACode{\AField{\_⊢⋯\_}}).
+
+  Compared to many extrinsically typed approaches, our framework also models
+  typing and type preservation. Compared to many intrinsically typed approaches, our
+  framework gracefully extends to polymorphic scenarios.
+
+  The main limitation of our framework is the shape of typing relations, similarly
+  as it is the case with approaches based on intrinsic typing: we can
+  only model classical ternary typing relations.
+  To adapt our framework to more complicated typing relations, the
+  machinery for untyped substitution can be reused, but the
+  abstractions related to typing need to be modified.
+  We found that this works surprisingly well in practice,
+  where we have already made extended typing abstractions that support
+  linear typing ala Wood and Atkey\cite{DBLP:journals/corr/abs-2005-02247} and
+  general substructural typing ala Licata et al\cite{DBLP:conf/rta/LicataSR17}.
+  In both cases, the typing relation is extended with a fourth
+  component that models usage restrictions on the type context.
+
   While intrinsic typing allows to unify definitions with their type
@@
   substitutions across different syntactic categories, as we have demonstrated.
-  We believe this makes extrinsic typing more suitable for polymorphic languages:
-  with extrinsic typing, the definitions and lemmas for untyped
-  substitutions can be automated for virtually any object language,
-  whereas with intrinsic typing, it seems impossible to unify
-  substitutions across syntactic categories, because an intrinsically
-  typed multi-sorted syntax would require to index terms with terms,
-  which is the same problem that occurs with giving intrinsic types to a dependently typed language.
-
-  Our framework is able to automatically derive definitions and lemmas
-  about untyped substitution, and provides an abstraction for proving
-  type preservation of substitution with a single lemma (\ACode{\AField{\_⊢⋯\_}}).
-
-  Similarly to frameworks with intrinsic typing, the main limitation in our work is the shape of
-  typing relations: our typing kits can only model classical ternary typing relations.
-  To adapt our framework to more complicated typing relations, the machinery for untyped substitution
-  can be reused, but the typing kits and typing traversals need to be modified.
-  We found that this works surprisingly well in practice,
-  where we have already made extended typing kits to support
-  linear typing ala Wood and Atkey\cite{DBLP:journals/corr/abs-2005-02247} and
-  general substructural typing ala Licata et al\cite{DBLP:conf/rta/LicataSR17}.
-  In both cases, the typing relation has a fourth component which models usage restrictions
-  on the type context.
+  We believe this makes our framework particularly suited for polymorphic languages,
+  where the downside of extrinsic typing is automated away, and the benefits
+  of a unified substitution bears fruits.
+  % Furthermore, extrinsic typing allows us
+  % to avoid the ``subst-hell'', which typically appears in formalizations
+  % of logical relations for polymorphic languages.
+  % We believe this makes extrinsic typing more suitable for polymorphic languages:
+  % with extrinsic typing, the definitions and lemmas for untyped
+  % substitutions can be automated for virtually any object language,
+  % whereas with intrinsic typing, it seems impossible to unify
+  % substitutions across syntactic categories, because an intrinsically
+  % typed multi-sorted syntax would require to index terms with terms,
+  % which is the same problem that occurs with giving intrinsic types to a dependently typed language.
 

### 2024-03-18 paper/paper.tex
Update.

@@
   We believe this makes our framework particularly suited for polymorphic languages,
-  where the downside of extrinsic typing is automated away, and the benefits
-  of a unified substitution bears fruits.
+  where the downside of extrinsic typing is automated away, and where we have variables
+  across multiple syntactic categories, so the benefits of a unified
+  substitution bear fruits.
   % Furthermore, extrinsic typing allows us

### 2024-03-18 paper/paper.tex
Update.

@@
 
-%\nolinenumbers %uncomment to disable line numbering
+% \nolinenumbers %uncomment to disable line numbering
 
@@
   which includes the simplified framework (368 lines of code) and the
-  System F formalization (221 lines of code).
+  System F formalization (212 lines of code, of which 79 can be derived in the full framework).
   The full framework including the case studies is available on Github:

### 2024-03-18 paper/paper.tex
Update.

@@
   which includes the simplified framework (368 lines of code) and the
-  System F formalization (212 lines of code, of which 79 can be derived in the full framework).
+  System F formalization (212 lines of code, where 79 can be derived in the full framework).
   The full framework including the case studies is available on Github:

### 2024-03-18 paper/paper.tex
Update.

@@
   space. The omitted proofs can be found in our supplementary material,
-  which includes the simplified framework (368 lines of code) and the
-  System F formalization (212 lines of code, where 79 can be derived in the full framework).
+  which includes the simplified framework (365 lines of code) and the
+  System F formalization (212 lines of code, where 79 can be derived by using the full framework).
   The full framework including the case studies is available on Github:

### 2024-03-18 paper/agdatex.py
Split paper and presentation in own repos.

@@ -1,313 +1,0 @@
-#!/usr/bin/env python3
-
-import os
-import sys
-from pathlib import Path
-from argparse import ArgumentParser
-import tempfile
-import subprocess
-import shutil
-import json
-import hashlib
-
-# Parse Arguments
-
-ap = ArgumentParser(
-    prog="agdatex",
-    description="Extract LaTeX-Macros from .agda files guided by comment annotations",
-)
-
-ap.add_argument("-o", "--outputdir", metavar="PATH", default="latex",
-                help="Output directory for agda's LaTeX backend. Forwarded as --latex-dir to agda. Default: 'latex'.")
-
-ap.add_argument("-e", "--exportfile", metavar="PATH",
-                help="This file will \\input all generated .tex files. Both .tex and .sty are supported. Default: 'OUTPUTDIR/agda-generated.sty'.")
-
-ap.add_argument("-t", "--tempdir", metavar="PATH",
-                help="Temporary directory to copy the project root to. Default: fresh system-dependent temporary dir.")
-
-ap.add_argument("-k", "--keeptempdir", action='store_true',
-                help="Keep temporary directory for debugging.")
-
-ap.add_argument("-r", "--root", metavar="PATH",
-                help="Project root. Default: Search for .git directory.")
-
-ap.add_argument("-i", "--index", metavar="PATH",
-                help="Write the list of generated macros to this file.")
-
-ap.add_argument("-v", "--verbose", action='store_true',
-                     help="Turn on verbosity")
-
-ap.add_argument("-n", "--noop", action='store_true',
-                    help="No action: dry run")
-
-ap.add_argument ("-c", "--clear", action='store_true',
-                     help="Clear cashes to force rebuild all")
-
-ap.add_argument("sources", metavar="SRC_PATH", nargs="+",
-                help="Path to an annotated .agda-file.")
-
-args = ap.parse_args()
-
-
-# Resolve Project Root
-
-if args.root is not None:
-    root = Path(args.root)
-else:
-    # Search for .git directory.
-    root = Path(os.getcwd())
-    found_git_dir = False
-    while True:
-        for p in root.iterdir():
-            if p.is_dir() and p.name == ".git":
-                found_git_dir = True
-                break
-        if found_git_dir or len(root.parents) == 0:
-            break
-        root = root.parents[0]
[…]
